### 2009-02-19 paper/paper.tex
Added the paper to the repository.

%\documentclass[10pt,twocolumn]{article}
\documentclass[12pt, conference]{IEEEtran}
%\IEEEoverridecommandlockouts
%\documentclass{article}
%\documentclass[12pt]{IEEEtran}
%\documentclass{elsart5p}

% Make sure we know its a draft for now
%\usepackage{draftwatermark}
\usepackage{verbatim}
%\documentclass[12pt]{article}
\usepackage{epsfig}
\usepackage[hypertexnames=false,bookmarksopenlevel=1,bookmarksopen,bookmarksnumbered,colorlinks,plainpages=false,linktocpage]{hyperref}
%\usepackage{natbib}
%\pagestyle{empty}
%\psdraft
%\baselineskip=20pt %Sets line spacing to 1 unit
%\bibliographystyle{elsart-num-names}
\bibliographystyle{IEEEtran}
\usepackage{cite}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1:}]}{\end{trivlist}}

%\begin{frontmatter}
\begin{document}
\title{The Forensic Interchange Format - a file format for digital evidence}
\author{M. I. Cohen\footnote{scudette@gmail.com} and Bradley Schatz}
%\author{M. I. Cohen}
%\address{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
%\baselineskip=20pt %Sets line spacing to 1 unit
%\end{frontmatter}
%\thanks{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}%
\maketitle

\begin{abstract}
Forensics analysis requires the acquisition of management of many
different types of evidence material. Often the same evidence is
reviewed by several different tools. We propose an open, extensible
file format for storing of evidence and interchanging analysis results
among different tools. The Forensic Interchange Format (FIF) was
designed to be simple to implement, relying on the well supported Zip
specifications for bit level file access.
\end{abstract}

\section{Introduction}

\section{FIF Specifications}
The FIF specifications distinct between bit level specifications and
a high level schema. The bit level specifications refer to how the
files are to be actually represented on disk. The schema specifies a
stream based abstraction which is implemented using the bit level
specification.

\subsection{FIF bit level specification}
The FIF specifications only require the underlying storage mechanism
be capable of storing multiple named segments of data. The current
specification uses the Zip file format as the underlying storage
mechanism, however in future versions of the specifications other
storage layers may be also be defined.

The Zip file specifications are publicly available and widely
implemented. The Zip64 extensions are provided to enable Zip archives
to grow beyond a two gigabyte limit by using 64 bit
references. Although Zip64 also defines encryption and authentication
extensions, we do not use these in this specification due to the
restrictions imposed on their use. Rather, we specify stream based
encryption scheme for use in FIF files.

The basic structure of a Zip archive is shown in Figure
\ref{zip_structure}. For bit level details see the Zip file
Application notes \cite{zipspecs}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
  \caption{The basic structure of a Zip archive}
  \label{zip_structure}
  \end{center}
\end{figure}

As can be seen, the archive consists of a {\em Central Directory (CD)}
located at the end of the archive. The CD is a list of pointers to
individual {\em File header} structures located within the body of the
archive. Headers are then followed by the file data, after it has been
compressed by the appropriate compression method (as specified in the
header). Each archived file is optionally followed by a {\em Data
Descriptor} describing the length and CRC of the archived file. Using
the data descriptor field allows implementations to write archives
without needing to seek in the output file. This allows Zip files to
be written to pipes for example, sending an image over the network
using netcat or ssh.

\subsection{FIF Definitions}
The FIF specification defines a number of basic terms:
\begin{itemize}
\item {\em A FIF Volume} is a single ZIP archive containing {\em
segments}.

\item {\em A segment} is a single unit of data written to the Zip
archive as an archive file. The only properties of segments of
interest to the FIF specification are the segment name and the
timestamp.

\item {\em The segment name} is the name under which a segment is
stored within the archive. This consists of a directory part named
{\em a stream} and an {\em chunk name}. For example the segment named
``default/000001'' refers to the stream called ``default'' and has a
chunk name of ``000001''. Although the ZIP specification does not
specify the encoding of segment names, the FIF specification requires
all segment names to be encoded using UTF8. If stream name is
unspecified, implementations should open the stream named ``default''.

\item {\em Attributes} are a list of key/value pairs stored at each
stream's ``properties'' segment (e.g. ``stream/properties''). The
``properties'' file is a text file which stores attributes one on each
line. Attribute lines consist of the key, immediately followed by
``='' for text values or ``:'' for binary values. This is immediately
followed by the encoded value. Values containing text MUST be encoded
using UTF8, while values containing binary data MAY be encoded using
Base64. If keys are repeated, their relative order MUST be preserved
by implementations, but different keys may be presented in any order.

Attributes MUST be specified at the top level of each volume
(i.e. must have a segment named ``properties'')

%about the stream. A properties file also exists at the top
%level of the ZIP hierarchy to store global information such as the
%volumes UUID, the FIF version and other volumes which make up the

\item {\em A FIF file} is a collection of volumes with the same
UUID. The UUID is generated according to RFC4122 for each new FIF
file. As new volumes are created for the same file, the UUID is copied
to all related volumes. When a FIF file is opened, all its volumes are
loaded and merged in such a way that all the segments from all the
volumes are visible as if they were all found in the same volume.

\end{itemize}

\subsection{FIF Volumes and segment merging}
There are a number of aspects in the Zip file format specifications
which we do not utilize. For example, the Zip64 specification allows
for a multitude of encryption and hashing algorithms. Some of these
features however, seem to be patent encumbered. Since the FIF
specification aims to separate high level information such as hashes,
cryptography and signatures from the low level bit level
representation we implement those features at a higher level.

In many forensic applications, individual segments within the volume
need to be updated \cite{AFF}. These updates can cause the volume to
become fragmented as holes appear within the middle of the volume, and
updated segments get added to the end.

When we wish to add a new segment to an existing volume, the new
segment header SHOULD overwrite the volume central directory, and the
central directory should be rewritten including the new entry
immediately after the last file in the archive. If it is not possible
to overwrite the file, the new segment can be written at the end,
followed by an updated copy of the central directory (in that case the
old central directory remains in the file). This is illustrated in
Figure \ref{zip_structure}.

A segment within the archive is considered to be updated when a new
segment is added to the archive with the same name and a later
timestamp. Implementations MUST satisfy read requests for duplicate
segment names with those segments with the later timestamp, or if the
two timestamps are the same, the segment which appears at a later
offset in the volume. This effectively allows for a segment to be
updated without leading to internal volume fragmentation.

The fact that the older version of all updated segments remains intact
within each volume can be used for auditing, or to allow a rollback to
previous versions of the segment. The {\em frepack} tool is provided
to repack, split and manage FIF volumes by creating new volumes and
redistributing segments among them.

Note that we do not rely on Zip64's built-in support for splitting
archives at all (although implementations should be able to read such
files). Each volume is a complete and stand alone ZIP file. It is the
FIF implementation which considers the segments contained within it to
belong to a larger collection. The reason behind this decision is that
often it is needed to add new volumes to an existing set without
modifying the existing set at all.

Consider the typical usage scenario depicted in Figure \ref{usage}, of
a FIF file which contains a disk image. This volume is distributed to
two independent analysts, which wish to add new streams to the FIF
file (for example the result of their analysis). In this case they can
each create a new volume which extends to original volume and save
their analysis on this new volume. Now they only need to share this
new volume with other analysts who also have a copy of the old volume
to interchange their findings.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}
  \caption{A typical usage scenario. Both Alice and Bob receive a FIF
volume 1 but work independently. Rather than modifying Volume 1, they
each create another volume and save their results to the new
volume. They can now exchange the new volume and effectively merge
their results into the same FIF set.}
  \label{usage}
  \end{center}
\end{figure}

This is made possible because each volume is independent of one
another, and can be loosely collated into a bigger FIF file
set. However, considering the above scenario, is it clear that
although volume 1 can be used by itself, both volume 2 and 3 might
refer to data in volume 1 and require it loaded too. We specify a
volume attribute named ``volume'' which links to the location of other
volumes.

\subsection{FIF Streams}
The above section explained how FIF file sets can be split into
volumes, each containing segments. When the entire FIF file set is
loaded, segments may be fetched from any volume. The present section
describes how these individual segments are used to form {\em stream
abstractions}.

A stream is a high level abstraction of a single source of data. For
example, a stream may contain a disk image or a network
capture. Although we do not specify the exact API, implementations
MUST make a stream opened for reading have at a minimum a read, seek
and tell methods. Seeking MUST support the whence argument as
described in the standard C implementation.

Streams opened for writing MAY only support the write method as
seeking in compressed streams may not be possible. Applications should
avoid needing to seek in output streams, and try to structure writes
serially. It is usually possible to reorder a sequence of non
overlapping random writes into a linear stream and a map which is
described below.

All streams within the FIF file MUST have a properties segment. The
properties segment describes the stream attributes as specified
previously. Streams must have the following mandatory attributes:

\begin{itemize}
\item {\em type} is the kind of stream.

\item {\em size} is the size of the stream in bytes.
\end{itemize}

The type attribute is used to interpret the other segments within the
stream in a stream specific way.

\section{Standard FIF segments}
The FIF specification defines a number of basic types, but extensions
are encouraged. Future versions of the specification might standardize
further types. Implementations MUST implement the basic types.

\subsection{The Image Stream}
The {\em Image} stream contains a single read only forensic data
source. For example, this Stream might contains a hard disk image, a
memory image or a network capture (in PCAP format). The Image stream
provides random read access into the stream (i.e. allows for randomly
seeking the stream to any point).

The image itself is stored in chunks within the FIF file. Chunks are
segments named as an 8 digit, zero padded decimal integer
representation of their chunk id (e.g. ``default/00000032''). Image
streams must specify the {\em chunk\_size} attribute, as the number of
image bytes each chunk contains. Chunks may be compressed at the
underlying Zip container level, using any of the compression
algorithms specified by the zip specifications.

%% Talk about signing and hashing here....

\subsection{The Map Stream}
Linear transformations of data are commonplace in forensic
analysis. For example, a file is often simply a collection of bytes
drawn from an image. A TCP/IP stream is simply a collection of
payloads from selected network packets. Sometimes the same data may be
viewed in a number of ways - for example a Virtual Address Space is a
mapping of the Physical Address Space through a page table
transformation. It is, therefore, beneficial to define a way in which
the underlying storage system can apply transformations to existing
data to produce transformed data.

For example, Zero Storage Carving \cite{Meijer2006}, is a way of
specifying carved files in terms of a sequence of blocks taken from
the image. \cite{Cohen2007} extended this concept to an arbitrary
mapping function. This is the approach chosen in this specification
due to its flexibility.

The Map stream defines a mapping function, as a piecewise continuous
linear transform of one or more {\em target} streams to produce a new
{\em stream}. Target streams are specified by using ``target''
attribute, which may be used many times to provide a list of targets.

The mapping function file consists of a series of lines, each
containing three 64 bit integers separated by comma, and encoded using
decimal notation. The integers represent stream offset, target offset
and target number. Target numbers are an index to the list of targets
given in the attribute section (list index starts at zero for the
first entry).

Denoting the stream offset by $x$, and the target offset by $y$, the
map specifies a set of points $(X_i,Y_i,T_i)$. Read requests for a
byte at a mapped stream offset $x$ can then be satisfied by reading a
byte from target $T_i$ at offset $y$ given by:
\begin{eqnarray}
y = (x - X_i) + Y_i & &
\forall x \in \left [X_i, X_i+1 \right )
\end{eqnarray}

For example, consider the following map:
\begin{verbatim*}
0,0,0
4096,10000,0
8192,5000,0
\end{verbatim*}

To read this stream we satisfy read requests of offsets between 0 and
4095 in the stream from offset 0 in target 0. Requests for bytes
between 4096 and 8191 are fetched from target 0 at offset
10000. Finally bytes after 8192 (until the specified size of the
stream) are fetched from offset 5000 in target 0.

In order to efficiently express periodic maps such as those found in
RAID arrays, the Map stream may be provided with two optional
parameters a {\em target\_period} ($T_p$), and {\em stream\_period}
($S_p$). If specified, the above relation becomes:
\begin{eqnarray*}
p &:=& floor\left (\frac{x}{S_p} \right) \\
x' &:=& mod(x ,S_p)  \\   \label{eq:no1}
y &:=& (x'-X_i) + Y_i + p \times T_p
\end{eqnarray*}

Where $mod$ is the modulus function, while $floor$ signifies integer
division. For example consider the following, which corresponds to a 3
disk RAID-5 array:
\begin{verbatim*}
stream_period=393216
target_period=196608

0,0,1
65536,0,0
131072,65536,2
196608,65536,1
262144,131072,0
327680,131072,2
\end{verbatim*}

\subsection{The Overlay stream}
The Overlay stream is a variant of the {\em Image} stream in that it
allows for images to be stored in (possibly compressed) chunks. The
main utility of the Overlay stream is in providing transparent access
to other forensic formats, such as EWF, AFF. These formats also use
chunk based compression to provide the ability to randomly seek while
reading. To access these file formats we only need to have an index of
each chunk offset, and then directly use the overlayed files. The {\em
ewf2fif} tool provides such an overlay feature for EWF volumes by
employing libewf to scan the internal data structures.

Therefore, the {\em chunk\_size} attribute must be specified. In
addition the {\em compression\_type} attribute MAY be specified. The
attribute should contain an integer corresponding to a suitable Zip
compression scheme (e.g. ZIP\_DEFLATE). If not specified it is assumed
that no compression is used.

The ``target'' stream attribute may be specified multiple times and
refer to external files using the {\em file://} scheme. There can be a
number of overlay segments named ``stream name/overlay.00'' which are
all merged to a single overlay. If more than one overlay is present,
the attribute {\em overlays} MUST be specify how many.

The overlay segment contains a series of lines with comma delimited,
decimal encoded, 64 bit integers representing chunk number, target
offset, target length and target number. Target number is an index
into the targets defined in the attributes (count starts witn zero).
Chunks are then fetched from the specified target and decompressed if
necessary.

A target number of -1 signifies the current file (the zip archive
itself). This can be useful for modifying an existing EWF file to be a
FIF file. Since Zip files are normally read from the end of the file,
but EWF and AFF are both read from the front of the file it may be
possible to append a FIF file to the end of an EWF file without
interfering with the overlayed file. In this case the converted file
can still be used as an EWF file without change, but it can also be
used as a FIF file.

In image streams some chunks may compress very well. In that case, the
overheads introduced by the Zip File Header for each chunk in the
Image stream may become unacceptably high. In that case its possible
to use an {\em Overlay stream} to coalesce chunks into larger
segments. This approach is illustrated in Figure \ref{overlay}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{overlay.eps}}
  \caption{The use of an Overlay Stream to overlay an image with
coalesced chunks.}
  \label{overlay}
  \end{center}
\end{figure}

In the above figure a segment is written to the FIF volume which
contains a number of chunks back to back. An overlay stream then uses
direct references to each compressed chunk within the coalesced
segment. Although this technique is valid, it is generally
discouraged since a reorganization of the Zip file might break the
references within the overlay file. This representation also breaks
the appealing simplicity of the standard Image Stream, since it is no
longer possible to simply unzip all the chunks and concatenate them
together to recover the original image.

\subsection{The Private stream}

\subsection{Encrypted Streams}

\section{Implementation Recommendations}
Since the individual chunks are simply files in a Zip archive, any zip
library can be utilized for reading these chunks. However, this may
not be efficient since simple Zip implementations might rescan the
Central Directory for each segment request. Since in practice there
can be a large number of chunks in each volume, it is recommended that
implementations build an index of offsets and length of each chunk
into the file. This kind of index is very similar to the strategy
other file formats use \cite{EWF}. The index may then be saved to the
file by use of a Private stream.

Zip64 specifies duplicate data in the Central Directory entry for each
segment and its File Header (such as filename, size, CRC etc). Many
implementations only populate this information in one of these
places. In the interest of robustness, it is recommended that FIF
implementations populate both places at the slight expense of higher
overheads, because this can be used to recover and correct damaged
files. If the central directory is lost, it is possible to scan
through the zip file and repair the central directory from the File
headers.

The underlying Zip implementation should support multiple writers
simultaneously by interleaving their writes to the file. For example,
for two streams being written concurrently, the new segments can be
added to the same archive in any order. In addition, the volume may be
closed and a new volume opened without writers being aware of
this. Since all segments from all volumes are merged, streams can be
split arbitrarily across volumes in any way. It is recommended that
stream properties be written to each volume to ensure at least some of
the stream can be used even if some volumes are lost.

\bibliography{IEEEabrv,paper}
\end{document}

### 2009-02-21 paper/paper.tex
Updated paper to talk about hashing and encryption.

@@
 %\documentclass[10pt,twocolumn]{article}
-\documentclass[12pt, conference]{IEEEtran}
+\documentclass[10pt, conference]{IEEEtran}
 %\IEEEoverridecommandlockouts
@@
 % Make sure we know its a draft for now
-%\usepackage{draftwatermark}
+\usepackage{draftwatermark}
 \usepackage{verbatim}
@@
 \begin{abstract}
-Forensics analysis requires the acquisition of management of many
+Forensics analysis requires the acquisition and management of many
 different types of evidence material. Often the same evidence is
@@
 restrictions imposed on their use. Rather, we specify stream based
-encryption scheme for use in FIF files.
+encryption scheme for use in FIF files (See section
+\ref{crypted_stream}).
 
@@
 Attributes MUST be specified at the top level of each volume
-(i.e. must have a segment named ``properties'')
+(i.e. must have a segment named ``properties''). The volume attributes
+are used to specify the volume UUID.
 
@@
 
+\label{target_reference}
+{\em A target reference} is a way of referencing objects. This is used
+in a number of places which will be highlighted below. Using
+references it is possible for the FIF file to refer to objects
+embedded within itself, or within other FIF files. Its is also
+possible to refer to external flat files. A number of usage scenarios
+are described in Section \ref{usage_scenarios}. The reference is
+specified with the following conventions:
+
+\begin{itemize}
+\item A reference starting with {\em file://} refers to an external
+file. The external file reference must be sanitised (to remove
+.. directory traversal) and rooted at the current directory.
+
+\item A reference starting with {\em http://} refers to a HTTP
+object which will be fetched. Support for this is optional.
+
+\item A reference in the format {\em fif://UUID/stream/segment} refers
+to an external FIF file set (identified by its UUID). If the segment
+is missing (i.e. the reference ends with a ``/''), the reference is
+for the entire stream. We do not specify a way of actually resolving
+the UUID into a physical locations. This can achieved by
+implementations using many methods, such as Active Directory, LDAP,
+DNS, flat files or SQL databases.
+
+\item A references ending with a ``/'' signify a reference to a
+stream in the current FIF file. The stream will be opened in its
+entirety. If the reference does not end with a ``/'', the reference is
+for a single segment.
+\end{itemize}
+
 \subsection{FIF Volumes and segment merging}
@@
 specification aims to separate high level information such as hashes,
-cryptography and signatures from the low level bit level
-representation we implement those features at a higher level.
+cryptography and signatures from the low bit level representation, we
+implement those features at a higher level.
 
@@
 become fragmented as holes appear within the middle of the volume, and
-updated segments get added to the end.
+updated segments are added at the end.
 
@@
 old central directory remains in the file). This is illustrated in
-Figure \ref{zip_structure}.
+Figure \ref{zip_structure}. Implementations may choose to delay
+writing the central directory until the file is closed.
 
@@
   \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}
+
   \caption{A typical usage scenario. Both Alice and Bob receive a FIF
-volume 1 but work independently. Rather than modifying Volume 1, they
-each create another volume and save their results to the new
-volume. They can now exchange the new volume and effectively merge
-their results into the same FIF set.}
+  volume but work independently. Rather than modifying the volume,
+  they each create another volume and save their results to the new
+  volume. They can now exchange the smaller new volumes and
+  effectively merge their results into the same FIF set.}
+
   \label{usage}
@@
 another, and can be loosely collated into a bigger FIF file
-set. However, considering the above scenario, is it clear that
-although volume 1 can be used by itself, both volume 2 and 3 might
-refer to data in volume 1 and require it loaded too. We specify a
-volume attribute named ``volume'' which links to the location of other
-volumes.
+set.
 
@@
 and tell methods. Seeking MUST support the whence argument as
-described in the standard C implementation.
+described in the standard C implementation (i.e. allow seeking to the
+end of the stream).
 
-Streams opened for writing MAY only support the write method as
+Streams opened for writing MAY limit support for seeking, as efficient
 seeking in compressed streams may not be possible. Applications should
@@
 serially. It is usually possible to reorder a sequence of non
-overlapping random writes into a linear stream and a map which is
-described below.
+overlapping random writes into an Image stream and a Map Stream (These
+are discussed in Sections \ref{image_stream} and
+\ref{map_stream}). Implementations are free to hide the inability to
+seek in writing streams by automatically implementing such a scheme,
+but this is not mandated by the specification.
 
@@
 properties segment describes the stream attributes as specified
-previously. Streams must have the following mandatory attributes:
-
-\begin{itemize}
-\item {\em type} is the kind of stream.
-
-\item {\em size} is the size of the stream in bytes.
-\end{itemize}
-
-The type attribute is used to interpret the other segments within the
-stream in a stream specific way.
+previously. Streams must have the {\em type} attribute which allows
+implementations to determine how they should be processed.
 
@@
 \subsection{The Image Stream}
-The {\em Image} stream contains a single read only forensic data
-source. For example, this Stream might contains a hard disk image, a
-memory image or a network capture (in PCAP format). The Image stream
-provides random read access into the stream (i.e. allows for randomly
-seeking the stream to any point).
+\label{image_stream}
+The {\em Image} stream ({\em type=image}) contains a single read only
+forensic data source. For example, this Stream might contains a hard
+disk image, a memory image or a network capture (in PCAP format). The
+Image stream provides random read access into the stream (i.e. allows
+for randomly seeking the stream to any point). The {\em size}
+attribute MUST be specified to determine how large the stream is.
 
@@
 
-%% Talk about signing and hashing here....
+Chunks can optionally be individually hashed, allowing for detection
+of modifications. The attribute {\em hash\_type} specifies the type of
+hash used (implementations SHOULD support md5,sha1 and sha256, but can
+define others). The hashes are stored in any number of segments within
+the stream, of arbitrary name. The hash segment names is provided by
+using the {\em hash} stream attribute.
+
+Each hash segment MUST be a list of a chunk id encoded in 32 bit
+little endian integer format followed by the hash (the length of which
+depends on the hash itself). This approach is illustrated in Figure
+\ref{hash_index}.
+
+Hash segments can contain any number of chunk hashes in them. It is
+recommended that a hash segment be created when a volume is closed to
+cover all the chunks within that volume. This allows for hashes to be
+verified even if some volumes are missing. It is permitted for a hash
+segment to update a hash of a chunk previously set by an earlier hash
+segment - this may be required when a chunk is updated. The hash
+segment can then be signed to assure the integrity of the updated
+segment.
+
+\begin{figure}[tb]
+  \begin{center}
+  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
+
+  \caption{The structure of a hashing segment. The attributes {\em
+  hash} and {\em hash\_type} declare the hash segment. The segment
+  contains a list of integer encoded chunk id followed by an unencoded
+  hash for each chunk.}
+
+  \label{hash_index}
+  \end{center}
+\end{figure}
+
+This design differs from AFF, which creates a separate segment for
+each hash. The Zip headers are more verbose than the AFF segment
+headers, and typically FIF chunks are much smaller (default 32kb, vs
+16Mb for AFF). Therefore, creating a new segment for each hash value
+results in more overheads in the FIF format. By packing the hashes
+into a few segments it is possible to reduce this overhead to a
+minimum.
+
+Hashes are calculated in an identical way to that defined by AFF,
+thats is the full name of the segment (including stream component) is
+followed by a single NULL byte, followed by the (uncompressed) content
+of the chunk. The inclusion of the segment name ensure that even
+identical chunks in content will have a different hash.
 
 \subsection{The Map Stream}
+\label{map_stream}
 Linear transformations of data are commonplace in forensic
@@
 main utility of the Overlay stream is in providing transparent access
-to other forensic formats, such as EWF, AFF. These formats also use
+to other forensic formats, such as EWF or AFF. These formats also use
 chunk based compression to provide the ability to randomly seek while
@@
   \begin{center}
-  \mbox{\epsfxsize=0.8\columnwidth \epsffile{overlay.eps}}
+  \mbox{\epsfxsize=0.6\columnwidth \epsffile{overlay.eps}}
   \caption{The use of an Overlay Stream to overlay an image with
@@
 \subsection{The Private stream}
+Often forensic software needs to store internal state, such as the
+current state of the GUI, or internal data structures which may be
+required for caching. Usually, software create additional cache files
+to maintain this information. It is advantageous to be able to store
+these in the evidence file itself. The evidence file can then be
+re-opened by the software at a later stage and private data can be
+retrieved and used directly. Such private application data can be
+stored in a Private stream. There is no specification of what can be
+stored in the private stream, and applications with private streams
+should be able to store arbitrary segments in any format at all.
+
+It is recommended that applications name their private segment
+sufficiently accurately so as not to be confused with other
+applications, or different versions of the same application
+(e.g. ``pyflag/0.87/cache/'').
 
 \subsection{Encrypted Streams}
+Encryption is an important property in an evidence file format
+\cite{AFF}. In particular with the FIF format, multiple streams may be
+present in the file set, and often different access levels are
+desired. For example, for evidence set containing both network
+captures and disk images it may be desirable to limit access to
+streams based on legal authorizations.
+
+Although the Zip64 standard specifies encryption, it is not suitable
+for our purposes since it encrypts each segment separately, and does
+not specify a sufficiently flexible scheme (e.g. support for PKI, PGP
+keys). Segment based encryption may lead to information leakage when
+segments are compressed, as the uncompressed size of the segment may
+be deduced.
+
+The Encrypted Stream ({\em type=encrypted}) SHOULD be implemented by
+FIF implementations, but may be omitted with an obvious loss of
+functionality (Encryption may be prohibited or restricted in some
+jurisdictions).
+
+The Encrypted Stream is similar to the Image Stream above in that data
+is written out in chunks (although its probably not effective to
+compress these chunks at all). Each chunk is to be encrypted upon
+writing and transparently decrypted using a scheme declared in the
+{\em scheme} attribute.
+
+The Encrypted Stream may contain any data at all. It is useful
+however, to store an actual FIF volume within the Encrypted
+stream. This provides block level encryption for the contained FIF
+volume. It is recommended that Encrypted Streams containing a FIF
+volume set an attribute ({\em content-type=application/x-fif-volume})
+to allow implementations to automatically recognize volumes. This
+approach is illustrated in Figure \ref{crypted_fif}.
+
+\begin{figure}[tb]
+  \begin{center}
+  \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
+  \caption{Embedding an encrypted FIF volume within an Encrypted Stream.}
+  \label{crypted_fif}
+  \end{center}
+\end{figure}
+
+The result is that a number of FIF volume are used as {\em Container
+Volumes} to provide storage for Encrypted Streams. The main {\em
+Embedded Volume}, which actually contains data is stored within the
+Encrypted Stream, effectively distributed throughout the container
+volumes. Note that the Container Volume may contain several Encrypted
+Streams and therefore contain multiple Encrypted FIF
+Volumes. Container Volumes may contain non encrypted streams as well,
+and may implement different encryption schemes and keys for each
+Encrypted scheme. This effectively allows arbitrary access policies to
+be implemented as only volumes which can be accessed can be merged.
+
+Implementations should create embedded FIF volumes as extensions to
+the container volumes (i.e. with the same UUID). Thus when embedded
+volumes are opened, their streams become automatically viewable as
+part of the evidence set.
+
+\subsection{Encryption Schemes}
+We specify a number of encryption schemes, but extra schemes may be
+defined by the implementation.
+
+\subsubsection{Null Encryption}
+This scheme ({\em scheme=null}) specifies no encryption at all. This
+may be useful for testing an implementation, but obviously provides
+no real security.
+
+\subsubsection{AES-SHA-PSK}
+This scheme ({\em scheme=aes-sha-psk}) uses AES for block level
+encryption, with the following SHA based ESSIV scheme, and Pre-Shared
+Key (passphrase).
+
+When creating a new volume, a master key is generated using the first
+128 bits of a SHA1 hash of the Pre-Shared-Key appended to a 64 bit
+random salt:
+\begin{eqnarray}
+Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
+\end{eqnarray}
+
+For each chunk, the chunk IV is obtained by taking the first 128 bits
+of the SHA1 hash of the chunk id encoded as a 32 bit little endian
+integer, followed by the master key:
+\begin{eqnarray}
+IV = \left | SHA1(chunk\_id + Key_{Master}) \right | _{128}
+\end{eqnarray}
+
+The block is then encrypted or decrypted using AES with this IV.
+
+Implementations MUST specify the salt as a stream attribute named {\em
+salt} in base 64 encoding. Implementations are free to specify any API
+for passing the Pre-Shared Key, for example the PSK can be specified
+in a configuration file, typed in at the terminal, passed in using
+process environment variables, or encoded within the FIF filename
+itself.
+
+\subsection{Optional features}
+Implementations may implement the following features to simplify the
+interfaces.
+
+\subsubsection{Automated volume loading}
+FIF volumes are effectively independent, as each volume does not need
+to know anything about other volumes. This makes it necessary to
+specify in advance all volumes for the implementation to merge them
+all. In particular it may be necessary to specify encrypted volumes
+(which may not be read if the user has no authorization for the
+Encrypted Stream).
+
+As an optional enhancement FIF specifies a volume attribute named {\em
+volume} (i.e. an attribute present in the top level ``properties''
+segment). This is a target reference (see section
+\ref{target_reference}) for other volumes. 
+Implementations should attempt to open external references (e.g. using
+the {\em file://} convention) before internal references. As new
+volumes are merged their volume attributes must be scanned for
+additional references. Each volume must list all other known volumes
+as volume attributes.
+
+In this way, encrypted volumes, and external volumes can be located
+and found automatically.
 
@@
 
+\section{Usage Scenarios}
+In this section we describe how a FIF file format may be used in
+various situations. Many of those can be solved using other forensic
+file formats, but often in a more awkward way.
+
+\subsubsection{Rapidly converting a set of DD images}
+Many hardware devices are available to acquire hard disks in the
+fields. These often produce a set of uncompressed images split at a
+certain size. If the images are split at exactly the same size its
+possible to create a FIF archive consisting of a single Overlay
+Stream, with a chunk\_size set to the size of each of the images. The
+images can then be listed in order as external targets (e.g. {\em
+target=file://disk1.dd}). This overlay can be created instantly
+without needing to re-compress any of the images. In addition, digital
+signatures can be generated for the whole stream or for each segment.
+
+\subsubsection{Rapid conversion of EWF file sets}
+Much existing evidence has been acquired using the EWF file
+format. This format, similarly, consists of a series of chunks which
+may be compressed, with interdispersed indexes. A tool such as
+``ewf2fif'' is able to utilize the libewf library to create an Overlay
+stream of the original image without the need to re-compress it. This
+conversion is very quick and the resulting FIF file is very small -
+each volume in the EWF file set is referenced through an external
+target reference, and the Overlay Stream is built as an index of each
+chunk location.
+
+\subsubsection{Acquisition of RAID disks}
+Often disks in a system are grouped into RAID devices, commonly RAID-5
+or RAID-0. Previously, if disks were acquired independently, they
+would need to be analysed using a tool which was able to reassemble
+RAID devices.
+
+With the FIF format, each of the disks can be acquired as a separate
+Image Stream. Finally a tool such as PyFlag may be used to deduce the
+RAID map, which can be appended to the FIF file as a Map Stream. This
+Map Stream can then be opened by any tool without the tool needing to
+have explicit support for RAID reassembling.
+
+\subsubsection{Cryptographic management of evidence}
+A FIF archive may hold multiple encrypted volumes, each in its own
+Encrypted Stream. Each of those streams is encrypted using a different
+master key, and therefore can have different passphrases, and can be
+assigned to different users by encrypting the master key with
+different X509 certificates. Its is also possible for users to create
+non-encrypted volumes within the FIF file.
+
+This can be used to enforce access controls in line with current
+legislative requirements. For example, within the same investigation
+different material is often obtained under different warrants
+(e.g. wiretap authorizations are different from search
+warrants). Therefore, different investigators and analysts need
+different access to the different streams. However, the analysts may
+still store the results of their analysis in an unencrypted form, or
+assign others permissions to decrypt their analysis results, without
+providing access to the underlying data.
+
+This can be used in sharing meta data (e.g. Map Streams of files of
+interest) between analysts, without needing to provide access to the
+underlying data.
+
+\subsubsection{Forensic Application State}
+Often forensic applications need to store files other than the
+evidence itself - for example, they might need to store internal data,
+annotations, keyword hits etc. Currently these applications store the
+data in a proprietary external file or database. This makes it
+difficult to archive because the evidence itself may become separated
+from the case file.
+
+It is advantageous for these applications to store their state within
+the evidence file itself using a Private Stream. Then when opening the
+evidence file again, the results of their analysis will become
+available.
+
 \bibliography{IEEEabrv,paper}

### 2009-02-21 paper/paper.tex
Added Overlay stream type.

@@
 file. The external file reference must be sanitised (to remove
-.. directory traversal) and rooted at the current directory.
+.. directory traversal) and rooted at the current directory. A
+filename of ``.'' refers to the current FIF volume.
 
@@
 attribute should contain an integer corresponding to a suitable Zip
-compression scheme (e.g. ZIP\_DEFLATE). If not specified it is assumed
+compression scheme (e.g. DEFLATE). If not specified it is assumed
 that no compression is used.
@@
 The ``target'' stream attribute may be specified multiple times and
-refer to external files using the {\em file://} scheme. There can be a
-number of overlay segments named ``stream name/overlay.00'' which are
-all merged to a single overlay. If more than one overlay is present,
-the attribute {\em overlays} MUST be specify how many.
+refer to external files (See section \ref{target_reference}). There
+can be a number of overlay segments named ``stream name/overlay.00''
+which are all merged to a single overlay. If more than one overlay is
+present, the attribute {\em overlays} MUST be specify how many.
 
@@
 offset, target length and target number. Target number is an index
-into the targets defined in the attributes (count starts witn zero).
+into the targets defined in the attributes (count starts with zero).
 Chunks are then fetched from the specified target and decompressed if
@@
 
-A target number of -1 signifies the current file (the zip archive
-itself). This can be useful for modifying an existing EWF file to be a
-FIF file. Since Zip files are normally read from the end of the file,
-but EWF and AFF are both read from the front of the file it may be
-possible to append a FIF file to the end of an EWF file without
+It is possible to refer to the current file by providing a target of
+``file://.''. This can be useful for modifying an existing EWF file to
+be a FIF file. Since Zip files are normally read from the end of the
+file, but EWF and AFF are both read from the front of the file it may
+be possible to append a FIF file to the end of an EWF file without
 interfering with the overlayed file. In this case the converted file
@@
 direct references to each compressed chunk within the coalesced
-segment. Although this technique is valid, it is generally
-discouraged since a reorganization of the Zip file might break the
-references within the overlay file. This representation also breaks
-the appealing simplicity of the standard Image Stream, since it is no
+segment. The target is then specified as the segment name. Although
+this technique is valid, it is generally discouraged since it breaks
+the appealing simplicity of the standard Image Stream.  It is no
 longer possible to simply unzip all the chunks and concatenate them

### 2009-02-21 fif.py

@@ -263,8 +263,8 @@ class FIFFile(zipfile.ZipFile):
                 fileobj = open(fileobj,'r+b')
             except IOError:
                 ## it refers to an external file
-                if fileobj.startswith("file:///"):
-                    return open(fileobj[len("file:///"):],'r+b')
+                if fileobj.startswith("file://"):
+                    return open(fileobj[len("file://"):],'r+b')
                 else:
                     ## Maybe its a stream within this current FIF
                     ## fileset
@@ -498,7 +498,8 @@ class FIFFile(zipfile.ZipFile):
             self.fp.seek(self.readptr)
             zipfile.ZipFile.close(self)
 
-    def create_stream_for_writing(self, stream_name, stream_type='Image',
+    def create_stream_for_writing(self, stream_name='default',
+                                  stream_type='Image',
                                   **args):
         """ This method creates a new stream for writing in the
         current FIF archive.
@@ -685,7 +686,7 @@ class Image(FIFFD):
         self.mode = mode
         self.name = self.stream_name = stream_name
         self.chunk_id = 0
-        self.chunksize = int(properties.get('chunksize', 32*1024))
+        self.chunksize = int(properties.get('chunk_size', 32*1024))
         self.size = int(properties.get('size',0))
         self.readptr = 0
         self.Store = Store()
@@ -700,7 +701,7 @@ class Image(FIFFD):
         if length==None: length=sys.maxint
         result = ''
         length = min(self.size - self.readptr, length)
-        while length>=0:
+        while length>0:
             data= self.partial_read(length)
             if len(data)==0: break
             
@@ -759,6 +760,90 @@ class Image(FIFFD):
             
         FIFFD.close(self)
 
+class Overlay(Image):
+    """ The overlay stream allows for external chunk references so we
+    can just piggy back on top of files like EWF.
+    """
+    type = 'Overlay'
+    def __init__(self, mode='w', *args, **kwargs):
+        self.chunks = {}
+        self.filenames = {}
+        self.inverted_filenames = {}
+        self.overlay_count = 0
+        Image.__init__(self, mode=mode, *args, **kwargs)
+        if mode=='r':
+            self.open_overlay()
+
+    def open_overlay(self):
+        self.targets = []
+        for t in self.properties.getarray('target'):
+            self.targets.append(self.fd.resolve_url(t))
+        
+        for segment in self.properties.getarray('overlay'):
[…]

### 2009-02-21 test6.py

@@ -1,0 +1,63 @@
+""" This tests the Overlay Stream.
+
+We generate an Overlay from test.E01 in the current directory and then
+use SK to read a file from it.
+
+We expect to have a patched ./ewfinfo that prints the chunk layout to
+stdout. We also need the sk python binding.
+
+"""
+import pyewf,subprocess, fif,os
+
+FILENAME = "test6.zip"
+
+try:
+    os.unlink(FILENAME)
+except: pass
+
+fiffile = fif.FIFFile()
+fiffile.create_new_volume(FILENAME)
+
+fd = fiffile.create_stream_for_writing(stream_type='Overlay')
+
+p = subprocess.Popen(['./ewfinfo','-c','test.E01'], stdout=subprocess.PIPE)
+
+for line in p.stdout.xreadlines():
+    line = line.strip()
+    if "," in line:
+        id, offset, size, compression, filename = line.split(",",5)
+        id = int(id)
+        offset = long(offset)
+        size = long(size)
+        compression = int(compression)
+        fd.set_chunk(id, offset, size, compression, filename)
+    elif "=" in line:
+        key, value = line.split("=",1)
+        if key=='size':
+            fd.size = int(value)
+        fd.properties.set(key, value)
+
+fd.close()
+
+test_fd = fiffile.open_stream("default")
+fd = pyewf.open(["test.E01",])
+
+#test_fd.seek(-10,2)
+#fd.seek(-10,2)
+
+#print len(test_fd.read(10))
+#print len(fd.read(10))
+
+## Do big reads - sk is very slow with compressed ntfs files
+BLOCKSIZE = 1024*1024
+while 1:
+    test_data = test_fd.read(BLOCKSIZE)
+    if len(test_data)==0: break
+
+    data = fd.read(BLOCKSIZE)
+    if data!=test_data:
+        print len(data), len(test_data)
+    assert(data == test_data)
+
+
+fiffile.close()

### 2009-02-21 patches/ewfinfo.c
Added a patch to ewfinfo to disclose the list of chunk offsets. We can now build a FIF overlay on top of an EWF file.

@@ -1,0 +1,648 @@
+/*
+ * ewfinfo
+ * Shows information stored in an EWF file
+ *
+ * Copyright (c) 2006-2008, Joachim Metz <forensics@hoffmannbv.nl>,
+ * Hoffmann Investigations. All rights reserved.
+ *
+ * Refer to AUTHORS for acknowledgements.
+ *
+ * Redistribution and use in source and binary forms, with or without modification,
+ * are permitted provided that the following conditions are met:
+ *
+ * - Redistributions of source code must retain the above copyright notice,
+ *   this list of conditions and the following disclaimer.
+ * - Redistributions in binary form must reproduce the above copyright notice,
+ *   this list of conditions and the following disclaimer in the documentation
+ *   and/or other materials provided with the distribution.
+ * - Neither the name of the creator, related organisations, nor the names of
+ *   its contributors may be used to endorse or promote products derived from
+ *   this software without specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDER, COMPANY AND CONTRIBUTORS
+ * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO,
+ * THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include <common.h>
[…]

### 2009-02-23 paper/paper.tex
Added an initial C implementation. So far we have a mostly working Zip library.

@@
 
+\subsubsection{Merging of evidence set}
+In this common usage case, Bob, goes on site to image one hard disk,
+while Alice independently images a second hard disk. After returning
+to the office they both realise their images belong to the same
+set. They merge their images, by updating the UUIDs on one of the
+volume set to now belong to the other volume set - the merge is almost
+instantaneous as they only need to write an extra ``properties'' file
+on the merged volumes.
+
 \bibliography{IEEEabrv,paper}

### 2009-02-23 facquire.py

@@ -76,7 +76,7 @@ while 1:
         new_name = "%s.%02d.zip" % (args[1], count)
         print "Creating new volume %s" % new_name
         ## Make sure that the old volume knows about the new one:
-        basefif.properties['volume'] = "file:///" + new_name
+        basefif.properties['volume'] = "file://" + new_name
 
         basefif.close()
         basefif.create_new_volume(new_name)

### 2009-02-23 fif.py

@@ -1,6 +1,6 @@
 import zipfile, struct, zlib, re
 import time, binascii
-import bisect, os, uuid, sys
+import bisect, os, uuid, sys, sha
 
 ## The version of this implementation of FIF
 VERSION = "FIF1.0"
@@ -175,7 +175,7 @@ class FIFFile(zipfile.ZipFile):
         if type(filenames)==str:
             filenames = [filenames]
             
-        self.file_offsets = {}
+        self.file_offsets = {}        
         self.Store = Store()
         self.zipfiles = []
         self.properties = properties()
@@ -301,7 +301,7 @@ class FIFFile(zipfile.ZipFile):
                     
                 ## Ok we add it to our volume set
                 if type(fileobj)==file:
-                    name = "file:///%s" % os.path.basename(fileobj.name)
+                    name = "file://%s" % os.path.basename(fileobj.name)
                 else:
                     name = fileobj.name
 
@@ -492,7 +492,7 @@ class FIFFile(zipfile.ZipFile):
                 
             self.writestr("properties", self.properties)
             self.flush()
-            
+
             ## Call our base class to close us - this will dump out
             ## the CD
             self.fp.seek(self.readptr)

### 2009-02-23 finfo.py

@@ -1,0 +1,17 @@
+""" This program displays detailed information about FIF filesets.
+
+We can also verify signatures and test the fileset integrity.
+"""
+import fif, optparse, sys, os
+
+parser = optparse.OptionParser()
+(options, args) = parser.parse_args()
+
+if len(args)==0:
+    print "You must specify a source and a destination"
+    sys.exit(1)
+
+## Open all the args as a fif fileset
+fiffile = fif.FIFFile(args)
+for f in fiffile.file_offsets.keys():
+    print f

### 2009-02-23 src/include/class.h

@@ -1,0 +1,410 @@
+/***************************************************
+  Classes and objects in C
+
+  This file makes it easy to implement classes and objects in C. To
+  define a class we need to perform three steps:
+
+  Define the class prototype. This is suitable to go in a .h file for
+  general use by other code.
+
+  Note all classes extend Object.
+
+  Example::
+
+CLASS(Foo, Object)
+    int x;
+    int y;
+
+    //This declares a method of a class Foo, called Con returning a
+    //Foo object. In other words it is a constructor.
+    Foo METHOD(Foo, Con, int x, int y);
+    int METHOD(Foo, add);
+
+END_CLASS
+
+Now we need to define some functions for the constructor and
+methods. Note that the constuctor is using ALLOCATE_CLASS to allocate
+space for the class structures. Callers may call with self==NULL to
+force allocation of a new class. Note that we do not call the
+constructor of our superclass implicitly here. (Calling the sperclass
+constructor is optional, but ALLOCATE_CLASS is not.).
+
+Foo Foo_Con(Foo self,int x,int y) {
+  self->x = x;
+  self->y = y;
+
+  return self;
+};
+
+int Foo_add(Foo this) {
+  return (this->x + this->y);
+};
+
+Now we need to define the Virtual function table - These are those
+functions and attributes which are defined in this class (over its
+superclass). Basically these are all those things in the class
+definition above, with real function names binding them. (Note that by
+convention we preceed the name of the method with the name of the
+class):
+
+VIRTUAL(Foo,Object)
+   VMETHOD(Con) = Foo_Con;
+   VMETHOD(add) = Foo_add;
+END_VIRTUAL
+
+We can use inheritance too:
+
+CLASS(Bar, Foo)
+   Bar METHOD(Bar, Con, char *something)
+END_CLASS
+
+Here Bar extends Foo and defines a new constructor with a different prototype:
+
+VIRTUAL(Bar,Foo)
+   VMETHOD(Con) = Bar_Con
+END_VIRTUAL
+
+If there is a function which expects a Foo, we will need to over ride
+the Foo constructor in the Bar, so the function will not see the
+difference between the Foo and Bar:
+
+CLASS(Bar,Foo)
+  int bar_attr;
+END_CLASS
+
+Foo Bar_Con(Foo self, int x, int y) {
+...
+}
+
+VIRTUAL(Bar, Foo)
+  VMETHOD(super.Con) = Bar_Con
+END_VIRTUAL
+
[…]

### 2009-02-23 src/include/crypto/sha1.h

@@ -1,0 +1,10 @@
+typedef struct {
+  unsigned long state[5];
+  unsigned long count[2];
+  unsigned char buffer[64];
+} SHA_CTX;
+
+void SHA1_Transform(unsigned long state[5], unsigned char buffer[64]);
+int SHA1_Init(SHA_CTX* context);
+int SHA1_Update(SHA_CTX* context, unsigned char* data, unsigned int len);
+int SHA1_Final(unsigned char digest[20], SHA_CTX* context);

### 2009-02-23 src/include/fif.h

@@ -1,0 +1,11 @@
+#include "class.h"
+#include "talloc.h"
+#include "zip.h"
+
+// This is the main FIF class - it manages the Zip archive
+CLASS(FIFFile, Object)
+
+     FIFFile METHOD(FIFFile, Con );
+
+END_CLASS
+

### 2009-02-23 src/include/list.h

@@ -1,0 +1,276 @@
+#ifndef __LIST_H
+#define __LIST_H
+
+/* This file is from Linux Kernel (include/linux/list.h) 
+ * and modified by simply removing hardware prefetching of list items. 
+ * Here by copyright, credits attributed to wherever they belong.
+ * Kulesh Shanmugasundaram (kulesh [squiggly] isis.poly.edu)
+ */
+
+/*
+ * Simple doubly linked list implementation.
+ *
+ * Some of the internal functions ("__xxx") are useful when
+ * manipulating whole lists rather than single entries, as
+ * sometimes we already know the next/prev entries and we can
+ * generate better code by using them directly rather than
+ * using the generic single-entry routines.
+ */
+
+struct list_head {
+	struct list_head *next, *prev;
+};
+
+#define LIST_HEAD_INIT(name) { &(name), &(name) }
+
+#define LIST_HEAD(name) \
+	struct list_head name = LIST_HEAD_INIT(name)
+
+#define INIT_LIST_HEAD(ptr) do { \
+	(ptr)->next = (ptr); (ptr)->prev = (ptr); \
+} while (0)
+
+/*
+ * Insert a new entry between two known consecutive entries. 
+ *
+ * This is only for internal list manipulation where we know
+ * the prev/next entries already!
+ */
+static inline void __list_add(struct list_head *new,
+			      struct list_head *prev,
+			      struct list_head *next)
+{
+	next->prev = new;
+	new->next = next;
+	new->prev = prev;
+	prev->next = new;
+}
+
+/**
+ * list_add - add a new entry
+ * @new: new entry to be added
+ * @head: list head to add it after
+ *
+ * Insert a new entry after the specified head.
+ * This is good for implementing stacks.
+ */
+static inline void list_add(struct list_head *new, struct list_head *head)
+{
+	__list_add(new, head, head->next);
+}
+
+/**
+ * list_add_tail - add a new entry
+ * @new: new entry to be added
+ * @head: list head to add it before
+ *
+ * Insert a new entry before the specified head.
+ * This is useful for implementing queues.
+ */
+static inline void list_add_tail(struct list_head *new, struct list_head *head)
+{
+	__list_add(new, head->prev, head);
+}
+
+/*
+ * Delete a list entry by making the prev/next entries
+ * point to each other.
+ *
+ * This is only for internal list manipulation where we know
+ * the prev/next entries already!
+ */
[…]

### 2009-02-23 src/include/talloc.h

@@ -1,0 +1,176 @@
+#ifndef _TALLOC_H_
+#define _TALLOC_H_
+/* 
+   Unix SMB/CIFS implementation.
+   Samba temporary memory allocation functions
+
+   Copyright (C) Andrew Tridgell 2004-2005
+   Copyright (C) Stefan Metzmacher 2006
+   
+     ** NOTE! The following LGPL license applies to the talloc
+     ** library. This does NOT imply that all of Samba is released
+     ** under the LGPL
+   
+   This library is free software; you can redistribute it and/or
+   modify it under the terms of the GNU Lesser General Public
+   License as published by the Free Software Foundation; either
+   version 2 of the License, or (at your option) any later version.
+
+   This library is distributed in the hope that it will be useful,
+   but WITHOUT ANY WARRANTY; without even the implied warranty of
+   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+   Lesser General Public License for more details.
+
+   You should have received a copy of the GNU Lesser General Public
+   License along with this library; if not, write to the Free Software
+   Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+*/
+
+/* this is only needed for compatibility with the old talloc */
+typedef void TALLOC_CTX;
+#include <sys/types.h>
+#include <stdarg.h>
+#include <stdlib.h>
+#include <string.h>
+#include <malloc.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+/*
+  this uses a little trick to allow __LINE__ to be stringified
+*/
+#define _STRING_LINE_(s)    #s
+#define _STRING_LINE2_(s)   _STRING_LINE_(s)
+#define __LINESTR__       _STRING_LINE2_(__LINE__)
+#define __location__ __FILE__ ":" __LINESTR__
+
+#ifndef TALLOC_DEPRECATED
+#define TALLOC_DEPRECATED 0
+#endif
+
+#ifndef PRINTF_ATTRIBUTE
+#if (__GNUC__ >= 3)
+/** Use gcc attribute to check printf fns.  a1 is the 1-based index of
+ * the parameter containing the format, and a2 the index of the first
+ * argument. Note that some gcc 2.x versions don't handle this
+ * properly **/
[…]

### 2009-02-23 src/include/zip.h

@@ -1,0 +1,152 @@
+#include "misc.h"
+#include "class.h"
+#include "talloc.h"
+#include "list.h"
+#include <zlib.h>
+
+#define BUFF_SIZE 4096
+
+
+/** These are ZipFile structures */
+struct EndCentralDirectory {
+  uint32_t magic;
+  uint16_t number_of_this_disk;
+  uint16_t disk_with_cd;
+  uint16_t total_entries_in_cd_on_disk;
+  uint16_t total_entries_in_cd;
+  uint32_t size_of_cd;
+  uint32_t offset_of_cd;
+  uint16_t comment_len;
+}__attribute__((packed));
+
+struct CDFileHeader {
+  uint32_t magic;
+  uint16_t version_made_by;
+  uint16_t version_needed;
+  uint16_t flags;
+  uint16_t compression_method;
+  uint16_t dostime;
+  uint16_t dosdate;
+  uint32_t crc32;
+  uint32_t compress_size;
+  uint32_t file_size;
+  uint16_t file_name_length;
+  uint16_t extra_field_len;
+  uint16_t file_comment_length;
+  uint16_t disk_number_start;
+  uint16_t internal_file_attr;
+  uint32_t external_file_attr;
+  uint32_t relative_offset_local_header;
+  char *filename;
+  char *extra_field;
+  char *file_comment;
+}__attribute__((packed));
+
+struct ZipFileHeader {
+  uint32_t magic;
+  uint16_t version;
+  uint16_t flags;
+  uint16_t compression_method;
+  uint16_t lastmodtime;
+  uint16_t lastmoddate;
+  uint32_t crc32;
+  uint32_t compress_size;
+  uint32_t file_size;
+  uint16_t file_name_length;
+  uint16_t extra_field_len;
+}__attribute__((packed));
+
+// Base class for file like objects
+CLASS(FileLikeObject, Object)
+     int64_t readptr;
+     uint64_t size;
+     char *name;
+     
+     uint64_t METHOD(FileLikeObject, seek, int64_t offset, int whence);
+     int METHOD(FileLikeObject, read, char *buffer, int length);
+     int METHOD(FileLikeObject, write, char *buffer, int length);
+     uint64_t METHOD(FileLikeObject, tell);
+     void METHOD(FileLikeObject, close);
+END_CLASS
+
+// This file like object is backed by a real disk file:
+CLASS(FileBackedObject, FileLikeObject)
+     int fd;
+
[…]

### 2009-02-23 src/lib/class.c

@@ -1,0 +1,83 @@
+/******************************************************
+# Copyright 2004: Commonwealth of Australia.
+#
+# Developed by the Computer Network Vulnerability Team,
+# Information Security Group.
+# Department of Defence.
+#
+# Michael Cohen <scudette@users.sourceforge.net>
+#
+# ******************************************************
+#  Version: FLAG  $Version: 0.87-pre1 Date: Thu Jun 12 00:48:38 EST 2008$
+# ******************************************************
+#
+# * This program is free software; you can redistribute it and/or
+# * modify it under the terms of the GNU General Public License
+# * as published by the Free Software Foundation; either version 2
+# * of the License, or (at your option) any later version.
+# *
+# * This program is distributed in the hope that it will be useful,
+# * but WITHOUT ANY WARRANTY; without even the implied warranty of
+# * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# * GNU General Public License for more details.
+# *
+# * You should have received a copy of the GNU General Public License
+# * along with this program; if not, write to the Free Software
+# * Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA  02111-1307, USA.
+# ******************************************************/
+#include "class.h"
+
+/** This is used for error reporting.
+ */
+char *__error_str;
+enum _error_type _global_error;
+#define ERROR_BUFFER_SIZE 1024
+char _error_buff[ERROR_BUFFER_SIZE];
+char _traceback[ERROR_BUFFER_SIZE];
+
+void *raise_errors(enum _error_type t, char *reason, ...) {
+  if(reason) {
+    va_list ap;
+    va_start(ap, reason);
+    vsnprintf(_error_buff, ERROR_BUFFER_SIZE-1,reason,ap);
+    _error_buff[ERROR_BUFFER_SIZE]=0;
+    va_end(ap);
+  };
+
+  _global_error = t;
+
+  return NULL;
+};
+
+// Noone should instantiate Object directly. this should be already
+// allocated therefore:
+
+inline void Object_Alloc(Object this) {
+  this->__class__ = &__Object;
+  this->__super__ = NULL;
+};
+
+inline void Object_init() {
+  Object_Alloc(&__Object);
+};
+
+struct Object __Object = {
+  .__class__ = &__Object,
+  .__super__ = &__Object,
+  .__size = sizeof(struct Object)
+};
+
[…]

### 2009-02-23 src/lib/fifinfo.c

@@ -1,0 +1,37 @@
+#include "fif.h"
+#include "class.h"
+#include <stdio.h>
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <fcntl.h>
+
+int main() {
+  ZipFile zip = CONSTRUCT(ZipFile, ZipFile, Con, NULL, NULL);
+  FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con, zip, "new_test.zip", 'w');
+  FileBackedObject out_fd;
+
+  // Create a new Zip volume for writing
+  CALL(zip, create_new_volume, (FileLikeObject)fd);
+
+  // Open the member foobar for writing
+  out_fd = CALL(zip, open_member, "foobar", 'w',
+			       ZIP_DEFLATE);
+
+  // It worked - now copy /bin/ls into it
+  if(out_fd) {
+    char buffer[BUFF_SIZE];
+    int fd=open("/bin/ls",O_RDONLY);
+    int length;
+    while(1) {
+      length = read(fd, buffer, BUFF_SIZE);
+      if(length == 0) break;
+
+      CALL(out_fd, write, buffer, length);
+    };
+
+    CALL(out_fd, close);
+    CALL(zip, close);
+  };
+
+  PrintError();
+};

### 2009-02-23 src/lib/sha1.c

@@ -1,0 +1,213 @@
+/*
+SHA-1 in C
+By Steve Reid <steve@edmweb.com>
+100% Public Domain
+
+Test Vectors (from FIPS PUB 180-1)
+"abc"
+  A9993E36 4706816A BA3E2571 7850C26C 9CD0D89D
+"abcdbcdecdefdefgefghfghighijhijkijkljklmklmnlmnomnopnopq"
+  84983E44 1C3BD26E BAAE4AA1 F95129E5 E54670F1
+A million repetitions of "a"
+  34AA973C D4C4DAA4 F61EEB2B DBAD2731 6534016F
+*/
+
+/* #define LITTLE_ENDIAN * This should be #define'd if true. */
+/* #define SHA1HANDSOFF * Copies data before messing with it. */
+
+#include <stdio.h>
+#include <string.h>
+#include "crypto/sha1.h"
+
+#define rol(value, bits) (((value) << (bits)) | ((value) >> (32 - (bits))))
+
+/* blk0() and blk() perform the initial expand. */
+/* I got the idea of expanding during the round function from SSLeay */
+#ifdef LITTLE_ENDIAN
+#define blk0(i) (block->l[i] = (rol(block->l[i],24)&0xFF00FF00) \
+		 |(rol(block->l[i],8)&0x00FF00FF))
+#else
+#define blk0(i) block->l[i]
+#endif
+#define blk(i) (block->l[i&15] = rol(block->l[(i+13)&15]^block->l[(i+8)&15] \
+				     ^block->l[(i+2)&15]^block->l[i&15],1))
+
+/* (R0+R1), R2, R3, R4 are the different operations used in SHA1 */
+#define R0(v,w,x,y,z,i) z+=((w&(x^y))^y)+blk0(i)+0x5A827999+rol(v,5);w=rol(w,30);
+#define R1(v,w,x,y,z,i) z+=((w&(x^y))^y)+blk(i)+0x5A827999+rol(v,5);w=rol(w,30);
+#define R2(v,w,x,y,z,i) z+=(w^x^y)+blk(i)+0x6ED9EBA1+rol(v,5);w=rol(w,30);
+#define R3(v,w,x,y,z,i) z+=(((w|x)&y)|(w&x))+blk(i)+0x8F1BBCDC+rol(v,5);w=rol(w,30);
+#define R4(v,w,x,y,z,i) z+=(w^x^y)+blk(i)+0xCA62C1D6+rol(v,5);w=rol(w,30);
+
+
[…]

### 2009-02-23 src/lib/talloc.c

@@ -1,0 +1,1462 @@
+/* 
+   Samba Unix SMB/CIFS implementation.
+
+   Samba trivial allocation library - new interface
+
+   NOTE: Please read talloc_guide.txt for full documentation
+
+   Copyright (C) Andrew Tridgell 2004
+   Copyright (C) Stefan Metzmacher 2006
+   
+     ** NOTE! The following LGPL license applies to the talloc
+     ** library. This does NOT imply that all of Samba is released
+     ** under the LGPL
+   
+   This library is free software; you can redistribute it and/or
+   modify it under the terms of the GNU Lesser General Public
+   License as published by the Free Software Foundation; either
+   version 2 of the License, or (at your option) any later version.
+
+   This library is distributed in the hope that it will be useful,
+   but WITHOUT ANY WARRANTY; without even the implied warranty of
+   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+   Lesser General Public License for more details.
+
+   You should have received a copy of the GNU Lesser General Public
+   License along with this library; if not, write to the Free Software
+   Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+*/
+
+/*
+  inspired by http://swapped.cc/halloc/
+*/
+
+#ifdef _SAMBA_BUILD_
+#include "version.h"
+#if (SAMBA_VERSION_MAJOR<4)
+#include "includes.h"
+/* This is to circumvent SAMBA3's paranoid malloc checker. Here in this file
+ * we trust ourselves... */
+#ifdef malloc
+#undef malloc
+#endif
+#ifdef realloc
+#undef realloc
+#endif
+#define _TALLOC_SAMBA3
+#endif /* (SAMBA_VERSION_MAJOR<4) */
+#endif /* _SAMBA_BUILD_ */
+
+#ifndef _TALLOC_SAMBA3
+//#include "replace.h"
+#include "talloc.h"
+#endif /* not _TALLOC_SAMBA3 */
+
+/* use this to force every realloc to change the pointer, to stress test
+   code that might not cope */
+#define ALWAYS_REALLOC 0
+
+
+#define MAX_TALLOC_SIZE 0x10000000
+#define TALLOC_MAGIC 0xe814ec70
[…]

### 2009-02-23 src/lib/zip.c

@@ -1,0 +1,599 @@
+/** This file implements the basic Zip handling code. We allow
+    concurrent read/write.
+
+    FIXME: The FileBackedObject needs to be tailored for windows.
+*/
+#include "zip.h"
+#include <errno.h>
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <fcntl.h>
+#include "list.h"
+#include <unistd.h>
+#include <zlib.h>
+#include <time.h>
+
+int close_fd(void *self) {
+  FileBackedObject this = (FileBackedObject)self;
+  close(this->fd);
+  return 0;
+};
+
+/** Implementation of FileBackedObject */
+FileBackedObject FileBackedObject_con(FileBackedObject self, 
+				      char *filename, char mode) {
+  int flags;
+
+  switch(mode) {
+  case 'w':
+    flags =  O_CREAT|O_RDWR|O_TRUNC|O_BINARY;
+    break;
+  case 'a':
+    flags =  O_CREAT|O_RDWR|O_APPEND|O_BINARY;
+    break;
+  case 'r':
+    flags = O_BINARY | O_RDONLY;
+    break;
+  default:
+    RaiseError(EInvalidParameter, "Unknown mode '%c'", mode);
+    return NULL;
+  };
+  self->fd = open(filename, flags, S_IRWXU | S_IRWXG | S_IRWXO);
+  if(self->fd<0){
+    RaiseError(EIOError, "Can't open %s (%s)", filename, strerror(errno));
+    return NULL;
+  };
+
+  // Make sure that we close the file when we free ourselves - we dont
+  // want to leak file handles
+  talloc_set_destructor((void *)self, close_fd);
+
+  // Work out what the file size is:
+  self->super.size = lseek(self->fd, 0, SEEK_END);
+  self->super.name = talloc_strdup(self, filename);
+
+  return self;
+};
+
+uint64_t FileLikeObject_seek(FileLikeObject self, int64_t offset, int whence) {
+  if(whence==SEEK_SET) {
+    self->readptr = offset;
+  } else if(whence==SEEK_CUR) {
+    self->readptr += offset;
+    if(self->readptr<0) self->readptr=0;
+  } else if(whence==SEEK_END) {
+    self->readptr = self->size + offset;
+  };
+
+  if(self->readptr < 0) {
+    self->readptr=0;
+
+  } else if(self->readptr > self->size) {
+    self->readptr = self->size;
+  };
+
+  return self->readptr;
+};
+  
+/** 
[…]

### 2009-02-23 src/include/misc.h
Forgot some files.

@@ -1,0 +1,13 @@
+#ifndef MISC_H
+#define MISC_H
+#include "config.h"
+
+#ifdef HAVE_INTTYPES_H
+#include <inttypes.h>
+#endif
+
+#ifndef O_BINARY
+#define O_BINARY 0
+#endif
+
+#endif

### 2009-02-23 src/lib/fifinfo.c

@@ -8,7 +8,7 @@
 int main() {
   ZipFile zip = CONSTRUCT(ZipFile, ZipFile, Con, NULL, NULL);
   FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con, zip, "new_test.zip", 'w');
-  FileBackedObject out_fd;
+  FileLikeObject out_fd;
 
   // Create a new Zip volume for writing
   CALL(zip, create_new_volume, (FileLikeObject)fd);

### 2009-02-23 src/lib/fifinfo.c
Added more commenting.

@@ -29,9 +29,12 @@ int main() {
       CALL(out_fd, write, buffer, length);
     };
 
+    // Close the member (finalises the member)
     CALL(out_fd, close);
+    // Close the archive
     CALL(zip, close);
   };
 
   PrintError();
+  talloc_free(zip);
 };

### 2009-02-23 src/lib/zip.c

@@ -13,13 +13,17 @@
 #include <zlib.h>
 #include <time.h>
 
+/** Implementation of FileBackedObject.
+
+FileBackedObject is a FileLikeObject which uses a read file to back
+itself.
+*/
 int close_fd(void *self) {
   FileBackedObject this = (FileBackedObject)self;
   close(this->fd);
   return 0;
 };
 
-/** Implementation of FileBackedObject */
 FileBackedObject FileBackedObject_con(FileBackedObject self, 
 				      char *filename, char mode) {
   int flags;
@@ -270,6 +274,7 @@ ZipFile ZipFile_Con(ZipFile self, FileLikeObject fd) {
  error_reason:
     RaiseError(EInvalidParameter, "%s is not a zip file", self->fd->name);
  error:
+    talloc_free(self);
     return NULL;
 };
 
@@ -494,6 +499,8 @@ ZipFileStream ZipFileStream_Con(ZipFileStream self, ZipFile zip,
     struct tm *now = localtime(&epoch_time);
     
     cd->magic = 0x2014b50;
+    cd->version_made_by = 0x317;
+    cd->version_needed = 0x14;
     cd->compression_method = compression;
     cd->crc32 = crc32(0L, Z_NULL, 0);
     cd->dosdate = (now->tm_year + 1900 - 1980) << 9 | (now->tm_mon + 1) << 5 | now->tm_mday;
@@ -506,6 +513,7 @@ ZipFileStream ZipFileStream_Con(ZipFileStream self, ZipFile zip,
   };
 
   return self;
+
  error:
   talloc_free(self);
   return NULL;
@@ -523,14 +531,16 @@ int ZipFileStream_write(FileLikeObject self, char *buffer, int length) {
 				       (unsigned char*)buffer,
 				       length);
 
+  /** Position our write pointer */
   CALL(this->zinfo->fd, seek, this->zip_offset + self->readptr, SEEK_SET);
   if(this->zinfo->cd_header.compression_method == ZIP_DEFLATE) {
     unsigned char compressed[BUFF_SIZE];
 
-    // Rotate the data through the stream
+    /** Give the buffer to zlib */
     this->strm.next_in = buffer;
     this->strm.avail_in = length;
 
+    /** We spin here until zlib consumed all the data */
     do {
       int ret;
 
@@ -538,16 +548,23 @@ int ZipFileStream_write(FileLikeObject self, char *buffer, int length) {
       this->strm.next_out = compressed;
       
       ret = deflate(&this->strm, Z_NO_FLUSH);
[…]

### 2009-02-23 test.py

@@ -8,7 +8,7 @@ This test requires the sk python bindings and pyewf bindings because
 we test access speed to each other.
 """
 
-import sk, fif, pyewf, pyaff
+import sk, fif, pyewf
 import time, sys, glob
 
 def test_time(fd):

### 2009-02-23 facquire.py
Implemented a cache for ZipFile.

@@ -37,9 +37,9 @@ if os.access(args[1], os.F_OK) and not options.append:
     print "%s already exists - you must specify -a to append to an existing file" % args[1]
     sys.exit(1)
 
-infd = open(args[0],'r')
-mode = 'w'
-if options.append: mode='a'
+infd = open(args[0],'rb')
+mode = 'wb'
+if options.append: mode='ab'
 
 ## Get a handle to the FIF file
 basefif = fiffile = fif.FIFFile()

### 2009-02-23 src/include/class.h

@@ -394,6 +394,7 @@ void *raise_errors(enum _error_type t, char *string,  ...);
 extern char _traceback[];
 
 #define ZSTRING_NO_NULL(str) str , (strlen(str))
+#define ZSTRING(str) str , (strlen(str)+1)
 #define RaiseError(t, ...)			\
   do {						\
     snprintf(_traceback, ERROR_BUFFER_SIZE, "%s:%d - %s", __FILE__,	\

### 2009-02-23 src/include/zip.h

@@ -87,6 +87,11 @@ CLASS(ZipInfo, Object)
      // Each ZipInfo object remembers which file it came from:
      FileLikeObject fd;
 
+     // ZipInfo items may hold a cache of the decompressed data - they
+     // belong to the cache list and may hold the data
+     struct list_head cache_list;
+     char *cached_data;
+
      ZipInfo METHOD(ZipInfo, Con, FileLikeObject fd);
 END_CLASS
 
@@ -107,12 +112,19 @@ CLASS(ZipFile, Object)
      */
      ZipInfo *hash_table;
 
+     /** This is a chunk cache. Its a sorted list in order of usage
+	 (newer pages to the front) 
+     */
+     int cache_size;
+     int max_cache_size;
+     ZipInfo cache_head;
+
      /** A zip file is opened on a file like object */
      ZipFile METHOD(ZipFile, Con, FileLikeObject file);
 
 // Fetch a member as a string - this is suitable for small memebrs
 // only as we allocate memory for it.
-     char * METHOD(ZipFile, read_member, char *filename);
+     int METHOD(ZipFile, read_member, char *filename, char **buffer, int *len);
 
 // This method is called to create a new volume on the FileLikeObject
      void METHOD(ZipFile, create_new_volume, FileLikeObject file);

### 2009-02-23 src/lib/fifinfo.c

@@ -4,9 +4,16 @@
 #include <sys/types.h>
 #include <sys/stat.h>
 #include <fcntl.h>
+#include <sys/time.h>
+#include <time.h>
 
-int main() {
+
+/** First test builds a new zip file from /bin/ls */
+void test1() {
+  // Make a new zip file
   ZipFile zip = CONSTRUCT(ZipFile, ZipFile, Con, NULL, NULL);
+
+  // Make a new file
   FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con, zip, "new_test.zip", 'w');
   FileLikeObject out_fd;
 
@@ -15,7 +22,7 @@ int main() {
 
   // Open the member foobar for writing
   out_fd = CALL(zip, open_member, "foobar", 'w',
-			       ZIP_DEFLATE);
+		ZIP_DEFLATE);
 
   // It worked - now copy /bin/ls into it
   if(out_fd) {
@@ -37,4 +44,42 @@ int main() {
 
   PrintError();
   talloc_free(zip);
+};
+
+#define TIMES 1000
+
+void test2() {
+  // Open the file for reading
+  FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con, NULL, "new_test.zip", 'r');
+  // Open the zip file
+  ZipFile zip = CONSTRUCT(ZipFile, ZipFile, Con, fd, (FileLikeObject)fd);
+
+  char *result=NULL;
+  int length=0;
+  int i;
+  struct timeval epoch_time;
+  struct timeval new_time;
+  int diff;
+
+  gettimeofday(&epoch_time, NULL);
+
+  for(i=0;i<TIMES;i++) {
+    if(CALL(zip, read_member, "foobar", &result, &length)<0) break;
+  };
+
+  gettimeofday(&new_time, NULL);
+  diff = (new_time.tv_sec * 1000 + new_time.tv_usec/1000) -
+    (epoch_time.tv_sec * 1000 + epoch_time.tv_usec/1000);
+  printf("Decompressed foobar %d times in %d mseconds (%f)\n", 
+	 TIMES,  diff,
+	 ((float)diff)/TIMES);
+  PrintError();
+  talloc_free(fd);
+};
+
+int main() {
+  ClearError();
+  test1();
+  ClearError();
+  test2();
 };

### 2009-02-23 src/lib/zip.c

@@ -144,6 +144,7 @@ int ZipInfo_destructor(void *self) {
 ZipInfo ZipInfo_Con(ZipInfo self, FileLikeObject fd) {
   self->fd = fd;
   INIT_LIST_HEAD(&self->list);
+  INIT_LIST_HEAD(&self->cache_list);
 
   return self;
 };
@@ -185,11 +186,12 @@ void add_to_hash(ZipFile self, ZipInfo zip) {
 */
 ZipInfo fetch_from_hashtable(ZipInfo *hashtable, char *filename) {
   ZipInfo zip;
-  int index = hash(ZSTRING_NO_NULL(filename));
+  int index = hash(ZSTRING(filename));
 
   if(hashtable[index]) {
     list_for_each_entry(zip, &(hashtable[index]->list), list) {
-      if(zip->cd_header.filename && !strcmp(filename, zip->cd_header.filename))
+      if(zip->cd_header.filename && !memcmp(filename, zip->cd_header.filename,
+					    zip->cd_header.file_name_length))
 	return zip;
     };
   };
@@ -205,6 +207,11 @@ ZipFile ZipFile_Con(ZipFile self, FileLikeObject fd) {
 
   // Initialise the hash table:
   self->hash_table = talloc_array(self, ZipInfo, HASH_TABLE_SIZE);
+
+  // Make a new cache
+  self->max_cache_size = 50;
+  self->cache_size = 0;
+  self->cache_head = CONSTRUCT(ZipInfo, ZipInfo, Con, self, NULL);
 
   // If we were not given something to read, we are done.
   if(!fd) return self;
@@ -278,21 +285,38 @@ ZipFile ZipFile_Con(ZipFile self, FileLikeObject fd) {
     return NULL;
 };
 
-char *ZipFile_read_member(ZipFile self, char *filename) {
+int ZipFile_read_member(ZipFile self, char *filename, 
+			char **buffer, int *length) {
   ZipInfo zip = fetch_from_hashtable(self->hash_table, filename);
   char *result;
-  char compression_method = zip->cd_header.compression_method;
+  char compression_method;
   struct ZipFileHeader header;
-  int length = sizeof(header);
+  int read_length = sizeof(header);
+  if(!zip) return -1;
 
-  if(!zip) return NULL;
-  // We only expect to read a small amount of data here:
-  result = talloc_size(self, zip->cd_header.file_size);
+  // We know how large we would like the buffer so we make it that big
+  if(*length < zip->cd_header.file_size) {
[…]

### 2009-02-23 src/include/class.h
Started to implement Streams.

@@ -318,12 +318,25 @@ super.add as well.
 				   (virt_class)_talloc_memdup(context, &__ ## class, sizeof(struct class),  __location__ "(" #class ")"), \
 				   ## __VA_ARGS__) )
 
+/** This variant is useful when all we have is a class reference
+    (GETCLASS(Foo)) or &__Foo
+*/
+#define CONSTRUCT_FROM_REFERENCE(class, constructor, context, ... )	\
+  ( class->constructor(							\
+		       (void *)_talloc_memdup(context, class, ((Object)class)->__size,  __location__ "(" #class ")"), \
+		      ## __VA_ARGS__) )
+
 #else
 #define CONSTRUCT(class, virt_class, constructor, context, ... )		\
   (class)( class ## _init(), virt_class ## _init(),			\
 	   __ ## class.constructor(					\
 				   (virt_class)talloc_memdup(context, &__ ## class, sizeof(struct class)), \
 				   ## __VA_ARGS__) )
+
+#define CONSTRUCT_FROM_REFERENCE(class, constructor, context, ... )	\
+  ( class->constructor(							\
+		       (void *)_talloc_memdup(context, class, ((Object)class)->__size,  __location__ "(" #class ")"), \
+		      ## __VA_ARGS__) )
 
 #endif
 

### 2009-02-23 src/include/fif.h

@@ -2,10 +2,28 @@
 #include "talloc.h"
 #include "zip.h"
 
+/* A AFFFD is a special FileLikeObject for openning AFF2 streams */
+CLASS(AFFFD, FileLikeObject)
+     AFFFD METHOD(AFFFD, Con, char *stream_name);
+END_CLASS
+
+/** The Image Stream represents an Image in chunks */
+CLASS(Image, AFFFD)
+END_CLASS
+
+struct dispatch_t {
+  char *type;
+  AFFFD class_ptr;
+};
+
+// A dispatch table for instantiating the right classes according to
+// the stream_type
+extern struct dispatch_t dispatch[];
+
 // This is the main FIF class - it manages the Zip archive
-CLASS(FIFFile, Object)
+CLASS(FIFFile, ZipFile)
 
-     FIFFile METHOD(FIFFile, Con );
-
+     AFFFD METHOD(FIFFile, create_stream_for_writing, char *stream_name,
+		  char *stream_type);
 END_CLASS
 

### 2009-02-23 src/lib/fifinfo.c

@@ -42,12 +42,12 @@ void test1() {
     CALL(zip, close);
   };
 
-  PrintError();
   talloc_free(zip);
 };
 
 #define TIMES 1000
 
+/** This tests the cache for reading zip members */
 void test2() {
   // Open the file for reading
   FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con, NULL, "new_test.zip", 'r');
@@ -73,13 +73,30 @@ void test2() {
   printf("Decompressed foobar %d times in %d mseconds (%f)\n", 
 	 TIMES,  diff,
 	 ((float)diff)/TIMES);
-  PrintError();
+
   talloc_free(fd);
+};
+
+/** We we try to create a new stream */
+void test3() {
+  // Open the file for reading
+  FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con, NULL, "new_test.zip", 'r');
+  // Open the zip file
+  FIFFile fif = CONSTRUCT(FIFFile, ZipFile, super.Con, fd, (FileLikeObject)fd);
+  AFFFD image = CALL(fif, create_stream_for_writing, "default", "Image");
+  
 };
 
 int main() {
   ClearError();
   test1();
+  PrintError();
+
   ClearError();
   test2();
+  PrintError();
+
+  ClearError();
+  test3();
+  PrintError();
 };

### 2009-02-23 src/lib/libfif.c

@@ -1,0 +1,58 @@
+#include "fif.h"
+
+/** This is a dispatches of stream classes depending on their name */
+struct dispatch_t dispatch[] = {
+  { "Image", (AFFFD)(&__Image) },
+  { NULL, NULL}
+};
+
+/** We need to call these initialisers explicitely or the class
+    references wont work.
+*/
+static void init_streams(void) {
+  FileLikeObject_init();
+  AFFFD_init();
+  Image_init();
+};
+
+AFFFD FIFFile_create_stream_for_writing(FIFFile self, char *stream_name,
+					char *stream_type) {
+  int i;
+
+  init_streams();
+
+  for(i=0; dispatch[i].type !=NULL; i++) {
+    if(!strcmp(dispatch[i].type, stream_type)) {
+      AFFFD result;
+
+      // A special constructor from a class reference
+      result = CONSTRUCT_FROM_REFERENCE(dispatch[i].class_ptr, 
+					Con, self, stream_name);
+      return result;
+    };
+  };
+
+  RaiseError(ERuntimeError, "No handler for stream_type '%s'", stream_type);
+  return NULL;
+};
+
+VIRTUAL(FIFFile, ZipFile)
+     VMETHOD(create_stream_for_writing) = FIFFile_create_stream_for_writing;
+END_VIRTUAL
+
+AFFFD AFFFD_Con(AFFFD self, char *stream_name) {
+  return self;
+};
+
+VIRTUAL(AFFFD, FileLikeObject)
+     VMETHOD(Con) = AFFFD_Con;
+END_VIRTUAL
+
+AFFFD Image_Con(AFFFD self, char *stream_name) {
+  printf("Constructing an image %s\n", stream_name);
+  return self;
+};
+
+VIRTUAL(Image, AFFFD)
+     VMETHOD(super.Con) = Image_Con;
+END_VIRTUAL

### 2009-02-24 src/include/fif.h
More work towards implementing the new Image stream.

@@ -1,14 +1,73 @@
+#include "config.h"
 #include "class.h"
 #include "talloc.h"
 #include "zip.h"
+#include "stringio.h"
+
+struct FIFFile;
+
+/** Properties are key value pairs. We preserve their order though */
+CLASS(Properties, Object)
+     struct list_head list;
+     char *key;
+     char *value;
+     uint32_t date;
+     
+     /** We can create a properties object from a text file which we
+	 can parse.
+     */
+     Properties METHOD(Properties, Con);
+
+     /** Add new key value to the list, returns the value.
+     */
+     char *METHOD(Properties, add, char *key, char *value, uint32_t date);
+
+     /* An iterator over the properties list to return the list of values
+	of the given key.
+     */
+     char *METHOD(Properties, iter_next, Properties *current, char *key);
+
+     /** Parses the text file and append to our list */
+     int METHOD(Properties, parse, char *text, uint32_t date);
+
+     /** Dumps the current properties array into a string. Callers own the memory. */
+     char *METHOD(Properties, str);
+END_CLASS
 
 /* A AFFFD is a special FileLikeObject for openning AFF2 streams */
 CLASS(AFFFD, FileLikeObject)
-     AFFFD METHOD(AFFFD, Con, char *stream_name);
+// We store properties for the stream here
+     Properties props;
+
+     // This is FIFFile who owns.
+     struct FIFFile *parent;
+  
+     AFFFD METHOD(AFFFD, Con, char *stream_name, 
+		  Properties props, struct FIFFile *parent);
 END_CLASS
 
 /** The Image Stream represents an Image in chunks */
 CLASS(Image, AFFFD)
+// Data is divided into segments, when a segment is completed (it
+// contains chunks_in_segment chunks) we dump it to the archive. These
+// are stream attributes
+     int chunks_in_segment;   // Default 2048
+     int chunk_size;          // Default 32kb   
+                              // -> Default segment size 64Mb
+
+     // Writes get queued here until full chunks are available
+     StringIO chunk_buffer;
+
+     // The segment is written here until its complete and then it
+     // gets flushed
+     StringIO segment_buffer;
+     int chunk_count;
+
+     int segment_count;
+
+     // An array of indexes into the segment where chunks are stored
+     uint32_t *chunk_indexes;
+
 END_CLASS
 
 struct dispatch_t {
[…]

### 2009-02-24 src/include/misc.h

@@ -10,4 +10,6 @@
 #define O_BINARY 0
 #endif
 
+#define BUFF_SIZE 4096
+
 #endif

### 2009-02-24 src/include/stringio.h

@@ -1,0 +1,61 @@
+#ifndef _STRINGIO_H
+#define _STRINGIO_H
+#include "config.h"			       
+#include "class.h"
+#include <sys/types.h>
+#include <stdint.h>
+
+CLASS(StringIO,Object)
+  /** This is the size of the internal buffer */
+  int size;
+  /** Current readptr */
+  uint64_t readptr;
+  char *data;
+  
+  /** constructor */
+  StringIO METHOD(StringIO, Con);
+
+  /** Writes data into the string_io at the current offset, growing the
+    string_io if needed **/
+  int METHOD(StringIO, write, char *data, int len);
+
+  /** Write a format string into the stringio **/
+  int METHOD(StringIO, sprintf, char *fmt, ...);
+
+  /** Reads data from the current string location into the buffer (We
+      presume it is large enough. We return how much data was actually
+      read */
+  int METHOD(StringIO, read, char *data, int len);
+
+  /** These allow us to read and write to StringIOs rather than direct
+    buffers */
+  int METHOD(StringIO, write_stream, StringIO stream, int length);
+  int METHOD(StringIO, read_stream, StringIO stream, int length);
+
+  /** The seek method */
+  uint64_t METHOD(StringIO, seek, int64_t offset, int whence);
+
+  /** get_buffer: Returns a pointer/length to the buffer (relative to readptr) */
+  void METHOD(StringIO, get_buffer, char **data, int *len);
+
+  /** Return true if we are at the end of the file */
+  int METHOD(StringIO, eof);
+
+  /** Truncates the end of the stream to this size */
+  void METHOD(StringIO, truncate, int len);
+
+  /** Removes the first len bytes from the start of the stream. The
+      stream is repositioned at its start */
+  void METHOD(StringIO, skip, int len);
+
+  /** find a substring, returns a pointer inside data */
+  char *METHOD(StringIO, find, char *string);
+
+  /** case insensitive version of find */
+  char *METHOD(StringIO, ifind, char *string);
+
+  /** Destructor */
+  void METHOD(StringIO, destroy);
+END_CLASS
+
+#endif

### 2009-02-24 src/include/zip.h

@@ -4,9 +4,6 @@
 #include "list.h"
 #include <zlib.h>
 
-#define BUFF_SIZE 4096
-
-
 /** These are ZipFile structures */
 struct EndCentralDirectory {
   uint32_t magic;
@@ -140,6 +137,10 @@ CLASS(ZipFile, Object)
 // This method flushes the central directory and finalises the
 // file. The file may still be accessed for reading after this.
      void METHOD(ZipFile, close);
+
+// A convenience function for storing a string as a new file
+     void METHOD(ZipFile, writestr, char *filename, char *data, int len, int compression);
+
 END_CLASS
 
 #define ZIP_STORED 0

### 2009-02-24 src/lib/fifinfo.c

@@ -77,17 +77,79 @@ void test2() {
   talloc_free(fd);
 };
 
-/** We we try to create a new stream */
+#if 0
+/** We try to create a new stream */
 void test3() {
   // Open the file for reading
   FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con, NULL, "new_test.zip", 'r');
   // Open the zip file
   FIFFile fif = CONSTRUCT(FIFFile, ZipFile, super.Con, fd, (FileLikeObject)fd);
-  AFFFD image = CALL(fif, create_stream_for_writing, "default", "Image");
+  AFFFD image = CALL(fif, create_stream_for_writing, "default", "Image", NULL);
   
 };
+#endif
+
+/** This tests the properties classes */
+void test4() {
+  Properties test = CONSTRUCT(Properties, Properties, Con, NULL);
+  Properties i=NULL;
+
+  CALL(test, add, "key", "value", 0);
+  CALL(test, add, "key", "value", 0);
+  CALL(test, add, "key", "something else", 0);
+  CALL(test, add, "key2", "value", 0);
+  CALL(test, parse, "key=foobar\nhello=world\n",100);
+
+  while(1) {
+    char *value=CALL(test, iter_next, &i, "key");
+    if(!value) break;
+    printf("Got value %s\n", value);
+  };
+
+  // Now dump the whole thing:
+  printf("'%s'", CALL(test, str));	 
+
+  talloc_free(test);
+};
+
+/** Test the Image class */
+void test5() {
+  // Make a new zip file
+  FIFFile fiffile = CONSTRUCT(FIFFile, ZipFile, super.Con, NULL, NULL);
+
+  // Make a new file
+  FileBackedObject fd = CONSTRUCT(FileBackedObject, FileBackedObject, con,
+				  fiffile, "new_test.zip", 'w');
+
+  // Create a new properties object
+  Properties props = CONSTRUCT(Properties, Properties, Con, fiffile);
+
+  FileLikeObject stream;
+  int out_fd = open("/bin/ls", O_RDONLY);
+  char buffer[BUFF_SIZE * 10];
+  int length;
+
+  // Create a new Zip volume for writing
+  fiffile->super.create_new_volume((ZipFile)fiffile, (FileLikeObject)fd);
+
+  // Make a new Image stream
+  stream = (FileLikeObject)CALL(fiffile, create_stream_for_writing, "default","Image", props);
[…]

### 2009-02-24 src/lib/libfif.c

@@ -1,4 +1,5 @@
 #include "fif.h"
+#include "time.h"
 
 /** This is a dispatches of stream classes depending on their name */
 struct dispatch_t dispatch[] = {
@@ -16,7 +17,7 @@ static void init_streams(void) {
 };
 
 AFFFD FIFFile_create_stream_for_writing(FIFFile self, char *stream_name,
-					char *stream_type) {
+					char *stream_type, Properties props) {
   int i;
 
   init_streams();
@@ -27,7 +28,7 @@ AFFFD FIFFile_create_stream_for_writing(FIFFile self, char *stream_name,
 
       // A special constructor from a class reference
       result = CONSTRUCT_FROM_REFERENCE(dispatch[i].class_ptr, 
-					Con, self, stream_name);
+					Con, self, stream_name, props, self);
       return result;
     };
   };
@@ -40,7 +41,7 @@ VIRTUAL(FIFFile, ZipFile)
      VMETHOD(create_stream_for_writing) = FIFFile_create_stream_for_writing;
 END_VIRTUAL
 
-AFFFD AFFFD_Con(AFFFD self, char *stream_name) {
+AFFFD AFFFD_Con(AFFFD self, char *stream_name, Properties props, FIFFile parent) {
   return self;
 };
 
@@ -48,11 +49,200 @@ VIRTUAL(AFFFD, FileLikeObject)
      VMETHOD(Con) = AFFFD_Con;
 END_VIRTUAL
 
-AFFFD Image_Con(AFFFD self, char *stream_name) {
+AFFFD Image_Con(AFFFD self, char *stream_name, Properties props, FIFFile parent) {
+  Image this = (Image)self;
+  char *tmp;
+
   printf("Constructing an image %s\n", stream_name);
+  if(!props) props = CONSTRUCT(Properties, Properties, Con, self);
+  
+  // Set the defaults
+  tmp = CALL(props, iter_next, NULL, "chunks_in_segment");
+  if(!tmp) tmp = CALL(props, add, "chunks_in_segment", "2048", time(NULL));
+  this->chunks_in_segment = strtol( tmp, NULL, 0);
+
+  tmp = CALL(props, iter_next, NULL, "chunk_size");
+  if(!tmp) tmp = CALL(props, add, "chunk_size", "0x8000", time(NULL));
+
+  this->chunk_size = strtol(tmp, NULL,0);
+    
[…]

### 2009-02-24 src/lib/stringio.c

@@ -1,0 +1,201 @@
+#include "stringio.h"
+#include "misc.h"
+
+StringIO StringIO_constructor(StringIO self) {
+  //Create a valid buffer to hold the data:
+  self->data = talloc(self,char);
+  self->size = 0;
+  self->readptr=0;
+  
+  return self;
+};
+
+int StringIO_write(StringIO self,char *data, int len) {
+  if(self->readptr+len > self->size) {
+    self->size = self->readptr + len;
+    
+    self->data = talloc_realloc_size(self,self->data,self->size+1);
+  };
+  
+  memcpy(self->data+self->readptr,data,len);
+  self->readptr+=len;
+  
+  return len;
+};
+
+int StringIO_sprintf(StringIO self, char *fmt, ...) {
+  va_list ap;
+  char *data;
+  int len;
+  
+  va_start(ap, fmt);
+  data = talloc_vasprintf(self, fmt, ap);
+  va_end(ap);
+  len = strlen(data);
+  
+  if(self->readptr+len > self->size) {
+    self->size = self->readptr + len;    
+    self->data = talloc_realloc_size(self,self->data,self->size+1);
+  };
+  
+  memcpy(self->data+self->readptr,data,len);
+  self->readptr+=len;
+  talloc_free(data);
+  return len;
+};
+
+int StringIO_read(StringIO self,char *data,int len) {
+  if(self->readptr+len > self->size) {
+    len = self->size-self->readptr;
+  };
+
+  memcpy(data,self->data+self->readptr,len);
+  self->readptr+=len;
+  return(len);
+};
+
+/** Writes into ourselves from a stream */
+int StringIO_read_stream(StringIO self, StringIO stream, int length) {
+  int len;
+  char buff[BUFF_SIZE];
+
+  while(length > 0) {
+    len = CALL(stream, read, buff, min(length, BUFF_SIZE));
+    if(len==0) break;
+
+    CALL(self, write, buff, len);
+    length -= len;
+  };
+
+  return length;
+
+  // This is too error prone if we have complex stringio classes.
+#if 0  
+  stream->get_buffer(stream,&data,&len);
+
+  //Only write whats available:
+  if(length>len) length=len;
+
+  self->write(self,data,length);
+
+  //Move the input stream by that many bytes:
+  stream->seek(stream,length,SEEK_CUR);
+
+  return length;
+#endif
+};
+
[…]

### 2009-02-24 src/lib/zip.c

@@ -500,12 +500,23 @@ void ZipFile_close(ZipFile self) {
   };
 };
 
+
+/** This is just a convenience function - real simple now */
+void ZipFile_writestr(ZipFile self, char *filename, 
+		      char *data, int len, int compression) {
+  FileLikeObject fd = CALL(self, open_member, filename, 'w', compression);
+
+  CALL(fd, write, data, len);
+  CALL(fd, close);
+};
+
 VIRTUAL(ZipFile, Object)
      VMETHOD(Con) = ZipFile_Con;
      VMETHOD(read_member) = ZipFile_read_member;
      VMETHOD(create_new_volume) = ZipFile_create_new_volume;
      VMETHOD(open_member) = ZipFile_open_member;
      VMETHOD(close) = ZipFile_close;
+     VMETHOD(writestr) = ZipFile_writestr;
 END_VIRTUAL
 
 

### 2009-03-09 src/lib/unittests.c
Finished the Map stream implementation.

@@ -341,8 +341,8 @@ void test_map_create() {
   map = (MapDriver)CALL(oracle, create, (AFFObject *)&__MapDriver);
   if(map) {
     CALL((AFFObject)map, set_property, "aff2:stored", volume);
-    CALL((AFFObject)map, set_property, "aff2:image_period", "3");
-    CALL((AFFObject)map, set_property, "aff2:file_period", "6");
+    CALL((AFFObject)map, set_property, "aff2:target_period", "3");
+    CALL((AFFObject)map, set_property, "aff2:image_period", "6");
     CALL((AFFObject)map, set_property, "aff2:blocksize", "64k");
     map = (MapDriver)CALL((AFFObject)map, finish);
   };
@@ -361,12 +361,23 @@ void test_map_create() {
   CALL(map, add, 5, 2, D2);
 
   fd = (FileLikeObject)CALL(oracle, open, NULL, D1);
-  CALL((AFFObject)map, set_property, "size", from_int(fd->size * 2));
+  CALL((AFFObject)map, set_property, "aff2:size", from_int(fd->size * 2));
   CALL(oracle, cache_return, (AFFObject)fd);
 
   CALL(map, save_map);
   CALL((FileLikeObject)map, close);
   CALL(oracle, cache_return, (AFFObject)map);
+
+  // Make a link to the map:
+  {
+    Link link;
+    
+    link = (Link)CALL(oracle, create, (AFFObject *)&__Link);
+    // The link will be stored in this zipfile
+    CALL(link, link, oracle, volume, URNOF(map), "map");
+    CALL((AFFObject)link, finish);
+    CALL(oracle, cache_return, (AFFObject)link);
+  };
 
  exit:
   zipfile = (ZipFile)CALL(oracle, open, NULL, volume);
@@ -374,8 +385,37 @@ void test_map_create() {
   CALL(oracle, cache_return, (AFFObject)zipfile);
 };
 
+#define TEST_BUFF_SIZE 63*1024
+void test_map_read() {
+  MapDriver map;
+  int outfd, length;
+  char buff[TEST_BUFF_SIZE];
+  ZipFile zipfile = CONSTRUCT(ZipFile, ZipFile, Con, NULL, "file://" TEST_FILE);
+
+  CALL(oracle, cache_return, (AFFObject)zipfile);
+  map  = (MapDriver)CALL(oracle, open, NULL, "map");
+  if(!map) return;
+
+  outfd = creat("output.dd", 0644);
[…]

### 2009-03-10 src/include/misc.h
Added a libcurl implemented HTTP File Like object. Now we can automatically work of remove http archives.

@@ -2,12 +2,27 @@
 #define MISC_H
 #include "config.h"
 
-#ifdef HAVE_INTTYPES_H
-#include <inttypes.h>
+#ifdef WINDOWS
+#include <winsock2.h>
+#include <ws2tcpip.h>
+#include <io.h>
+#include <stdio.h>
+#define MSG_NOSIGNAL 0
+typedef  unsigned long int in_addr_t;
+#else
+#include <sys/socket.h>
+#include <netinet/tcp.h>
+#include <netinet/in.h>
+#include <netdb.h>
+#include <arpa/inet.h>
+#include <stdio.h>
+#include <unistd.h>
+
+#define O_BINARY 0
 #endif
 
-#ifndef O_BINARY
-#define O_BINARY 0
+#ifdef HAVE_INTTYPES_H
+#include <inttypes.h>
 #endif
 
 #define BUFF_SIZE 4096

### 2009-03-10 src/include/zip.h

@@ -141,6 +141,16 @@ CLASS(FileBackedObject, FileLikeObject)
      FileBackedObject METHOD(FileBackedObject, Con, char *filename, char mode);
 END_CLASS
 
+#include <curl/curl.h>
+
+// This is implemented in HTTP
+CLASS(HTTPObject, FileLikeObject)
+// The socket
+     CURL *curl;
+     StringIO buffer;
+
+     HTTPObject METHOD(HTTPObject, Con, char *url);
+END_CLASS
 
      /** The resolver is at the heart of the AFF2 specification - its
 	 responsible with returning various objects from a globally

### 2009-03-10 src/lib/http.c

@@ -1,0 +1,124 @@
+#include "zip.h"
+
+/******************************************************************
+   This is an implementation of a simple HTTP File Like Object.
+
+We basically use libcurl to do the heavy lifting so we might actually
+get to use ftp://, http://, https:// and more all for free.
+
+******************************************************************/
+static int close_fd(void *self) {
+  HTTPObject this = (HTTPObject)self;
+  curl_easy_cleanup(this->curl);
+
+  return 0;
+};
+
+static size_t write_callback(void *ptr, size_t size, size_t nmemb, void *self){ 
+  HTTPObject this = (HTTPObject)self;
+
+  CALL(this->buffer, write, ptr, size*nmemb);
+  return size*nmemb;
+};
+
+// We want to get this header, Content-Range: bytes 0-100/3388
+static size_t parse_header_callback(void *ptr, size_t size, size_t nmemb, void *self) {
+  HTTPObject this = (HTTPObject)self;
+  int length = size * nmemb;
+  char *string;
+  char *x;
+  static char *content_range = "Content-Range: bytes ";
+  uint64_t tmp,tmp_size;
+
+  string = talloc_size(self, length +1);
+  // Null terminate the string
+  memcpy(string, ptr, length);
+  x=strstr(string, content_range);
+  if(x && x==string &&
+     3 == sscanf(string + strlen(content_range), "%lld-%lld/%lld", 
+		 &tmp, &tmp, &tmp_size)) {
+    ((FileLikeObject)this)->size = tmp_size;
+  };
+
+  talloc_free(string);
+  return length;
+};
+
+static AFFObject HTTPObject_AFFObject_Con(AFFObject self, char *uri) {
+  HTTPObject this = (HTTPObject)self;
+
+  return (AFFObject)CALL(this, Con, uri);
+};
+
+
+static HTTPObject HTTPObject_Con(HTTPObject self,
+				 char *url) {
+  // Parse out the url
+  CURL *handle = self->curl = curl_easy_init();
+  self->buffer = CONSTRUCT(StringIO, StringIO, Con, self);
+  URNOF(self) = talloc_strdup(self, url);
+
+  //  curl_easy_setopt( handle, CURLOPT_VERBOSE, 1L ); 
+  curl_easy_setopt( handle, CURLOPT_URL, url);
+
[…]

### 2009-03-10 src/lib/resolver.c

@@ -16,6 +16,10 @@ static struct dispatch_t dispatch[] = {
   { "image", (AFFObject)&__Image },
   { "map", (AFFObject)&__MapDriver},
   { "file://", (AFFObject)&__FileBackedObject },
+  // All handled by libcurl
+  { "http://", (AFFObject)&__HTTPObject },
+  { "https://", (AFFObject)&__HTTPObject },
+  { "ftp://", (AFFObject)&__HTTPObject },
   { NULL, NULL}
 };
 
@@ -34,6 +38,7 @@ void AFF2_Init(void) {
   Blob_init();
   Resolver_init();
   Link_init();
+  HTTPObject_init();
 
   init_luts();
 

### 2009-03-10 src/lib/unittests.c

@@ -386,11 +386,11 @@ void test_map_create() {
 };
 
 #define TEST_BUFF_SIZE 63*1024
-void test_map_read() {
+void test_map_read(char *filename) {
   MapDriver map;
   int outfd, length;
   char buff[TEST_BUFF_SIZE];
-  ZipFile zipfile = CONSTRUCT(ZipFile, ZipFile, Con, NULL, "file://" TEST_FILE);
+  ZipFile zipfile = CONSTRUCT(ZipFile, ZipFile, Con, NULL,filename);
 
   CALL(oracle, cache_return, (AFFObject)zipfile);
   map  = (MapDriver)CALL(oracle, open, NULL, "map");
@@ -411,6 +411,14 @@ void test_map_read() {
  error:
   CALL(oracle, cache_return, (AFFObject)map);
   return;
+};
+
+void test_http_handle() {
+  FileLikeObject http = CONSTRUCT(HTTPObject, HTTPObject, Con,NULL, "http://127.0.0.1/test.c");
+  char buff[BUFF_SIZE];
+
+  CALL(http, read, buff, 100);
+
 };
 
 int main() {
@@ -441,11 +449,16 @@ int main() {
   printf("\n*******************\ntest 5\n********************\n");
   test_map_create();
   PrintError();
+
+  AFF2_Init();
+  ClearError();
+  test_map_read( "file://" TEST_FILE);
+  PrintError();
  */
 
   AFF2_Init();
   ClearError();
-  test_map_read();
+  test_map_read("http://127.0.0.1/" TEST_FILE);
   PrintError();
 
   return 0;

### 2009-03-10 paper/paper.tex
simson1: Ignore-this: cd3848f5f55f9e0c9da0835e6bc0d37c
Initial work with the new section describing changes to AFF.

@@
 %\documentclass{elsart5p}
+
+% DFRWS Call Information:
+% http://www.dfrws.org/2009/cfp.shtml
+% Important Dates
+% Submission deadline: March 16, 2009 (any time zone)
+% Author notification: April 28, 2009
+% Final draft due: May 19, 2009
+% Pre-conference Workshops: August 16, 2009
+% Conference dates: August 17-19, 2009
+
+% Publication Criteria
+% Research papers must be original contributions, not substantially
+% duplicate previous work, and must not be under simultaneous
+% publication review elsewhere. The review process will be
+% ``double-blind'' (the reviewers will not know who the authors are, and
+% the authors will not know who the reviewers are). Therefore, the
+% version submitted for review should not contain the names or
+% affiliations of the authors. When referring to their own previous
+% work, authors should use the third person instead of the first person
+% (i.e. ``Smith and Jones [2] previously determined...'' instead of ``We
+% [2] previously determined..''). Authors are expected to present their
+% work in person at the workshop and must have at least one registration
+% per paper in order to be included in the proceedings.
 
@@
 \begin{document}
-\title{The Forensic Interchange Format - a file format for digital evidence}
-\author{M. I. Cohen\footnote{scudette@gmail.com} and Bradley Schatz}
+\title{Extending the Advanced Forensic Format to accommodate Multiple
+  Data Sources, Logical Evidence and Forensic Workflow}
+\author{Blinded for review}
+%\author{M. I. Cohen\footnote{scudette@gmail.com}, Simson Garfinkel, and Bradley Schatz}
 %\author{M. I. Cohen}
@@
 %\end{frontmatter}
-%\thanks{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}%
+%\thanks{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
 \maketitle
@@
 Forensics analysis requires the acquisition and management of many
-different types of evidence material. Often the same evidence is
-reviewed by several different tools. We propose an open, extensible
+different types of evidence material, including individual disk
+drives, RAID sets, network packets, memory images, and extracted files. Often the same evidence is
+reviewed by several different tools or examiners in different
+locations. We propose a backwards-compatible evolutionary redesign of the Advanced
+Forensic Format an open, extensible
 file format for storing of evidence and interchanging analysis results
@@
 \section{Introduction}
+
+Unlike other kinds of evidence, digital evidence can be instantly lost
+or corrupted if technical measures are not immediately taken to
+preserve its existence and integrity. As a result, preserving digital
+evidence is an important part of most digital
+investigations\cite{carrier:event-based}.
+
+In recent years there has been a steady and growing interest in the
+actual file formats and containers used to store this digital
+evidence. Early practitioners created exact bit-for-bit copies (``mirror disks'') employed
+proprietary software systems (e.g. \cite{safeback,ilook,encase}) for
+making and authenticating ``images'' of digital evidence. PyFlag\cite{pyflag}
+introduced a ``seekable gzip'' format that allowed disk images to be
+stored in a form that was compressed but allowed the random-access to
+evidence data necessary for forensic analysis. The Advanced Forensic
+Format (AFF) expanded on this idea with a forensic file format that
+allowed both data and arbitrary metadata to be stored in a single
+digital archive\cite{garfinkel:aff}. Recently the open source AFF
+implementation library (AFFLIB) was expanded to support both
+encryption and digital signatures of digital evidence\cite{garfinkel:affcrypto}.
+
+This paper extends the previous work by adapting AFF to the multiple
+heterogeneous data types that might arise in a modern digital
+investigation, including data from multiple data storage devices, new
+data types (including network packets and memory images), extracted
+logical evidence, and forensic workflow. We do this in a manner that
+is at once upwards compatible with AFF while adopting the AFF
+bit-level specification so that AFF files can be processed using
+standard non-forensic tools.
+
+\section{A New Forensic Format}
+
+AFF's flexibility comes from the fact that forensic data and metadata
+is stored as arbitrary name/value pairs called \emph{segments}. For example, the first 16MB
+of a disk image is stored in a segment called \texttt{page0}, the
+second 16MB in a segment called \texttt{page1}, \emph{etc.} Because of
+this flexibility, it was
+relatively easy for Garfinkel to extend AFF to support encryption,
+digital signatures, and the storage of new kinds of metadata such as
+chain-of-custody information\cite{garfinkel:affcrypto}. .
+
+\subsection{AFF Limitations}
+After imaging literally thousands of devices using Garfinkel's AFF
+tools and adding support for AFF to numerous open source programs, we
+observed numerous problems in the underlying standard
+and Garfinkel's AFFLIB implementation:
+
+\begin{itemize}
+\item While AFF's design stores a single disk image in each evidence
+  file, modern digital investigations typically involve many seized
+  computers or pieces of media. 
+\item AFF has no provision for storing memory images or intercepted
+  network packets.
+\item AFF has no provisions for storing extracted files that is
+  analogous to the EnCase ``Logical Evidence File'' (L01) format.
+\item AFF's encryption system leaks information about the contents of
+  an evidence file because segment names are not encrypted.
+\item AFF's default compression page size of 16MB can impose significant overhead
+  when accessing NTFS Master File Tables (MFT), as these structures
+  tend to be highly fragmented on systems that have seen significant
+  use. 
+\item Although the AFF specification calls for a ``table of contents'' similar
+  to the ZIP\cite{zip-format} ``central directory'' that is stored at the end of AFF
+  files, Garfinkel never implemented this directory in the publicly
+  released AFFLIB. As a result, every header of every segment in an
+  AFF file needs to be read when a file is opened. It practice this
+  can take 10-30 seconds the first time a large AFF file is
+  opened. (Once the file is opened, the sectors corresponding to the
+  segment headers are stored in the computer's disk cache.)
+\item AFF's bit-level specification is essentially a simple container
+  file specification. Given the that there are other container file
+  specifications that are much more widely supported with both
+  developer and end-user tools, it seemed reasonable to migrate AFF
+  from its home-grown format to one of the existing standards.
+\end{itemize}
+
+We have addressed these issues with a three step process. First, we
+have extended the AFF schema to support multiple forensic images
+(which are now called \emph{streams}) in a single evidence
+file. Second, we have changed the on disk format, jettisoning AFF1.0's
+unique container format and replacing it with the ZIP64 file
+format. Third, we are adopting AFFLIB to implement the new schema and
+file format, which will allow existing AFF-aware programs to access
+files written in the new format without modification. We call this new
+format AFF2.
+
+\subsection{AFF2: New Features}
+AFF2 is designed to include the following functionality:
+
+\begin{itemize}
+\item Evidence is stored in \emph{volumes}. A volume can consist of a
+  single file (\emph{e.g.} a file with the \texttt{.aff} extension), multiple files (\texttt{.afd}), an
+  object storage system (such as Amazon S3), or a hybrid (for example,
+  metadata stored in an SQL database with physical data stored in
+  large files in a file system.)
+\item A single volume can contain evidence from one or more physical
+  devices.
+\item RAID information can be directly encoded, allowing data within a
+  RAID to be logically constructed at read time.
+\item Although the page size remains 16MB by default, each 16MB page
+  can be built from multiple compressed chunks (8K size by default),
+  significantly decreasing the overhead to access a few blocks inside
+  a compressed page.
+\item The bit-level container file format has changed from AFFs's
+  original container file format to ZIP64, allowing improved
+  integration with existing developer tools.
+\item Transparent upwards compatibility, so that programs linked with
+  the AFFLIB implementation or the pyaff.py python library can
+  interchangeably read either AFF1 or AFF2 files. 
+\end{itemize}
 

### 2009-03-11 src/lib/unittests.c
Updated a figure in the paper.

@@ -53,7 +53,7 @@ void test1() {
 
 When creating a new AFFObject we:
 
-1) Ask the oracle to create it (prividing the class pointer).
+1) Ask the oracle to create it (providing the class pointer).
 
 2) Set all the required and optional parameters.
 

### 2009-03-11 paper/paper.tex
simson2 Ignore-this: ecbd6ae9f39c5531d3004f8a6bfb7faf
Simson is taking a typing break. This commit is in the middle of a
massive rewrite and refactoring of the paper. Please don't make
substantive changes yet.

@@
 \usepackage{draftwatermark}
+\SetWatermarkFontSize{3cm}
 \usepackage{verbatim}
@@
 \usepackage{epsfig}
-\usepackage[hypertexnames=false,bookmarksopenlevel=1,bookmarksopen,bookmarksnumbered,colorlinks,plainpages=false,linktocpage]{hyperref}
+\usepackage[hypertexnames=false,bookmarksopenlevel=1,bookmarksopen,bookmarksnumbered,colorlinks,plainpages=false,linktocpage,linkcolor=black,citecolor=black,filecolor=black,urlcolor=black]{hyperref}
 %\usepackage{natbib}
@@
 bit-level specification so that AFF files can be processed using
-standard non-forensic tools.
+standard non-forensic tools.  We call the new system AFF4, and use the
+phrase AFF1 to refer to the legacy system developed by
+Garfinkel.\footnote{Although Garfinkel never changed the AFF bit-level
+specification, Garfinkel released AFFLIB implementations with major
+version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid confusion.}
 
-\section{A New Forensic Format}
+\begin{figure}
+\fbox{
+\begin{minipage}{.95\linewidth}
+\begin{itemize}
+\item The ability to store evidence from one or more physical
+  devices.
+\item Schema allowing evidence files to contain memory
+  images, network packet intercepts, Netflow data, and extracted
+  files.
+\item Direct encoding of RAID configuration, allowing RAID volumes
+  to be automatically reassembled from evidence files made of the
+  physical RAID devices.
+\item A revised compression system that allows individual chunks
+  within a 16MB segment to be decompressed without decompressing the
+  entire segment. 
+\item Support for Global Distributed Evidence Management.
+\end{itemize}
+\end{minipage}}
+\caption{Specific design goals for the next generation forensic file
+  format.}
+\end{figure}
+\section{The Need for an Improved Forensic Format}
 
-AFF's flexibility comes from the fact that forensic data and metadata
-is stored as arbitrary name/value pairs called \emph{segments}. For example, the first 16MB
-of a disk image is stored in a segment called \texttt{page0}, the
-second 16MB in a segment called \texttt{page1}, \emph{etc.} Because of
-this flexibility, it was
-relatively easy for Garfinkel to extend AFF to support encryption,
-digital signatures, and the storage of new kinds of metadata such as
-chain-of-custody information\cite{garfinkel:affcrypto}. .
+AFF1's flexibility came from the fact that forensic data and metadata
+is stored as arbitrary name/value pairs called \emph{segments}. For
+example, the first 16MB of a disk image is stored in a segment called
+\texttt{page0}, the second 16MB in a segment called \texttt{page1},
+\emph{etc.} Because of this flexibility, it was relatively easy for
+Garfinkel to extend AFF1 to support encryption, digital signatures, and
+the storage of new kinds of metadata such as chain-of-custody
+information\cite{garfinkel:affcrypto}.
 
 \subsection{AFF Limitations}
-After imaging literally thousands of devices using Garfinkel's AFF
-tools and adding support for AFF to numerous open source programs, we
+After imaging literally thousands of devices using Garfinkel's AFF1
+tools and adding support for AFF1 to numerous open source programs, we
 observed numerous problems in the underlying standard
@@
 \begin{itemize}
-\item While AFF's design stores a single disk image in each evidence
+\item While AFF1's design stores a single disk image in each evidence
   file, modern digital investigations typically involve many seized
   computers or pieces of media. 
-\item AFF has no provision for storing memory images or intercepted
+\item AFF1 has no provision for storing memory images or intercepted
   network packets.
-\item AFF has no provisions for storing extracted files that is
-  analogous to the EnCase ``Logical Evidence File'' (L01) format.
-\item AFF's encryption system leaks information about the contents of
+\item AFF1 has no provisions for storing extracted files that is
+  analogous to the EnCase ``Logical Evidence File'' (L01) format, or
+  for linking evidence to web pages.
+\item AFF1's encryption system leaks information about the contents of
   an evidence file because segment names are not encrypted.
-\item AFF's default compression page size of 16MB can impose significant overhead
+\item AFF1's default compression page size of 16MB can impose significant overhead
   when accessing NTFS Master File Tables (MFT), as these structures
@@
   use. 
-\item Although the AFF specification calls for a ``table of contents'' similar
+\item Although the AFF1 specification calls for a ``table of contents'' similar
   to the ZIP\cite{zip-format} ``central directory'' that is stored at the end of AFF
   files, Garfinkel never implemented this directory in the publicly
-  released AFFLIB. As a result, every header of every segment in an
+  released AFF1 implementation, AFFLIB. As a result, every header of every segment in an
   AFF file needs to be read when a file is opened. It practice this
-  can take 10-30 seconds the first time a large AFF file is
+  can take 10--30 seconds the first time a large AFF file is
   opened. (Once the file is opened, the sectors corresponding to the
   segment headers are stored in the computer's disk cache.)
-\item AFF's bit-level specification is essentially a simple container
+\item AFF1's bit-level specification is essentially a simple container
   file specification. Given the that there are other container file
@@
 
-We have addressed these issues with a three step process. First, we
-have extended the AFF schema to support multiple forensic images
-(which are now called \emph{streams}) in a single evidence
-file. Second, we have changed the on disk format, jettisoning AFF1.0's
-unique container format and replacing it with the ZIP64 file
-format. Third, we are adopting AFFLIB to implement the new schema and
-file format, which will allow existing AFF-aware programs to access
-files written in the new format without modification. We call this new
-format AFF2.
+\subsection{Global Distributed Evidence Management}
+While Garfinkel designed AFF for use on a single machine that could
+both image evidence and perform analysis, the main use of AFF in
+practice has been in distributed environments in which imaging and
+analysis takes place in multiple locations and is performed by
+multiple individuals.
 
-\subsection{AFF2: New Features}
-AFF2 is designed to include the following functionality:
+Managing evidence in a globally distributed system requires the use of
+globally unique identifiers. The original AFF specification assigned
+each piece of evidence a unique 128-bit identifier called a GID but did
+not make it clear when this identifier should be changed and when it
+should remain the same.
 
-\begin{itemize}
-\item Evidence is stored in \emph{volumes}. A volume can consist of a
-  single file (\emph{e.g.} a file with the \texttt{.aff} extension), multiple files (\texttt{.afd}), an
-  object storage system (such as Amazon S3), or a hybrid (for example,
-  metadata stored in an SQL database with physical data stored in
-  large files in a file system.)
-\item A single volume can contain evidence from one or more physical
-  devices.
-\item RAID information can be directly encoded, allowing data within a
-  RAID to be logically constructed at read time.
-\item Although the page size remains 16MB by default, each 16MB page
-  can be built from multiple compressed chunks (8K size by default),
-  significantly decreasing the overhead to access a few blocks inside
-  a compressed page.
-\item The bit-level container file format has changed from AFFs's
-  original container file format to ZIP64, allowing improved
-  integration with existing developer tools.
-\item Transparent upwards compatibility, so that programs linked with
-  the AFFLIB implementation or the pyaff.py python library can
-  interchangeably read either AFF1 or AFF2 files. 
-\end{itemize}
+In practice, however, global distributed evidence management requires
+more than simply tracking the movement of disk images: it requires
+approaches for sharing evidence to multiple disconnected
+evidence, allowing offline work, and then seamlessly recombining the
+work products of the analysts in a third security domain.
 
-\section{FIF Specifications}
-The FIF specifications distinct between bit level specifications and
-a high level schema. The bit level specifications refer to how the
-files are to be actually represented on disk. The schema specifies a
-stream based abstraction which is implemented using the bit level
-specification.
+Consider the typical usage scenario depicted in Figure \ref{usage}, of
+an file which contains a disk image. This volume is distributed to two
+independent analysts, Alice and Bob. Alice may find and extract
+individual files, while Bob may correlate information in the evidence
+file with other data that is available on departmental
+servers. Although in some environments Alice and Bob may be able to
+work on a shared file that is located on a server, in other
+environments there will not be sufficient bandwidth or
+connectivity. Instead, each analyst will be required to store the
+information in their own evidence file; these files will then be
+recombined at a later point in time.
 
-\subsection{FIF bit level specification}
-The FIF specifications only require the underlying storage mechanism
-be capable of storing multiple named segments of data. The current
-specification uses the Zip file format as the underlying storage
-mechanism, however in future versions of the specifications other
-storage layers may be also be defined.
+In this case they can
+each create a new volume which extends to original volume and save
+their analysis on this new volume. Now they only need to share this
+new volume with other analysts who also have a copy of the old volume
+to interchange their findings.
 
-The Zip file specifications are publicly available and widely
-implemented. The Zip64 extensions are provided to enable Zip archives
-to grow beyond a two gigabyte limit by using 64 bit
-references. Although Zip64 also defines encryption and authentication
-extensions, we do not use these in this specification due to the
-restrictions imposed on their use. Rather, we specify stream based
-encryption scheme for use in FIF files (See section
-\ref{crypted_stream}).
-
-The basic structure of a Zip archive is shown in Figure
-\ref{zip_structure}. For bit level details see the Zip file
-Application notes \cite{zipspecs}.
+This is made possible because each volume is independent of one
+another, but can still be recombined into a bigger evidence set.
 
@@
   \begin{center}
-  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
-  \caption{The basic structure of a Zip archive}
-  \label{zip_structure}
+  \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}
+  \caption{A typical usage scenario. Both Alice and Bob receive an AFF
+  volume but work independently. Rather than modifying the volume,
+  they each create their own local volumes and save their results into
+  those files. They can now exchange the smaller new volumes and
+  effectively merge their results into the same AFF set when they are finished.}
+  \label{usage}
   \end{center}
@@
 
-As can be seen, the archive consists of a {\em Central Directory (CD)}
+\subsection{Migration from AFF1 to AFF4}
+
+We have addressed these issues in three steps. First, we have extended
+the AFF schema to support multiple forensic images (which are now
+called \emph{streams}) in a single evidence file. Second, we replaced
+AFF's ``bit-level'' format with an improved format that is more
+flexible and has better existing support. Third, we are adopting
+AFFLIB to implement the new schema and file format, which will allow
+existing AFF-aware programs to access files written in either the old
+or the new format without modification.
+
+We call the result of our work effort AFF4.
+
+\section{Introducing AFF4}
+
+\subsection{AFF4 Terminology and New Concepts}
+
+This section discusses the AFF4 terminology.
+
+\begin{itemize}
+\item \emph{An Evidence Volume} is a single file with a \texttt{.aff}
+  extension. 
+\item \emph{An Evidence Set} is a collection of one or more Evidence
+  Volumes. Evidence Sets can also be stored in object storage system (such as
+  Amazon S3), or a hybrid---for example, metadata stored in an SQL
+  database with physical data stored in large files in a file
+  system. 
+\item \emph{An Evidence Set UUID} is a unique identifier generated
+  according to RFC4122\cite{rfc4122} for each new Evidence
+  Set. Multiple Evidence Volumes that make up the same Evidence Set
+  should all have the same UUID. When an Evidence Set is opened all its volumes are
+  loaded, their UUIDs are checked, and all of their contents are
+  logically merged so as to be visible as if they were all found in the same volume.
+\item \emph{A segment} is a single unit of data written to volume. AFF4
+  segments have a \emph{segment name}, a \emph{segment timestamp}, and
+  a \emph{segment contents}.
+
+\item \emph{A stream} is a collection of segments that can be
+  concatenated together to form a single data object.
+
+\item \emph{An image stream} is a stream that is dedicated to storing evidence from a single
+  source of data.  For example, a stream may contain a disk image or a network
+  capture. Whereas AFF could only support a single stream in an evidence
+  file, AFF4 allows any number of streams to be stored. Each stream in
+  the archive is  given a unique identifier.
+
+\item \emph{A bevy} is a specific segment that is used to hold
+  forensic image data, such as disk sectors, memory pages, or network
+  packets. AFF1 used the term ``pages'' to describe bevys, but this
+  term caused confusion when AFF was used to hold memory images, hence the
+  new term.
+
+\item \emph{A segment name} is the name under which a segment is
+  stored within the archive. AFF4 segments have two-part names that
+  typically contain a \emph{stream name} and \emph{bevy number}. For
+  example the segment named \texttt{default/000001} refers to the stream
+  called \texttt{default} and has a chunk name of \texttt{000001}. All segment
+  names are encoded using UTF8. AFFLIB is being modified to allow any
+  stream to be opened using the \texttt{af\_open} function; if a
+  stream name is not specified, AFF4 implementations will open the
+  stream named ``default.''
+
+\item \emph{The segment timestamp} is a timestamp for the segment
+  recorded in GMT.
+
+\item \emph{Attributes} are a list of key/value pairs stored in each
+stream's ``properties'' segment (e.g. ``stream/properties''). The
+``properties'' file is a text file which stores attributes one on each
+line. Attribute lines consist of the key, immediately followed by
+``='' for text values or ``:'' for binary values. This is immediately
+followed by the encoded value. Values containing text are encoded
+using UTF8, while values containing binary data are encoded using
+Base64. There may be multiple values for each attribute; in these
+cases the order is preserved
+by implementations. Implements may present different keys may be
+presented in any order, however.
+\vspace{1ex}
+Each AFF4 archive must also have a top-level set of attributes---that
+is, there is always a segment named ``properties.'' These volume
+attributes are used to specify a Unique Identifier (UUID). (AFF1
+stored each property in its own named segment. This created
+significant overhead in many situations.)
+
+\item \emph{A Target Reference} is a way of referencing objects. This is used
+in a number of places which will be highlighted below. Using
+references it is possible for the FIF file to refer to objects
+embedded within itself, or within other FIF files. Its is also
+possible to refer to external flat files. A number of usage scenarios
+are described in Section \ref{usage_scenarios}. 
+\end{itemize}
+
+The Target Reference is an important AFF4 innovation: this mechanism
+gives AFF4 files a way for describing evidence that is stored in
+different evidence volumes, in the host computer's file system, or on
+the Internet. We have defined the following target reference schema:
+
+\label{target_reference}
+
+\begin{itemize}
+\item A reference in the format {\em aff://UUID/stream/segment} refers
+to an external Evidence file set (identified by its UUID). If the segment
+is missing (i.e. the reference ends with a ``/''), the reference is
+for the entire stream. We do not specify a way of actually resolving
+the UUID into a physical locations. This can achieved by
+implementations using many methods, such as Active Directory, LDAP,
+DNS, flat files or SQL databases.
+
+\item A reference starting with {\em file://} refers to an external
+file. The external file reference must be sanitised (to remove
+.. directory traversal) and rooted at the current directory. A
+filename of ``.'' refers to the current FIF volume.
+
+\item A reference starting with {\em http://} refers to a HTTP
+object which will be fetched. Support for this is optional.
+
+\item A references ending with a ``/'' signifies a reference to a
+stream in the current FIF file. The stream will be opened in its
+entirety. If the reference does not end with a ``/'', the reference is
+for a single segment.
+\end{itemize}
+
+\subsection{AFF4 Streams}
+As discussed above, individual segments in an AFF4 evidence set can be
+combined into a single logical object called a \emph{stream}. AFF4
+supports an arbitrary number of streams per volume. The AFFLIB library
+allows the segments in a stream to be operated on as if they were a
+single file inside a file system by supporting the traditional
+POSIX-like functionality of \texttt{open()}, \texttt{seek()} and
+\texttt{read()}.
+
+AFFLIB also supports the \emph{writing} of streams, a feature that is
+used by Garfinkel's \texttt{aimage} disk imaging program. In general
+arbitrary writing to streams can be expensive, especially when writing
+is combined with seek operations. The current \texttt{aimage} program
+will create AFF files with gaps in them during its error recovery
+process (when \texttt{aimage} seeks to the end of a disk and reads it
+backwards in an attempt to get around media failures). We are working
+on a rewrite of \texttt{aimage} which performs imaging and recovery
+on bevy boundaries, rather than using the current arbitrary
+``high-water-mark'' and ``low-water-mark.''
+
+All streams within an Evidence Set \emph{must} have a properties segment. The
+properties segment describes what the stream contains and determines
+how the stream is processed.
+
+\subsection{The Image Stream}
+\label{image_stream}
+The AFF4 \emph{Image} stream contains a single read-only forensic data
+set. For example, this stream might contains a hard
+disk image, a memory image or a network capture (in PCAP
+format). Random read access into the stream is provided with
+AFFLIB. (AFFLIB also provided the ability to randomly \emph{write}
+into the evidence stream for sanitization purposes, although this
+feature may be removed at a later date.)
+
+Image streams must have at several attributes. First, they must have
+\texttt{type=image} to indicate that they are image streams. Streams
+must have a \texttt{size} attribute (specified in 8-bit bytes) to
+indicate the last addressable byte in the stream. Streams can have
+holes in them; in this case, attempts to read data from within the
+holes can generate an error or return NULLs.
+
+The image itself is stored in \emph{pages} within the AFF4 file. Pages are
+segments named as an 8 digit, zero padded decimal integer
+representation of their chunk id (e.g. ``default/00000032''). Image
+streams must specify the \texttt{page\_size} attribute, as the number of
+image bytes each chunk contains. Pages may be compressed at the
+page level, or may be broken up into multiple chunks (typically 8KB),
+and individually compressed. In this case the beginning of the page
+segment will contain an offset table indicating the location of each
+compression chunk. This is similar to the technique used by the Expert
+Witness file format used by EnCase\cite{encase-3.0} and implemented by
+the open source libewf\cite{libewf} package.
+
+Pages can optionally be individually hashed, allowing for detection
+of modifications. The attribute \texttt{hash\_type} specifies the type of
+hash used (implementations will typically support md5,sha1 and sha256, but can
+define others). The hashes are stored in any number of segments within
+the stream, of arbitrary name. The hash segment names is provided by
+using the \texttt{hash\_name} stream attribute.
+
+Hashes are calculated in an identical way to that defined by AFF1's
+crypto layer\cite{garfinkel:affcrypto}---that is, the full name of the
+segment (including stream component) is followed by a single NULL
+byte, followed by the (uncompressed) content of the chunk. The
+inclusion of the segment name ensure that even identical chunks in
+content will have a different hash and prevents reordering attacks.
+
+Each hash segment is a list of a chunk id encoded in 32 bit
+little endian integer format followed by the hash (the length of which
+depends on the hash itself). This approach is illustrated in Figure
+\ref{hash_index}.
+
+Hash segments can contain any number of chunk hashes in them. It is
+recommended that a hash segment be created when a volume is closed to
+cover all the chunks within that volume. This allows for hashes to be
+verified even if some volumes are missing. The AFFLIB implementation
+may update a hash of a chunk previously set by an earlier hash
+segment---this would be required when a chunk is updated. The hash
+segment can then be signed to assure the integrity of the updated segment.
+
+\begin{figure}[tb]
+  \begin{center}
+  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
+  \caption{The structure of a hashing segment. The attributes {\em
+  hash} and {\em hash\_type} declare the hash segment. The segment
+  contains a list of integer encoded chunk id followed by an unencoded
+  hash for each chunk.}
+  \label{hash_index}
+  \end{center}
+\end{figure}
+
+This design differs from AFF1, which creates a separate segment for
+each hash. AFF4 uses headers that are more verbose than the AFF
+segment headers. Therefore, creating a new segment for each hash value
+would result in too much overhead: Packing the hashes into a few
+segments it is possible to reduce this overhead to a minimum.
+
+\subsection{ZIP64}
+For AFF4, we have decided to change AFF1's
+original container file format to ZIP64\cite{zipspecs}. There are many
+reasons for this decision:
+
+\begin{itemize}
+\item There is already wide support for the ZIP and ZIP64 formats. By
+  migrating to these formats, we can take advantage of the rich number
+  of user and developer tools already available.
+\item ZIP64 includes a CRC32 and timestamp for each segment. We
+  believe that the additional integrity checks and timestamps will be
+  useful in a forensic environment.
+\item ZIP64 already includes support for the ``table of contents''
+  envisioned but never implemented by Garfinkel; ZIP64 calls this
+  table the ``Central Directory.'' It is written at the end of the
+  ZIP64 volume.
+\end{itemize}
+
+Figure~\ref{zip_structure} shows the basic structure of a Zip
+archive. As can be seen, the archive consists of a {\em Central Directory} (CD)
 located at the end of the archive. The CD is a list of pointers to
@@
 
-\subsection{FIF Definitions}
-The FIF specification defines a number of basic terms:
+\begin{figure}[tb]
+  \begin{center}
+  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
+  \caption{The basic structure of a Zip archive}
+  \label{zip_structure}
+  \end{center}
+\end{figure}
+
+It is important to note that AFF4 only requires that the underlying storage mechanism
+be capable of storing multiple named segments of data. Although our
+AFF4 implementation will use the ZIP64 file format as an underlying storage
+mechanism, our system also supports legacy AFF1 volumes as well as
+segments that are stored in other storage systems---for example,
+Amazon.com's Simple Storage System (S3)\cite{s2-aws-home-page-money}.
+
+\subsubsection{Unused ZIP64 Features}
+
+There are a number of aspects in the Zip file format specifications
+which we do not utilize. 
 \begin{itemize}
-\item {\em A FIF Volume} is a single ZIP archive containing {\em
-segments}.
+\item We ignore Zip64's built-in support for splitting
+archives into multiple ZIP files. Instead, our implementation treats
+each volume as complete and stand-alone ZIP file. The AFF4
+implementation then considers the segments contained within as 
+belonging to a larger collection. The reason behind this decision is that
+often it is needed to add new volumes to an existing set without
+modifying the existing set at all. Zip64's built-in support for
+splitting an archive renders each individual volume in the set
+unusable unless all are available.
 
-\item {\em A segment} is a single unit of data written to the Zip
-archive as an archive file. The only properties of segments of
-interest to the FIF specification are the segment name and the
-timestamp.
-
-\item {\em The segment name} is the name under which a segment is
-stored within the archive. This consists of a directory part named
-{\em a stream} and an {\em chunk name}. For example the segment named
-``default/000001'' refers to the stream called ``default'' and has a
-chunk name of ``000001''. Although the ZIP specification does not
-specify the encoding of segment names, the FIF specification requires
-all segment names to be encoded using UTF8. If stream name is
-unspecified, implementations should open the stream named ``default''.
-
-\item {\em Attributes} are a list of key/value pairs stored at each
-stream's ``properties'' segment (e.g. ``stream/properties''). The
-``properties'' file is a text file which stores attributes one on each
-line. Attribute lines consist of the key, immediately followed by
-``='' for text values or ``:'' for binary values. This is immediately
-followed by the encoded value. Values containing text MUST be encoded
-using UTF8, while values containing binary data MAY be encoded using
-Base64. If keys are repeated, their relative order MUST be preserved
-by implementations, but different keys may be presented in any order.
-
-Attributes MUST be specified at the top level of each volume
-(i.e. must have a segment named ``properties''). The volume attributes
-are used to specify the volume UUID.
-
-%about the stream. A properties file also exists at the top
-%level of the ZIP hierarchy to store global information such as the
-%volumes UUID, the FIF version and other volumes which make up the
-
-\item {\em A FIF file} is a collection of volumes with the same
-UUID. The UUID is generated according to RFC4122 for each new FIF
-file. As new volumes are created for the same file, the UUID is copied
-to all related volumes. When a FIF file is opened, all its volumes are
-loaded and merged in such a way that all the segments from all the
-volumes are visible as if they were all found in the same volume.
+\item Although Zip64 also defines encryption and authentication
+extensions, we do not use them due to the
+restrictions imposed on their use and because they lack of functionality that is
+important for forensic user. Instead, we use AFF1's digital signature
+facilities for integrity and non-repudiation, and we introduce a new  stream based
+encryption scheme for ensuring data privacy (Section~\ref{crypted_stream}).
 
@@
 
-\label{target_reference}
-{\em A target reference} is a way of referencing objects. This is used
-in a number of places which will be highlighted below. Using
-references it is possible for the FIF file to refer to objects
-embedded within itself, or within other FIF files. Its is also
-possible to refer to external flat files. A number of usage scenarios
-are described in Section \ref{usage_scenarios}. The reference is
-specified with the following conventions:
+\subsubsection{A Forensicly Sound Strategy For Updating ZIP64 Files}
 
+In many forensic applications, individual segments within the volume
+need to be updated. To this end, Garfinkel implemented several
+``update'' functions in AFF1\cite{garfinkel:aff}.
+
+There were two problems with the original AFF updating approach:
 \begin{itemize}
-\item A reference starting with {\em file://} refers to an external
-file. The external file reference must be sanitised (to remove
-.. directory traversal) and rooted at the current directory. A
-filename of ``.'' refers to the current FIF volume.
-
-\item A reference starting with {\em http://} refers to a HTTP
-object which will be fetched. Support for this is optional.
-
-\item A reference in the format {\em fif://UUID/stream/segment} refers
-to an external FIF file set (identified by its UUID). If the segment
-is missing (i.e. the reference ends with a ``/''), the reference is
-for the entire stream. We do not specify a way of actually resolving
-the UUID into a physical locations. This can achieved by
-implementations using many methods, such as Active Directory, LDAP,
-DNS, flat files or SQL databases.
-
-\item A references ending with a ``/'' signify a reference to a
-stream in the current FIF file. The stream will be opened in its
-entirety. If the reference does not end with a ``/'', the reference is
-for a single segment.
+\item Updates can cause the volume to
+become fragmented as holes appear within the middle of the volume, and
+updated segments are added at the end.
+\item Updates could be performed silently, raising questions of
+  forensic soundness.
 \end{itemize}
 
-\subsection{FIF Volumes and segment merging}
-There are a number of aspects in the Zip file format specifications
-which we do not utilize. For example, the Zip64 specification allows
-for a multitude of encryption and hashing algorithms. Some of these
-features however, seem to be patent encumbered. Since the FIF
-specification aims to separate high level information such as hashes,
-cryptography and signatures from the low bit level representation, we
-implement those features at a higher level.
-
-In many forensic applications, individual segments within the volume
-need to be updated \cite{AFF}. These updates can cause the volume to
-become fragmented as holes appear within the middle of the volume, and
-updated segments are added at the end.
-
-When we wish to add a new segment to an existing volume, the new
-segment header SHOULD overwrite the volume central directory, and the
+In AFF4, when a new segment is added to an existing volume, the new
+segment header overwrites the volume central directory, and the
 central directory should be rewritten including the new entry
@@
 old central directory remains in the file). This is illustrated in
-Figure \ref{zip_structure}. Implementations may choose to delay
-writing the central directory until the file is closed.
+Figure \ref{zip_structure}. Implementations may delay
+writing the central directory until the file is closed; should the
+system crash before the central directory is written, it can be
+recovered by scanning the entire ZIP file.
 
@@
 segment is added to the archive with the same name and a later
-timestamp. Implementations MUST satisfy read requests for duplicate
+timestamp. Implementations satisfy read requests for duplicate
 segment names with those segments with the later timestamp, or if the
@@
 within each volume can be used for auditing, or to allow a rollback to
-previous versions of the segment. The {\em frepack} tool is provided
-to repack, split and manage FIF volumes by creating new volumes and
-redistributing segments among them.
+previous versions of the segment. We have created a new tool,
+\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
+creating new volumes and redistributing segments among them.
 
-Note that we do not rely on Zip64's built-in support for splitting
-archives at all (although implementations should be able to read such
-files). Each volume is a complete and stand alone ZIP file. It is the
-FIF implementation which considers the segments contained within it to
-belong to a larger collection. The reason behind this decision is that
-often it is needed to add new volumes to an existing set without
-modifying the existing set at all.
+\subsubsection{Segment merging}
 
-Consider the typical usage scenario depicted in Figure \ref{usage}, of
-a FIF file which contains a disk image. This volume is distributed to
-two independent analysts, which wish to add new streams to the FIF
-file (for example the result of their analysis). In this case they can
-each create a new volume which extends to original volume and save
-their analysis on this new volume. Now they only need to share this
-new volume with other analysts who also have a copy of the old volume
-to interchange their findings.
-
-\begin{figure}[tb]
-  \begin{center}
-  \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}
-
-  \caption{A typical usage scenario. Both Alice and Bob receive a FIF
-  volume but work independently. Rather than modifying the volume,
-  they each create another volume and save their results to the new
-  volume. They can now exchange the smaller new volumes and
-  effectively merge their results into the same FIF set.}
-
-  \label{usage}
-  \end{center}
-\end{figure}
-
-This is made possible because each volume is independent of one
-another, and can be loosely collated into a bigger FIF file
-set.
-
-\subsection{FIF Streams}
-The above section explained how FIF file sets can be split into
-volumes, each containing segments. When the entire FIF file set is
-loaded, segments may be fetched from any volume. The present section
-describes how these individual segments are used to form {\em stream
-abstractions}.
-
-A stream is a high level abstraction of a single source of data. For
-example, a stream may contain a disk image or a network
-capture. Although we do not specify the exact API, implementations
-MUST make a stream opened for reading have at a minimum a read, seek
-and tell methods. Seeking MUST support the whence argument as
-described in the standard C implementation (i.e. allow seeking to the
-end of the stream).
-
-Streams opened for writing MAY limit support for seeking, as efficient
-seeking in compressed streams may not be possible. Applications should
-avoid needing to seek in output streams, and try to structure writes
-serially. It is usually possible to reorder a sequence of non
-overlapping random writes into an Image stream and a Map Stream (These
-are discussed in Sections \ref{image_stream} and
-\ref{map_stream}). Implementations are free to hide the inability to
-seek in writing streams by automatically implementing such a scheme,
-but this is not mandated by the specification.
-
-All streams within the FIF file MUST have a properties segment. The
-properties segment describes the stream attributes as specified
-previously. Streams must have the {\em type} attribute which allows
-implementations to determine how they should be processed.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Standard FIF segments}
@@
 further types. Implementations MUST implement the basic types.
-
-\subsection{The Image Stream}
-\label{image_stream}
-The {\em Image} stream ({\em type=image}) contains a single read only
-forensic data source. For example, this Stream might contains a hard
-disk image, a memory image or a network capture (in PCAP format). The
-Image stream provides random read access into the stream (i.e. allows
-for randomly seeking the stream to any point). The {\em size}
-attribute MUST be specified to determine how large the stream is.
-
-The image itself is stored in chunks within the FIF file. Chunks are
-segments named as an 8 digit, zero padded decimal integer
-representation of their chunk id (e.g. ``default/00000032''). Image
-streams must specify the {\em chunk\_size} attribute, as the number of
-image bytes each chunk contains. Chunks may be compressed at the
-underlying Zip container level, using any of the compression
-algorithms specified by the zip specifications.
-
-Chunks can optionally be individually hashed, allowing for detection
-of modifications. The attribute {\em hash\_type} specifies the type of
-hash used (implementations SHOULD support md5,sha1 and sha256, but can
-define others). The hashes are stored in any number of segments within
-the stream, of arbitrary name. The hash segment names is provided by
-using the {\em hash} stream attribute.
-
-Each hash segment MUST be a list of a chunk id encoded in 32 bit
-little endian integer format followed by the hash (the length of which
-depends on the hash itself). This approach is illustrated in Figure
-\ref{hash_index}.
-
-Hash segments can contain any number of chunk hashes in them. It is
-recommended that a hash segment be created when a volume is closed to
-cover all the chunks within that volume. This allows for hashes to be
-verified even if some volumes are missing. It is permitted for a hash
-segment to update a hash of a chunk previously set by an earlier hash
-segment - this may be required when a chunk is updated. The hash
-segment can then be signed to assure the integrity of the updated
-segment.
-
-\begin{figure}[tb]
-  \begin{center}
-  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
-
-  \caption{The structure of a hashing segment. The attributes {\em
-  hash} and {\em hash\_type} declare the hash segment. The segment
-  contains a list of integer encoded chunk id followed by an unencoded
-  hash for each chunk.}
-
-  \label{hash_index}
-  \end{center}
-\end{figure}
-
-This design differs from AFF, which creates a separate segment for
-each hash. The Zip headers are more verbose than the AFF segment
-headers, and typically FIF chunks are much smaller (default 32kb, vs
-16Mb for AFF). Therefore, creating a new segment for each hash value
-results in more overheads in the FIF format. By packing the hashes
-into a few segments it is possible to reduce this overhead to a
-minimum.
-
-Hashes are calculated in an identical way to that defined by AFF,
-thats is the full name of the segment (including stream component) is
-followed by a single NULL byte, followed by the (uncompressed) content
-of the chunk. The inclusion of the segment name ensure that even
-identical chunks in content will have a different hash.
 
@@
 
+\section{Conclusion and Future Work}
+\subsection{Expanding AFFLIB for AFF4}
+Although we do not specify the exact API, implementations
+MUST make a stream opened for reading have at a minimum a read, seek
+and tell methods. Seeking MUST support the whence argument as
+described in the standard C implementation (i.e. allow seeking to the
+end of the stream).
+
 \bibliography{IEEEabrv,paper}

### 2009-03-11 paper/paper.tex
simson3 Ignore-this: 1c077e496d797e18ac61488f2b7b568a
Significant work done by Simson. Going to sleep now.

@@
 \SetWatermarkFontSize{3cm}
-\usepackage{verbatim}
+\usepackage{fancyvrb}
 %\documentclass[12pt]{article}
@@
 
-\subsection{AFF4 Terminology and New Concepts}
+\subsection{New Concepts and Terminology}
 
@@
 
-\subsection{AFF4 Streams}
+\subsection{Streams}
 As discussed above, individual segments in an AFF4 evidence set can be
@@
 how the stream is processed.
+
+\subsection{ZIP64}
+For AFF4, we have changed AFF1's original container file format to
+ZIP64\cite{zipspecs}. There are many reasons for this decision:
+
+\begin{itemize}
+\item There is already wide support for the ZIP and ZIP64 formats. By
+  migrating to these formats, we can take advantage of the rich number
+  of user and developer tools already available.
+\item ZIP64 includes a CRC32 and timestamp for each segment. We
+  believe that the additional integrity checks and timestamps will be
+  useful in a forensic environment.
+\item ZIP64 already includes support for the ``table of contents''
+  envisioned but never implemented by Garfinkel; ZIP64 calls this
+  table the ``Central Directory.'' It is written at the end of the
+  ZIP64 volume.
+\end{itemize}
+
+Figure~\ref{zip_structure} shows the basic structure of a Zip
+archive. As can be seen, the archive consists of a {\em Central Directory} (CD)
+located at the end of the archive. The CD is a list of pointers to
+individual {\em File header} structures located within the body of the
+archive. Headers are then followed by the file data, after it has been
+compressed by the appropriate compression method (as specified in the
+header). Each archived file is optionally followed by a {\em Data
+Descriptor} describing the length and CRC of the archived file. Using
+the data descriptor field allows implementations to write archives
+without needing to seek in the output file. This allows Zip files to
+be written to pipes for example, sending an image over the network
+using netcat or ssh.
+
+\begin{figure}[tb]
+  \begin{center}
+  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
+  \caption{The basic structure of a Zip archive}
+  \label{zip_structure}
+  \end{center}
+\end{figure}
+
+It is important to note that AFF4 only requires that the underlying storage mechanism
+be capable of storing multiple named segments of data. Although our
+AFF4 implementation will use the ZIP64 file format as an underlying storage
+mechanism, our system also supports legacy AFF1 volumes as well as
+segments that are stored in other storage systems---for example,
+Amazon.com's Simple Storage System (S3)\cite{s2-aws-home-page-money}.
+
+\subsubsection{Unused ZIP64 Features}
+
+There are a number of aspects in the Zip file format specifications
+which we do not utilize. 
+\begin{itemize}
+\item We ignore Zip64's built-in support for splitting
+archives into multiple ZIP files. Instead, our implementation treats
+each volume as complete and stand-alone ZIP file. The AFF4
+implementation then considers the segments contained within as 
+belonging to a larger collection. The reason behind this decision is that
+often it is needed to add new volumes to an existing set without
+modifying the existing set at all. Zip64's built-in support for
+splitting an archive renders each individual volume in the set
+unusable unless all are available.
+
+\item Although Zip64 also defines encryption and authentication
+extensions, we do not use them due to the
+restrictions imposed on their use and because they lack of functionality that is
+important for forensic user. Instead, we use AFF1's digital signature
+facilities for integrity and non-repudiation, and we introduce a new  stream based
+encryption scheme for ensuring data privacy (Section~\ref{crypted_stream}).
+
+\end{itemize}
+
+\subsubsection{A Forensically Sound Strategy For Updating ZIP64 Files}
+
+In many forensic applications, individual segments within the volume
+need to be updated. To this end, Garfinkel implemented several
+``update'' functions in AFF1\cite{garfinkel:aff}.
+
+There were two problems with the original AFF updating approach:
+\begin{itemize}
+\item Updates can cause the volume to
+become fragmented as holes appear within the middle of the volume, and
+updated segments are added at the end.
+\item Updates could be performed silently, raising questions of
+  forensic soundness.
+\end{itemize}
+
+In AFF4, when a new segment is added to an existing volume, the new
+segment header overwrites the volume central directory, and the
+central directory should be rewritten including the new entry
+immediately after the last file in the archive. If it is not possible
+to overwrite the file, the new segment can be written at the end,
+followed by an updated copy of the central directory (in that case the
+old central directory remains in the file). This is illustrated in
+Figure \ref{zip_structure}. Implementations may delay
+writing the central directory until the file is closed; should the
+system crash before the central directory is written, it can be
+recovered by scanning the entire ZIP file.
+
+A segment within the archive is considered to be updated when a new
+segment is added to the archive with the same name and a later
+timestamp. Implementations satisfy read requests for duplicate
+segment names with those segments with the later timestamp, or if the
+two timestamps are the same, the segment which appears at a later
+offset in the volume. This effectively allows for a segment to be
+updated without leading to internal volume fragmentation.
+
+The fact that the older version of all updated segments remains intact
+within each volume can be used for auditing, or to allow a rollback to
+previous versions of the segment. We have created a new tool,
+\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
+creating new volumes and redistributing segments among them.
+
+\section{Standard AFF4 Streams}
+In this section we discuss specific stream types that we have created
+in our initial AFF4 implementation.
 
@@
 
-\subsection{ZIP64}
-For AFF4, we have decided to change AFF1's
-original container file format to ZIP64\cite{zipspecs}. There are many
-reasons for this decision:
-
-\begin{itemize}
-\item There is already wide support for the ZIP and ZIP64 formats. By
-  migrating to these formats, we can take advantage of the rich number
-  of user and developer tools already available.
-\item ZIP64 includes a CRC32 and timestamp for each segment. We
-  believe that the additional integrity checks and timestamps will be
-  useful in a forensic environment.
-\item ZIP64 already includes support for the ``table of contents''
-  envisioned but never implemented by Garfinkel; ZIP64 calls this
-  table the ``Central Directory.'' It is written at the end of the
-  ZIP64 volume.
-\end{itemize}
-
-Figure~\ref{zip_structure} shows the basic structure of a Zip
-archive. As can be seen, the archive consists of a {\em Central Directory} (CD)
-located at the end of the archive. The CD is a list of pointers to
-individual {\em File header} structures located within the body of the
-archive. Headers are then followed by the file data, after it has been
-compressed by the appropriate compression method (as specified in the
-header). Each archived file is optionally followed by a {\em Data
-Descriptor} describing the length and CRC of the archived file. Using
-the data descriptor field allows implementations to write archives
-without needing to seek in the output file. This allows Zip files to
-be written to pipes for example, sending an image over the network
-using netcat or ssh.
-
-\begin{figure}[tb]
-  \begin{center}
-  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
-  \caption{The basic structure of a Zip archive}
-  \label{zip_structure}
-  \end{center}
-\end{figure}
-
-It is important to note that AFF4 only requires that the underlying storage mechanism
-be capable of storing multiple named segments of data. Although our
-AFF4 implementation will use the ZIP64 file format as an underlying storage
-mechanism, our system also supports legacy AFF1 volumes as well as
-segments that are stored in other storage systems---for example,
-Amazon.com's Simple Storage System (S3)\cite{s2-aws-home-page-money}.
-
-\subsubsection{Unused ZIP64 Features}
-
-There are a number of aspects in the Zip file format specifications
-which we do not utilize. 
-\begin{itemize}
-\item We ignore Zip64's built-in support for splitting
-archives into multiple ZIP files. Instead, our implementation treats
-each volume as complete and stand-alone ZIP file. The AFF4
-implementation then considers the segments contained within as 
-belonging to a larger collection. The reason behind this decision is that
-often it is needed to add new volumes to an existing set without
-modifying the existing set at all. Zip64's built-in support for
-splitting an archive renders each individual volume in the set
-unusable unless all are available.
-
-\item Although Zip64 also defines encryption and authentication
-extensions, we do not use them due to the
-restrictions imposed on their use and because they lack of functionality that is
-important for forensic user. Instead, we use AFF1's digital signature
-facilities for integrity and non-repudiation, and we introduce a new  stream based
-encryption scheme for ensuring data privacy (Section~\ref{crypted_stream}).
-
-\end{itemize}
-
-\subsubsection{A Forensicly Sound Strategy For Updating ZIP64 Files}
-
-In many forensic applications, individual segments within the volume
-need to be updated. To this end, Garfinkel implemented several
-``update'' functions in AFF1\cite{garfinkel:aff}.
-
-There were two problems with the original AFF updating approach:
-\begin{itemize}
-\item Updates can cause the volume to
-become fragmented as holes appear within the middle of the volume, and
-updated segments are added at the end.
-\item Updates could be performed silently, raising questions of
-  forensic soundness.
-\end{itemize}
-
-In AFF4, when a new segment is added to an existing volume, the new
-segment header overwrites the volume central directory, and the
-central directory should be rewritten including the new entry
-immediately after the last file in the archive. If it is not possible
-to overwrite the file, the new segment can be written at the end,
-followed by an updated copy of the central directory (in that case the
-old central directory remains in the file). This is illustrated in
-Figure \ref{zip_structure}. Implementations may delay
-writing the central directory until the file is closed; should the
-system crash before the central directory is written, it can be
-recovered by scanning the entire ZIP file.
-
-A segment within the archive is considered to be updated when a new
-segment is added to the archive with the same name and a later
-timestamp. Implementations satisfy read requests for duplicate
-segment names with those segments with the later timestamp, or if the
-two timestamps are the same, the segment which appears at a later
-offset in the volume. This effectively allows for a segment to be
-updated without leading to internal volume fragmentation.
-
-The fact that the older version of all updated segments remains intact
-within each volume can be used for auditing, or to allow a rollback to
-previous versions of the segment. We have created a new tool,
-\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
-creating new volumes and redistributing segments among them.
-
-\subsubsection{Segment merging}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Standard FIF segments}
-The FIF specification defines a number of basic types, but extensions
-are encouraged. Future versions of the specification might standardize
-further types. Implementations MUST implement the basic types.
-
 \subsection{The Map Stream}
@@
 analysis. For example, a file is often simply a collection of bytes
-drawn from an image. A TCP/IP stream is simply a collection of
-payloads from selected network packets. Sometimes the same data may be
-viewed in a number of ways - for example a Virtual Address Space is a
+drawn from an image, while a TCP/IP stream is simply a collection of
+payloads from selected network packets. 
+Sometimes the same data may be
+viewed in a number of ways---for example a Virtual Address Space is a
 mapping of the Physical Address Space through a page table
-transformation. It is, therefore, beneficial to define a way in which
-the underlying storage system can apply transformations to existing
-data to produce transformed data.
+transformation.  Zero Storage Carving \cite{Meijer2006} is a way of
+specifying carved files in terms of a sequence of blocks taken from
+the image; Cohen extended this concept to an arbitrary
+mapping function\cite{Cohen2007} .
 
-For example, Zero Storage Carving \cite{Meijer2006}, is a way of
-specifying carved files in terms of a sequence of blocks taken from
-the image. \cite{Cohen2007} extended this concept to an arbitrary
-mapping function. This is the approach chosen in this specification
-due to its flexibility.
+AFF4 defines a common way in
+which the underlying storage system can apply transformations to
+existing data to produce transformed data. This transformation is
+performed using a \emph{map stream}.
 
@@
 Denoting the stream offset by $x$, and the target offset by $y$, the
-map specifies a set of points $(X_i,Y_i,T_i)$. Read requests for a
+Map specifies a set of points $(X_i,Y_i,T_i)$. Read requests for a
 byte at a mapped stream offset $x$ can then be satisfied by reading a
@@
 
-Where $mod$ is the modulus function, while $floor$ signifies integer
-division. For example consider the following, which corresponds to a 3
-disk RAID-5 array:
-\begin{verbatim*}
+Where $mod$ is the modulus function and $floor$ signifies integer
+division. For example consider Figure~\ref{map}, which corresponds to a 3
+disk RAID-5 array.
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
 stream_period=393216
@@
 327680,131072,2
-\end{verbatim*}
+\end{Verbatim}
+\caption{A Map stream that corresponds to a 3 disk RAID-5 array}\label{ref}
+\end{figure}
 
-\subsection{The Overlay stream}
-The Overlay stream is a variant of the {\em Image} stream in that it
-allows for images to be stored in (possibly compressed) chunks. The
-main utility of the Overlay stream is in providing transparent access
-to other forensic formats, such as EWF or AFF. These formats also use
-chunk based compression to provide the ability to randomly seek while
-reading. To access these file formats we only need to have an index of
+\subsection{The Overlay Stream}
+The Overlay Stream is a variant of the {\em Image} stream that
+provides a layer of indirection between requests made for specific
+bytes (or chunks or bevys) and the actual location where the requested
+information is stored. The main utility of the Overlay stream is in
+providing transparent access to legacy forensic formats, such as EWF.
+
+To access these legacy file formats we only need to have an index of
 each chunk offset, and then directly use the overlayed files.  The {\em
-ewf2fif} tool provides such an overlay feature for EWF volumes by
+ewf2aff} tool provides such an overlay feature for EWF volumes by
 employing libewf to scan the internal data structures.
 
-Therefore, the {\em chunk\_size} attribute must be specified. In
-addition the {\em compression\_type} attribute MAY be specified. The
+Overlay stream require that a \texttt{chunk\_size} attribute be specified. In
+addition the {\em compression\_type} attribute may be specified. The
 attribute should contain an integer corresponding to a suitable Zip
@@
 ``file://.''. This can be useful for modifying an existing EWF file to
-be a FIF file. Since Zip files are normally read from the end of the
-file, but EWF and AFF are both read from the front of the file it may
-be possible to append a FIF file to the end of an EWF file without
+be a AFF4 file. Since Zip files are normally read from the end of the
+file, but EWF are both read from the front of the file, it may
+be possible to append a AFF4 Central Directory to the end of an EWF file without
 interfering with the overlayed file. In this case the converted file
-can still be used as an EWF file without change, but it can also be
-used as a FIF file.
+can still be used as an EWF file without change, but could also be
+used as an AFF4 file.
 
@@
 
-In the above figure a segment is written to the FIF volume which
+In the above figure a segment is written to the AFF4 volume which
 contains a number of chunks back to back. An overlay stream then uses
@@
 
-\subsection{The Private stream}
-Often forensic software needs to store internal state, such as the
-current state of the GUI, or internal data structures which may be
-required for caching. Usually, software create additional cache files
-to maintain this information. It is advantageous to be able to store
-these in the evidence file itself. The evidence file can then be
-re-opened by the software at a later stage and private data can be
-retrieved and used directly. Such private application data can be
-stored in a Private stream. There is no specification of what can be
-stored in the private stream, and applications with private streams
-should be able to store arbitrary segments in any format at all.
-
-It is recommended that applications name their private segment
-sufficiently accurately so as not to be confused with other
-applications, or different versions of the same application
-(e.g. ``pyflag/0.87/cache/'').
-
 \subsection{Encrypted Streams}
-Encryption is an important property in an evidence file format
-\cite{AFF}. In particular with the FIF format, multiple streams may be
-present in the file set, and often different access levels are
-desired. For example, for evidence set containing both network
-captures and disk images it may be desirable to limit access to
-streams based on legal authorizations.
+Encryption is an important property in an evidence file format. In
+particular, multiple streams may be present in the file set, and often
+different access levels are desired. For example, for evidence set
+containing both network captures and disk images it may be desirable
+to limit access to streams based on legal authorizations.
 
@@
 
-The Encrypted Stream ({\em type=encrypted}) SHOULD be implemented by
-FIF implementations, but may be omitted with an obvious loss of
-functionality (Encryption may be prohibited or restricted in some
-jurisdictions).
-
-The Encrypted Stream is similar to the Image Stream above in that data
-is written out in chunks (although its probably not effective to
-compress these chunks at all). Each chunk is to be encrypted upon
-writing and transparently decrypted using a scheme declared in the
-{\em scheme} attribute.
+AFF4 therefore introduces a new encryption scheme, the Encrypted
+Stream.  The Encrypted Stream is a stream that encapsulates other
+streams on a bevy-by-bevy basis. Each bevy transparently encrypted upon
+writing and decrypted on reading. The choice of algorithm and mode is declared in the
+Stream's \texttt{scheme} attribute.
 
 The Encrypted Stream may contain any data at all. It is useful
-however, to store an actual FIF volume within the Encrypted
-stream. This provides block level encryption for the contained FIF
-volume. It is recommended that Encrypted Streams containing a FIF
-volume set an attribute ({\em content-type=application/x-fif-volume})
+however, to store an actual AFF4 volume within the Encrypted
+stream. This provides block level encryption for the contained AFF4
+volume. It is recommended that Encrypted Streams containing a AFF
+volume set an attribute (\texttt{content-type=application/x-aff-volume})
 to allow implementations to automatically recognize volumes. This
-approach is illustrated in Figure \ref{crypted_fif}.
+approach is illustrated in Figure~\ref{crypted_fif}.
 
@@
   \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
-  \caption{Embedding an encrypted FIF volume within an Encrypted Stream.}
+  \caption{Embedding an encrypted AFF4 volume within an Encrypted Stream.}
   \label{crypted_fif}
@@
 
-The result is that a number of FIF volume are used as {\em Container
+The result is that a number of AFF4 volumes are used as {\em Container
 Volumes} to provide storage for Encrypted Streams. The main {\em
@@
 Encrypted Stream, effectively distributed throughout the container
-volumes. Note that the Container Volume may contain several Encrypted
-Streams and therefore contain multiple Encrypted FIF
+volumes. Note that the outer Volume may contain several Encrypted
+Streams and therefore contain multiple AFF4 Encrypted 
 Volumes. Container Volumes may contain non encrypted streams as well,
@@
 be implemented as only volumes which can be accessed can be merged.
-
-Implementations should create embedded FIF volumes as extensions to
-the container volumes (i.e. with the same UUID). Thus when embedded
-volumes are opened, their streams become automatically viewable as
-part of the evidence set.
 
@@
 in a configuration file, typed in at the terminal, passed in using
-process environment variables, or encoded within the FIF filename
+process environment variables, or encoded within the AFF4 filename
 itself.
 
-\subsection{Optional features}
-Implementations may implement the following features to simplify the
-interfaces.
+\subsection{Private Streams}
+Often forensic software needs to store internal state, such as the
+current state of the GUI, or internal data structures which may be
+required for caching. Usually, software create additional cache files
+to maintain this information. It is advantageous to be able to store
+these in the evidence file itself. The evidence file can then be
+re-opened by the software at a later stage and private data can be
+retrieved and used directly. Such private application data can be
+stored in a Private stream. There is no specification of what can be
+stored in the private stream, and applications with private streams
+should be able to store arbitrary segments in any format at all.
 
-\subsubsection{Automated volume loading}
-FIF volumes are effectively independent, as each volume does not need
-to know anything about other volumes. This makes it necessary to
-specify in advance all volumes for the implementation to merge them
-all. In particular it may be necessary to specify encrypted volumes
-(which may not be read if the user has no authorization for the
-Encrypted Stream).
+It is recommended that applications name their private segment
+sufficiently accurately so as not to be confused with other
+applications, or different versions of the same application
+(e.g. ``pyflag/0.87/cache/'').
 
-As an optional enhancement FIF specifies a volume attribute named {\em
-volume} (i.e. an attribute present in the top level ``properties''
-segment). This is a target reference (see section
-\ref{target_reference}) for other volumes. 
-Implementations should attempt to open external references (e.g. using
-the {\em file://} convention) before internal references. As new
-volumes are merged their volume attributes must be scanned for
-additional references. Each volume must list all other known volumes
-as volume attributes.
+\section{Implementing AFF4 in AFFLIB}
+Although there are numerous ZIP implementations available
+today, we have created our own implementation using a combination of
+available open source technology and our own novel contributions. We
+are in the process of integrating this code into the AFFLIB library
+and performing extensive regression testing.
 
-In this way, encrypted volumes, and external volumes can be located
-and found automatically.
+There are many reasons to develop our own ZIP64 implementation for
+AFF4:
 
-\section{Implementation Recommendations}
-Since the individual chunks are simply files in a Zip archive, any zip
-library can be utilized for reading these chunks. However, this may
-not be efficient since simple Zip implementations might rescan the
+\begin{itemize}
+\item Most ZIP implementations do not implement the ZIP64
+  extensions. These extensions are required to support Evidence
+  Volumes larger than 2GB.
+\item Simple Zip implementations might rescan the
 Central Directory for each segment request. Since in practice there
-can be a large number of chunks in each volume, it is recommended that
-implementations build an index of offsets and length of each chunk
-into the file. This kind of index is very similar to the strategy
-other file formats use \cite{EWF}. The index may then be saved to the
-file by use of a Private stream.
+can be a large number of segments in a volume, it is advisable to have
+a ZIP64 implementation that is optimized to storing thousands (or even
+hundreds of thousands) of segments in an efficient data structure. 
+\item While the ZIP specification duplicates data found in the Central
+  Directory entry in each File Header (such as filename, size, CRC
+  etc), many implementations that we have examined only populate this
+  information in one of these places. In the interest of robustness,
+  we wanted to assure that data stored in both locations would be
+  populated to allow recovery of at least \emph{some} evidence that
+  might exist in damaged volumes. If central directory is lost, it is
+  possible to scan through the zip file and repair the central
+  directory from the File headers.
 
-Zip64 specifies duplicate data in the Central Directory entry for each
-segment and its File Header (such as filename, size, CRC etc). Many
-implementations only populate this information in one of these
-places. In the interest of robustness, it is recommended that FIF
-implementations populate both places at the slight expense of higher
-overheads, because this can be used to recover and correct damaged
-files. If the central directory is lost, it is possible to scan
-through the zip file and repair the central directory from the File
-headers.
-
-The underlying Zip implementation should support multiple writers
-simultaneously by interleaving their writes to the file. For example,
-for two streams being written concurrently, the new segments can be
-added to the same archive in any order. In addition, the volume may be
-closed and a new volume opened without writers being aware of
-this. Since all segments from all volumes are merged, streams can be
-split arbitrarily across volumes in any way. It is recommended that
-stream properties be written to each volume to ensure at least some of
-the stream can be used even if some volumes are lost.
+\item Our implementation supports simultaneous access by multiple
+  readers and writers. Our system allows concurrency by locking the
+  file and interleaving write operations. Managing simultaneous
+  writers is simplified through our update strategy which only writes
+  new segments to the file, never modifying old segments. 
+\item AFF4 ``properties'' have specific semantics above and beyond
+  normal segments. For example, AFF4 calls for all of the property
+  segments associated with a file to be merged together. Furthermore,
+  when updating the properties in a multi-volume Set, it is useful to
+  write the same properties to each volume to ensure at least some of
+the stream can be used even if some volumes are lost. Such
+functionality is not present in standard ZIP64 implementations.
+\end{itemize}
 
@@
 \section{Conclusion and Future Work}
-\subsection{Expanding AFFLIB for AFF4}
-Although we do not specify the exact API, implementations
-MUST make a stream opened for reading have at a minimum a read, seek
-and tell methods. Seeking MUST support the whence argument as
-described in the standard C implementation (i.e. allow seeking to the
-end of the stream).
 
@@
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2009-03-11 src/include/class.h
Implemented a simple imager to create AFF2 files.

@@ -422,4 +422,8 @@ void *raise_errors(enum _error_type t, char *string,  ...);
 #define PrintError()				\
   do {if(_global_error) printf("%s",_traceback); fflush(stdout); }while(0);
 
+#define CheckError(error)			\
+  (_global_error == error)
+
+
 #endif

### 2009-03-11 src/include/zip.h

@@ -25,6 +25,11 @@ char *unescape_filename(char *filename);
     of the list, and when memory pressure increases we expire data
     from the front of the list.
 */
+enum Cache_policy {
+  CACHE_EXPIRE_FIRST,
+  CACHE_EXPIRE_LEAST_USED
+};
+
 CLASS(Cache, Object)
 // The key which is used to access the data
      void *key;
@@ -46,6 +51,7 @@ CLASS(Cache, Object)
 
      // This is a pointer to the head of the cache
      struct Cache *cache_head;
+     enum Cache_policy policy;
 
      // The current number of objects managed by this cache
      int cache_size;
@@ -205,6 +211,9 @@ CLASS(Resolver, Object)
 
      // Exports all the properties to do with uri - user owns the buffer.
      char *METHOD(Resolver, export, char *uri);
+
+     // Exports all the properties to do with uri - user owns the buffer.
+     char *METHOD(Resolver, export_all);
 
 // Deletes the attribute from the resolver
      void METHOD(Resolver, del, char *uri, char *attribute);

### 2009-03-11 src/lib/image.c

@@ -159,7 +159,7 @@ static int dump_chunk(Image this, char *data, uint32_t length, int force) {
     // Push one more offset to the index to cover the last chunk
     this->chunk_indexes[this->chunk_count + 1] = this->segment_buffer->readptr;
 
-    printf("Dumping segment %s\n", tmp);
+    printf("Dumping segment %s (%d bytes)\n", tmp, this->segment_buffer->readptr);
 
     // Store the entire segment in the zip file
     CALL((ZipFile)parent, writestr,

### 2009-03-11 src/lib/resolver.c

@@ -165,8 +165,10 @@ static Cache Cache_get(Cache self, void *key) {
     if(!CALL(i, cmp, key)) {
       // Thats it - we remove it from where its in and put it on the
       // tail:
-      list_move(&i->cache_list, &self->cache_list);
-      list_move(&i->hash_list, &hash_list_head->hash_list);
+      if(self->policy == CACHE_EXPIRE_LEAST_USED) {
+	list_move(&i->cache_list, &self->cache_list);
+	list_move(&i->hash_list, &hash_list_head->hash_list);
+      };
       //printf("Getting %p\n", i);
       return i;
     };
@@ -435,6 +437,7 @@ VIRTUAL(Resolver, Object)
      VMETHOD(resolve) = Resolver_resolve;
      VMETHOD(add)  = Resolver_add;
      VMETHOD(export) = Resolver_export;
+     VMETHOD(export_all) = Resolver_export_all;
      VMETHOD(open) = Resolver_open;
      VMETHOD(cache_return) = Resolver_return;
      VMETHOD(set) = Resolver_set;

### 2009-03-11 src/lib/zip.c

@@ -398,7 +398,7 @@ static ZipFile ZipFile_Con(ZipFile self, char *fd_urn) {
 	  char *text = CALL(self, read_member, self, filename, &len);
 	  char *context = dirname(filename);
 
-	  printf("Found property file %s\n%s", filename, text);
+	  //	  printf("Found property file %s\n%s", filename, text);
 
 	  if(text) {
 	    CALL(oracle, parse, context, text, len);

### 2009-03-11 src/tools/aff2imager.c

@@ -1,0 +1,250 @@
+#include <getopt.h>
+#include "zip.h"
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <fcntl.h>
+#include <errno.h>
+
+void help() {
+  printf("aff2imager tool\n\n"
+	 "--open, -o \tOpen this file to pre-populate the resolver \n"
+	 "           \t(Can be specified multiple times).\n"
+	 "--info, -i \tList information about all known objects.\n"
+	 );
+};
+
+void list_info() {
+  char *result =  CALL(oracle, export_all);
+  printf("%s", result);
+
+  talloc_free(result);
+};
+
+void aff2_open_volume(char *urn) {
+  // Try the different volume implementations in turn until one works:
+  ZipFile volume;
+  char *filename;
+
+  ClearError();
+  volume = CONSTRUCT(ZipFile, ZipFile, Con, NULL, urn);
+  if(CheckError(ERuntimeError)) {
+    ClearError();
+    RaiseError(ERuntimeError, "Could not open %s - will try to use file protocol.", urn);
+    volume = CONSTRUCT(ZipFile, ZipFile, Con, NULL, talloc_asprintf(NULL, "file://%s",urn));    
+  };
+
+  PrintError();
+
+  CALL(oracle, cache_return, (AFFObject)volume);
+};
+
+void aff2_print_info() {
+  Cache i,j;
+
+  list_for_each_entry(i, &oracle->urn->cache_list, cache_list) {
+    char *urn = (char *)i->key;
+    Cache attributes = (Cache)i->data;
+
+    printf("\n******** Object %s ***********\n",urn);
+    list_for_each_entry(j, &attributes->cache_list, cache_list) {
+      printf("\t%s = %s\n", (char *)j->key, (char *)j->data);
+    };
+  };
+};
+
+int aff2_image(char *driver, char *output_file, char *stream_name, 
+	       char *chunks_in_segment,
+	       char *append,
+	       char *source) {
+  ZipFile zipfile = (ZipFile)CALL(oracle, create, (AFFObject *)&__ZipFile);
+  char *output = talloc_strdup(zipfile, output_file);
+  Image image;
+  char buffer[BUFF_SIZE];
+  int in_fd;
+  int length;
+  Link link;
+  char zipfile_urn[BUFF_SIZE];
+
+  if(!zipfile) goto error;
+
+  in_fd=open(source,O_RDONLY);
+  if(in_fd<0) {
[…]

### 2009-03-11 paper/paper.tex
More changes in the paper.

@@
   files.
-\item Direct encoding of RAID configuration, allowing RAID volumes
-  to be automatically reassembled from evidence files made of the
-  physical RAID devices.
+
+%% Raid is only one example of the application of the map stream - the
+%% most important might be the ability to use sleuthkit to generate
+%% block allocation tables for all the files - thereby allowing
+%% arbitrary programs to read individual files without using sleuthkit
+\item Direct encoding of arbitrary mapping transformations of one or
+more sources to allow multiple logical views. For example, allowing
+RAID volumes to be automatically reassembled from evidence files made
+of the physical RAID devices.
 \item A revised compression system that allows individual chunks
@@
 
+%% Since this is the itemised definition its probably better to
+%% restrict it to the basic definitions and then move most of the
+%% discussion a bit later. I want to show how concepts build on one
+%% another - maybe even put a UML or class diagram
+
 \begin{itemize}
-\item \emph{An Evidence Volume} is a single file with a \texttt{.aff}
-  extension. 
-\item \emph{An Evidence Set} is a collection of one or more Evidence
-  Volumes. Evidence Sets can also be stored in object storage system (such as
-  Amazon S3), or a hybrid---for example, metadata stored in an SQL
-  database with physical data stored in large files in a file
-  system. 
-\item \emph{An Evidence Set UUID} is a unique identifier generated
-  according to RFC4122\cite{rfc4122} for each new Evidence
-  Set. Multiple Evidence Volumes that make up the same Evidence Set
-  should all have the same UUID. When an Evidence Set is opened all its volumes are
-  loaded, their UUIDs are checked, and all of their contents are
-  logically merged so as to be visible as if they were all found in the same volume.
-\item \emph{A segment} is a single unit of data written to volume. AFF4
-  segments have a \emph{segment name}, a \emph{segment timestamp}, and
-  a \emph{segment contents}.
+\item \emph{An AFF Object} is the basic building block of our
+file format. AFF Objects have a globally unique name (URN) as described in
+\cite{RFC1737}. The name is defined within the aff2 namespace, and is
+made unique by use of a unique identifier generated as per
+RFC4122\cite{rfc4122}.
 
-\item \emph{A stream} is a collection of segments that can be
-  concatenated together to form a single data object.
+\item \emph{A Relation} is a statement about an AFF Object which
+expresses some property of the object. The relation comprises of a
+triple of (Subject, Attribute, Value). All metadata is reduced to the
+triple notation.
 
-\item \emph{An image stream} is a stream that is dedicated to storing evidence from a single
-  source of data.  For example, a stream may contain a disk image or a network
-  capture. Whereas AFF could only support a single stream in an evidence
-  file, AFF4 allows any number of streams to be stored. Each stream in
-  the archive is  given a unique identifier.
+\item \emph{An Evidence Volume} is a type of AFF Object which is
+responsible for providing storage to AFF segments. Volumes may be
+implemented in a number of ways, but must provide a mechanism for
+storing and retrieving segments by their URN. We discuss several
+Volume implementations below.
 
-\item \emph{A bevy} is a specific segment that is used to hold
-  forensic image data, such as disk sectors, memory pages, or network
-  packets. AFF1 used the term ``pages'' to describe bevys, but this
-  term caused confusion when AFF was used to hold memory images, hence the
-  new term.
+\item \emph{A segment} is a single unit of data written to a volume. AFF4
+  segments have a \emph{segment name} provided by their URN, a
+  \emph{segment timestamp} in GMT, and the \emph{segment contents}.
 
-\item \emph{A segment name} is the name under which a segment is
-  stored within the archive. AFF4 segments have two-part names that
-  typically contain a \emph{stream name} and \emph{bevy number}. For
-  example the segment named \texttt{default/000001} refers to the stream
-  called \texttt{default} and has a chunk name of \texttt{000001}. All segment
-  names are encoded using UTF8. AFFLIB is being modified to allow any
-  stream to be opened using the \texttt{af\_open} function; if a
-  stream name is not specified, AFF4 implementations will open the
-  stream named ``default.''
+\item \emph{A stream} is an AFF Object which provides the ability to
+seek and read random data. Stream objects are free to implement their storage
+in a number of ways, but must provide clients the stream like
+interface. We discuss a number of stream types below.
 
-\item \emph{The segment timestamp} is a timestamp for the segment
-  recorded in GMT.
+% Is it important to define this? do we use it?
+%\item \emph{An Evidence Set} is a loose collection of one or more Evidence
+%Volumes.
 
-\item \emph{Attributes} are a list of key/value pairs stored in each
-stream's ``properties'' segment (e.g. ``stream/properties''). The
-``properties'' file is a text file which stores attributes one on each
-line. Attribute lines consist of the key, immediately followed by
-``='' for text values or ``:'' for binary values. This is immediately
-followed by the encoded value. Values containing text are encoded
-using UTF8, while values containing binary data are encoded using
-Base64. There may be multiple values for each attribute; in these
-cases the order is preserved
-by implementations. Implements may present different keys may be
-presented in any order, however.
-\vspace{1ex}
-Each AFF4 archive must also have a top-level set of attributes---that
-is, there is always a segment named ``properties.'' These volume
-attributes are used to specify a Unique Identifier (UUID). (AFF1
-stored each property in its own named segment. This created
-significant overhead in many situations.)
+%Evidence Sets can also be stored in object storage system (such as
+%  Amazon S3), or a hybrid---for example, metadata stored in an SQL
+%  database with physical data stored in large files in a file
+%  system.
 
-\item \emph{A Target Reference} is a way of referencing objects. This is used
-in a number of places which will be highlighted below. Using
-references it is possible for the FIF file to refer to objects
-embedded within itself, or within other FIF files. Its is also
-possible to refer to external flat files. A number of usage scenarios
-are described in Section \ref{usage_scenarios}. 
+% I would describe this process after the definitions.
+
+%\item \emph{An Evidence Set UUID} is a unique identifier generated
+%  according to RFC4122\cite{rfc4122} for each new Evidence
+%  Set. Multiple Evidence Volumes that make up the same Evidence Set
+%  should all have the same UUID. When an Evidence Set is opened all its volumes are
+%  loaded, their UUIDs are checked, and all of their contents are
+%  logically merged so as to be visible as if they were all found in
+% the same volume.
+
+%% I Wonder if this explaination should be postposed to the section we
+%% actually discuss it - otherwise we duplicate it again
+%% \item \emph{An image stream} is a stream that is dedicated to storing evidence from a single
+%%  source of data.  For example, a stream may contain a disk image or a network
+%%  capture. Whereas AFF could only support a single stream in an evidence
+%%  file, AFF4 allows any number of streams to be stored. Each stream in
+%%  the archive is  given a unique identifier.
+
+%%\item \emph{A bevy} is a specific segment that is used to hold
+%%  forensic image data, such as disk sectors, memory pages, or network
+%%  packets. AFF1 used the term ``pages'' to describe bevys, but this
+%%  term caused confusion when AFF was used to hold memory images, hence the
+%%  new term.
+
+%% This is no longer correct - segment names are URNs and are encoded
+%% according to RFC1737 as mentioned above.
+%%\item \emph{A segment name} is the name under which a segment is
+%%  stored within the archive. AFF4 segments have two-part names that
+%%  typically contain a \emph{stream name} and \emph{bevy number}. For
+%%  example the segment named \texttt{default/000001} refers to the stream
+%%  called \texttt{default} and has a chunk name of \texttt{000001}. All segment
+%%  names are encoded using UTF8. AFFLIB is being modified to allow any
+%%  stream to be opened using the \texttt{af\_open} function; if a
+%%  stream name is not specified, AFF4 implementations will open the
+%%  stream named ``default.''
+
+%%\item \emph{The segment timestamp} is a timestamp for the segment
+%%  recorded in GMT.
+
+%% This should be discussed as a way we are storing the RDF triples,
+%% rather than as a definition.
+%% \item \emph{Attributes} are a list of key/value pairs stored in each
+%% stream's ``properties'' segment (e.g. ``stream/properties''). The
+%% ``properties'' file is a text file which stores attributes one on each
+%%line. Attribute lines consist of the key, immediately followed by
+%%``='' for text values or ``:'' for binary values. This is immediately
+%%followed by the encoded value. Values containing text are encoded
+%%using UTF8, while values containing binary data are encoded using
+%%Base64. There may be multiple values for each attribute; in these
+%%cases the order is preserved
+%%by implementations. Implements may present different keys may be
+%%presented in any order, however.
+%%\vspace{1ex}
+
+%% This is not correct any more as volumes must store their own URN in
+%% a different way (the zip implementation uses an archive comment for
+%% this).
+
+%% Each AFF4 archive must also have a top-level set of attributes---that
+%% is, there is always a segment named ``properties.'' These volume
+%% attributes are used to specify a Unique Identifier (UUID). (AFF1
+%% stored each property in its own named segment. This created
+%% significant overhead in many situations.)
+
+%% Since target references are all just URNs now the resolving is automatic.
+
+\item \emph{A Target Reference} is a way of referencing objects by use
+of a Uniform Resource Identifier (URI). The URI can be another AFF
+Object URN or may be a more general Uniform Resource Locator (URL),
+such as for example a HTTP or FTP object.
+
+\item \emph{The Resolver} is a central data store which collects and resolves
+attributes for the different AFF Objects. We discuss the Resolver in
+detail in Section \ref{resolver}.
+
+%%in a number of places which will be highlighted below. Using
+%%references it is possible for the FIF file to refer to objects
+%%embedded within itself, or within other FIF files. Its is also
+%%possible to refer to external flat files. A number of usage scenarios
+%%are described in Section \ref{usage_scenarios}. 
 \end{itemize}
 
-The Target Reference is an important AFF4 innovation: this mechanism
-gives AFF4 files a way for describing evidence that is stored in
-different evidence volumes, in the host computer's file system, or on
-the Internet. We have defined the following target reference schema:
+Since AFF Objects have globally unique URNs it is possible for an
+Object in one volume to refer to an object in another volume. The
+resolver has visibility of all URNs and their current physical
+location.
+
+%The Target Reference is an important AFF4 innovation: this mechanism
+%gives AFF4 files a way for describing evidence that is stored in
+%different evidence volumes, in the host computer's file system, or on
+%the Internet. We have defined the following target reference schema:
 
@@
 
-\begin{itemize}
-\item A reference in the format {\em aff://UUID/stream/segment} refers
-to an external Evidence file set (identified by its UUID). If the segment
-is missing (i.e. the reference ends with a ``/''), the reference is
-for the entire stream. We do not specify a way of actually resolving
-the UUID into a physical locations. This can achieved by
-implementations using many methods, such as Active Directory, LDAP,
-DNS, flat files or SQL databases.
+%\begin{itemize}
+%\item A reference in the format {\em aff://UUID/stream/segment} refers
+%to an external Evidence file set (identified by its UUID). If the segment
+%is missing (i.e. the reference ends with a ``/''), the reference is
+%for the entire stream. We do not specify a way of actually resolving
+%the UUID into a physical locations. This can achieved by
+%implementations using many methods, such as Active Directory, LDAP,
+%DNS, flat files or SQL databases.
 
-\item A reference starting with {\em file://} refers to an external
-file. The external file reference must be sanitised (to remove
-.. directory traversal) and rooted at the current directory. A
-filename of ``.'' refers to the current FIF volume.
+%\item A reference starting with {\em file://} refers to an external
+%file. The external file reference must be sanitised (to remove
+%.. directory traversal) and rooted at the current directory. A
+%filename of ``.'' refers to the current FIF volume.
 
-\item A reference starting with {\em http://} refers to a HTTP
-object which will be fetched. Support for this is optional.
+%\item A reference starting with {\em http://} refers to a HTTP
+%object which will be fetched. Support for this is optional.
 
-\item A references ending with a ``/'' signifies a reference to a
-stream in the current FIF file. The stream will be opened in its
-entirety. If the reference does not end with a ``/'', the reference is
-for a single segment.
-\end{itemize}
+%% This is no longer correct as all segments, streams and volumes have
+%% a unique URN so we dont need to make these distinctions.
+
+%% \item A references ending with a ``/'' signifies a reference to a
+%% stream in the current FIF file. The stream will be opened in its
+%% entirety. If the reference does not end with a ``/'', the reference is
+%% for a single segment.
+%%\end{itemize}
 

### 2009-03-12 paper/paper.tex
More work on paper.

@@
 %\documentclass{elsart5p}
+\usepackage{listings}
+\lstset{language=Python,
+	frame=single,
+	captionpos=b,
+	frameround=tttt,
+	basicstyle=\small,          % print whole listing small
+	keywordstyle=\color{black}\bfseries,
+                            % underlined bold black keywords
+	identifierstyle=,           % nothing happens
+	commentstyle=\color{white}, % white comments
+	stringstyle=\ttfamily,      % typewriter type for strings
+	showstringspaces=false}     % no special string spaces
 
@@
 \item \emph{An Evidence Volume} is a type of AFF Object which is
-responsible for providing storage to AFF segments. Volumes may be
-implemented in a number of ways, but must provide a mechanism for
-storing and retrieving segments by their URN. We discuss several
-Volume implementations below.
+responsible for providing storage to AFF segments. Volumes must
+provide a mechanism for storing and retrieving segments by their
+URN. We discuss two volume implementations below, namely the {\em
+Zip64 based volume} and the {\em Directory} based volume.
 
@@
 \item \emph{A stream} is an AFF Object which provides the ability to
-seek and read random data. Stream objects are free to implement their storage
-in a number of ways, but must provide clients the stream like
-interface. We discuss a number of stream types below.
+seek and read random data. Stream objects implement abstracted
+storage, but must provide clients the stream like interface. For
+example, we discuss the {\em Image stream} used to store large images,
+the {\em Map stream} used to create transformations and the {\em
+Encrypted stream} used to provide encryption.
 
@@
 \subsection{ZIP64}
-For AFF4, we have changed AFF1's original container file format to
-ZIP64\cite{zipspecs}. There are many reasons for this decision:
+As described above, AFF4 volumes are only required to store and
+retrieve segments by their URN. For AFF4, we have changed the default
+volume container file format to ZIP64\cite{zipspecs}. There are many
+reasons for this decision:
 
@@
   migrating to these formats, we can take advantage of the rich number
-  of user and developer tools already available.
-\item ZIP64 includes a CRC32 and timestamp for each segment. We
-  believe that the additional integrity checks and timestamps will be
-  useful in a forensic environment.
+  of user and developer tools already available. The volume may be
+inspected using any number of commercial zip application (e.g. Windows
+Explorer natively supports Zip files).
+
+%% In not sure how this is relevant
+%\item ZIP64 includes a CRC32 and timestamp for each segment. We
+%  believe that the additional integrity checks and timestamps will be
+%  useful in a forensic environment.
+
 \item ZIP64 already includes support for the ``table of contents''
@@
   ZIP64 volume.
+
+\item ZIP64 libraries are readily available making proprietary implementations of
+interfaces to the AFF4 volume format simple to write. For example a
+simple python program to dump out an Image stream is illustrated in
+Listing \ref{python_code}.
+
 \end{itemize}
+
+\begin{lstlisting}[
+	float=tb, label=python_code, caption=Sample Python code to
+	dump out an Image Stream. As can be seen the chunk index
+	segment is used to slice the data segment into chunks. The
+	chunks are decompressed and written to the output file.]
+volume=zipfile.ZipFile(FILE)
+outfd = open("dump.dd","w")
+bevy_count = 0
+
+while 1:
+    idx_segment = volume.read(STREAM + 
+		"%08d.idx" % bevy_count)
+    segment = volume.read(STREAM + 
+		"%08d" % bevy_count)
+    indexes = struct.unpack("<" + "l" * \
+	(len(idx_segment)/4),idx_segment)
+
+    for i in range(len(indexes)-1):
+	chunk = segment[indexes[i]:
+	                indexes[i+1]]
+        outfd.write(zlib.decompress(chunk))
+
+    bevy_count += 1
+\end{lstlisting}
 
@@
 
-\begin{figure}
-\begin{Verbatim}[frame=single]
-stream_period=393216
-target_period=196608
-
-0,0,1
-65536,0,0
-131072,65536,2
-196608,65536,1
-262144,131072,0
-327680,131072,2
-\end{Verbatim}
-\caption{A Map stream that corresponds to a 3 disk RAID-5 array}\label{ref}
-\end{figure}
+\begin{lstlisting}[
+	float=tb, caption=A Map stream that corresponds to a 3 disk
+	RAID-5 array. The targets are URNs for the respective
+        disks. Note that map coordinates are given in multiples of 
+	block size.,
+	label=ref, language=]
+aff4:block_size=64k 
+aff4:stream_period=6 
+aff4:target_period=3
+
+0,0,disk1
+1,0,disk0
+2,1,disk2
+3,1,disk1
+4,2,disk0
+5,2,disk2
+\end{lstlisting}
 
@@
 fields. These often produce a set of uncompressed images split at a
-certain size. If the images are split at exactly the same size its
+certain size. It is possible to construct a Map Stream which
+seamlessly reassembles the logical image from all the individual disk
+images. The new stream may be kept in its own volume and may be very
+small.
+
+If the images are split at exactly the same size its
 possible to create a FIF archive consisting of a single Overlay

### 2009-03-13 paper/paper.tex
More work - added a section describing the resolver.

@@
 	frameround=tttt,
-	basicstyle=\small,          % print whole listing small
+	basicstyle=\scriptsize,
 	keywordstyle=\color{black}\bfseries,
@@
 
-\subsection{Global Distributed Evidence Management}
+\section{Global Distributed Evidence Management}
 While Garfinkel designed AFF for use on a single machine that could
@@
 \section{Introducing AFF4}
-
-\subsection{New Concepts and Terminology}
 
@@
 
-\subsection{Streams}
-As discussed above, individual segments in an AFF4 evidence set can be
-combined into a single logical object called a \emph{stream}. AFF4
-supports an arbitrary number of streams per volume. The AFFLIB library
-allows the segments in a stream to be operated on as if they were a
-single file inside a file system by supporting the traditional
-POSIX-like functionality of \texttt{open()}, \texttt{seek()} and
-\texttt{read()}.
-
-AFFLIB also supports the \emph{writing} of streams, a feature that is
-used by Garfinkel's \texttt{aimage} disk imaging program. In general
-arbitrary writing to streams can be expensive, especially when writing
-is combined with seek operations. The current \texttt{aimage} program
-will create AFF files with gaps in them during its error recovery
-process (when \texttt{aimage} seeks to the end of a disk and reads it
-backwards in an attempt to get around media failures). We are working
-on a rewrite of \texttt{aimage} which performs imaging and recovery
-on bevy boundaries, rather than using the current arbitrary
-``high-water-mark'' and ``low-water-mark.''
-
-All streams within an Evidence Set \emph{must} have a properties segment. The
-properties segment describes what the stream contains and determines
-how the stream is processed.
+\section{Metadata and the Universal Resolver}
+The AFF4 model treats metadata as an abstract concept which may exists
+independently from the data itself. We term {\em metadata} to be a set
+of statements about objects, written in triple notations (Subject,
+Attribute, Value), where {\em Subject} is the URN of the object the
+statement is made about. Using this system we are able to store
+arbitrary attributes about any object in the AFF4 universe.
+
+The AFF4 design extends beyond the management of a single volume,
+stream or image to a universal system for managing data of many
+types. This necessarily means that a single running instance is
+generally unable to have visibility of the entire AFF4 universe. For
+example, if a volume is opened which contains a Map Stream targeting a
+stream stored in a different volume, its not generally possible to
+tell in which volume the stream exists, and where that volume is
+actually stored.
+
+To provide this global visibility of metadata we define a central
+metadata management entity, named the {\em Universal Resolver}. The
+Universal Resolver contains all the metadata about the AFF4 universe,
+that is to say it is able to resolve queries about any attribute about
+any URN in the universe.
+
+Although the resolver has complete visibility of all attributes, it is
+still useful to store metadata within the volume itself, particularly
+data pertaining to the volume itself. If we did not store the metadata
+within the volume itself, then the volume would not be accessible to
+implementations which do not have this metadata.
+
+To this end we define a way for serializing metadata statements (or
+triples) into a standard format which implementations can load into
+their respective resolvers when parsing the volume. Triples can be
+stored in segments having a URN ending with ``{\em properties}''. The
+AFFLIB implementation loads these segments automatically into the
+Universal Resolver.
+
+Triples are stored within the properties segment one per line, with
+the subject URN (encoded according to RFC1737), followed by whitespace
+and the attribute name. This is then followed by the equal sign and
+the UTF8 encoding of the value. An example properties file for an
+Image Stream is shown in Listing \ref{image_metadata}.
+
+It is important to stress that the properties file is simply a
+serialization of statements into volume segments. The statements may
+exist without being stored in a volume (for example, being stored on
+an external SQL server). Alternatively, these statements may be stored
+in some other way inside or outside the volume (e.g. SQLite
+databases).
+
+When the volume is loaded, the AFFLIB implementation automatically
+loads any properties files and populates its Universal Resolver with
+the information it has visibility of. AFFLIB provides a mechanism to
+use an external resolver stored on a MySQL database to provide a
+Universal Resolver that shares information between different instances
+on the same network.
+
+\begin{lstlisting}[
+	caption=Example properties files for several AFF4 objects (URNs are
+shortened for illustration).,
+	label=image_metadata,
+	float=tb,
+	 ]
+Directory Volume:
+  urn:aff2:f901be8e-d4b2 aff2:stored=http://../case1/
+  urn:aff2:f901be8e-d4b2 aff2:type=directory
+
+ZipFile Volume:
+  urn:aff2:98a6dad6-4918 aff2:stored=file:///file.zip
+  urn:aff2:98a6dad6-4918 aff2:type=zip
+
+Image Stream:
+  urn:aff2:83a3d6db-85d5 aff2:stored=urn:aff2:f901be8e-d4b2
+  urn:aff2:83a3d6db-85d5 aff2:chunks_in_segment=256
+  urn:aff2:83a3d6db-85d5 aff2:chunk_size=32k
+  urn:aff2:83a3d6db-85d5 aff2:type=image
+  urn:aff2:83a3d6db-85d5 aff2:size=5242880
+
+Map Stream:
+  urn:aff2:ed8f1e7a-94aa aff2:target_period=3
+  urn:aff2:ed8f1e7a-94aa aff2:image_period=6
+  urn:aff2:ed8f1e7a-94aa aff2:blocksize=64k
+  urn:aff2:ed8f1e7a-94aa aff2:stored=urn:aff2:83a3d6db-85d5
+  urn:aff2:ed8f1e7a-94aa aff2:type=map
+  urn:aff2:ed8f1e7a-94aa aff2:size=0xA00000
+
+Link Object:
+  map aff2:target=urn:aff2:ed8f1e7a-94aa
+  map aff2:type=link
+\end{lstlisting}
+
+\section{Volumes}
+The volume object is responsible for providing storage for
+segments. Segments are stored and retrieved using their URNs. We
+describe two different implementations of volume objects, namely the
+{\em Directory Volume} and the {\em ZipFile Volume}. It is possible to
+convert from one implementation to another easily, without effecting
+any external references. This is because external references are to
+the Volume URN itself, so its safe to convert from one representation
+to another. In this way its possible for example, to unzip a Zip64
+volume into a directory and vice versa.
+
+\subsection{Directory Volumes}
+The Directory Volume is the simplest type of volume. It simply stores
+different segments based on their URNs in a single directory. Since
+some filesystems are unable to represent URNs accurately (e.g. Windows
+has many limitations on the types of characters allowed for a
+filename), the Directory Volume encodes URNs according to RFC1738
+\cite{RFC1738}; non-printable characters are escaped with a \%
+followed by the ASCII ordinal of the character.
+
+The Directory Volume uses the {\em aff2:stored} attribute to provide a
+base URL. The URL for each segment is then constructed by appending
+the escaped segment URN to the base URL. Note that there is no
+restriction on what type of URL this can be, so it may be a location
+on a filesystem (e.g. {\em file://some\_directory/}) or a location on a
+HTTP server (e.g. {\em http://intranet.server/some\_path}). In this
+way its possible to move the entire volume from a filesystem to a web
+server transparently.
+
+The Directory Volume stores its own URN in a special segment named
+``{\em \_\_URN\_\_}'' at the base of the directory.
 
 \subsection{ZIP64}
-As described above, AFF4 volumes are only required to store and
-retrieve segments by their URN. For AFF4, we have changed the default
-volume container file format to ZIP64\cite{zipspecs}. There are many
-reasons for this decision:
+For AFF4, we have changed the default volume container file format to
+ZIP64\cite{zipspecs}. There are many reasons for this decision:
 
@@
 \item ZIP64 libraries are readily available making proprietary implementations of
-interfaces to the AFF4 volume format simple to write. For example a
+interfaces to the AFF4 volume format simple to write. For example, a
 simple python program to dump out an Image stream is illustrated in
@@
 	chunks are decompressed and written to the output file.]
-volume=zipfile.ZipFile(FILE)
-outfd = open("dump.dd","w")
-bevy_count = 0
+volume=zipfile.ZipFile(INPUT_FILE)
+outfd = open(OUTPUT_FILE,"w")
+count = 0
 
 while 1:
-    idx_segment = volume.read(STREAM + 
-		"%08d.idx" % bevy_count)
-    segment = volume.read(STREAM + 
-		"%08d" % bevy_count)
-    indexes = struct.unpack("<" + "l" * \
-	(len(idx_segment)/4),idx_segment)
+    idx_segment = volume.read(STREAM+"/%08d.idx" % count)
+    bevy = volume.read(STREAM+"/%08d" % count)
+    indexes = struct.unpack("<" + "L" * 
+	(len(idx_segment)/4), idx_segment)
 
     for i in range(len(indexes)-1):
-	chunk = segment[indexes[i]:
-	                indexes[i+1]]
+	chunk = bevy[indexes[i]:indexes[i+1]]
         outfd.write(zlib.decompress(chunk))
 
-    bevy_count += 1
+    count += 1
 \end{lstlisting}
@@
 
-\begin{figure}[tb]
+\begin{figure}[]
   \begin{center}
@@
 
-It is important to note that AFF4 only requires that the underlying storage mechanism
-be capable of storing multiple named segments of data. Although our
-AFF4 implementation will use the ZIP64 file format as an underlying storage
+It is important to note that AFF4 only requires that the volume be
+capable of storing multiple named segments of data. Although our AFF4
+implementation uses the ZIP64 file format as an underlying storage
 mechanism, our system also supports legacy AFF1 volumes as well as
@@
 
-\subsubsection{Unused ZIP64 Features}
-
+\subsection{Unused ZIP64 Features}
 There are a number of aspects in the Zip file format specifications
@@
 implementation then considers the segments contained within as
-belonging to a larger collection. The reason behind this decision is that
-often it is needed to add new volumes to an existing set without
-modifying the existing set at all. Zip64's built-in support for
-splitting an archive renders each individual volume in the set
-unusable unless all are available.
+belonging to the universal collection. We provide the ability to split
+a stream across volumes automatically, as each segment may be stored
+in a different volume, and the Universal Resolver is used to find it.
 
 \item Although Zip64 also defines encryption and authentication
-extensions, we do not use them due to the
-restrictions imposed on their use and because they lack of functionality that is
-important for forensic user. Instead, we use AFF1's digital signature
-facilities for integrity and non-repudiation, and we introduce a new  stream based
-encryption scheme for ensuring data privacy (Section~\ref{crypted_stream}).
+extensions, we do not use them due to the restrictions imposed on
+their use and because they lack of functionality that is important for
+forensic user. Instead, we use AFF1's digital signature facilities for
+integrity and non-repudiation, and we introduce a new stream based
+encryption scheme for ensuring data privacy
+(Section~\ref{crypted_stream}).
 
@@
 
-\subsubsection{A Forensically Sound Strategy For Updating ZIP64 Files}
-
-In many forensic applications, individual segments within the volume
-need to be updated. To this end, Garfinkel implemented several
-``update'' functions in AFF1\cite{garfinkel:aff}.
-
-There were two problems with the original AFF updating approach:
-\begin{itemize}
-\item Updates can cause the volume to
-become fragmented as holes appear within the middle of the volume, and
-updated segments are added at the end.
-\item Updates could be performed silently, raising questions of
-  forensic soundness.
-\end{itemize}
-
-In AFF4, when a new segment is added to an existing volume, the new
-segment header overwrites the volume central directory, and the
-central directory should be rewritten including the new entry
-immediately after the last file in the archive. If it is not possible
-to overwrite the file, the new segment can be written at the end,
-followed by an updated copy of the central directory (in that case the
-old central directory remains in the file). This is illustrated in
-Figure \ref{zip_structure}. Implementations may delay
-writing the central directory until the file is closed; should the
-system crash before the central directory is written, it can be
-recovered by scanning the entire ZIP file.
-
-A segment within the archive is considered to be updated when a new
-segment is added to the archive with the same name and a later
-timestamp. Implementations satisfy read requests for duplicate
-segment names with those segments with the later timestamp, or if the
-two timestamps are the same, the segment which appears at a later
-offset in the volume. This effectively allows for a segment to be
-updated without leading to internal volume fragmentation.
-
-The fact that the older version of all updated segments remains intact
-within each volume can be used for auditing, or to allow a rollback to
-previous versions of the segment. We have created a new tool,
-\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
-creating new volumes and redistributing segments among them.
-
-\section{Standard AFF4 Streams}
-In this section we discuss specific stream types that we have created
-in our initial AFF4 implementation.
+%% Im not sure this is really that relevant here. Byte range updates
+%% can be achieved by an append and a map operation, which might be a
+%% better option for aimager working in dd_rescue mode.
+
+%\subsubsection{A Forensically Sound Strategy For Updating ZIP64 Files}
+%In many forensic applications, individual segments within the volume
+%need to be updated. To this end, Garfinkel implemented several
+%``update'' functions in AFF1\cite{garfinkel:aff}.
+
+%There were two problems with the original AFF updating approach:
+%\begin{itemize}
+%\item Updates can cause the volume to
+%become fragmented as holes appear within the middle of the volume, and
+%updated segments are added at the end.
+%\item Updates could be performed silently, raising questions of
+%  forensic soundness.
+%\end{itemize}
+
+%In AFF4, when a new segment is added to an existing volume, the new
+%segment header overwrites the volume central directory, and the
+%central directory should be rewritten including the new entry
+%immediately after the last file in the archive. If it is not possible
+%to overwrite the file, the new segment can be written at the end,
+%followed by an updated copy of the central directory (in that case the
+%old central directory remains in the file). This is illustrated in
+%Figure \ref{zip_structure}. Implementations may delay
+%writing the central directory until the file is closed; should the
+%system crash before the central directory is written, it can be
+%recovered by scanning the entire ZIP file.
+
+%A segment within the archive is considered to be updated when a new
+%segment is added to the archive with the same name and a later
+%timestamp. Implementations satisfy read requests for duplicate
+%segment names with those segments with the later timestamp, or if the
+%two timestamps are the same, the segment which appears at a later
+%offset in the volume. This effectively allows for a segment to be
+%updated without leading to internal volume fragmentation.
+
+%The fact that the older version of all updated segments remains intact
+%within each volume can be used for auditing, or to allow a rollback to
+%previous versions of the segment. We have created a new tool,
+%\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
+%creating new volumes and redistributing segments among them.
+
+\section{Streams}
+The Stream object provides random access to an abstract representation
+of a body of data. Our implementation allows the segments in a stream
+to be operated on as if they were a single file by supporting the
+traditional POSIX-like functionality of
+\texttt{open()}, \texttt{seek()}, \texttt{write()}, and \texttt{read()}.
+
+The following describes a number of types of streams. It is important
+to note that clients of our implementation do not care how a
+particular stream is implemented. Streams are opened by their URNs,
+and the library itself ensure they provide the Stream interface. So
+for example, users do not care if a stream is a Map Stream or an Image
+Stream - the data provided is the same.
+
+%% Not sure how its relevant - maybe we can move it to a new section
+%% about randomly writing stream.
+
+%AFFLIB also supports the \emph{writing} of streams, a feature that is
+%used by Garfinkel's \texttt{aimage} disk imaging program. In general
+%arbitrary writing to streams can be expensive, especially when writing
+%is combined with seek operations. The current \texttt{aimage} program
+%will create AFF files with gaps in them during its error recovery
+%process (when \texttt{aimage} seeks to the end of a disk and reads it
+%backwards in an attempt to get around media failures). We are working
+%on a rewrite of \texttt{aimage} which performs imaging and recovery
+%on bevy boundaries, rather than using the current arbitrary
+%``high-water-mark'' and ``low-water-mark.''
+
+%All streams within an Evidence Set \emph{must} have a properties segment. The
+%properties segment describes what the stream contains and determines
+%how the stream is processed.
 
@@
 \label{image_stream}
-The AFF4 \emph{Image} stream contains a single read-only forensic data
-set. For example, this stream might contains a hard
-disk image, a memory image or a network capture (in PCAP
-format). Random read access into the stream is provided with
-AFFLIB. (AFFLIB also provided the ability to randomly \emph{write}
-into the evidence stream for sanitization purposes, although this
-feature may be removed at a later date.)
+The AFF4 \emph{Image Stream} stores a single read-only forensic data
+set. For example, this stream might contains a hard disk image, a
+memory image or a network capture (in PCAP format).
+
+Storage for the data is done by using multiple data segments stored on
+various volumes. Data segment URNs are derived by appending an 8
+digit, zero padded decimal integer representation of an incrementing
+id to the stream URN (e.g. ``urn:aff2:83a3d6db-85d5/00000032'').
+
+Image streams must specify the
+\texttt{page\_size} attribute, as the number of image bytes each chunk
+contains. Pages may be compressed at the page level, or may be broken
+up into multiple chunks (typically 8KB), and individually
+compressed. In this case the beginning of the page segment will
+contain an offset table indicating the location of each compression
+chunk. This is similar to the technique used by the Expert Witness
+file format used by EnCase\cite{encase-3.0} and implemented by the
+open source libewf\cite{libewf} package.
+
+Pages can optionally be individually hashed, allowing for detection
+of modifications. The attribute \texttt{hash\_type} specifies the type of
+hash used (implementations will typically support md5,sha1 and sha256, but can
+define others). The hashes are stored in any number of segments within
+the stream, of arbitrary name. The hash segment names is provided by
+using the \texttt{hash\_name} stream attribute.
 
@@
 holes can generate an error or return NULLs.
-
-The image itself is stored in \emph{pages} within the AFF4 file. Pages are
-segments named as an 8 digit, zero padded decimal integer
-representation of their chunk id (e.g. ``default/00000032''). Image
-streams must specify the \texttt{page\_size} attribute, as the number of
-image bytes each chunk contains. Pages may be compressed at the
-page level, or may be broken up into multiple chunks (typically 8KB),
-and individually compressed. In this case the beginning of the page
-segment will contain an offset table indicating the location of each
-compression chunk. This is similar to the technique used by the Expert
-Witness file format used by EnCase\cite{encase-3.0} and implemented by
-the open source libewf\cite{libewf} package.
-
-Pages can optionally be individually hashed, allowing for detection
-of modifications. The attribute \texttt{hash\_type} specifies the type of
-hash used (implementations will typically support md5,sha1 and sha256, but can
-define others). The hashes are stored in any number of segments within
-the stream, of arbitrary name. The hash segment names is provided by
-using the \texttt{hash\_name} stream attribute.
 
@@
 be a AFF4 file. Since Zip files are normally read from the end of the
-file, but EWF are both read from the front of the file, it may
-be possible to append a AFF4 Central Directory to the end of an EWF file without
-interfering with the overlayed file. In this case the converted file
-can still be used as an EWF file without change, but could also be
-used as an AFF4 file.
+file, but EWF are both read from the front of the file, it may be
+possible to append a AFF4 Central Directory to the end of an EWF file
+without interfering with the overlayed file. In this case the
+converted file can still be used as an EWF file without change, but
+could also be used as an AFF4 file.
 

### 2009-03-13 paper/paper.tex
More work on paper.

@@
 
+Each URN within the AFF4 universe must have an ``\texttt{type}''
+attribute to denote the type of the Object.
+
 \begin{lstlisting}[
@@
 traditional POSIX-like functionality of
-\texttt{open()}, \texttt{seek()}, \texttt{write()}, and \texttt{read()}.
+\texttt{open()}, \texttt{seek()}, \texttt{write()}, and
+\texttt{read()}. All streams must have the ``\texttt{size}'' attribute
+to denote the last bytes addressible within the stream. This is
+required in order to support the POSIX whence attribute which may
+require seeking from the end of the stream.
 
@@
 set. For example, this stream might contains a hard disk image, a
-memory image or a network capture (in PCAP format).
+memory image or a network capture (in PCAP format). Image streams have
+an \texttt{aff2:type} attribute of \texttt{image}.
 
@@
 digit, zero padded decimal integer representation of an incrementing
-id to the stream URN (e.g. ``urn:aff2:83a3d6db-85d5/00000032'').
-
-Image streams must specify the
-\texttt{page\_size} attribute, as the number of image bytes each chunk
-contains. Pages may be compressed at the page level, or may be broken
-up into multiple chunks (typically 8KB), and individually
-compressed. In this case the beginning of the page segment will
-contain an offset table indicating the location of each compression
-chunk. This is similar to the technique used by the Expert Witness
-file format used by EnCase\cite{encase-3.0} and implemented by the
-open source libewf\cite{libewf} package.
-
-Pages can optionally be individually hashed, allowing for detection
-of modifications. The attribute \texttt{hash\_type} specifies the type of
-hash used (implementations will typically support md5,sha1 and sha256, but can
-define others). The hashes are stored in any number of segments within
-the stream, of arbitrary name. The hash segment names is provided by
-using the \texttt{hash\_name} stream attribute.
-
-Image streams must have at several attributes. First, they must have
-\texttt{type=image} to indicate that they are image streams. Streams
-must have a \texttt{size} attribute (specified in 8-bit bytes) to
-indicate the last addressable byte in the stream. Streams can have
-holes in them; in this case, attempts to read data from within the
-holes can generate an error or return NULLs.
-
-Hashes are calculated in an identical way to that defined by AFF1's
-crypto layer\cite{garfinkel:affcrypto}---that is, the full name of the
-segment (including stream component) is followed by a single NULL
-byte, followed by the (uncompressed) content of the chunk. The
-inclusion of the segment name ensure that even identical chunks in
-content will have a different hash and prevents reordering attacks.
-
-Each hash segment is a list of a chunk id encoded in 32 bit
-little endian integer format followed by the hash (the length of which
-depends on the hash itself). This approach is illustrated in Figure
-\ref{hash_index}.
-
-Hash segments can contain any number of chunk hashes in them. It is
-recommended that a hash segment be created when a volume is closed to
-cover all the chunks within that volume. This allows for hashes to be
-verified even if some volumes are missing. The AFFLIB implementation
-may update a hash of a chunk previously set by an earlier hash
-segment---this would be required when a chunk is updated. The hash
-segment can then be signed to assure the integrity of the updated segment.
+id to the stream URN (e.g. ``urn:aff2:83a3d6db-85d5/00000032''). Each
+data segment is called a bevy and stores a number of compressed chunks
+back to back. The chunk index segment is a segment containing a list
+of relative offsets to the beginning of each chunks within the
+bevy. The chunk index segment URN is derived  by appending the bevy
+URN with ``.idx''. This is illustrated in Figure \ref{image_stream}.
 
@@
   \begin{center}
-  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
-  \caption{The structure of a hashing segment. The attributes {\em
-  hash} and {\em hash\_type} declare the hash segment. The segment
-  contains a list of integer encoded chunk id followed by an unencoded
-  hash for each chunk.}
-  \label{hash_index}
+  \mbox{\epsfxsize=0.6\columnwidth \epsffile{image_segments.eps}}
+  \caption{The structure of Image Stream Bevies. Each bevy is a
+collection of compressed chunks stored back to back. Relative chunk
+offsets are stored in the chunk index segment.}
+  \label{image_stream}
   \end{center}
@@
 
-This design differs from AFF1, which creates a separate segment for
-each hash. AFF4 uses headers that are more verbose than the AFF
-segment headers. Therefore, creating a new segment for each hash value
-would result in too much overhead: Packing the hashes into a few
-segments it is possible to reduce this overhead to a minimum.
+Image streams specify the \texttt{chunk\_size} attribute, as the
+number of image bytes each chunk contains (typically chunk sizes are
+32kb). Also specified is the \texttt{chunks\_per\_segment} attribute
+which specifies how many chunks are stored in each bevy. Each chunk is
+compressed individually using the zlib compress algorithm. This
+general structure of storing chunks within larger segments similar to
+the technique used by the Expert Witness file format used by
+EnCase\cite{encase-3.0} and implemented by the open source
+libewf\cite{libewf} package. The advantages of keeping smaller sized
+chunks is the better match between requested size and the minimum size
+required for decompression. This means that less data is needed to be
+decompressed unnecessarily.
+
+%% FIXME - hashing may apply to any segment not just bevies
+%Bevies can optionally be individually hashed, allowing for detection
+%of modifications. The attribute \texttt{hash\_type} specifies the type
+%of hash used (implementations will typically support md5,sha1 and
+%sha256, but can define others). The hashes are stored in any number of
+%segments within the stream, of arbitrary name. The hash segment names
+%is provided by using the \texttt{hash\_name} stream attribute.
+
+%Hashes are calculated in an identical way to that defined by AFF1's
+%crypto layer\cite{garfinkel:affcrypto}---that is, the full name of the
+%segment (including stream component) is followed by a single NULL
+%byte, followed by the (uncompressed) content of the chunk. The
+%inclusion of the segment name ensure that even identical chunks in
+%content will have a different hash and prevents reordering attacks.
+
+%Each hash segment is a list of a chunk id encoded in 32 bit
+%little endian integer format followed by the hash (the length of which
+%depends on the hash itself). This approach is illustrated in Figure
+%\ref{hash_index}.
+
+%Hash segments can contain any number of chunk hashes in them. It is
+%recommended that a hash segment be created when a volume is closed to
+%cover all the chunks within that volume. This allows for hashes to be
+%verified even if some volumes are missing. The AFFLIB implementation
+%may update a hash of a chunk previously set by an earlier hash
+%segment---this would be required when a chunk is updated. The hash
+%segment can then be signed to assure the integrity of the updated segment.
+
+%\begin{figure}[tb]
+%  \begin{center}
+%  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
+%  \caption{The structure of a hashing segment. The attributes {\em
+%  hash} and {\em hash\_type} declare the hash segment. The segment
+%  contains a list of integer encoded chunk id followed by an unencoded
+%  hash for each chunk.}
+%  \label{hash_index}
+%  \end{center}
+%\end{figure}
+
+%This design differs from AFF1, which creates a separate segment for
+%each hash. AFF4 uses headers that are more verbose than the AFF
+%segment headers. Therefore, creating a new segment for each hash value
+%would result in too much overhead: Packing the hashes into a few
+%segments it is possible to reduce this overhead to a minimum.
 
@@
 the image; Cohen extended this concept to an arbitrary
-mapping function\cite{Cohen2007} .
-
-AFF4 defines a common way in
-which the underlying storage system can apply transformations to
-existing data to produce transformed data. This transformation is
-performed using a \emph{map stream}.
-
-The Map stream defines a mapping function, as a piecewise continuous
-linear transform of one or more {\em target} streams to produce a new
-{\em stream}. Target streams are specified by using ``target''
-attribute, which may be used many times to provide a list of targets.
-
-The mapping function file consists of a series of lines, each
-containing three 64 bit integers separated by comma, and encoded using
-decimal notation. The integers represent stream offset, target offset
-and target number. Target numbers are an index to the list of targets
-given in the attribute section (list index starts at zero for the
-first entry).
+mapping function\cite{Cohen2007} which can be used to describe
+arbitrary mappings of carved files within a single image.
+
+In this work we extend the mapping function concept to allow a single
+map to draw data from arbitrary streams (called {\em targets}). This
+transform is implemented via the {\em Map stream}.
+
+The mapping function is described in a segment named by appending a
+``/map'' to the stream URN. The segment data consists of a series of
+lines, each containing a stream offset, followed by a comma, and a
+target offset. Offsets are encoded using decimal notation, and are
+followed by a target URN.
 
@@
 For example, consider the following map:
-\begin{verbatim*}
-0,0,0
-4096,10000,0
-8192,5000,0
-\end{verbatim*}
+\begin{lstlisting}
+0,0,urn:aff2:83a3d6db-85d5
+4096,10000,urn:aff2:f901be8e-d4b2
+8192,5000,urn:aff2:83a3d6db-85d5
+\end{lstlisting}
 
 To read this stream we satisfy read requests of offsets between 0 and
-4095 in the stream from offset 0 in target 0. Requests for bytes
-between 4096 and 8191 are fetched from target 0 at offset
-10000. Finally bytes after 8192 (until the specified size of the
-stream) are fetched from offset 5000 in target 0.
+4095 in the stream from offset 0 to offset 4095 in
+\emph{urn:aff2:83a3d6db-85d5}. Requests for bytes between 4096 and 8191 are
+fetched from \emph{urn:aff2:f901be8e-d4b2} at offset 10000. Finally
+bytes after 8192 (until the specified size of the stream) are fetched
+from offset 5000 in \emph{urn:aff2:83a3d6db-85d5}.
 
@@
 Where $mod$ is the modulus function and $floor$ signifies integer
-division. For example consider Figure~\ref{map}, which corresponds to a 3
+division. For example consider Listing~\ref{map}, which corresponds to a 3
 disk RAID-5 array.
@@
 	block size.,
-	label=ref, language=]
+	label=map, language=]
 aff4:block_size=64k 
@@
 
-\subsection{The Overlay Stream}
-The Overlay Stream is a variant of the {\em Image} stream that
-provides a layer of indirection between requests made for specific
-bytes (or chunks or bevys) and the actual location where the requested
-information is stored. The main utility of the Overlay stream is in
-providing transparent access to legacy forensic formats, such as EWF.
-
-To access these legacy file formats we only need to have an index of
-each chunk offset, and then directly use the overlayed files.  The {\em
-ewf2aff} tool provides such an overlay feature for EWF volumes by
-employing libewf to scan the internal data structures.
-
-Overlay stream require that a \texttt{chunk\_size} attribute be specified. In
-addition the {\em compression\_type} attribute may be specified. The
-attribute should contain an integer corresponding to a suitable Zip
-compression scheme (e.g. DEFLATE). If not specified it is assumed
-that no compression is used.
-
-The ``target'' stream attribute may be specified multiple times and
-refer to external files (See section \ref{target_reference}). There
-can be a number of overlay segments named ``stream name/overlay.00''
-which are all merged to a single overlay. If more than one overlay is
-present, the attribute {\em overlays} MUST be specify how many.
-
-The overlay segment contains a series of lines with comma delimited,
-decimal encoded, 64 bit integers representing chunk number, target
-offset, target length and target number. Target number is an index
-into the targets defined in the attributes (count starts with zero).
-Chunks are then fetched from the specified target and decompressed if
-necessary.
-
-It is possible to refer to the current file by providing a target of
-``file://.''. This can be useful for modifying an existing EWF file to
-be a AFF4 file. Since Zip files are normally read from the end of the
-file, but EWF are both read from the front of the file, it may be
-possible to append a AFF4 Central Directory to the end of an EWF file
-without interfering with the overlayed file. In this case the
-converted file can still be used as an EWF file without change, but
-could also be used as an AFF4 file.
-
-In image streams some chunks may compress very well. In that case, the
-overheads introduced by the Zip File Header for each chunk in the
-Image stream may become unacceptably high. In that case its possible
-to use an {\em Overlay stream} to coalesce chunks into larger
-segments. This approach is illustrated in Figure \ref{overlay}.
-
-\begin{figure}[tb]
-  \begin{center}
-  \mbox{\epsfxsize=0.6\columnwidth \epsffile{overlay.eps}}
-  \caption{The use of an Overlay Stream to overlay an image with
-coalesced chunks.}
-  \label{overlay}
-  \end{center}
-\end{figure}
-
-In the above figure a segment is written to the AFF4 volume which
-contains a number of chunks back to back. An overlay stream then uses
-direct references to each compressed chunk within the coalesced
-segment. The target is then specified as the segment name. Although
-this technique is valid, it is generally discouraged since it breaks
-the appealing simplicity of the standard Image Stream.  It is no
-longer possible to simply unzip all the chunks and concatenate them
-together to recover the original image.
+%\Subsection{The Overlay Stream}
+%The Overlay Stream Is A Variant Of The {\Em Image} Stream That
+%Provides A Layer Of Indirection Between Requests Made For Specific
+%Bytes (Or Chunks Or Bevys) And The Actual Location Where The Requested
+%Information Is Stored. The Main Utility Of The Overlay Stream Is In
+%Providing Transparent Access To Legacy Forensic Formats, Such As Ewf.
+
+%To Access These Legacy File Formats We Only Need To Have An Index Of
+%Each Chunk Offset, And Then Directly Use The Overlayed Files.  The {\Em
+%Ewf2aff} Tool Provides Such An Overlay Feature For Ewf Volumes By
+%Employing Libewf To Scan The Internal Data Structures.
+
+%Overlay Stream Require That A \Texttt{Chunk\_Size} Attribute Be Specified. In
+%Addition The {\Em Compression\_Type} Attribute May Be Specified. The
+%Attribute Should Contain An Integer Corresponding To A Suitable Zip
+%Compression Scheme (E.G. Deflate). If Not Specified It Is Assumed
+%That No Compression Is Used.
+
+%The ``Target'' Stream Attribute May Be Specified Multiple Times And
+%Refer To External Files (See Section \Ref{Target_Reference}). There
+%Can Be A Number Of Overlay Segments Named ``Stream Name/Overlay.00''
+%Which Are All Merged To A Single Overlay. If More Than One Overlay Is
+%Present, The Attribute {\Em Overlays} Must Be Specify How Many.
+
+%The Overlay Segment Contains A Series Of Lines With Comma Delimited,
+%Decimal Encoded, 64 Bit Integers Representing Chunk Number, Target
+%Offset, Target Length And Target Number. Target Number Is An Index
+%Into The Targets Defined In The Attributes (Count Starts With Zero).
+%Chunks Are Then Fetched From The Specified Target And Decompressed If
+%Necessary.
+
+%It Is Possible To Refer To The Current File By Providing A Target Of
+%``File://.''. This Can Be Useful For Modifying An Existing Ewf File To
+%Be A Aff4 File. Since Zip Files Are Normally Read From The End Of The
+%File, But Ewf Are Both Read From The Front Of The File, It May Be
+%Possible To Append A Aff4 Central Directory To The End Of An Ewf File
+%Without Interfering With The Overlayed File. In This Case The
+%Converted File Can Still Be Used As An Ewf File Without Change, But
+%Could Also Be Used As An Aff4 File.
+
+%In Image Streams Some Chunks May Compress Very Well. In That Case, The
+%Overheads Introduced By The Zip File Header For Each Chunk In The
+%Image Stream May Become Unacceptably High. In That Case Its Possible
+%To Use An {\Em Overlay Stream} To Coalesce Chunks Into Larger
+%Segments. This Approach Is Illustrated In Figure \Ref{Overlay}.
+
+%\Begin{Figure}[Tb]
+%  \Begin{Center}
+%  \Mbox{\Epsfxsize=0.6\Columnwidth \Epsffile{Overlay.Eps}}
+%  \Caption{The Use Of An Overlay Stream To Overlay An Image With
+%Coalesced Chunks.}
+%  \Label{Overlay}
+%  \End{Center}
+%\End{Figure}
+
+%In The Above Figure A Segment Is Written To The Aff4 Volume Which
+%Contains A Number Of Chunks Back To Back. An Overlay Stream Then Uses
+%Direct References To Each Compressed Chunk Within The Coalesced
+%Segment. The Target Is Then Specified As The Segment Name. Although
+%This Technique Is Valid, It Is Generally Discouraged Since It Breaks
+%The Appealing Simplicity Of The Standard Image Stream.  It Is No
+%Longer Possible To Simply Unzip All The Chunks And Concatenate Them
+%Together To Recover The Original Image.
 
 \subsection{Encrypted Streams}
+\label{crypted_stream}
 Encryption is an important property in an evidence file format. In
@@
 containing both network captures and disk images it may be desirable
-to limit access to streams based on legal authorizations.
+to limit access to streams based on legal authorizations, even though
+the same set is distributed to a number of people.
 
@@
 AFF4 therefore introduces a new encryption scheme, the Encrypted
-Stream.  The Encrypted Stream is a stream that encapsulates other
-streams on a bevy-by-bevy basis. Each bevy transparently encrypted upon
-writing and decrypted on reading. The choice of algorithm and mode is declared in the
-Stream's \texttt{scheme} attribute.
+Stream.  The Encrypted Stream provides transparent encryption and
+decryption onto a single target stream. The target stream actually
+stores the encrypted data, and read requests from the stream are
+satisfied by decrypting the relevant data from this backing
+stream.
 
@@
 stream. This provides block level encryption for the contained AFF4
-volume. It is recommended that Encrypted Streams containing a AFF
-volume set an attribute (\texttt{content-type=application/x-aff-volume})
-to allow implementations to automatically recognize volumes. This
-approach is illustrated in Figure~\ref{crypted_fif}.
+volume. It is recommended that Encrypted Streams containing an AFF4
+volume set an attribute \texttt{content-type=application/x-aff-volume}
+on the container stream, to allow implementations to automatically
+recognize volumes. This approach is illustrated in
+Figure~\ref{crypted_fif}.
 
@@
   \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
-  \caption{Embedding an encrypted AFF4 volume within an Encrypted Stream.}
+  \caption{Embedding an encrypted AFF4 volume within an Encrypted
+Stream. The container volume contains an encrypted stream backed by an
+image stream. Once the encrypted stream is opened, the volume stored
+on its image stream is accessible. Now it is possible to see the
+secret image stream stored within the volume.}
   \label{crypted_fif}
@@
 \subsection{Encryption Schemes}
-We specify a number of encryption schemes, but extra schemes may be
-defined by the implementation.
+Our implementation provides a number of encryption schemes.
 
@@
 
-When creating a new volume, a master key is generated using the first
-128 bits of a SHA1 hash of the Pre-Shared-Key appended to a 64 bit
-random salt:
+When creating a new encryption stream, a master key is generated using
+the first 128 bits of a SHA1 hash of the Pre-Shared-Key appended to a
+64 bit random salt:
 \begin{eqnarray}
@@
 
-The block is then encrypted or decrypted using AES with this IV.
+The block is then encrypted or decrypted using AES in CBC mode with
+this IV.
 
@@
   new segments to the file, never modifying old segments.
+
 \item AFF4 ``properties'' have specific semantics above and beyond

### 2009-03-13 paper/paper.tex
Started to implement encrypted stream.

@@
 \item Our implementation supports simultaneous access by multiple
-  readers and writers. Our system allows concurrency by locking the
-  file and interleaving write operations. Managing simultaneous
-  writers is simplified through our update strategy which only writes
-  new segments to the file, never modifying old segments.
-
-\item AFF4 ``properties'' have specific semantics above and beyond
-  normal segments. For example, AFF4 calls for all of the property
-  segments associated with a file to be merged together. Furthermore,
-  when updating the properties in a multi-volume Set, it is useful to
-  write the same properties to each volume to ensure at least some of
-the stream can be used even if some volumes are lost. Such
-functionality is not present in standard ZIP64 implementations.
+  readers and writers. Since our system requires all metadata to be
+  shared through the Universal Resolver, this lends itself to
+  providing Universal Locking on a per Object basis. So for example,
+  if one process wants to add a new segment into a Zip volume, they
+  can lock it via the Resolver, add the segment and unlock the volume
+  object in the resolver.
 \end{itemize}
@@
 \section{Usage Scenarios}
-In this section we describe how a FIF file format may be used in
-various situations. Many of those can be solved using other forensic
-file formats, but often in a more awkward way.
+In this section we describe how AFF4 may be used in various
+situations. Many of those can be solved using other forensic file
+formats, but often in a more awkward way.
 
@@
 seamlessly reassembles the logical image from all the individual disk
-images. The new stream may be kept in its own volume and may be very
-small.
-
-If the images are split at exactly the same size its
-possible to create a FIF archive consisting of a single Overlay
-Stream, with a chunk\_size set to the size of each of the images. The
-images can then be listed in order as external targets (e.g. {\em
-target=file://disk1.dd}). This overlay can be created instantly
-without needing to re-compress any of the images. In addition, digital
-signatures can be generated for the whole stream or for each segment.
-
-\subsubsection{Rapid conversion of EWF file sets}
-Much existing evidence has been acquired using the EWF file
-format. This format, similarly, consists of a series of chunks which
-may be compressed, with interdispersed indexes. A tool such as
-``ewf2fif'' is able to utilize the libewf library to create an Overlay
-stream of the original image without the need to re-compress it. This
-conversion is very quick and the resulting FIF file is very small -
-each volume in the EWF file set is referenced through an external
-target reference, and the Overlay Stream is built as an index of each
-chunk location.
+images. The map stream may be kept in its own volume, or appended to
+one (or all) of the image fragments.
 
@@
 
-With the FIF format, each of the disks can be acquired as a separate
-Image Stream. Finally a tool such as PyFlag may be used to deduce the
-RAID map, which can be appended to the FIF file as a Map Stream. This
-Map Stream can then be opened by any tool without the tool needing to
-have explicit support for RAID reassembling.
+With the AFF4 format, each of the disks can be acquired as a separate
+Image Stream. Finally a tool such as PyFlag \cite{pyflag_raid} may be
+used to deduce the RAID map, which can be appended to the AFF4 file as
+a Map Stream. This Map Stream can then be opened by any tool to get a
+logical view of the RAID, without the tool needing to have explicit
+support for RAID reassembling.
+
+\subsubsection{In-place encryption of AFF4 volumes}
+With the large size of a typical forensic image it is often difficult
+to have to create a second copy of a volume which needs to be
+encrypted. It is usually more convenient to encrypt the file in place,
+overwriting the plain text data with the equivalent encrypted data.
+
+This can be achieved using the technique shown in
+Figure~\ref{inplace_crypto}. The first step is to encrypt the first
+block from the front of the file and add it to the end of the archive
+as a new encrypted segment. Then a ZipFile header is written to the
+front which encapsulates the rest of the plain text volume. The plain
+text volume is then encrypted in place. Finally a map stream is used
+to re-map the first and second fragments into place, and a
 
 \subsubsection{Cryptographic management of evidence}
-A FIF archive may hold multiple encrypted volumes, each in its own
+An AFF4 archive may hold multiple encrypted volumes, each in its own
 Encrypted Stream. Each of those streams is encrypted using a different

### 2009-03-13 src/include/class.h

@@ -275,7 +275,7 @@ extern char *_traceback;
 #define VMETHOD(method)				\
   (this)->method
 #define VATTR(attribute)			\
-  (this)->attribute
+  ((this)->attribute)
 
 #define INIT_CLASS(class)					\
   if(!__ ## class ## _initialised) { class ## _init(); }

### 2009-03-13 src/include/zip.h

@@ -7,7 +7,7 @@
 #include <zlib.h>
 
 #define HASH_TABLE_SIZE 256
-#define CACHE_SIZE 5
+#define CACHE_SIZE 15
 
 // A helper to access the URN or an object.
 #define URNOF(x)  ((AFFObject)x)->urn
@@ -339,6 +339,26 @@ CLASS(MapDriver, FileLikeObject)
      void METHOD(MapDriver, save_map);
 END_CLASS
 
+/************************************************************
+  An implementation of the encrypted Stream.
+
+  This stream encrypts and decrypts data from a target stream in
+  blocks determined by the "block_size" attribute. The IV consists of
+  the block number (as 32 bit int in little endian) appended to an 8
+  byte salt.
+*************************************************************/
+CLASS(Encrypted, FileLikeObject)
+     StringIO block_buffer;
+     char *key;
+     char *salt;
+     char *target_urn;
+     // Our volume urn
+     char *volume;
+
+     // The block size for CBC mode
+     int block_size;
+END_CLASS
+
 // A blob is a single lump of data
 CLASS(Blob, AFFObject)
      char *data;
@@ -467,5 +487,7 @@ END_CLASS
 // This is the main FIF class - it manages the Zip archive
 CLASS(FIFFile, ZipFile)
 END_CLASS
+
+void dump_stream_properties(FileLikeObject self, char *volume);
 
 #endif

### 2009-03-13 src/lib/encrypt.c

@@ -1,0 +1,90 @@
+#include "openssl/aes.h"
+#include "zip.h"
+
+AFFObject Encrypted_AFFObject_Con(AFFObject self, char *urn) {
+  Encrypted this = (Encrypted)self;
+  int value;
+
+  if(urn) {
+    URNOF(self) = talloc_strdup(self, urn);
+
+    value = resolver_get_with_default(oracle, self->urn, "aff2:block_size", "4k");
+    this->block_size = parse_int(value);
+
+    this->target_urn = CALL(oracle, resolve, URNOF(self), "aff2:target");
+    if(!this->target_urn) {
+      RaiseError(ERuntimeError, "No target stream specified");
+      goto error;
+    };
+
+    this->volume = CALL(oracle, resolve, URNOF(self), "aff2:stored");
+    if(!this->volume) {
+      RaiseError(ERuntimeError, "No idea where im stored?");
+      goto error;
+    };
+
+    CALL(self, set_property, "aff2:type", "encrypted");
+    this->block_buffer = CONSTRUCT(StringIO, StringIO, Con, self);
+    // FIXME
+    this->key = "hello world";
+    this->salt = "12345678";
+  } else {
+    this->__super__->super.Con(self, urn);
+  };
+
+  return self;
+
+ error:
+  talloc_free(self);
+  return NULL;
+};
+
+AFFObject Encrypted_finish(AFFObject self) {
+  return self->Con(self, URNOF(self));
+};
+
+int Encrypted_read(FileLikeObject self, char *buffer, unsigned long int length) {
+  Encrypted this = (Encrypted)self;
+  FileLikeObject target = (FileLikeObject)CALL(oracle, open, NULL, this->target_urn);
+  int result;
+
+  result = CALL(target, read, buffer, length);
+  CALL(oracle, cache_return, (AFFObject)target);
+
+  self->readptr += result;
+
+  return result;
+};
+
+int Encrypted_write(FileLikeObject self, char *buffer, unsigned long int length) {
+  Encrypted this = (Encrypted)self;
+  FileLikeObject target = (FileLikeObject)CALL(oracle, open, NULL, this->target_urn);
+  int result;
+
+  result = CALL(target, write, buffer, length);
+  CALL(oracle, cache_return, (AFFObject)target);
+
+  self->readptr += result;
+  self->size = max(self->size, self->readptr);
+  return result;
+};
+
[…]

### 2009-03-13 src/lib/image.c

@@ -227,29 +227,10 @@ static int Image_write(FileLikeObject self, char *buffer, unsigned long int leng
 
 static void Image_close(FileLikeObject self) {
   Image this = (Image)self;
-  char tmp[BUFF_SIZE];
-  char *properties;
 
   // Write the last chunk
   dump_chunk(this, this->chunk_buffer, this->chunk_buffer_readptr, 1);
-
-  // Set the stream size
-  snprintf(tmp, BUFF_SIZE, "%lld", self->size);
-  CALL(oracle, set, ((AFFObject)self)->urn, "aff2:size", tmp);
-
-  // Write out a properties file
-  properties = CALL(oracle, export, URNOF(self));
-  if(properties) {
-    ZipFile zipfile = (ZipFile)CALL(oracle, open, self, this->parent_urn);
-
-    snprintf(tmp, BUFF_SIZE, "%s/properties", ((AFFObject)self)->urn);
-    CALL((ZipFile)zipfile, writestr, tmp, ZSTRING_NO_NULL(properties),
-	 NULL, 0, ZIP_STORED);
-
-    talloc_free(properties);
-    // Done with zipfile
-    CALL(oracle, cache_return, (AFFObject)zipfile);
-  };
+  dump_stream_properties(self, this->parent_urn);
 
   this->__super__->close(self);
 };
@@ -425,6 +406,30 @@ static int Image_read(FileLikeObject self, char *buffer, unsigned long int lengt
 
   return len;
 };
+
+void dump_stream_properties(FileLikeObject self, char *volume) {
+  char tmp[BUFF_SIZE];
+  char *properties;
+
+  // Set the stream size
+  snprintf(tmp, BUFF_SIZE, "%lld", self->size);
+  CALL(oracle, set, ((AFFObject)self)->urn, "aff2:size", tmp);
+
+  // Write out a properties file
+  properties = CALL(oracle, export, URNOF(self));
+  if(properties) {
+    ZipFile zipfile = (ZipFile)CALL(oracle, open, self, volume);
+
+    snprintf(tmp, BUFF_SIZE, "%s/properties", ((AFFObject)self)->urn);
+    CALL((ZipFile)zipfile, writestr, tmp, ZSTRING_NO_NULL(properties),
+	 NULL, 0, ZIP_STORED);
+
+    talloc_free(properties);
+    // Done with zipfile
+    CALL(oracle, cache_return, (AFFObject)zipfile);
[…]

### 2009-03-13 src/lib/resolver.c

@@ -15,6 +15,7 @@ static struct dispatch_t dispatch[] = {
   { "link", (AFFObject)&__Link },
   { "image", (AFFObject)&__Image },
   { "map", (AFFObject)&__MapDriver},
+  { "encrypted", (AFFObject)&__Encrypted},
   { "file://", (AFFObject)&__FileBackedObject },
   // All handled by libcurl
   { "http://", (AFFObject)&__HTTPObject },
@@ -39,6 +40,7 @@ void AFF2_Init(void) {
   Resolver_init();
   Link_init();
   HTTPObject_init();
+  Encrypted_init();
 
   init_luts();
 
@@ -290,7 +292,7 @@ static void Resolver_return(Resolver self, AFFObject obj) {
 
 static void Resolver_add(Resolver self, char *uri, char *attribute, char *value) {
   Cache tmp;
-  //  printf("Adding to resolver: %s %s=%s\n", uri, attribute, (char *)value);
+  printf("Adding to resolver: %s %s=%s\n", uri, attribute, (char *)value);
   
   tmp = CALL(self->urn, get_item, uri);
   if(!tmp) {
@@ -470,7 +472,7 @@ static AFFObject AFFObject_Con(AFFObject self, char *uri) {
 
 
 static void AFFObject_set_property(AFFObject self, char *attribute, char *value) {
-  CALL(oracle, add,
+  CALL(oracle, set,
        self->urn,
        attribute, value);
 };

### 2009-03-13 src/lib/unittests.c

@@ -421,6 +421,83 @@ void test_http_handle() {
 
 };
 
+void test_encrypted(char *filename) {
+  ZipFile container = (ZipFile)CALL(oracle, create, (AFFObject *)&__ZipFile);
+  FileLikeObject encrypted_stream, embedded_stream;
+  ZipFile embedded_volume;
+  char *container_urn = talloc_strdup(NULL, URNOF(container));
+  char *encrypted_stream_uri, *embedded_stream_uri, *embedded_volume_uri;
+
+  CALL((AFFObject)container, set_property, "aff2:stored", "file://" TEST_FILE);
+
+  if(!CALL((AFFObject)container, finish))
+    return;
+
+  // Create an encrypted stream:
+  encrypted_stream = (FileLikeObject)CALL(oracle, create, (AFFObject *)&__Encrypted);
+  // The encrypted stream is in the container
+  CALL((AFFObject)encrypted_stream, set_property, "aff2:stored", URNOF(container));
+
+  // Create an embedded Image stream inside the encrypted stream:
+  embedded_stream = (FileLikeObject)CALL(oracle, create, (AFFObject *)&__Image);
+  CALL((AFFObject)embedded_stream, set_property, "aff2:stored", URNOF(container));
+  CALL((AFFObject)encrypted_stream, set_property, "aff2:target", URNOF(embedded_stream));
+
+  if(!CALL((AFFObject)encrypted_stream, finish)) {
+    CALL(oracle, cache_return, (AFFObject)container);
+    return;
+  };
+
+  if(!CALL((AFFObject)embedded_stream, finish)) {
+    CALL(oracle, cache_return, (AFFObject)encrypted_stream);
+    CALL(oracle, cache_return, (AFFObject)container);
+    return;
+  };
+
+  // Finished with those
+  CALL(oracle, cache_return, (AFFObject)container);
+  CALL(oracle, cache_return, (AFFObject)embedded_stream);
+  CALL(oracle, cache_return, (AFFObject)encrypted_stream);
+
+  // Create a new volume inside the embedded stream
+  embedded_volume = (ZipFile)CALL(oracle, create, (AFFObject *)&__ZipFile);
+  CALL((AFFObject)embedded_volume, set_property, "aff2:stored", URNOF(encrypted_stream));
+
+  if(!CALL((AFFObject)embedded_volume, finish)) {
[…]

### 2009-03-13 src/lib/zip.c

@@ -152,6 +152,7 @@ static uint64_t FileLikeObject_tell(FileLikeObject self) {
 static void FileLikeObject_close(FileLikeObject self) {
   //talloc_free(self);
 };
+
 static void FileBackedObject_close(FileLikeObject self) {
   FileBackedObject this=(FileBackedObject)self;
 

### 2009-03-14 paper/paper.tex
More work on the paper. Removed compression from encrypted streams.

@@
 \usepackage{listings}
+\usepackage{times}
 \lstset{language=Python,
@@
 %\bibliographystyle{elsart-num-names}
-\bibliographystyle{IEEEtran}
 \usepackage{cite}
@@
 %\begin{frontmatter}
+\newcommand{\omitt}[1]{}
 \begin{document}
 \title{Extending the Advanced Forensic Format to accommodate Multiple
-  Data Sources, Logical Evidence and Forensic Workflow}
+  Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow}
 \author{Blinded for review}
@@
 different types of evidence material, including individual disk
-drives, RAID sets, network packets, memory images, and extracted files. Often the same evidence is
-reviewed by several different tools or examiners in different
-locations. We propose a backwards-compatible evolutionary redesign of the Advanced
-Forensic Format an open, extensible
-file format for storing of evidence and interchanging analysis results
-among different tools. The Forensic Interchange Format (FIF) was
-designed to be simple to implement, relying on the well supported Zip
-specifications for bit level file access.
+drives, RAID sets, network packets, memory images, and extracted
+files. Often the same evidence is reviewed by several different tools
+or examiners in different locations. We propose a backwards-compatible
+evolutionary redesign of the Advanced Forensic Format an open,
+extensible file format for storing and sharing of evidence, arbitrary
+case related information and analysis results among different
+tools. The new specification was designed to be simple to implement,
+allowing the use of the well supported Zip File format specifications
+for bit level file access.
 \end{abstract}
@@
 \section{Introduction}
-
 Unlike other kinds of evidence, digital evidence can be instantly lost
@@
 actual file formats and containers used to store this digital
-evidence. Early practitioners created exact bit-for-bit copies (``mirror disks'') employed
-proprietary software systems (e.g. \cite{safeback,ilook,encase}) for
-making and authenticating ``images'' of digital evidence. PyFlag\cite{pyflag}
-introduced a ``seekable gzip'' format that allowed disk images to be
-stored in a form that was compressed but allowed the random-access to
-evidence data necessary for forensic analysis. The Advanced Forensic
-Format (AFF) expanded on this idea with a forensic file format that
-allowed both data and arbitrary metadata to be stored in a single
-digital archive\cite{garfinkel:aff}. Recently the open source AFF
-implementation library (AFFLIB) was expanded to support both
-encryption and digital signatures of digital evidence\cite{garfinkel:affcrypto}.
+evidence. Early practitioners created exact bit-for-bit copies
+(``mirror disks'') employed proprietary software systems
+(e.g. \cite{safeback,ilook,encase}) for making and authenticating
+``images'' of digital evidence. PyFlag\cite{pyflag} introduced a
+``seekable gzip'' format that allowed disk images to be stored in a
+form that was compressed but allowed the random-access to evidence
+data necessary for forensic analysis. The \emph{sgzip} format does not
+store any metadata at all.
+
+The Advanced Forensic Format (AFF) expanded on this idea with a
+forensic file format that allowed both data and arbitrary metadata to
+be stored in a single digital archive\cite{garfinkel:aff}. The EWF
+file format also provides for a small number of predefined metadata
+fields to be stored within the file format.
+
+Both of these approaches are limited in their expressiveness, by
+nature of their (property, value) data model. The subject of the
+properties is implicitly the single image contained within the
+evidence container. Clearly an implicit subject metadata model is
+inadequate for storing and managing multiple evidence sources
+simultaneously.
+
+The Sealed Digital Evidence Bags architecture of Schatz proposed an
+approach to digital evidence containers that facilitated composition
+of evidence and arbitrary evidence related information, through a
+simple data model and globally unique referencing
+scheme\cite{schatz:sdeb}.
+
+The AFF4 metadata model described in this work, is based on (subject,
+property, value) statements. An explicitly defined subject (identified
+by an AFF4 URN) allows one to not only associate traditional EWF style
+metadata with a contained image, but to also define new information
+entities representing both real world and virtual entities related to
+the matter. For example, an AFF4 compliant hard disk acquisition tool
+should, in addition to storing an image, store entities representing
+and describing the tool itself, the operator of the tool, the user
+supplied parameters supplied to the tool, and the drive being imaged.
+
+Recently the open source AFF implementation library (AFFLIB) was
+expanded to support both encryption and digital signatures of digital
+evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
+encryption and authentication is critical for forensic file formats.
 
@@
 specification, Garfinkel released AFFLIB implementations with major
-version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid confusion.}
+version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid
+confusion.}
 
@@
 
-%% Raid is only one example of the application of the map stream - the
-%% most important might be the ability to use sleuthkit to generate
-%% block allocation tables for all the files - thereby allowing
-%% arbitrary programs to read individual files without using sleuthkit
 \item Direct encoding of arbitrary mapping transformations of one or
-more sources to allow multiple logical views. For example, allowing
+more sources to allow multiple logical views. This allows 
 RAID volumes to be automatically reassembled from evidence files made
-of the physical RAID devices.
+of the physical RAID devices. Maps can be created using Sleuthkit and
+stored in the evidence file, thereafter allowing arbitrary programs to directly
+read individual files stored in a disk image.
+
 \item A revised compression system that allows individual chunks
@@
   entire segment.
+
 \item Support for Global Distributed Evidence Management.
@@
   computers or pieces of media.
+
+\item The data model of AFF1 enabled storing metadata related to the 
+  contained image as (property, value) pairs. This data model does
+  not, however, support expressing arbitrary information about more
+  than one entity.
+
 \item AFF1 has no provision for storing memory images or intercepted
   network packets.
+
 \item AFF1 has no provisions for storing extracted files that is
@@
   for linking evidence to web pages.
+
 \item AFF1's encryption system leaks information about the contents of
   an evidence file because segment names are not encrypted.
+
 \item AFF1's default compression page size of 16MB can impose significant overhead
@@
   use.
+
 \item Although the AFF1 specification calls for a ``table of contents'' similar
-  to the ZIP\cite{zip-format} ``central directory'' that is stored at the end of AFF
-  files, Garfinkel never implemented this directory in the publicly
-  released AFF1 implementation, AFFLIB. As a result, every header of every segment in an
-  AFF file needs to be read when a file is opened. It practice this
-  can take 10--30 seconds the first time a large AFF file is
-  opened. (Once the file is opened, the sectors corresponding to the
-  segment headers are stored in the computer's disk cache.)
+  to the Zip\cite{zip-format} ``central directory'' that is stored at
+  the end of AFF files, Garfinkel never implemented this directory in
+  the publicly released AFF1 implementation, AFFLIB. As a result,
+  every header of every segment in an AFF file needs to be read when a
+  file is opened. It practice this can take 10--30 seconds the first
+  time a large AFF file is opened. (Once the file is opened, the
+  sectors corresponding to the segment headers are stored in the
+  computer's disk cache.)
+
 \item AFF1's bit-level specification is essentially a simple container
@@
   from its home-grown format to one of the existing standards.
+
 \end{itemize}
@@
 analysis takes place in multiple locations and is performed by
-multiple individuals.
-
-Managing evidence in a globally distributed system requires the use of
-globally unique identifiers. The original AFF specification assigned
-each piece of evidence a unique 128-bit identifier called a GID but did
-not make it clear when this identifier should be changed and when it
-should remain the same.
-
-In practice, however, global distributed evidence management requires
+multiple individuals. This is a state of affairs which mirrors common
+practice in the field.
+
+Global distributed evidence management requires
 more than simply tracking the movement of disk images: it requires
@@
 work products of the analysts in a third security domain.
+
+Managing evidence in a globally distributed system requires the use of
+globally unique identifiers. AFF1 assigns
+each piece of evidence a unique 128-bit identifier called a GID but did
+not make it clear when this identifier should be changed and when it
+should remain the same.
 
@@
 
-In this case they can
-each create a new volume which extends to original volume and save
-their analysis on this new volume. Now they only need to share this
-new volume with other analysts who also have a copy of the old volume
-to interchange their findings.
+In this case they can each create a new volume which extends to
+original volume and save their analysis on this new volume. Now they
+only need to share this new volume with other analysts who also have a
+copy of the old volume to interchange their findings.
 
@@
 \subsection{Migration from AFF1 to AFF4}
-
-We have addressed these issues in three steps. First, we have extended
-the AFF schema to support multiple forensic images (which are now
-called \emph{streams}) in a single evidence file. Second, we replaced
-AFF's ``bit-level'' format with an improved format that is more
-flexible and has better existing support. Third, we are adopting
-AFFLIB to implement the new schema and file format, which will allow
-existing AFF-aware programs to access files written in either the old
-or the new format without modification.
+We have addressed these issues in a number of ways. First, we have
+extended the AFF schema to support multiple forensic images (which are
+now called \emph{streams}). It is therefore possible now to store
+multiple related streams in the same volume, or to acquire mutiple
+streams at the same time (e.g. network and disk data).
+
+We replaced AFF's ``bit-level'' format with an improved format that is
+more flexible and has better existing support, namely the Zip64 file
+format is not adopted as a default storage format. This enhanced our
+compatibility with other tools and simplifies implementations for
+compatible systems.
+
+We have also defined an information model which addresses uniquely
+identifying evidence data, expressing arbitrary information and
+integrating information and image data from multiple evidence volumes
+at a time. This allows us to extend evidence collection and management
+to a truely universal scale - allowing creation and management of a
+large corpus.
+
+We have also adopted AFFLIB to implement the new schema and file
+format, which will allow existing AFF-aware programs to access images
+written in either the old or the new format without modification. We
+are extending the AFFLIB API to support the new information mode.
 
@@
 
-This section discusses the AFF4 terminology.
-
-%% Since this is the itemised definition its probably better to
-%% restrict it to the basic definitions and then move most of the
-%% discussion a bit later. I want to show how concepts build on one
-%% another - maybe even put a UML or class diagram
+This section discusses the AFF4 terminology. The AFF4 design is object
+oriented, in that a few generic objects are presented with externaly
+accessible behaviour. We discuss a number of implementations of these
+high level concepts and show how these can be put together in common
+usage cases.
 
@@
 made unique by use of a unique identifier generated as per
-RFC4122\cite{rfc4122}.
+RFC4122\cite{RFC4122}.
 
@@
 
-% Is it important to define this? do we use it?
-%\item \emph{An Evidence Set} is a loose collection of one or more Evidence
-%Volumes.
-
-%Evidence Sets can also be stored in object storage system (such as
-%  Amazon S3), or a hybrid---for example, metadata stored in an SQL
-%  database with physical data stored in large files in a file
-%  system.
-
-% I would describe this process after the definitions.
-
-%\item \emph{An Evidence Set UUID} is a unique identifier generated
-%  according to RFC4122\cite{rfc4122} for each new Evidence
-%  Set. Multiple Evidence Volumes that make up the same Evidence Set
-%  should all have the same UUID. When an Evidence Set is opened all its volumes are
-%  loaded, their UUIDs are checked, and all of their contents are
-%  logically merged so as to be visible as if they were all found in
-% the same volume.
-
-%% I Wonder if this explaination should be postposed to the section we
-%% actually discuss it - otherwise we duplicate it again
-%% \item \emph{An image stream} is a stream that is dedicated to storing evidence from a single
-%%  source of data.  For example, a stream may contain a disk image or a network
-%%  capture. Whereas AFF could only support a single stream in an evidence
-%%  file, AFF4 allows any number of streams to be stored. Each stream in
-%%  the archive is  given a unique identifier.
-
-%%\item \emph{A bevy} is a specific segment that is used to hold
-%%  forensic image data, such as disk sectors, memory pages, or network
-%%  packets. AFF1 used the term ``pages'' to describe bevys, but this
-%%  term caused confusion when AFF was used to hold memory images, hence the
-%%  new term.
-
-%% This is no longer correct - segment names are URNs and are encoded
-%% according to RFC1737 as mentioned above.
-%%\item \emph{A segment name} is the name under which a segment is
-%%  stored within the archive. AFF4 segments have two-part names that
-%%  typically contain a \emph{stream name} and \emph{bevy number}. For
-%%  example the segment named \texttt{default/000001} refers to the stream
-%%  called \texttt{default} and has a chunk name of \texttt{000001}. All segment
-%%  names are encoded using UTF8. AFFLIB is being modified to allow any
-%%  stream to be opened using the \texttt{af\_open} function; if a
-%%  stream name is not specified, AFF4 implementations will open the
-%%  stream named ``default.''
-
-%%\item \emph{The segment timestamp} is a timestamp for the segment
-%%  recorded in GMT.
-
-%% This should be discussed as a way we are storing the RDF triples,
-%% rather than as a definition.
-%% \item \emph{Attributes} are a list of key/value pairs stored in each
-%% stream's ``properties'' segment (e.g. ``stream/properties''). The
-%% ``properties'' file is a text file which stores attributes one on each
-%%line. Attribute lines consist of the key, immediately followed by
-%%``='' for text values or ``:'' for binary values. This is immediately
-%%followed by the encoded value. Values containing text are encoded
-%%using UTF8, while values containing binary data are encoded using
-%%Base64. There may be multiple values for each attribute; in these
-%%cases the order is preserved
-%%by implementations. Implements may present different keys may be
-%%presented in any order, however.
-%%\vspace{1ex}
-
-%% This is not correct any more as volumes must store their own URN in
-%% a different way (the zip implementation uses an archive comment for
-%% this).
-
-%% Each AFF4 archive must also have a top-level set of attributes---that
-%% is, there is always a segment named ``properties.'' These volume
-%% attributes are used to specify a Unique Identifier (UUID). (AFF1
-%% stored each property in its own named segment. This created
-%% significant overhead in many situations.)
-
-%% Since target references are all just URNs now the resolving is automatic.
-
 \item \emph{A Target Reference} is a way of referencing objects by use
@@
 Object URN or may be a more general Uniform Resource Locator (URL),
-such as for example a HTTP or FTP object.
+such as for example a HTTP or FTP object. This innovation allows
+objects in one volume to refer to objects in different volumes,
+facilitating data fusion and cross referencing.
 
 \item \emph{The Resolver} is a central data store which collects and resolves
-attributes for the different AFF Objects. We discuss the Resolver in
-detail in Section \ref{resolver}.
-
-%%in a number of places which will be highlighted below. Using
-%%references it is possible for the FIF file to refer to objects
-%%embedded within itself, or within other FIF files. Its is also
-%%possible to refer to external flat files. A number of usage scenarios
-%%are described in Section \ref{usage_scenarios}. 
+attributes for the different AFF Objects. The Resolver has universal
+visibility of objects from all volumes, and therefore guides
+implementations in resolving external references. We discuss the
+Resolver in detail in Section \ref{resolver}.
 \end{itemize}
 
-Since AFF Objects have globally unique URNs it is possible for an
-Object in one volume to refer to an object in another volume. The
-resolver has visibility of all URNs and their current physical
-location.
-
-%The Target Reference is an important AFF4 innovation: this mechanism
-%gives AFF4 files a way for describing evidence that is stored in
-%different evidence volumes, in the host computer's file system, or on
-%the Internet. We have defined the following target reference schema:
-
-\label{target_reference}
-
-%\begin{itemize}
-%\item A reference in the format {\em aff://UUID/stream/segment} refers
-%to an external Evidence file set (identified by its UUID). If the segment
-%is missing (i.e. the reference ends with a ``/''), the reference is
-%for the entire stream. We do not specify a way of actually resolving
-%the UUID into a physical locations. This can achieved by
-%implementations using many methods, such as Active Directory, LDAP,
-%DNS, flat files or SQL databases.
-
-%\item A reference starting with {\em file://} refers to an external
-%file. The external file reference must be sanitised (to remove
-%.. directory traversal) and rooted at the current directory. A
-%filename of ``.'' refers to the current FIF volume.
-
-%\item A reference starting with {\em http://} refers to a HTTP
-%object which will be fetched. Support for this is optional.
-
-%% This is no longer correct as all segments, streams and volumes have
-%% a unique URN so we dont need to make these distinctions.
-
-%% \item A references ending with a ``/'' signifies a reference to a
-%% stream in the current FIF file. The stream will be opened in its
-%% entirety. If the reference does not end with a ``/'', the reference is
-%% for a single segment.
-%%\end{itemize}
-
 \section{Metadata and the Universal Resolver}
-The AFF4 model treats metadata as an abstract concept which may exists
+\label{resolver}
+One of the fundamental tasks of forensics is in ensuring the
+provenance of any particular piece of digital evidence, for example by
+employing a chain of custody process. To assist in this, acquisition
+methodologies emphasise the process of individuating evidence, and
+storing it in a manner that it is easily identifiable.
+
+Management of evidence requires on effective identififaction, with
+practitioners currently employing acquisition time metadata such as
+case identifiers and description fields in the EWF file format; file
+and directory naming schemes, and labeling of evidence container hard
+drives. Evidence may also be referred to by external means in an
+inconsistent way. For example, in an investigator's case note a disk
+image may be referred to by the name of the suspect (e.g. Joe's hard
+disk), the case number or dates.
+
+Such individuation schemes may be, however, problematic when
+automatically managing evidence. For example, at acquisition time a
+suitably unique individuator may not be selected. If that occured, at
+analysis time evidence container files may need to be renamed to avoid
+name collisions.
+
+The AFF4 design adopts a scheme of globally unique identifiers for
+identifying and referring to all evidence. We define an AFF4 specific
+URN namespace\cite{RFC1737} for our naming scheme (\emph{urn:aff4}).
+URNs can then be made unique by use of a unique identifier generated
+as per RFC4122\cite{RFC4122}. For example, an AFF4 URN might be
+\emph{urn:aff4:bcc02ea5-eeb3-40ce-90cf-7315daf2505e}
+
+In this way we can identify uniquely any evidence without the change
+for name collisions. Once a universal identification regime is
+created, it is now possible to collect metadata about each object. The
+AFF4 model treats metadata as an abstract concept which may exists
 independently from the data itself. We term {\em metadata} to be a set
@@
 loads any properties files and populates its Universal Resolver with
-the information it has visibility of. AFFLIB provides a mechanism to
-use an external resolver stored on a MySQL database to provide a
-Universal Resolver that shares information between different instances
-on the same network.
-
-Each URN within the AFF4 universe must have an ``\texttt{type}''
+the information visibile to it. AFFLIB provides a mechanism to use an
+external resolver stored on a MySQL database to provide a Universal
+Resolver that shares information between different instances on the
+same network.
+
+Each URN within the AFF4 universe must have an ``\texttt{aff4:type}''
 attribute to denote the type of the Object.
@@
 Directory Volume:
-  urn:aff2:f901be8e-d4b2 aff2:stored=http://../case1/
-  urn:aff2:f901be8e-d4b2 aff2:type=directory
+  urn:aff4:f901be8e-d4b2 aff4:stored=http://../case1/
+  urn:aff4:f901be8e-d4b2 aff4:type=directory
 
 ZipFile Volume:
-  urn:aff2:98a6dad6-4918 aff2:stored=file:///file.zip
-  urn:aff2:98a6dad6-4918 aff2:type=zip
+  urn:aff4:98a6dad6-4918 aff4:stored=file:///file.zip
+  urn:aff4:98a6dad6-4918 aff4:type=zip
 
 Image Stream:
-  urn:aff2:83a3d6db-85d5 aff2:stored=urn:aff2:f901be8e-d4b2
-  urn:aff2:83a3d6db-85d5 aff2:chunks_in_segment=256
-  urn:aff2:83a3d6db-85d5 aff2:chunk_size=32k
-  urn:aff2:83a3d6db-85d5 aff2:type=image
-  urn:aff2:83a3d6db-85d5 aff2:size=5242880
+  urn:aff4:83a3d6db-85d5 aff4:stored=urn:aff4:f901be8e-d4b2
+  urn:aff4:83a3d6db-85d5 aff4:chunks_in_segment=256
+  urn:aff4:83a3d6db-85d5 aff4:chunk_size=32k
+  urn:aff4:83a3d6db-85d5 aff4:type=image
+  urn:aff4:83a3d6db-85d5 aff4:size=5242880
 
 Map Stream:
-  urn:aff2:ed8f1e7a-94aa aff2:target_period=3
-  urn:aff2:ed8f1e7a-94aa aff2:image_period=6
-  urn:aff2:ed8f1e7a-94aa aff2:blocksize=64k
-  urn:aff2:ed8f1e7a-94aa aff2:stored=urn:aff2:83a3d6db-85d5
-  urn:aff2:ed8f1e7a-94aa aff2:type=map
-  urn:aff2:ed8f1e7a-94aa aff2:size=0xA00000
+  urn:aff4:ed8f1e7a-94aa aff4:target_period=3
+  urn:aff4:ed8f1e7a-94aa aff4:image_period=6
+  urn:aff4:ed8f1e7a-94aa aff4:blocksize=64k
+  urn:aff4:ed8f1e7a-94aa aff4:stored=urn:aff4:83a3d6db-85d5
+  urn:aff4:ed8f1e7a-94aa aff4:type=map
+  urn:aff4:ed8f1e7a-94aa aff4:size=0xA00000
 
 Link Object:
-  map aff2:target=urn:aff2:ed8f1e7a-94aa
-  map aff2:type=link
+  map aff4:target=urn:aff4:ed8f1e7a-94aa
+  map aff4:type=link
 \end{lstlisting}
@@
 
-The Directory Volume uses the {\em aff2:stored} attribute to provide a
+The Directory Volume uses the {\em aff4:stored} attribute to provide a
 base URL. The URL for each segment is then constructed by appending
@@
 
-\subsection{ZIP64}
+\subsection{Zip64 Volumes}
 For AFF4, we have changed the default volume container file format to
-ZIP64\cite{zipspecs}. There are many reasons for this decision:
+Zip64\cite{zip-format}. There are many reasons for this decision:
 
 \begin{itemize}
-\item There is already wide support for the ZIP and ZIP64 formats. By
+\item There is already wide support for the Zip and Zip64 formats. By
   migrating to these formats, we can take advantage of the rich number
   of user and developer tools already available. The volume may be
-inspected using any number of commercial zip application (e.g. Windows
-Explorer natively supports Zip files).
-
-%% In not sure how this is relevant
-%\item ZIP64 includes a CRC32 and timestamp for each segment. We
-%  believe that the additional integrity checks and timestamps will be
-%  useful in a forensic environment.
-
-\item ZIP64 already includes support for the ``table of contents''
-  envisioned but never implemented by Garfinkel; ZIP64 calls this
+  inspected using any number of commercial zip application
+  (e.g. Windows Explorer natively supports Zip files).
+
+\item Zip64 already includes support for the ``table of contents''
+  envisioned but never implemented by Garfinkel; Zip64 calls this
   table the ``Central Directory.'' It is written at the end of the
-  ZIP64 volume.
-
-\item ZIP64 libraries are readily available making proprietary implementations of
+  Zip64 volume.
+
+\item Zip64 libraries are readily available making proprietary implementations of
 interfaces to the AFF4 volume format simple to write. For example, a
-simple python program to dump out an Image stream is illustrated in
-Listing \ref{python_code}.
+simple python program to dump out an Image stream
+(section~\ref{image_stream}) is illustrated in Listing
+\ref{python_code}.
 
@@
 
-\begin{figure}[]
+\begin{figure}[tbp]
   \begin{center}
   \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
-  \caption{The basic structure of a Zip archive}
+  \caption{The basic structure of a Zip archive. Also can be seen how
+new archive members are added to an existing Zip File. The
+Central Directory gets overwritten by the new member, and a new
+Central Directory is written on the end.  }
   \label{zip_structure}
@@
 capable of storing multiple named segments of data. Although our AFF4
-implementation uses the ZIP64 file format as an underlying storage
+implementation uses the Zip64 file format as an underlying storage
 mechanism, our system also supports legacy AFF1 volumes as well as
 segments that are stored in other storage systems---for example,
-Amazon.com's Simple Storage System (S3)\cite{s2-aws-home-page-money}.
-
-\subsection{Unused ZIP64 Features}
+Amazon.com's Simple Storage System (S3)\cite{s3-aws-home-page-money}.
+
+\subsection{Unused Zip64 Features}
 There are a number of aspects in the Zip file format specifications
@@
 \item We ignore Zip64's built-in support for splitting
-archives into multiple ZIP files. Instead, our implementation treats
-each volume as complete and stand-alone ZIP file. The AFF4
+archives into multiple Zip files. Instead, our implementation treats
+each volume as a complete and stand-alone Zip file. The AFF4
 implementation then considers the segments contained within as
@@
 extensions, we do not use them due to the restrictions imposed on
-their use and because they lack of functionality that is important for
-forensic user. Instead, we use AFF1's digital signature facilities for
-integrity and non-repudiation, and we introduce a new stream based
-encryption scheme for ensuring data privacy
+their use and because they lack the functionality that is important
+for a forensic user. Instead, we use AFF1's digital signature
+facilities for integrity and non-repudiation, and we introduce a new
+stream based encryption scheme for ensuring data privacy
 (Section~\ref{crypted_stream}).
@@
 
-%% Im not sure this is really that relevant here. Byte range updates
-%% can be achieved by an append and a map operation, which might be a
-%% better option for aimager working in dd_rescue mode.
-
-%\subsubsection{A Forensically Sound Strategy For Updating ZIP64 Files}
-%In many forensic applications, individual segments within the volume
-%need to be updated. To this end, Garfinkel implemented several
-%``update'' functions in AFF1\cite{garfinkel:aff}.
-
-%There were two problems with the original AFF updating approach:
-%\begin{itemize}
-%\item Updates can cause the volume to
-%become fragmented as holes appear within the middle of the volume, and
-%updated segments are added at the end.
-%\item Updates could be performed silently, raising questions of
-%  forensic soundness.
-%\end{itemize}
-
-%In AFF4, when a new segment is added to an existing volume, the new
-%segment header overwrites the volume central directory, and the
-%central directory should be rewritten including the new entry
-%immediately after the last file in the archive. If it is not possible
-%to overwrite the file, the new segment can be written at the end,
-%followed by an updated copy of the central directory (in that case the
-%old central directory remains in the file). This is illustrated in
-%Figure \ref{zip_structure}. Implementations may delay
-%writing the central directory until the file is closed; should the
-%system crash before the central directory is written, it can be
-%recovered by scanning the entire ZIP file.
-
-%A segment within the archive is considered to be updated when a new
-%segment is added to the archive with the same name and a later
-%timestamp. Implementations satisfy read requests for duplicate
-%segment names with those segments with the later timestamp, or if the
-%two timestamps are the same, the segment which appears at a later
-%offset in the volume. This effectively allows for a segment to be
-%updated without leading to internal volume fragmentation.
-
-%The fact that the older version of all updated segments remains intact
-%within each volume can be used for auditing, or to allow a rollback to
-%previous versions of the segment. We have created a new tool,
-%\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
-%creating new volumes and redistributing segments among them.
+\subsection{Implementing Zip64 in AFFLIB}
+Although there are numerous Zip implementations available
+today, we have created our own implementation using a combination of
+available open source technology and our own novel contributions. We
+are in the process of integrating this code into the AFFLIB library
+and performing extensive regression testing.
+
+There are many reasons to develop our own Zip64 implementation for
+AFF4:
+
+\begin{itemize}
+\item Most Zip implementations do not implement the Zip64
+  extensions. These extensions are required to support Evidence
+  Volumes larger than 2GB.
+
+\item Simple Zip implementations might rescan the
+  Central Directory for each segment request. Since in practice there
+  can be a large number of segments in a volume, it is advisable to
+  have a Zip64 implementation that is optimized to storing thousands
+  (or even hundreds of thousands) of segments in an efficient data
+  structure. In fact our implementation uses the Universal Resolver to
+  store the parsed central directory information, which means that in
+  most cases we dont even need to scan the Central Directory at all.
+
+\item While the Zip specification duplicates data found in the Central
+  Directory entry in each File Header (such as filename, size, CRC
+  etc), many implementations that we have examined only populate this
+  information in one of these places. In the interest of robustness,
+  we wanted to ensure that data stored in both locations would be
+  populated to allow recovery of at least \emph{some} evidence that
+  might exist in damaged volumes. If the central directory is lost, it
+  is possible to scan through the volume, and locate all the Zip64
+  file headers. Then its possible to repair and reconstruct the
+  central directory.
+
+\item Our implementation supports simultaneous access by multiple
+  readers and writers. Since our system requires all metadata to be
+  shared through the Universal Resolver, this lends itself to
+  providing Universal Locking on a per Object basis. So for example,
+  if one process wants to add a new segment into a Zip volume, they
+  can lock it via the Resolver, add the segment and unlock the volume
+  object in the resolver, stopping concurrent access by other
+  programs, even on different machines.
+\end{itemize}
 
 \section{Streams}
-The Stream object provides random access to an abstract representation
+The Stream system provides random access to an abstract representation
 of a body of data. Our implementation allows the segments in a stream
@@
 \texttt{open()}, \texttt{seek()}, \texttt{write()}, and
-\texttt{read()}. All streams must have the ``\texttt{size}'' attribute
-to denote the last bytes addressible within the stream. This is
-required in order to support the POSIX whence attribute which may
-require seeking from the end of the stream.
-
-The following describes a number of types of streams. It is important
-to note that clients of our implementation do not care how a
+\texttt{read()}. All streams also have a ``\texttt{size}'' attribute
+to denote the last byte addressable within the stream. This is
+required in order to support the POSIX \emph{whence} attribute which
+may require seeking from the end of the stream.
+
+The following sections describe a number of types of streams. It is
+important to note that clients of our implementation do not care how a
 particular stream is implemented. Streams are opened by their URNs,
-and the library itself ensure they provide the Stream interface. So
+and the library itself ensures they provide the Stream interface. So
 for example, users do not care if a stream is a Map Stream or an Image
-Stream - the data provided is the same.
-
-%% Not sure how its relevant - maybe we can move it to a new section
-%% about randomly writing stream.
-
-%AFFLIB also supports the \emph{writing} of streams, a feature that is
-%used by Garfinkel's \texttt{aimage} disk imaging program. In general
-%arbitrary writing to streams can be expensive, especially when writing
-%is combined with seek operations. The current \texttt{aimage} program
-%will create AFF files with gaps in them during its error recovery
-%process (when \texttt{aimage} seeks to the end of a disk and reads it
-%backwards in an attempt to get around media failures). We are working
-%on a rewrite of \texttt{aimage} which performs imaging and recovery
-%on bevy boundaries, rather than using the current arbitrary
-%``high-water-mark'' and ``low-water-mark.''
-
-%All streams within an Evidence Set \emph{must} have a properties segment. The
-%properties segment describes what the stream contains and determines
-%how the stream is processed.
+Stream---the interface provided is the same.
 
@@
 memory image or a network capture (in PCAP format). Image streams have
-an \texttt{aff2:type} attribute of \texttt{image}.
+an \texttt{aff4:type} attribute of \texttt{image}.
 
@@
 digit, zero padded decimal integer representation of an incrementing
-id to the stream URN (e.g. ``urn:aff2:83a3d6db-85d5/00000032''). Each
+id to the stream URN (e.g. ``urn:aff4:83a3d6db-85d5/00000032''). Each
 data segment is called a bevy and stores a number of compressed chunks
-back to back. The chunk index segment is a segment containing a list
-of relative offsets to the beginning of each chunks within the
-bevy. The chunk index segment URN is derived  by appending the bevy
-URN with ``.idx''. This is illustrated in Figure \ref{image_stream}.
+back to back.
+
+The chunk index segment is a segment containing a list of relative
+offsets to the beginning of each chunk within the bevy. The chunk
+index segment URN is derived by appending the bevy URN with
+``.idx''. This is illustrated in Figure \ref{image_stream_bevy}.
 
@@
 offsets are stored in the chunk index segment.}
-  \label{image_stream}
+  \label{image_stream_bevy}
   \end{center}
@@
 compressed individually using the zlib compress algorithm. This
-general structure of storing chunks within larger segments similar to
-the technique used by the Expert Witness file format used by
+general structure of storing chunks within larger segments is similar
+to the technique used by the Expert Witness file format (EWF) used by
 EnCase\cite{encase-3.0} and implemented by the open source
@@
 drawn from an image, while a TCP/IP stream is simply a collection of
-payloads from selected network packets. 
-Sometimes the same data may be
+payloads from selected network packets. Sometimes the same data may be
 viewed in a number of ways---for example a Virtual Address Space is a
 mapping of the Physical Address Space through a page table
-transformation.  Zero Storage Carving \cite{Meijer2006} is a way of
+transformation.  Zero Storage Carving\cite{Meijer2006} is a way of
 specifying carved files in terms of a sequence of blocks taken from
-the image; Cohen extended this concept to an arbitrary
-mapping function\cite{Cohen2007} which can be used to describe
+the image; Cohen extended this concept to an arbitrary mapping
+function\cite{1363239,Cohen2007} which can be used to describe
 arbitrary mappings of carved files within a single image.
@@
 
-The mapping function is described in a segment named by appending a
+The mapping function is described in a segment named by appending
 ``/map'' to the stream URN. The segment data consists of a series of
@@
 \begin{lstlisting}
-0,0,urn:aff2:83a3d6db-85d5
-4096,10000,urn:aff2:f901be8e-d4b2
-8192,5000,urn:aff2:83a3d6db-85d5
+0,0,urn:aff4:83a3d6db-85d5
+4096,10000,urn:aff4:f901be8e-d4b2
+8192,5000,urn:aff4:83a3d6db-85d5
 \end{lstlisting}
@@
 4095 in the stream from offset 0 to offset 4095 in
-\emph{urn:aff2:83a3d6db-85d5}. Requests for bytes between 4096 and 8191 are
-fetched from \emph{urn:aff2:f901be8e-d4b2} at offset 10000. Finally
+\emph{urn:aff4:83a3d6db-85d5}. Requests for bytes between 4096 and 8191 are
+fetched from \emph{urn:aff4:f901be8e-d4b2} from offset 10000. Finally
 bytes after 8192 (until the specified size of the stream) are fetched
-from offset 5000 in \emph{urn:aff2:83a3d6db-85d5}.
+from offset 5000 in \emph{urn:aff4:83a3d6db-85d5}.
 
@@
 
+\subsection{The HTTP Stream}
+Arguably the most ubiquitous protocol for information sharing is the
+HTTP protocol\cite{HTTP_RFC}. The protocol features mature
+authentication and auditing and is fast and easy to set up with
+numerous web server implementations available on the market. The HTTP
+protocol is also designed to operate across a wide range of network
+architectures and is therefore more deployable than traditional file
+sharing protocols.
+
+For these reasons it is desirable to allow the HTTP protocol to be
+used in facilitating the sharing of evidence files between
+investigators. Luckily, the HTTP protocol fits naturally within the
+URN based scheme adopted by AFF4, since the HTTP Universal Resource
+Locator (URL) scheme is a subset of the URN scheme.
+
+For this reason, URLs may be used interchangeably with a URN within
+the AFF4 universe. For example, the \emph{aff4:stored} attribute of a
+volume may be specified as a URL
+(e.g. \emph{http://intranet/123453/}).  AFFLIB provides transparent
+support for HTTP and FTP URLs by means of the curl HTTP
+library\cite{libcurl}. The HTTP Stream, therefore satisfies read
+requests by making HTTP requests to the web server. We use the
+\emph{Content-Range} HTTP header to request exactly the byte range the
+client is interested in. This allows efficient network transport as we
+do not need to download unnecessary data, we just request exactly the
+data the client application requires.
+
+Currently the HTTP Stream is read only, but in future we anticipate
+write support to be implemented, for example, using the WebDav
+extensions to HTTP\cite{webdav-rfc}. The HTTP stream also supports the
+File Transfer Protocol (FTP) and HTTPS (Secure Sockets Layer---SSL)
+protocols transparently. This support is provided automatically by the
+curl library.
+
 \subsection{Encrypted Streams}
@@
 for our purposes since it encrypts each segment separately, and does
-not specify a sufficiently flexible scheme (e.g. support for PKI, PGP
-keys). Segment based encryption may lead to information leakage when
-segments are compressed, as the uncompressed size of the segment may
-be deduced.
+not specify a sufficiently flexible scheme (e.g. support for PKI or
+PGP keys). Segment based encryption may lead to information leakage
+when segments are compressed, as the uncompressed size of the segment
+may be deduced.
 
@@
 satisfied by decrypting the relevant data from this backing
-stream.
-
-The Encrypted Stream may contain any data at all. It is useful
-however, to store an actual AFF4 volume within the Encrypted
-stream. This provides block level encryption for the contained AFF4
-volume. It is recommended that Encrypted Streams containing an AFF4
-volume set an attribute \texttt{content-type=application/x-aff-volume}
-on the container stream, to allow implementations to automatically
-recognize volumes. This approach is illustrated in
-Figure~\ref{crypted_fif}.
+stream. The encrypted stream itself does not store any data at all ---
+all data is stored on its target stream.
+
+The Encrypted Stream may contain any data at all, including disk
+images, network captures or memory images. It is useful however, to
+store an entire AFF4 volume within the Encrypted stream. This provides
+block level encryption for the contained AFF4 volume (which might
+contain arbitrary streams). It is recommended that Encrypted Streams
+containing an AFF4 volume set an attribute
+\emph{content-type=application/x-aff-volume} on the container stream,
+to allow implementations to automatically recognize volumes. This
+approach is illustrated in Figure~\ref{crypted_fif}.
 
@@
   \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
+
   \caption{Embedding an encrypted AFF4 volume within an Encrypted
-Stream. The container volume contains an encrypted stream backed by an
-image stream. Once the encrypted stream is opened, the volume stored
-on its image stream is accessible. Now it is possible to see the
-secret image stream stored within the volume.}
+  Stream. The container volume contains an encrypted stream backed by
+  an image stream which is also stored in the container. Once the
+  encrypted stream is opened, the volume stored on its image stream is
+  accessible. Now it is possible to see the secret image stream stored
+  within the volume.}
+
   \label{crypted_fif}
@@
 
-Implementations MUST specify the salt as a stream attribute named {\em
-salt} in base 64 encoding. Implementations are free to specify any API
-for passing the Pre-Shared Key, for example the PSK can be specified
-in a configuration file, typed in at the terminal, passed in using
-process environment variables, or encoded within the AFF4 filename
-itself.
+The salt is specified as a stream attribute named {\em aff4:salt} in
+base64 encoding. Implementations are free to specify any API for
+passing the Pre-Shared Key, for example the PSK can be specified in a
+configuration file, typed in at the terminal, passed in using process
+environment variables, or encoded within the AFF4 URL itself.
 
@@
 current state of the GUI, or internal data structures which may be
-required for caching. Usually, software create additional cache files
-to maintain this information. It is advantageous to be able to store
-these in the evidence file itself. The evidence file can then be
-re-opened by the software at a later stage and private data can be
-retrieved and used directly. Such private application data can be
-stored in a Private stream. There is no specification of what can be
-stored in the private stream, and applications with private streams
-should be able to store arbitrary segments in any format at all.
-
-It is recommended that applications name their private segment
-sufficiently accurately so as not to be confused with other
-applications, or different versions of the same application
-(e.g. ``pyflag/0.87/cache/'').
-
-\section{Implementing AFF4 in AFFLIB}
-Although there are numerous ZIP implementations available
-today, we have created our own implementation using a combination of
-available open source technology and our own novel contributions. We
-are in the process of integrating this code into the AFFLIB library
-and performing extensive regression testing.
-
-There are many reasons to develop our own ZIP64 implementation for
-AFF4:
-
-\begin{itemize}
-\item Most ZIP implementations do not implement the ZIP64
-  extensions. These extensions are required to support Evidence
-  Volumes larger than 2GB.
-\item Simple Zip implementations might rescan the
-Central Directory for each segment request. Since in practice there
-can be a large number of segments in a volume, it is advisable to have
-a ZIP64 implementation that is optimized to storing thousands (or even
-hundreds of thousands) of segments in an efficient data structure. 
-\item While the ZIP specification duplicates data found in the Central
-  Directory entry in each File Header (such as filename, size, CRC
-  etc), many implementations that we have examined only populate this
-  information in one of these places. In the interest of robustness,
-  we wanted to assure that data stored in both locations would be
-  populated to allow recovery of at least \emph{some} evidence that
-  might exist in damaged volumes. If central directory is lost, it is
-  possible to scan through the zip file and repair the central
-  directory from the File headers.
-
-\item Our implementation supports simultaneous access by multiple
-  readers and writers. Since our system requires all metadata to be
-  shared through the Universal Resolver, this lends itself to
-  providing Universal Locking on a per Object basis. So for example,
-  if one process wants to add a new segment into a Zip volume, they
-  can lock it via the Resolver, add the segment and unlock the volume
-  object in the resolver.
-\end{itemize}
+required for caching.
+
+Currently, software create additional cache files to maintain this
+information. It is advantageous to be able to store these in the
+evidence volume itself. The evidence volume can then be re-opened by
+the software at a later stage and private data can be retrieved and
+used directly. Such private application data can be stored in a
+Private stream. There is no specification of what can be stored in the
+private stream, and applications with private streams should be able
+to store arbitrary segments in any format at all. Segments URNs should
+be prefixed by the private stream URNs, and the stream type should be
+specified as ``private''.
+
+It is recommended that applications create attributes for the private
+stream so they can reliably recognize their own private stream, and
+identify the version of the application which created it. The AFFLIB
+API can be used to render access to private segments and private
+attributes.
 
@@
 
-\subsubsection{Rapidly converting a set of DD images}
+\subsection{Rapidly converting a set of DD images}
 Many hardware devices are available to acquire hard disks in the
@@
 
-\subsubsection{Acquisition of RAID disks}
+Similarly, each component can be compressed independently into its own
+stream. A single map stream can then be produced to combine all the
+component streams into a single logical stream. This approach can take
+advantage of multiple systems to actually do the compression as each
+component is compressed independently.
+
+\subsection{Acquisition of RAID disks}
 Often disks in a system are grouped into RAID devices, commonly RAID-5
@@
 logical view of the RAID, without the tool needing to have explicit
-support for RAID reassembling.
-
-\subsubsection{In-place encryption of AFF4 volumes}
+support for RAID reassembling. This approach enables parallel
+acquisition of RAID drives, a feature long desired to handle the vast
+quantities of data presented by RAID.
+
+\subsection{In-place encryption of AFF4 volumes}
 With the large size of a typical forensic image it is often difficult
@@
 text volume is then encrypted in place. Finally a map stream is used
-to re-map the first and second fragments into place, and a
-
-\subsubsection{Cryptographic management of evidence}
+to re-map the first and second fragments into place.
+
+\begin{figure}[tb]
+  \begin{center}
+  \mbox{\epsfxsize=0.8\columnwidth \epsffile{inplace.eps}}
+
+  \caption{The process of encrypting a volume in place. The first
+  block from the volume is encrypted and appended to the end of the
+  volume as a new segment (A). Now a zip file header can be written in
+  its place to form the remainder of the volume as segment B. The rest
+  of the volume is encrypted in-place, block by block (Step 4). Now we
+  create a Map stream which combines segment A and B into a single
+  logical stream (M). An Encrypted stream which is backed by this map
+  stream (E) is now added. Finally a Central Directory is appended to
+  the end of the new volume. We end up with a new Zip File which
+  contains our old volume stored within an encrypted stream.}
+
+  \label{inplace_crypto}
+  \end{center}
+\end{figure}
+
+\subsection{Cryptographic management of evidence}
 An AFF4 archive may hold multiple encrypted volumes, each in its own
@@
 
-\subsubsection{Forensic Application State}
+\subsection{Forensic Application State}
 Often forensic applications need to store files other than the
@@
 
-\subsubsection{Merging of evidence set}
-In this common usage case, Bob, goes on site to image one hard disk,
-while Alice independently images a second hard disk. After returning
-to the office they both realise their images belong to the same
-set. They merge their images, by updating the UUIDs on one of the
-volume set to now belong to the other volume set - the merge is almost
-instantaneous as they only need to write an extra ``properties'' file
-on the merged volumes.
+\subsection{Sharing evidence}
+When managing a large corpus of evidence it is common to require these
+to be shared across the network. Many laboratories create a
+centralized Network Attached Storage (NAS) facility to retain their
+evidence corpus. This is typically shared via some file sharing
+protocol, commonly Windows sharing (CIFS) or NFS.
+
+It is advantageous to be able to move evidence around between several
+servers and locations. For example, consider a disk image stored in a
+Volume with a given URN. Normally, when a user wishes to open the
+image, they will specify the path and filename of the volume which
+contains it. The library will then open the volume and retrieve the
+stream.
+
+The AFF4 model is different however. In AFF4 metadata is managed
+through a Universal Resolver (perhaps implemented in a database) and
+does not really need to exist within a volume. This means that in the
+above example, the user will query the resolver about the location of
+the volume. The resolver returns a URL to the current location of the
+volume. The user's library will then open a HTTP Stream instance to
+the location returned and the volume may be opened directly over the
+intra-web connection.
+
+If the volume needs to be moved, the volume's \emph{aff4:stored}
+attribute can be adjusted to the new location, and all clients will
+automatically know where to find it. Furthermore, if several copies of
+the same volume exist, the resolver may have multiple values for the
+\emph{aff4:stored} property, some locations may be faster to access than
+others. The Resolver can then return the most suitable attribute based
+on locality.
+
+Another interesting side effect of the sharing model is that it is far
+more efficient to allow remote analysts to peruse subsets of the
+evidence. For example, Alice is in the laboratory and wants Bob who is
+on a remote site to view several files she found on her
+image. Traditionally she would either need to copy the entire volume
+and send it to him. Or maybe extract the files she thinks are relevant
+and send those to him.
+
+With the new AFF4 architecture, Bob can open the volume directly,
+using the HTTP protocol, and view the files he needs. Very little
+extra data is actually transferred in excess of the data he wishes to
+see. This is a much more efficient way to allow remote analysts to
+collaborate.
 
@@
 
+This paper describes a significant enhancement to Garfinkel's Advanced
+Forensic Format (AFF1). The new version, AFF4, offers significant new
+features including the ability to store multiple kinds of evidence
+from multiple devices in a single archive, and an improved separation
+between the underlying storage mechanism and forensic software that
+makes use of evidence stored using AFF. This improved system will
+allow a single archive of evidence to be used in a plethora of
+modalities, including in a single evidence file, multiple evidence
+files stored on multiple workstations, and evidence stored in a
+relational database or object management system---all without making
+changes to forensic software.
+
+We have an initial implementation of AFF4 working in the lab and plan
+to integrate it into Garfinkel's. After integration and testing, we
+plan to release this system as freely usable open source software.
+
+\bibliographystyle{IEEEtran}
 \bibliography{IEEEabrv,paper}
@@
 
+% BLS - The evidence set as identified by a particular UUID is no longer really in play 
+% as below. Evidence can be grouped into a case by defining a case object and stating
+% what objects (or volumes) are a part of it. Or it could be grouped into someting more 
+% arbitraty depending on what you want - its all just objects and relationships.
+
+%\item \emph{An Evidence Set UUID} is a unique identifier generated
+%  according to RFC4122\cite{RFC4122} for each new Evidence
+%  Set. Multiple Evidence Volumes that make up the same Evidence Set
+%  should all have the same UUID. When an Evidence Set is opened all its volumes are
+%  loaded, their UUIDs are checked, and all of their contents are
+%  logically merged so as to be visible as if they were all found in
+% the same volume.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2009-03-14 src/include/zip.h

@@ -268,6 +268,9 @@ CLASS(Image, FileLikeObject)
      int32_t *chunk_indexes;
      char *parent_urn;
 
+     // This is the compression type
+     int compression;
+
 END_CLASS
 
 /** The map stream driver maps an existing stream using a

### 2009-03-14 src/lib/encrypt.c

@@ -3,7 +3,7 @@
 
 AFFObject Encrypted_AFFObject_Con(AFFObject self, char *urn) {
   Encrypted this = (Encrypted)self;
-  int value;
+  char *value;
 
   if(urn) {
     URNOF(self) = talloc_strdup(self, urn);

### 2009-03-14 src/lib/image.c

@@ -56,6 +56,9 @@ static AFFObject Image_Con(AFFObject self, char *uri) {
     // These are the essential properties:
     value = resolver_get_with_default(oracle, self->urn, "aff2:chunk_size", "32k");
     this->chunk_size = parse_int(value);
+
+    value = resolver_get_with_default(oracle, self->urn, "aff2:compression", "8");
+    this->compression = parse_int(value);
     
     value = resolver_get_with_default(oracle, self->urn, "aff2:chunks_in_segment", "2048");
     this->chunks_in_segment = parse_int(value);
@@ -120,7 +123,7 @@ static int dump_chunk(Image this, char *data, uint32_t length, int force) {
   int clength=2*compressBound(length);
 
   // Should we offer to store chunks uncompressed?
-  if(0) {
+  if(this->compression == 0) {
     memcpy(cbuffer, data, length);
     clength = length;
   } else {
@@ -346,13 +349,19 @@ static int partial_read(FileLikeObject self, StringIO result, int length) {
   CALL(fd, close);
   talloc_free(fd);
 
-  // Try to decompress it:
-  if(uncompress((unsigned char *)uncompressed_chunk, 
-		(unsigned long int *)&uncompressed_length, 
-		(unsigned char *)compressed_chunk,
-		(unsigned long int)compressed_length) != Z_OK ) {
-    RaiseError(ERuntimeError, "Unable to decompress chunk %d", chunk_id);
-    goto error;
+
+  if(this->compression == 8) {
+    // Try to decompress it:
+    if(uncompress((unsigned char *)uncompressed_chunk, 
+		  (unsigned long int *)&uncompressed_length, 
+		  (unsigned char *)compressed_chunk,
+		  (unsigned long int)compressed_length) != Z_OK ) {
+      RaiseError(ERuntimeError, "Unable to decompress chunk %d", chunk_id);
+      goto error;
+    };
+  } else {
+    memcpy(uncompressed_chunk, compressed_chunk, compressed_length);
+    uncompressed_length = compressed_length;
   };
   
   //  printf("%d\n", uncompressed_length);

### 2009-03-14 src/lib/unittests.c

@@ -441,6 +441,7 @@ void test_encrypted(char *filename) {
   // Create an embedded Image stream inside the encrypted stream:
   embedded_stream = (FileLikeObject)CALL(oracle, create, (AFFObject *)&__Image);
   CALL((AFFObject)embedded_stream, set_property, "aff2:stored", URNOF(container));
+  CALL((AFFObject)embedded_stream, set_property, "aff2:compression", "0");
   CALL((AFFObject)encrypted_stream, set_property, "aff2:target", URNOF(embedded_stream));
 
   if(!CALL((AFFObject)encrypted_stream, finish)) {

### 2009-03-14 paper/paper.tex
Added patches for the afflib patch.

@@
 
+\subsection{Prior Work}
 In recent years there has been a steady and growing interest in the
@@
 
+Recently the open source AFF implementation library (AFFLIB) was
+expanded to support both encryption and digital signatures of digital
+evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
+encryption and authentication is critical for forensic file formats.
+
+\subsection{This paper}
 The AFF4 metadata model described in this work, is based on (subject,
@@
 supplied parameters supplied to the tool, and the drive being imaged.
-
-Recently the open source AFF implementation library (AFFLIB) was
-expanded to support both encryption and digital signatures of digital
-evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
-encryption and authentication is critical for forensic file formats.
 

### 2009-03-14 patches/afflib-3.3.4.diff.sh

@@ -1,0 +1,25 @@
+diff -rupN afflib-3.3.4/configure.ac afflib-3.3.4_AFF2/configure.ac
+diff -rupN afflib-3.3.4/lib/afflib.cpp afflib-3.3.4_AFF2/lib/afflib.cpp
+diff -rupN afflib-3.3.4/lib/afflib_i.h afflib-3.3.4_AFF2/lib/afflib_i.h
+diff -rupN afflib-3.3.4/lib/libaff2/blob.c afflib-3.3.4_AFF2/lib/libaff2/blob.c
+diff -rupN afflib-3.3.4/lib/libaff2/class.c afflib-3.3.4_AFF2/lib/libaff2/class.c
+diff -rupN afflib-3.3.4/lib/libaff2/image.c afflib-3.3.4_AFF2/lib/libaff2/image.c
+diff -rupN afflib-3.3.4/lib/libaff2/include/class.h afflib-3.3.4_AFF2/lib/libaff2/include/class.h
+diff -rupN afflib-3.3.4/lib/libaff2/include/error.h afflib-3.3.4_AFF2/lib/libaff2/include/error.h
+diff -rupN afflib-3.3.4/lib/libaff2/include/list.h afflib-3.3.4_AFF2/lib/libaff2/include/list.h
+diff -rupN afflib-3.3.4/lib/libaff2/include/misc.h afflib-3.3.4_AFF2/lib/libaff2/include/misc.h
+diff -rupN afflib-3.3.4/lib/libaff2/include/stringio.h afflib-3.3.4_AFF2/lib/libaff2/include/stringio.h
+diff -rupN afflib-3.3.4/lib/libaff2/include/talloc.h afflib-3.3.4_AFF2/lib/libaff2/include/talloc.h
+diff -rupN afflib-3.3.4/lib/libaff2/include/zip.h afflib-3.3.4_AFF2/lib/libaff2/include/zip.h
+diff -rupN afflib-3.3.4/lib/libaff2/link.c afflib-3.3.4_AFF2/lib/libaff2/link.c
+diff -rupN afflib-3.3.4/lib/libaff2/Makefile.am afflib-3.3.4_AFF2/lib/libaff2/Makefile.am
+diff -rupN afflib-3.3.4/lib/libaff2/map.c afflib-3.3.4_AFF2/lib/libaff2/map.c
[…]

### 2009-03-15 paper/paper.tex
bls-small-changes Ignore-this: 77e80e0adaef6db3e24c147848a64636

@@
 
+An important advance of this work is the introduction of storage transformation
+functions to the forensic storage container. Prior works simply focused
+on forensically sound storage of bitstreams, leaving the necessary activities
+of translating low level storage into higher level abstractions at the 
+aggregate block (ie. RAID), volume, and filesystem layers in the domain of 
+analysis tools, as transiently constructed artifacts. AFF4, through its introduction 
+of Map Streams, enables artefacts at all these layers to be stored at zero cost within
+the forensic container, which is an important enabler for interoperable 
+forensic tools.
+
 This paper extends the previous work by adapting AFF to the multiple
@@
 bit-level specification so that AFF files can be processed using
-standard non-forensic tools.  We call the new system AFF4, and use the
+standard non-forensic tools. The  We call the new system AFF4, and use the
 phrase AFF1 to refer to the legacy system developed by
@@
 
-AFF1's flexibility came from the fact that forensic data and metadata
-is stored as arbitrary name/value pairs called \emph{segments}. For
+AFF1's flexibility came from a data model of forensic data and metadata
+stored as arbitrary name/value pairs called \emph{segments}. For
 example, the first 16MB of a disk image is stored in a segment called
@@
 identifying and referring to all evidence. We define an AFF4 specific
-URN namespace\cite{RFC1737} for our naming scheme (\emph{urn:aff4}).
-URNs can then be made unique by use of a unique identifier generated
+URN scheme, which we call the AFF4 URN. URN's of this scheme use the
+namespace\cite{RFC1737} \emph{\'aff4\'} and therefore begin with the 
+string(\emph{urn:aff4}). AFF4 URNs are then be made unique by use of 
+a unique identifier generated
 as per RFC4122\cite{RFC4122}. For example, an AFF4 URN might be
@@
 created, it is now possible to collect metadata about each object. The
-AFF4 model treats metadata as an abstract concept which may exists
+AFF4 model treats metadata as an abstract concept which may exist
 independently from the data itself. We term {\em metadata} to be a set
@@
 statement is made about. Using this system we are able to store
-arbitrary attributes about any object in the AFF4 universe.
+arbitrary attributes about any object in the AFF4 universe. Additionally,
+as these statements are universally scoped, they may be stored anywhere.
 
@@
 
-The AFF4 model is different however. In AFF4 metadata is managed
-through a Universal Resolver (perhaps implemented in a database) and
-does not really need to exist within a volume. This means that in the
+The AFF4 model is different however. In AFF4 metadata is managed in a location
+independant manner 
+through a Universal Resolver (perhaps implemented in a database). Relevant information
+is found not only in Volumes, but may additionally come from other sources.
+This means that in the
 above example, the user will query the resolver about the location of

### 2009-03-15 paper/paper.tex
Added a screenshot of windows explorer openning an AFF4 file.

@@
 \begin{abstract}
-Forensics analysis requires the acquisition and management of many
-different types of evidence material, including individual disk
-drives, RAID sets, network packets, memory images, and extracted
-files. Often the same evidence is reviewed by several different tools
-or examiners in different locations. We propose a backwards-compatible
-evolutionary redesign of the Advanced Forensic Format an open,
-extensible file format for storing and sharing of evidence, arbitrary
-case related information and analysis results among different
-tools. The new specification was designed to be simple to implement,
-allowing the use of the well supported Zip File format specifications
-for bit level file access.
+Forensic analysis requires the acquisition and management of many
+different types of evidence, including individual disk drives, RAID
+sets, network packets, memory images, and extracted files. Often the
+same evidence is reviewed by several different tools or examiners in
+different locations. We propose a backwards-compatible evolutionary
+redesign of the Advanced Forensic Format---an open, extensible file
+format for storing and sharing of evidence, arbitrary case related
+information and analysis results among different tools. The new
+specification was designed to be simple to implement, allowing the use
+of the well supported Zip File format specifications for bit level
+file access.
 \end{abstract}
@@
 evidence. Early practitioners created exact bit-for-bit copies
-(``mirror disks'') employed proprietary software systems
+(``mirror disks'') or employed proprietary software systems
 (e.g. \cite{safeback,ilook,encase}) for making and authenticating
@@
 ``seekable gzip'' format that allowed disk images to be stored in a
-form that was compressed but allowed the random-access to evidence
-data necessary for forensic analysis. The \emph{sgzip} format does not
+form that was compressed but allowed random-access to evidence data
+necessary for forensic analysis. The \emph{sgzip} format does not
 store any metadata at all.
@@
 file format also provides for a small number of predefined metadata
-fields to be stored within the file format.
+fields to be stored within the file format\cite{libewf}.
 
@@
 bit-level specification so that AFF files can be processed using
-standard non-forensic tools. The  We call the new system AFF4, and use the
+standard non-forensic tools. We call the new system AFF4, and use the
 phrase AFF1 to refer to the legacy system developed by
@@
   every header of every segment in an AFF file needs to be read when a
-  file is opened. It practice this can take 10--30 seconds the first
+  file is opened. In practice this can take 10--30 seconds the first
   time a large AFF file is opened. (Once the file is opened, the
@@
 \item AFF1's bit-level specification is essentially a simple container
-  file specification. Given the that there are other container file
+  file specification. Given that there are other container file
   specifications that are much more widely supported with both
@@
 Consider the typical usage scenario depicted in Figure \ref{usage}, of
-an file which contains a disk image. This volume is distributed to two
+a volume containing a disk image. This volume is distributed to two
 independent analysts, Alice and Bob. Alice may find and extract
@@
   \begin{center}
-  \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}
+  \mbox{\epsfxsize=0.9\columnwidth \epsffile{usage1.eps}}
   \caption{A typical usage scenario. Both Alice and Bob receive an AFF
@@
 more flexible and has better existing support, namely the Zip64 file
-format is not adopted as a default storage format. This enhanced our
+format is now adopted as a default storage format. This enhanced our
 compatibility with other tools and simplifies implementations for
@@
 written in either the old or the new format without modification. We
-are extending the AFFLIB API to support the new information mode.
+are extending the AFFLIB API to support the new information model.
 
@@
 file format. AFF Objects have a globally unique name (URN) as described in
-\cite{RFC1737}. The name is defined within the aff2 namespace, and is
+\cite{RFC1737}. The name is defined within the aff4 namespace, and is
 made unique by use of a unique identifier generated as per
@@
 
+\label{definitions}
 \item \emph{A segment} is a single unit of data written to a volume. AFF4
   segments have a \emph{segment name} provided by their URN, a
-  \emph{segment timestamp} in GMT, and the \emph{segment contents}.
+  \emph{segment timestamp} in GMT, and the \emph{segment
+contents}. Segments are suitable for storing small quantities of data,
+and still present a stream interface.
 
@@
 visibility of objects from all volumes, and therefore guides
-implementations in resolving external references. We discuss the
-Resolver in detail in Section \ref{resolver}.
+implementations in resolving external references. 
 \end{itemize}
@@
 methodologies emphasise the process of individuating evidence, and
-storing it in a manner that it is easily identifiable.
-
-Management of evidence requires on effective identififaction, with
+storing it in a manner that is easily identifiable.
+
+Management of evidence requires an effective identififaction, with
 practitioners currently employing acquisition time metadata such as
@@
 and directory naming schemes, and labeling of evidence container hard
-drives. Evidence may also be referred to by external means in an
+drives. Evidence may also be referenced by external means in an
 inconsistent way. For example, in an investigator's case note a disk
@@
 
-Such individuation schemes may be, however, problematic when
-automatically managing evidence. For example, at acquisition time a
-suitably unique individuator may not be selected. If that occured, at
-analysis time evidence container files may need to be renamed to avoid
-name collisions.
+Such individuation schemes may be problematic when automatically
+managing evidence. For example, at acquisition time a suitably unique
+individuator may not be selected. If that occured, at analysis time
+evidence container files may need to be renamed to avoid name
+collisions.
 
@@
 
-In this way we can identify uniquely any evidence without the change
+In this way we can identify uniquely any evidence without the chance
 for name collisions. Once a universal identification regime is
@@
   inspected using any number of commercial zip application
-  (e.g. Windows Explorer natively supports Zip files).
+  (e.g. Windows Explorer natively supports Zip files as can be seen in
+  Figure~\ref{explorer}).
 
@@
   \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
-  \caption{The basic structure of a Zip archive. Also can be seen how
-new archive members are added to an existing Zip File. The
-Central Directory gets overwritten by the new member, and a new
-Central Directory is written on the end.  }
+
+  \caption{The basic structure of a Zip archive. Also shows is how new
+  archive members are added to an existing Zip File. The Central
+  Directory gets overwritten by the new member, and a new Central
+  Directory is written on the end.  }
+
   \label{zip_structure}
@@
 \end{itemize}
+
+\begin{figure}[tbp]
+  \begin{center}
+  \mbox{\epsfxsize=0.8\columnwidth \epsffile{compressed_folders.eps}}
+
+  \caption{An Image stream browsed from Windows Explorer.  Basic
+  access to the evidence volume can be made using familiar tools
+  improving transparency.}
+
+  \label{explorer}
+  \end{center}
+\end{figure}
 
@@
 component streams into a single logical stream. This approach can take
-advantage of multiple systems to actually do the compression as each
-component is compressed independently.
+advantage of multiple systems to actually do the compression in
+parallel as each component is compressed independently.
 
@@
 
-\subsection{Forensic Application State}
-Often forensic applications need to store files other than the
-evidence itself - for example, they might need to store internal data,
-annotations, keyword hits etc. Currently these applications store the
-data in a proprietary external file or database. This makes it
-difficult to archive because the evidence itself may become separated
-from the case file.
-
-It is advantageous for these applications to store their state within
-the evidence file itself using a Private Stream. Then when opening the
-evidence file again, the results of their analysis will become
-available.
-
 \subsection{Sharing evidence}
@@
 collaborate.
+
+\subsection{Logical File Acquisition}
+Sometimes in a forensic investigation it is impossible to take a true
+forensic image of the underlying disk, and we must resort to acquiring
+discrete files instead. Such a case might occur for example, when
+investigating an incident on a very large network storage system or a
+critical system which can not be brought down for the duration of the
+image acquisition.
+
+It is still advantageous in this case to bring these files into the
+AFF4 evidence universe by acquiring each evidence file using a unique
+URN. As explained in \ref{definitions}, segments are AFF4 stream
+objects which are implemented by storing the data in Zip archive
+members. It follows, therefore, that a regular zip file containing
+files is also a valid AFF4 volume providing the segment names are
+globally unique.
+
+So a logical image of files, can be created by any regular Zip
+compression program in the field. Once brought into the lab these
+volumes can be given a volume URN and imported into the Universal
+Resolver to provide access to all the files within the archive.
 

### 2009-04-03 src/lib/misc.c
Updated imager.

@@ -61,7 +61,7 @@ char *escape_filename(char *filename) {
   
   for(i=0;i<length;i++) {
     char x=filename[i];
-    if(illegal_filename_lut[x] || x>128) {
+    if(x<0 || illegal_filename_lut[x]) {
       sprintf(buffer+j, "%%%02X", x);
       j+=3;
       if(j>BUFF_SIZE-10) break;

### 2009-04-03 src/lib/resolver.c

@@ -11,12 +11,15 @@ struct dispatch_t {
 
 static struct dispatch_t dispatch[] = {
   { "blob", (AFFObject)&__Blob },
+
   { "volume", (AFFObject)&__ZipFile },
+  { "directory", (AFFObject)&__DirVolume },
   { "link", (AFFObject)&__Link },
   { "image", (AFFObject)&__Image },
   { "map", (AFFObject)&__MapDriver},
   { "encrypted", (AFFObject)&__Encrypted},
   { "file://", (AFFObject)&__FileBackedObject },
+
   // All handled by libcurl
   { "http://", (AFFObject)&__HTTPObject },
   { "https://", (AFFObject)&__HTTPObject },

### 2009-04-03 src/lib/zip.c

@@ -705,6 +705,9 @@ static void ZipFile_close(ZipFile self) {
 
     // Make sure the lock is removed from this volume now:
     CALL(oracle, del, URNOF(self), "aff2volatile:write_lock");
+
+    // Close the fd
+    CALL(fd, close);
   };
 };
 

### 2009-04-03 src/tools/aff2imager.c

@@ -9,7 +9,9 @@ void help() {
   printf("aff2imager tool\n\n"
 	 "--open, -o \tOpen this file to pre-populate the resolver \n"
 	 "           \t(Can be specified multiple times).\n"
-	 "--info, -i \tList information about all known objects.\n"
+	 "--image, -i \tImaging mode\n"
+	 "--extract, -e stream_name\tExtract stream to output (set using --output or stdout)\n"
+	 "--passphrase, -p password\tUse this passphrase to encrypt/decrypt the image\n"
 	 );
 };
 
@@ -38,7 +40,7 @@ void aff2_open_volume(char *urn) {
   CALL(oracle, cache_return, (AFFObject)volume);
 };
 
-void aff2_print_info() {
+void aff2_print_info(int verbose) {
   Cache i,j;
 
   list_for_each_entry(i, &oracle->urn->cache_list, cache_list) {
@@ -47,9 +49,146 @@ void aff2_print_info() {
 
     printf("\n******** Object %s ***********\n",urn);
     list_for_each_entry(j, &attributes->cache_list, cache_list) {
-      printf("\t%s = %s\n", (char *)j->key, (char *)j->data);
+      char *key = (char *)j->key;
+      char *value = (char *)j->data;
+
+      // Ignore volatile data in default verbosity.
+      if(verbose==0 && !memcmp(key, ZSTRING_NO_NULL("aff2volatile")))
+	continue;
+
+      printf("\t%s = %s\n", key, value);
     };
   };
+};
+
+/** When we want to use encryption we must create a container volume,
+    A number of encrypted streams are created within the container
+    volume and a new AFF2 volume is created within the encrypted
+    volume. The following are the major components:
+
+    Container aff2:stored file://filename.zip
+    Image URN aff2:stored in Container
+    Encrypted URN aff2:target Image
+    Encrypted Volume aff2:stored in Encrypted URN
+    Encrypted Image aff2:stored in Encrypted Volume
+
+    The source is copied onto Encrypted Image
+*/
+int aff2_encrypted_image(char *driver, char *output_file, char *stream_name, 
+			 char *chunks_in_segment,
+			 char *append,
+			 char *source) {
+  // This is the container volume
+  ZipFile zipfile, volume;
[…]

### 2009-04-03 paper/paper.tex
Added write support via http (uses PUT). Its now possible to just image directly to a http URL.

@@
 % Make sure we know its a draft for now
-\usepackage{draftwatermark}
-\SetWatermarkFontSize{3cm}
+%\usepackage{draftwatermark}
+%\SetWatermarkFontSize{3cm}
 \usepackage{fancyvrb}
@@
 Many hardware devices are available to acquire hard disks in the
-fields. These often produce a set of uncompressed images split at a
+field. These often produce a set of uncompressed images split at a
 certain size. It is possible to construct a Map Stream which
@@
 different X509 certificates. Its is also possible for users to create
-non-encrypted volumes within the FIF file.
+non-encrypted volumes within the AFF4 volume.
 

### 2009-04-03 src/include/zip.h

@@ -155,6 +155,11 @@ CLASS(HTTPObject, FileLikeObject)
      CURL *curl;
      StringIO buffer;
 
+     CURL *send_handle;
+     StringIO send_buffer;
+
+     CURLM *multi_handle;
+
      HTTPObject METHOD(HTTPObject, Con, char *url);
 END_CLASS
 
@@ -487,8 +492,9 @@ CLASS(ZipFileStream, FileLikeObject)
 			  char mode);
 END_CLASS
 
-// This is the main FIF class - it manages the Zip archive
-CLASS(FIFFile, ZipFile)
+// A directory volume implementation - all elements live in a single
+// directory structure
+CLASS(DirVolume, ZipFile)
 END_CLASS
 
 void dump_stream_properties(FileLikeObject self, char *volume);

### 2009-04-03 src/lib/dirvol.c

@@ -1,0 +1,7 @@
+/** This is an implementation of the directory volume - all blobs are
+    stored as files here.
+*/
+#include "zip.h"
+
+VIRTUAL(DirVolume, ZipFile)
+END_VIRTUAL

### 2009-04-03 src/lib/http.c

@@ -7,18 +7,24 @@ We basically use libcurl to do the heavy lifting so we might actually
 get to use ftp://, http://, https:// and more all for free.
 
 ******************************************************************/
-static int close_fd(void *self) {
-  HTTPObject this = (HTTPObject)self;
-  curl_easy_cleanup(this->curl);
-
-  return 0;
-};
-
 static size_t write_callback(void *ptr, size_t size, size_t nmemb, void *self){ 
   HTTPObject this = (HTTPObject)self;
 
   CALL(this->buffer, write, ptr, size*nmemb);
   return size*nmemb;
+};
+
+static size_t read_callback(void *ptr, size_t size, size_t nmemb, void *self){ 
+  HTTPObject this = (HTTPObject)self;
+
+  int len;
+  CALL(this->send_buffer, seek, 0, 0);
+
+  len = CALL(this->send_buffer, read, ptr, size*nmemb);
+  CALL(this->send_buffer, skip, len);
+  CALL(this->send_buffer, seek, 0, 2);
+  //  if(len==0) return CURL_READFUNC_PAUSE;
+  return len;
 };
 
 // We want to get this header, Content-Range: bytes 0-100/3388
@@ -55,6 +61,7 @@ static HTTPObject HTTPObject_Con(HTTPObject self,
 				 char *url) {
   // Parse out the url
   CURL *handle = self->curl = curl_easy_init();
+
   self->buffer = CONSTRUCT(StringIO, StringIO, Con, self);
   URNOF(self) = talloc_strdup(self, url);
 
@@ -65,11 +72,6 @@ static HTTPObject HTTPObject_Con(HTTPObject self,
   curl_easy_setopt( handle, CURLOPT_WRITEDATA, self );
   curl_easy_setopt( handle, CURLOPT_HEADERFUNCTION, parse_header_callback);
   curl_easy_setopt( handle, CURLOPT_HEADERDATA, self);
-
-  /*
-  curl_easy_setopt( handle, CURLOPT_SEEKFUNCTION, noop_seek_callback );
-  curl_easy_setopt( handle, CURLOPT_SEEKDATA, NULL );  
-  */
 
   if(!self->curl) {
     RaiseError(ERuntimeError, "Unable to initialise curl");
@@ -109,14 +111,69 @@ static int HTTPObject_read(FileLikeObject self, char *buffer,
 
 static int HTTPObject_write(FileLikeObject self, char *buffer,
 			    unsigned long int length) {
[…]

### 2009-04-03 src/lib/resolver.c

@@ -44,6 +44,7 @@ void AFF2_Init(void) {
   Link_init();
   HTTPObject_init();
   Encrypted_init();
+  DirVolume_init();
 
   init_luts();
 

### 2009-04-03 src/lib/zip.c

@@ -248,8 +248,7 @@ static AFFObject ZipFile_finish(AFFObject self) {
        "aff2:type",            /* Attributte */
        "volume");              /* Value */
     
-  CALL((ZipFile)this, Con, file_urn);
-  return self;
+  return (AFFObject)CALL((ZipFile)this, Con, file_urn);
 };
 
 

### 2009-04-03 src/tools/aff2imager.c

@@ -196,7 +196,7 @@ int aff2_image(char *driver, char *output_file, char *stream_name,
 	       char *append,
 	       char *source) {
   ZipFile zipfile = (ZipFile)CALL(oracle, create, (AFFObject *)&__ZipFile);
-  char *output = talloc_strdup(zipfile, output_file);
+  char *output;
   Image image;
   char buffer[BUFF_SIZE];
   int in_fd;
@@ -207,13 +207,18 @@ int aff2_image(char *driver, char *output_file, char *stream_name,
 
   if(!zipfile) goto error;
 
+  output = talloc_strdup(zipfile, output_file);
   in_fd=open(source,O_RDONLY);
   if(in_fd<0) {
     RaiseError(ERuntimeError, "Could not open source: %s", strerror(errno));
     goto error;
   };
 
-  if(strstr(output_file, "file://") != output_file) {
+  // We assume that if the name does not contain a : its a filename
+  // and append a file:// reference to it. Of course we expect a fully
+  // qualified URI (e.g. http://foobar/), but this is a user friendly
+  // feature.
+  if(!strstr(output_file, ":")) {
     output = talloc_asprintf(zipfile, "file://%s", output_file);
   };
 
@@ -227,8 +232,9 @@ int aff2_image(char *driver, char *output_file, char *stream_name,
   };
 
   // Is it ok?
-  if(!CALL((AFFObject)zipfile, finish))
+  if(!CALL((AFFObject)zipfile, finish)) {
     goto error;
+  };
 
   // Done with that now
   CALL(oracle, cache_return, (AFFObject)zipfile);
@@ -301,8 +307,6 @@ int aff2_image(char *driver, char *output_file, char *stream_name,
   return 0;
 
  error:
-  if(zipfile)
-    talloc_free(zipfile);
   PrintError();
   return -1;
 };

### 2009-04-29 paper/paper.tex
Small bugfix with zip file time. Started to work on reviewer comments.

@@
 
-Currently the HTTP Stream is read only, but in future we anticipate
-write support to be implemented, for example, using the WebDav
-extensions to HTTP\cite{webdav-rfc}. The HTTP stream also supports the
-File Transfer Protocol (FTP) and HTTPS (Secure Sockets Layer---SSL)
-protocols transparently. This support is provided automatically by the
-curl library.
+Our implementation also enables direct writing to a HTTP URL using the
+WebDav extensions to HTTP\cite{webdav-rfc}. The HTTP stream also
+supports the File Transfer Protocol (FTP) and HTTPS (Secure Sockets
+Layer---SSL) protocols transparently. This support is provided
+automatically by the curl library.
 
@@
 In this section we describe how AFF4 may be used in various
-situations. Many of those can be solved using other forensic file
-formats, but often in a more awkward way.
+situations. Since the AFF4 framework implements a distributed evidence
+management system, we demonstrate its use by a fictitious
+multinational corporation with offices in Los Angeles and New
+York. Each office has its own computer forensics lab and is connected
+via a WAN.
+
+\subsection{Using distributed evidence}
+An investigation is conducted by the New York team. The case relates
+to a hard disk image stream stored inside a volume, in turn stored on
+the NY evidence server at URL http://ny.wan/evidence1.aff4 . The team
+requires an analyst (Bob) in LA to assist with their analysis. The LA
+analyst types\footnote{fls is the file listing command which is part
+of the Sleuthkit}:
+
+\begin{lstlisting}
+fls -i aff4 "NY case 1"
+\end{lstlisting}
+This command causes the local AFF4 implementation to:
+\begin{enumerate}
+\item Contact the universal resolver asking where \emph{``NY case 1''} is
+stored.
+\item The universal resolver replies that it is a symbolic link to a
+stream called \emph{``urn:aff4:1234''} stored within the volume
+\emph{``urn:aff4:9876''}. Further queries reveal that the volume is
+located at \emph{http://ny.wan/evidence1.aff4}.
+\item The local AFF4 library then directly accesses the volume at the
+given URL. Note that the entire volume is not copied, instead specific
+chunks are retrieved on an as needed basis.
+\end{enumerate}
+
+The overall effect is that the user in LA is able to directly access
+the disk image specified using a friendly name, and stored at a remote
+location easily.
+
+\subsection{Load redistribution}
+In the previous scenario, Bob becomes involved in this case, and
+wishes to download the entire image locally to
+\emph{http://la.wan/evidence1.aff4}. The Universal Resolver now has
+two possible locations for the same volume URN, since there are two
+copies in existence. Based on pre-determined distance metrics, the
+resolver directs requests from Bob to the LA copy, while Alice is
+redirected to the NY copy. This load redistribution can be used for
+optimal management of evidence storage in a transparent way. Analyst
+are not aware of where the evidence is physically stored, and it
+appears as though all evidence is always available.
+
+\subsection{Remote imaging}
+The NY IT security team has just responded to an incident on one of
+their servers. Alice, the responding officer, wishes to image the
+server. She types:
+\begin{lstlisting}
+aff4imager -i -o http://ny.wan/evidence2.aff4 \
+	-k http://ny.wan/alice.key \
+	-c http://ny.wan/alice.crt /dev/sda
+\end{lstlisting}
+
+This command requests an image be created directly on the evidence
+server (it will be uploaded using webdav). The image is signed using
+Alice's certificate and key (which might need to be unlocked). Note
+that Alice does not need any hardware to obtain the image as it is
+done over the network - she therefore can respond rapidly.
+
+Bob is an analyst in LA which specializes in filesystem analysis. As
+soon as the acquisition is complete, the image is available for Bob to
+examine. Bob does not have permissions to create volumes on the NY
+evidence server, so he types:
+\begin{lstlisting}
+fsbuilder -o http://la.wan/evidence3.aff4 \
+	http://ny.wan/evidence2.aff4
+\end{lstlisting}
+This creates a new volume on the LA server which contains a set of Map
+streams referring to the original evidence. The new volume is near
+zero cost but refers to the original image (which is still stored in
+NY).
 
@@
 \subsection{Logical File Acquisition}
-Sometimes in a forensic investigation it is impossible to take a true
-forensic image of the underlying disk, and we must resort to acquiring
-discrete files instead. Such a case might occur for example, when
-investigating an incident on a very large network storage system or a
-critical system which can not be brought down for the duration of the
-image acquisition.
-
-It is still advantageous in this case to bring these files into the
+Alice is responding to an incident on a critical corporate
+server. Since the system can not be taken down for forensic imaging,
+Alice must resort to acquiring discrete files instead. Alice is unable
+to install and run any acquisition software on the server due to
+policy restrictions.
+
+It is still advantageous in this case to bring discrete files into the
 AFF4 evidence universe by acquiring each evidence file using a unique
@@
 members. It follows, therefore, that a regular zip file containing
-files is also a valid AFF4 volume providing the segment names are
-globally unique.
+files is also a valid AFF4 volume.
 
@@
 compression program in the field. Once brought into the lab these
-volumes can be given a volume URN and imported into the Universal
-Resolver to provide access to all the files within the archive.
+volumes are given a volume URN and imported into the Universal
+Resolver to provide access to all the files within the archive. At
+this stage digital signatures can also be added for each logical file.
+
+Alice uses windows explorer to obtain a Zip file of the files of
+interest. After taking the file back to the lab, she then signs the
+files, and adds a volume URN, making the Zip file a fully compliant
+AFF4 volume.
 
@@
 
-This paper describes a significant enhancement to Garfinkel's Advanced
-Forensic Format (AFF1). The new version, AFF4, offers significant new
-features including the ability to store multiple kinds of evidence
+This paper describes a significant enhancement to the Advanced
+Forensic Format (AFF1). AFF4, extends beyond a file format to describe
+a universal framework for evidence management, offering significant
+new features such as the ability to store multiple kinds of evidence
 from multiple devices in a single archive, and an improved separation
 between the underlying storage mechanism and forensic software that
-makes use of evidence stored using AFF. This improved system will
-allow a single archive of evidence to be used in a plethora of
-modalities, including in a single evidence file, multiple evidence
-files stored on multiple workstations, and evidence stored in a
-relational database or object management system---all without making
-changes to forensic software.
-
-We have an initial implementation of AFF4 working in the lab and plan
-to integrate it into Garfinkel's. After integration and testing, we
-plan to release this system as freely usable open source software.
+makes use of evidence stored using AFF. This improved system allows a
+single archive of evidence to be used in a plethora of modalities,
+including in a single evidence file, multiple evidence files stored on
+multiple workstations, and evidence stored in a relational database or
+object management system---all without making changes to forensic
+software.
+
+We have developed an open source reference implementation, but the
+AFF4 framework is simple enough for competing implementations to be
+implemented commercially. We hope this simplicity enhances AFF4's
+acceptance and adoption as a standard evidence management platform.
 

### 2009-04-29 lib/zip.c

@@ -682,6 +682,8 @@ static FileLikeObject ZipFile_open_member(ZipFile self, char *filename, char mod
 					       URNOF(self), 
 					       AFF4_DIRECTORY_OFFSET));
     char *escaped_filename;
+    time_t epoch_time = time(NULL);
+    struct tm *now = localtime(&epoch_time);
 
     // Make sure the filename we write on the archive is relative, and
     // the filename we use with the resolver is fully qualified
@@ -721,6 +723,10 @@ static FileLikeObject ZipFile_open_member(ZipFile self, char *filename, char mod
     header.compression_method = compression;
     header.file_name_length = strlen(escaped_filename);
     header.extra_field_len = extra_field_len;
+    header.lastmoddate = (now->tm_year + 1900 - 1980) << 9 | 
+      (now->tm_mon + 1) << 5 | now->tm_mday;
+    header.lastmodtime = now->tm_hour << 11 | now->tm_min << 5 | 
+      now->tm_sec / 2;
 
     CALL(fd, write,(char *)&header, sizeof(header));
     CALL(fd, write, ZSTRING_NO_NULL(escaped_filename));

### 2009-04-29 tools/fsbuilder.c

@@ -7,7 +7,7 @@
 #include "config.h"
 #include "common.h"
 #include <tsk3/libtsk.h>
-#include <aff4/aff4.h>
+#include "aff4.h"
 #include "talloc.h"
 #include "zip.h"
 

### 2009-05-03 paper/paper.tex
Updated paper.

@@
   Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow}
-\author{Blinded for review}
-%\author{M. I. Cohen\footnote{scudette@gmail.com}, Simson Garfinkel, and Bradley Schatz}
+\author{M. I. Cohen\footnote{scudette@gmail.com}, Simson Garfinkel, and Bradley Schatz}
 %\author{M. I. Cohen}
@@
 same evidence is reviewed by several different tools or examiners in
-different locations. We propose a backwards-compatible evolutionary
-redesign of the Advanced Forensic Format---an open, extensible file
-format for storing and sharing of evidence, arbitrary case related
-information and analysis results among different tools. The new
-specification was designed to be simple to implement, allowing the use
-of the well supported Zip File format specifications for bit level
-file access.
+different locations. We propose a backwards-compatible redesign of the
+Advanced Forensic Format---an open, extensible file format for storing
+and sharing of evidence, arbitrary case related information and
+analysis results among different tools. The new specification, termed
+AFF4, was designed to be simple to implement, allowing the use of the
+well supported Zip File format specifications for bit level file
+access.
 \end{abstract}
@@
 \section{Introduction}
-Unlike other kinds of evidence, digital evidence can be instantly lost
-or corrupted if technical measures are not immediately taken to
-preserve its existence and integrity. As a result, preserving digital
-evidence is an important part of most digital
-investigations\cite{carrier:event-based}.
+Storing and managing digital evidence is becoming increasingly more
+difficult, as the volume and size of digital evidence increases. The
+type of what consitutes evidence has also evolved to include data
+other than disk images, such as memory images, network images and
+regular files. Preserving digital evidence is an important part of
+most digital investigations\cite{carrier:event-based}, and managing
+the evidence in a distributed organization is now emerging as a
+critical requirement.
+
+This paper presents a framework for managing and storing digital
+evidence. We first examine existing evidence management file formats
+and outline their strengths and limitations. We then explain how the
+proposed Advanced Forensics Format (AFF4) framework extends these
+efforts into a universal evidence management system. The detailed
+description of the AFF4 proposal is then followed by concerete real
+world use cases.
 
@@
 In recent years there has been a steady and growing interest in the
-actual file formats and containers used to store this digital
+actual file formats and containers used to store digital
 evidence. Early practitioners created exact bit-for-bit copies
-(``mirror disks'') or employed proprietary software systems
-(e.g. \cite{safeback,ilook,encase}) for making and authenticating
-``images'' of digital evidence. PyFlag\cite{pyflag} introduced a
+(commonly referred to as ``dd images''). More recently, proprietary
+software systems for making and authenticating ``images'' of digital
+evidence are commonly used
+(e.g. \cite{safeback,ilook,encase}). PyFlag\cite{pyflag} introduced a
 ``seekable gzip'' format that allowed disk images to be stored in a
@@
 forensic file format that allowed both data and arbitrary metadata to
-be stored in a single digital archive\cite{garfinkel:aff}. The EWF
-file format also provides for a small number of predefined metadata
-fields to be stored within the file format\cite{libewf}.
-
-Both of these approaches are limited in their expressiveness, by
-nature of their (property, value) data model. The subject of the
-properties is implicitly the single image contained within the
-evidence container. Clearly an implicit subject metadata model is
-inadequate for storing and managing multiple evidence sources
-simultaneously.
-
-The Sealed Digital Evidence Bags architecture of Schatz proposed an
-approach to digital evidence containers that facilitated composition
-of evidence and arbitrary evidence related information, through a
-simple data model and globally unique referencing
-scheme\cite{schatz:sdeb}.
-
-Recently the open source AFF implementation library (AFFLIB) was
-expanded to support both encryption and digital signatures of digital
-evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
-encryption and authentication is critical for forensic file formats.
+be stored in a single digital archive\cite{garfinkel:aff}.
+
+The Eye Witness Forensic (EWF) file format was originally developed
+for Encase\cite{encase}, but then adopted by other vendors
+\cite{libewf}. The EWF file format also compresses the image into 32kb
+chunks which are stored back to back in groupings inside the file. The
+format contains multiple tables containing relative indexes to these
+compressed chunks which are also stored within the volume. EWF volumes
+have a maximum size of 2Gb and therefore usually split an image across
+many volumes. EWF provides for a small number of predefined metadata
+fields to be stored within the file format.
+
+Both the AFF and EWF file formats are designed to store a single
+image, and any metadata implicitely refers to that image. Therefore,
+these formats in themselves have a limited metadata model which is
+unable to exist outside the context of the image itself.
+
+%Both Of These Approaches Are Limited In Their Expressiveness, By
+%Nature Of Their (Property, Value) Data Model. The Subject Of The
+%Properties Is Implicitly The Single Image Contained Within The
+%Evidence Container. Clearly An Implicit Subject Metadata Model Is
+%Inadequate For Storing And Managing Multiple Evidence Sources
+%Simultaneously.
+
+Schatz proposed a Sealed Digital Evidence Bags architecture,
+facilitating composition of evidence and arbitrary evidence related
+information, through a simple data model and globally unique
+referencing scheme\cite{schatz:sdeb}.
+
+%Recently the open source AFF implementation library (AFFLIB) was
+%expanded to support both encryption and digital signatures of digital
+%evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
+%encryption and authentication is critical for forensic file formats.
 
 \subsection{This paper}
-The AFF4 metadata model described in this work, is based on (subject,
-property, value) statements. An explicitly defined subject (identified
-by an AFF4 URN) allows one to not only associate traditional EWF style
-metadata with a contained image, but to also define new information
-entities representing both real world and virtual entities related to
-the matter. For example, an AFF4 compliant hard disk acquisition tool
-should, in addition to storing an image, store entities representing
-and describing the tool itself, the operator of the tool, the user
-supplied parameters supplied to the tool, and the drive being imaged.
-
-An important advance of this work is the introduction of storage transformation
-functions to the forensic storage container. Prior works simply focused
-on forensically sound storage of bitstreams, leaving the necessary activities
-of translating low level storage into higher level abstractions at the 
-aggregate block (ie. RAID), volume, and filesystem layers in the domain of 
-analysis tools, as transiently constructed artifacts. AFF4, through its introduction 
-of Map Streams, enables artefacts at all these layers to be stored at zero cost within
-the forensic container, which is an important enabler for interoperable 
-forensic tools.
-
-This paper extends the previous work by adapting AFF to the multiple
+The AFF4 metadata model described in this work, goes beyond merely
+specifying metadata as an addition to the forensic evidence
+itself. Metadata forms an essential part of the file format, with the
+same metadata system storing both essential information required to
+read the image itself, as well as non-essential metadata. For example,
+an AFF4 compliant hard disk acquisition tool should, in addition to
+storing an image, store entities representing and describing the tool
+itself, the operator of the tool, the user supplied parameters
+supplied to the tool, and the drive being imaged.
+
+An important advance of this work is the introduction of storage
+transformation functions to the forensic storage container. Prior
+works simply focused on forensically sound storage of bitstreams,
+leaving the necessary activities of translating low level storage into
+higher level abstractions at the aggregate block (ie. RAID), volume,
+and filesystem layers in the domain of analysis tools, as transiently
+constructed artifacts. In contrast AFF4 has mechanisms for describing
+transformation in a flexible and concise way, allowing users to view
+multiple transformations of the same data with virtually zero storage
+cost. This mechanism is an important enabler for interoperable
+forensic tools. For example, carved files may be described in terms of
+their block allocation sequences from an image, rather than requiring
+the carved file to be copied again.
+
+This paper extends previous work on the Advanced Forensic Format (AFF)
+by taking many of the concepts developed and designing a new
+specification and toolset. The AFF4 format is a complete redesign of
+the architecture. The new architecture is capable of storing multiple
 heterogeneous data types that might arise in a modern digital
@@
 data types (including network packets and memory images), extracted
-logical evidence, and forensic workflow. We do this in a manner that
-is at once upwards compatible with AFF while adopting the AFF
-bit-level specification so that AFF files can be processed using
-standard non-forensic tools. We call the new system AFF4, and use the
-phrase AFF1 to refer to the legacy system developed by
-Garfinkel.\footnote{Although Garfinkel never changed the AFF bit-level
-specification, Garfinkel released AFFLIB implementations with major
-version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid
-confusion.}
-
-\begin{figure}
-\fbox{
-\begin{minipage}{.95\linewidth}
-\begin{itemize}
-\item The ability to store evidence from one or more physical
-  devices.
-\item Schema allowing evidence files to contain memory
-  images, network packet intercepts, Netflow data, and extracted
-  files.
-
-\item Direct encoding of arbitrary mapping transformations of one or
-more sources to allow multiple logical views. This allows 
-RAID volumes to be automatically reassembled from evidence files made
-of the physical RAID devices. Maps can be created using Sleuthkit and
-stored in the evidence file, thereafter allowing arbitrary programs to directly
-read individual files stored in a disk image.
-
-\item A revised compression system that allows individual chunks
-  within a 16MB segment to be decompressed without decompressing the
-  entire segment.
-
-\item Support for Global Distributed Evidence Management.
-\end{itemize}
-\end{minipage}}
-\caption{Specific design goals for the next generation forensic file
-  format.}
-\end{figure}
+logical evidence, and forensic workflow. The AFF4 format extends to
+become a Global Distributed Evidence Management system, while
+maintaing compatibility with the old AFF file format.
+
+We call the new system AFF4, and use the phrase AFF1 to refer to the
+legacy system developed by Garfinkel et al.\footnote{Although Garfinkel
+never changed the AFF bit-level specification, Garfinkel released
+AFFLIB implementations with major version numbers 1, 2 and 3. We
+therefore call our system AFF4 to avoid confusion.}
+
+%\begin{figure}
+%\fbox{
+%\begin{minipage}{.95\linewidth}
+%\begin{itemize}
+%\item The ability to store evidence from one or more physical
+%  devices.
+%\item Schema allowing evidence files to contain memory
+%  images, network packet intercepts, Netflow data, and extracted
+%  files.
+
+%\item Direct encoding of arbitrary mapping transformations of one or
+%more sources to allow multiple logical views. This allows 
+%RAID volumes to be automatically reassembled from evidence files made
+%of the physical RAID devices. Maps can be created using Sleuthkit and
+%stored in the evidence file, thereafter allowing arbitrary programs to directly
+%read individual files stored in a disk image.
+
+%\item A revised compression system that allows individual chunks
+%  within a 16MB segment to be decompressed without decompressing the
+%  entire segment.
+
+%\item Support for Global Distributed Evidence Management.
+%\end{itemize}
+%\end{minipage}}
+%\caption{Specific design goals for the next generation forensic file
+%  format.}
+%\end{figure}
+
 \section{The Need for an Improved Forensic Format}
@@
 \texttt{page0}, the second 16MB in a segment called \texttt{page1},
-\emph{etc.} Because of this flexibility, it was relatively easy for
-Garfinkel to extend AFF1 to support encryption, digital signatures, and
-the storage of new kinds of metadata such as chain-of-custody
+\emph{etc.} Because of this flexibility, it was relatively easy 
+to extend AFF1 to support encryption, digital signatures, and the
+storage of new kinds of metadata such as chain-of-custody
 information\cite{garfinkel:affcrypto}.
@@
 \subsection{AFF Limitations}
-After imaging literally thousands of devices using Garfinkel's AFF1
-tools and adding support for AFF1 to numerous open source programs, we
-observed numerous problems in the underlying standard
-and Garfinkel's AFFLIB implementation:
+We observed a number of practical problems in the underlying AFF1
+standard and Garfinkel's AFFLIB implementation:
 
@@
   network packets.
+
+\item AFF1 (and the EWF format) are unable to recover from loss or
+  corruption. Due to the specialised nature of these formats,
+  specialised recovery software must be written to recover from
+  corruptions.
 
@@
   every header of every segment in an AFF file needs to be read when a
-  file is opened. In practice this can take 10--30 seconds the first
-  time a large AFF file is opened. (Once the file is opened, the
-  sectors corresponding to the segment headers are stored in the
-  computer's disk cache.)
+  file is opened. In practice this can take up to 10--30 seconds the
+  first time a large AFF file is opened.
+
+%(Once the file is opened, the
+%  sectors corresponding to the segment headers are stored in the
+%  computer's disk cache.)
 
@@
 \section{Global Distributed Evidence Management}
-While Garfinkel designed AFF for use on a single machine that could
-both image evidence and perform analysis, the main use of AFF in
-practice has been in distributed environments in which imaging and
-analysis takes place in multiple locations and is performed by
-multiple individuals. This is a state of affairs which mirrors common
-practice in the field.
-
-Global distributed evidence management requires
-more than simply tracking the movement of disk images: it requires
-approaches for sharing evidence to multiple disconnected
-evidence, allowing offline work, and then seamlessly recombining the
-work products of the analysts in a third security domain.
+While AFF1 was designed for use on a single machine that could both
+image evidence and perform analysis, many modern practitioners work in
+distributed environments in which imaging and analysis takes place in
+multiple locations and is performed by multiple individuals.
+
+%This is a
+%state of affairs which mirrors common practice in the field.
+
+Global distributed evidence management requires more than simply
+tracking the movement of disk images: it requires approaches for
+sharing evidence to multiple disconnected evidence, allowing offline
+work, and then seamlessly recombining the work products of the
+analysts in a third security domain.
 
 Managing evidence in a globally distributed system requires the use of
-globally unique identifiers. AFF1 assigns
-each piece of evidence a unique 128-bit identifier called a GID but did
-not make it clear when this identifier should be changed and when it
-should remain the same.
+globally unique identifiers to ensure no name collisions can occur
+with disconnected systems. AFF1 assigns each piece of evidence a
+unique 128-bit identifier called a GID but did not make it clear when
+this identifier should be changed and when it should remain the same.
 
@@
 work on a shared file that is located on a server, in other
-environments there will not be sufficient bandwidth or
-connectivity. Instead, each analyst will be required to store the
-information in their own evidence file; these files will then be
-recombined at a later point in time.
+environments there will not be sufficient connectivity. Instead, each
+analyst will be required to store the information in their own
+evidence file; these files will then be recombined at a later point in
+time.
 
@@
 
-We have also adopted AFFLIB to implement the new schema and file
-format, which will allow existing AFF-aware programs to access images
-written in either the old or the new format without modification. We
-are extending the AFFLIB API to support the new information model.
-
-We call the result of our work effort AFF4.
+We have also released a library and toolset, \emph{LIBAFF4}
+implementing the new schema and file format. This new library provides
+legacy support for existing AFF1 volumes.
 
@@
 
-This section discusses the AFF4 terminology. The AFF4 design is object
-oriented, in that a few generic objects are presented with externaly
-accessible behaviour. We discuss a number of implementations of these
-high level concepts and show how these can be put together in common
-usage cases.
+This section discusses the AFF4 terminology and architecture. The AFF4
+design is object oriented, in that a few generic objects are presented
+with externaly accessible behaviour. We discuss a number of
+implementations of these high level concepts and show how these can be
+put together in common usage cases.
 
@@
 \label{definitions}
+\item \emph{A stream} is an AFF Object which provides the ability to
+  seek and read random data. Stream objects implement abstracted
+  storage, but must provide clients the stream like interface. For
+  example, we discuss the {\em Image stream} used to store large
+  images, the {\em Map stream} used to create transformations and the
+  {\em Encrypted stream} used to provide encryption.
+
 \item \emph{A segment} is a single unit of data written to a volume. AFF4
@@
   \emph{segment timestamp} in GMT, and the \emph{segment
-contents}. Segments are suitable for storing small quantities of data,
-and still present a stream interface.
-
-\item \emph{A stream} is an AFF Object which provides the ability to
-seek and read random data. Stream objects implement abstracted
-storage, but must provide clients the stream like interface. For
-example, we discuss the {\em Image stream} used to store large images,
-the {\em Map stream} used to create transformations and the {\em
-Encrypted stream} used to provide encryption.
+  contents}. Segments are suitable for storing small quantities of
+  data, and still present a stream interface.
 
@@
 \label{resolver}
-One of the fundamental tasks of forensics is in ensuring the
-provenance of any particular piece of digital evidence, for example by
-employing a chain of custody process. To assist in this, acquisition
-methodologies emphasise the process of individuating evidence, and
-storing it in a manner that is easily identifiable.
+%One of the fundamental tasks of forensics is in ensuring the
+%provenance of any particular piece of digital evidence, for example by
+%employing a chain of custody process. To assist in this, acquisition
+%methodologies emphasise the process of individuating evidence, and
+%storing it in a manner that is easily identifiable.
 
@@
 
-In this way we can identify uniquely any evidence without the chance
-for name collisions. Once a universal identification regime is
-created, it is now possible to collect metadata about each object. The
-AFF4 model treats metadata as an abstract concept which may exist
+%In this way we can identify uniquely any evidence without the chance
+%for name collisions. Once a universal identification regime is
+%created, it is now possible to collect metadata about each object.
+
+The AFF4 model treats metadata as an abstract concept which may exist
 independently from the data itself. We term {\em metadata} to be a set
@@
 statement is made about. Using this system we are able to store
-arbitrary attributes about any object in the AFF4 universe. Additionally,
-as these statements are universally scoped, they may be stored anywhere.
+arbitrary attributes about any object in the AFF4
+universe. Additionally, as these statements are universally scoped,
+they may be stored anywhere.
 
@@
 example, if a volume is opened which contains a Map Stream targeting a
-stream stored in a different volume, its not generally possible to
+stream stored in a different volume, it is not generally possible to
 tell in which volume the stream exists, and where that volume is
@@
 Universal Resolver contains all the metadata about the AFF4 universe,
-that is to say it is able to resolve queries about any attribute about
+that is to say it is able to resolve queries for any attribute about
 any URN in the universe.
@@
 stored in segments having a URN ending with ``{\em properties}''. The
-AFFLIB implementation loads these segments automatically into the
+AFF4 implementation loads these segments automatically into the
 Universal Resolver.
@@
 
-When the volume is loaded, the AFFLIB implementation automatically
-loads any properties files and populates its Universal Resolver with
-the information visibile to it. AFFLIB provides a mechanism to use an
-external resolver stored on a MySQL database to provide a Universal
-Resolver that shares information between different instances on the
-same network.
+When the volume is loaded, the AFF4 implementation automatically loads
+any properties files and populates its Universal Resolver with the
+information visibile to it. AFF4 provides a mechanism to use an
+external resolver stored on a MySQL database to provide for a
+Universal Resolver that shares information between different instances
+on the same network.
 
@@
 {\em Directory Volume} and the {\em ZipFile Volume}. It is possible to
-convert from one implementation to another easily, without effecting
+convert from one implementation to another easily, without affecting
 any external references. This is because external references are to
-the Volume URN itself, so its safe to convert from one representation
-to another. In this way its possible for example, to unzip a Zip64
-volume into a directory and vice versa.
+the Volume URN itself, so it is safe to convert from one
+representation to another. In this way its possible for example, to
+unzip a Zip64 volume into a directory and vice versa.
+
+It is important to emphesise that Volumes are merely containers which
+provide storage for segments. There is no restriction of which
+segments can be stored by any particular volume. For example, the
+segments which make up a single Image stream may be stored in a number
+of volumes (splitting the image in some way among them). Similarly,
+the segments representing a number of streams may be stored in the
+same volume.
 
@@
   of user and developer tools already available. The volume may be
-  inspected using any number of commercial zip application
-  (e.g. Windows Explorer natively supports Zip files as can be seen in
-  Figure~\ref{explorer}).
+  inspected using any number of commercial or open source zip
+  application (e.g. Windows Explorer natively supports Zip files as
+  can be seen in Figure~\ref{explorer}, Zip64 is supported natively
+  by Java, Python and PERL).
 
@@
 
-  \caption{The basic structure of a Zip archive. Also shows is how new
+  \caption{The basic structure of a Zip archive. Also shown is how new
   archive members are added to an existing Zip File. The Central
-  Directory gets overwritten by the new member, and a new Central
+  Directory is overwritten by the new member, and a new Central
   Directory is written on the end.  }
@@
 their use and because they lack the functionality that is important
-for a forensic user. Instead, we use AFF1's digital signature
+for a forensic user. Instead, we use AFF4's digital signature
 facilities for integrity and non-repudiation, and we introduce a new
@@
 \subsection{Implementing Zip64 in AFFLIB}
-Although there are numerous Zip implementations available
-today, we have created our own implementation using a combination of
-available open source technology and our own novel contributions. We
-are in the process of integrating this code into the AFFLIB library
-and performing extensive regression testing.
-
-There are many reasons to develop our own Zip64 implementation for
-AFF4:
+Although there are numerous Zip implementations available today, we
+have created our own implementation. There are many reasons to develop
+our own Zip64 implementation for AFF4:
 
@@
 id to the stream URN (e.g. ``urn:aff4:83a3d6db-85d5/00000032''). Each
-data segment is called a bevy and stores a number of compressed chunks
-back to back.
+data segment is called a \emph{bevy} and stores a number of compressed
+chunks back to back.
 
@@
 EnCase\cite{encase-3.0} and implemented by the open source
-libewf\cite{libewf} package. The advantages of keeping smaller sized
-chunks is the better match between requested size and the minimum size
-required for decompression. This means that less data is needed to be
-decompressed unnecessarily.
+libewf\cite{libewf} package. This improvement from AFF1's 16MB segment
+size results in a better match between requested size and the minimum
+size required for decompression. This means that less data is needed
+to be decompressed unnecessarily where reading small sectors randomly.
 
@@
 ``/map'' to the stream URN. The segment data consists of a series of
-lines, each containing a stream offset, followed by a comma, and a
-target offset. Offsets are encoded using decimal notation, and are
-followed by a target URN.
+lines, each containing a stream offset, a target offset and a target
+URN. Offsets are encoded using decimal notation.
 
@@
 block level encryption for the contained AFF4 volume (which might
-contain arbitrary streams). It is recommended that Encrypted Streams
-containing an AFF4 volume set an attribute
-\emph{content-type=application/x-aff-volume} on the container stream,
-to allow implementations to automatically recognize volumes. This
-approach is illustrated in Figure~\ref{crypted_fif}.
+contain arbitrary streams). This approach is illustrated in
+Figure~\ref{crypted_fif}.
 
@@
 and may implement different encryption schemes and keys for each
-Encrypted scheme. This effectively allows arbitrary access policies to
+Encrypted stream. This effectively allows arbitrary access policies to
 be implemented as only volumes which can be accessed can be merged.
 
-\subsection{Encryption Schemes}
-Our implementation provides a number of encryption schemes.
-
-\subsubsection{Null Encryption}
-This scheme ({\em scheme=null}) specifies no encryption at all. This
-may be useful for testing an implementation, but obviously provides
-no real security.
-
-\subsubsection{AES-SHA-PSK}
-This scheme ({\em scheme=aes-sha-psk}) uses AES for block level
-encryption, with the following SHA based ESSIV scheme, and Pre-Shared
-Key (passphrase).
-
-When creating a new encryption stream, a master key is generated using
-the first 128 bits of a SHA1 hash of the Pre-Shared-Key appended to a
-64 bit random salt:
-\begin{eqnarray}
-Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
-\end{eqnarray}
-
-For each chunk, the chunk IV is obtained by taking the first 128 bits
-of the SHA1 hash of the chunk id encoded as a 32 bit little endian
-integer, followed by the master key:
-\begin{eqnarray}
-IV = \left | SHA1(chunk\_id + Key_{Master}) \right | _{128}
-\end{eqnarray}
-
-The block is then encrypted or decrypted using AES in CBC mode with
-this IV.
-
-The salt is specified as a stream attribute named {\em aff4:salt} in
-base64 encoding. Implementations are free to specify any API for
-passing the Pre-Shared Key, for example the PSK can be specified in a
-configuration file, typed in at the terminal, passed in using process
-environment variables, or encoded within the AFF4 URL itself.
+\subsection{The Link Object}
+Although the URN of a stream names it unambiguously in the AFF4
+universe it is difficult to use and communicate due to its random
+nature. Most investigators would prefer to use a shorter name which
+might well represent the image better in their minds (e.g. a case
+name or warrant number).
+
+A Link is an object which has a single attribute
+\emph{aff4:target}. When the Link object is opened, it returns the
+object named by this attribute. This allows images with complex names
+to be referred to via short, meaningful names. Note that in practice
+both Image Streams and Link Objects are created by the imaging tools,
+so as far as the operators are concerned they can always refer to the
+Image stream via the Link name.
+
+\subsection{Identity Object}
+AFF4 defines a \emph{Statement} as a collection of facts, or
+(subject, attribute, value) tuples. Listing \ref{statement}
+illustrates a collection of facts encoded in the standard AFF4
+notation (SHA254 hashes are base64 encoded).
+
+\begin{lstlisting}[
+	caption=An Example Statement  (URNs and hashes are
+shortened for illustration).,
+	label=statement,
+	float=tb,
+	 ]
+urn:aff4:34a62f06/00000000 aff4:sha256=+Xf4i....7rPCgo=
+urn:aff4:34a62f06/00000000.idx aff4:sha256=ptV7xOK6....C7R6Xs=
+urn:aff4:34a62f06/properties aff4:sha256=yoZ9dtv....eUignYMtk=
+urn:aff4:34a62f06 aff4:sha256=udajC5C...BVii7psU=
+\end{lstlisting}
+
+The statement expresses a set of attributes of other AFF4 objects, and
+in particular the attribute of SHA256 hash is expressed (but other
+attributes may also be expressed).
+
+Digital signatures have been used in previous forensic file formats
+(such as AFF1) to provide authentication and non-repudiation of
+forensic evidence. In essence, a when a person signs an object they
+are vouching for its authenticity. Similarly, when a person signs a
+\emph{Statement}, they are vouching for its authenticity. This concept
+is similar to the Bill of Material (BOM) from AFF1.
+
+An AFF4 Identity object represents an entity, currently described by
+way of an X509 certificate. The URN of an identity object is the
+certificate's fingerprint, and is therefore unique to the
+certificate. Identity objects contains \emph{aff4:statement}
+attributes which refer to AFF4 streams containing statements. The
+identity object also contains a copy of the certificate used to sign
+the statements.
+
+To verify the signatures, the AFF4 library loads the stored
+certificate, then checks the signature for each statement. If a
+statement is verified (i.e. deemed as correct according to the
+identity), the facts within it are checked. Note that it is possible
+for multiple identities to sign the same data.
+
+%\subsection{Encryption Schemes}
+%Our implementation provides a number of encryption schemes.
+
+%\subsubsection{Null Encryption}
+%This scheme ({\em scheme=null}) specifies no encryption at all. This
+%may be useful for testing an implementation, but obviously provides
+%no real security.
+
+%\subsubsection{AES-SHA-PSK}
+%This scheme ({\em scheme=aes-sha-psk}) uses AES for block level
+%encryption, with the following SHA based ESSIV scheme, and Pre-Shared
+%Key (passphrase).
+
+%When creating a new encryption stream, a master key is generated using
+%the first 128 bits of a SHA1 hash of the Pre-Shared-Key appended to a
+%64 bit random salt:
+%\begin{eqnarray}
+%Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
+%\end{eqnarray}
+
+%For each chunk, the chunk IV is obtained by taking the first 128 bits
+%of the SHA1 hash of the chunk id encoded as a 32 bit little endian
+%integer, followed by the master key:
+%\begin{eqnarray}
+%IV = \left | SHA1(chunk\_id + Key_{Master}) \right | _{128}
+%\end{eqnarray}
+
+%The block is then encrypted or decrypted using AES in CBC mode with
+%this IV.
+
+%The salt is specified as a stream attribute named {\em aff4:salt} in
+%base64 encoding. Implementations are free to specify any API for
+%passing the Pre-Shared Key, for example the PSK can be specified in a
+%configuration file, typed in at the terminal, passed in using process
+%environment variables, or encoded within the AFF4 URL itself.
 
@@
 redirected to the NY copy. This load redistribution can be used for
-optimal management of evidence storage in a transparent way. Analyst
+optimal management of evidence storage in a transparent way. Analysts
 are not aware of where the evidence is physically stored, and it
 appears as though all evidence is always available.
+
+If Alice's local NY copy is now lost, Alice's local AFF4 library will
+fail to open the NY URL, and will automatically fall back to the copy
+stored in LA. This will require access across the WAN, which will be
+slower, but provides a kind of distributed fail over capability.
 
@@
 
-\subsection{In-place encryption of AFF4 volumes}
-With the large size of a typical forensic image it is often difficult
-to have to create a second copy of a volume which needs to be
-encrypted. It is usually more convenient to encrypt the file in place,
-overwriting the plain text data with the equivalent encrypted data.
-
-This can be achieved using the technique shown in
-Figure~\ref{inplace_crypto}. The first step is to encrypt the first
-block from the front of the file and add it to the end of the archive
-as a new encrypted segment. Then a ZipFile header is written to the
-front which encapsulates the rest of the plain text volume. The plain
-text volume is then encrypted in place. Finally a map stream is used
-to re-map the first and second fragments into place.
-
-\begin{figure}[tb]
-  \begin{center}
-  \mbox{\epsfxsize=0.8\columnwidth \epsffile{inplace.eps}}
-
-  \caption{The process of encrypting a volume in place. The first
-  block from the volume is encrypted and appended to the end of the
-  volume as a new segment (A). Now a zip file header can be written in
-  its place to form the remainder of the volume as segment B. The rest
-  of the volume is encrypted in-place, block by block (Step 4). Now we
-  create a Map stream which combines segment A and B into a single
-  logical stream (M). An Encrypted stream which is backed by this map
-  stream (E) is now added. Finally a Central Directory is appended to
-  the end of the new volume. We end up with a new Zip File which
-  contains our old volume stored within an encrypted stream.}
-
-  \label{inplace_crypto}
-  \end{center}
-\end{figure}
+%\subsection{In-place encryption of AFF4 volumes}
+%With the large size of a typical forensic image it is often difficult
+%to have to create a second copy of a volume which needs to be
+%encrypted. It is usually more convenient to encrypt the file in place,
+%overwriting the plain text data with the equivalent encrypted data.
+
+%This can be achieved using the technique shown in
+%Figure~\ref{inplace_crypto}. The first step is to encrypt the first
+%block from the front of the file and add it to the end of the archive
+%as a new encrypted segment. Then a ZipFile header is written to the
+%front which encapsulates the rest of the plain text volume. The plain
+%text volume is then encrypted in place. Finally a map stream is used
+%to re-map the first and second fragments into place.
+
+%\begin{figure}[tb]
+%  \begin{center}
+%  \mbox{\epsfxsize=0.8\columnwidth \epsffile{inplace.eps}}
+
+%  \caption{The process of encrypting a volume in place. The first
+%  block from the volume is encrypted and appended to the end of the
+%  volume as a new segment (A). Now a zip file header can be written in
+%  its place to form the remainder of the volume as segment B. The rest
+%  of the volume is encrypted in-place, block by block (Step 4). Now we
+%  create a Map stream which combines segment A and B into a single
+%  logical stream (M). An Encrypted stream which is backed by this map
+%  stream (E) is now added. Finally a Central Directory is appended to
+%  the end of the new volume. We end up with a new Zip File which
+%  contains our old volume stored within an encrypted stream.}
+
+%  \label{inplace_crypto}
+%  \end{center}
+%\end{figure}
 
@@
 assigned to different users by encrypting the master key with
-different X509 certificates. Its is also possible for users to create
+different X509 certificates. It is also possible for users to create
 non-encrypted volumes within the AFF4 volume.
@@
 
-\subsection{Sharing evidence}
-When managing a large corpus of evidence it is common to require these
-to be shared across the network. Many laboratories create a
-centralized Network Attached Storage (NAS) facility to retain their
-evidence corpus. This is typically shared via some file sharing
-protocol, commonly Windows sharing (CIFS) or NFS.
-
-It is advantageous to be able to move evidence around between several
-servers and locations. For example, consider a disk image stored in a
-Volume with a given URN. Normally, when a user wishes to open the
-image, they will specify the path and filename of the volume which
-contains it. The library will then open the volume and retrieve the
-stream.
-
-The AFF4 model is different however. In AFF4 metadata is managed in a location
-independant manner 
-through a Universal Resolver (perhaps implemented in a database). Relevant information
-is found not only in Volumes, but may additionally come from other sources.
-This means that in the
-above example, the user will query the resolver about the location of
-the volume. The resolver returns a URL to the current location of the
-volume. The user's library will then open a HTTP Stream instance to
-the location returned and the volume may be opened directly over the
-intra-web connection.
-
-If the volume needs to be moved, the volume's \emph{aff4:stored}
-attribute can be adjusted to the new location, and all clients will
-automatically know where to find it. Furthermore, if several copies of
-the same volume exist, the resolver may have multiple values for the
-\emph{aff4:stored} property, some locations may be faster to access than
-others. The Resolver can then return the most suitable attribute based
-on locality.
-
-Another interesting side effect of the sharing model is that it is far
-more efficient to allow remote analysts to peruse subsets of the
-evidence. For example, Alice is in the laboratory and wants Bob who is
-on a remote site to view several files she found on her
-image. Traditionally she would either need to copy the entire volume
-and send it to him. Or maybe extract the files she thinks are relevant
-and send those to him.
-
-With the new AFF4 architecture, Bob can open the volume directly,
-using the HTTP protocol, and view the files he needs. Very little
-extra data is actually transferred in excess of the data he wishes to
-see. This is a much more efficient way to allow remote analysts to
-collaborate.
+%\subsection{Sharing evidence}
+%When managing a large corpus of evidence it is common to require these
+%to be shared across the network. Many laboratories create a
+%centralized Network Attached Storage (NAS) facility to retain their
+%evidence corpus. This is typically shared via some file sharing
+%protocol, commonly Windows sharing (CIFS) or NFS.
+
+%It is advantageous to be able to move evidence around between several
+%servers and locations. For example, consider a disk image stored in a
+%Volume with a given URN. Normally, when a user wishes to open the
+%image, they will specify the path and filename of the volume which
+%contains it. The library will then open the volume and retrieve the
+%stream.
+
+%The AFF4 model is different however. In AFF4 metadata is managed in a location
+%independant manner 
+%through a Universal Resolver (perhaps implemented in a database). Relevant information
+%is found not only in Volumes, but may additionally come from other sources.
+%This means that in the
+%above example, the user will query the resolver about the location of
+%the volume. The resolver returns a URL to the current location of the
+%volume. The user's library will then open a HTTP Stream instance to
+%the location returned and the volume may be opened directly over the
+%intra-web connection.
+
+%If the volume needs to be moved, the volume's \emph{aff4:stored}
+%attribute can be adjusted to the new location, and all clients will
+%automatically know where to find it. Furthermore, if several copies of
+%the same volume exist, the resolver may have multiple values for the
+%\emph{aff4:stored} property, some locations may be faster to access than
+%others. The Resolver can then return the most suitable attribute based
+%on locality.
+
+%Another interesting side effect of the sharing model is that it is far
+%more efficient to allow remote analysts to peruse subsets of the
+%evidence. For example, Alice is in the laboratory and wants Bob who is
+%on a remote site to view several files she found on her
+%image. Traditionally she would either need to copy the entire volume
+%and send it to him. Or maybe extract the files she thinks are relevant
+%and send those to him.
+
+%With the new AFF4 architecture, Bob can open the volume directly,
+%using the HTTP protocol, and view the files he needs. Very little
+%extra data is actually transferred in excess of the data he wishes to
+%see. This is a much more efficient way to allow remote analysts to
+%collaborate.
 
@@
 We have developed an open source reference implementation, but the
-AFF4 framework is simple enough for competing implementations to be
-implemented commercially. We hope this simplicity enhances AFF4's
-acceptance and adoption as a standard evidence management platform.
+AFF4 framework is simple enough for competing implementations. We hope
+this simplicity enhances AFF4's acceptance and adoption as a standard
+evidence management platform.
 

### 2009-05-13 paper/paper.tex
simson's edit of AFF4 Ignore-this: e1cda0fe54390f6f1560d4fba53d2232

@@
 analysis results among different tools. The new specification, termed
-AFF4, was designed to be simple to implement, allowing the use of the
-well supported Zip File format specifications for bit level file
-access.
+AFF4, was designed to be simple to implement, bult upon the well
+supported ZIP file format specification. AFF4 implements will have
+downward compatability with existing AFF files.
 \end{abstract}
@@
 form that was compressed but allowed random-access to evidence data
-necessary for forensic analysis. The \emph{sgzip} format does not
-store any metadata at all.
-
-The Advanced Forensic Format (AFF) expanded on this idea with a
-forensic file format that allowed both data and arbitrary metadata to
-be stored in a single digital archive\cite{garfinkel:aff}.
-
-The Eye Witness Forensic (EWF) file format was originally developed
+necessary for forensic analysis.
+
+The Expert Witness Forensic (EWF) file format was originally developed
 for Encase\cite{encase}, but then adopted by other vendors
@@
 
+The Advanced Forensic Format (AFF) expanded on this idea with a
+forensic file format that allowed both data and arbitrary metadata to
+be stored in a single digital archive\cite{garfinkel:aff}.
+
 Both the AFF and EWF file formats are designed to store a single
@@
 unable to exist outside the context of the image itself.
+But AFF went beyond merely
+specifying metadata as an addition to the forensic evidence
+itself. Metadata forms an essential part of the file format, with the
+same metadata system storing both essential information required to
+read the image itself, as well as non-essential metadata. AFF allows
+the use of arbitrary name/value pairs to store metadata, but it also
+stored critical information such as sector size and device serial
+number using the same metadata representation system. \verb+Aimage+,
+the AFF hard  disk acquisition tool should, not only stored the image,
+but also stored a description of the tool
+itself, the version of AFFLIB used to create the image, the computer
+on which the image was made, the operator of the tool, the user supplied parameters
+supplied to the tool, all in addition to the drive being imaged.
 
@@
 \subsection{This paper}
-The AFF4 metadata model described in this work, goes beyond merely
-specifying metadata as an addition to the forensic evidence
-itself. Metadata forms an essential part of the file format, with the
-same metadata system storing both essential information required to
-read the image itself, as well as non-essential metadata. For example,
-an AFF4 compliant hard disk acquisition tool should, in addition to
-storing an image, store entities representing and describing the tool
-itself, the operator of the tool, the user supplied parameters
-supplied to the tool, and the drive being imaged.
-
 An important advance of this work is the introduction of storage
@@
 transformation in a flexible and concise way, allowing users to view
-multiple transformations of the same data with virtually zero storage
+multiple transformations of the same data with little additional storage
 cost. This mechanism is an important enabler for interoperable
@@
 data types (including network packets and memory images), extracted
-logical evidence, and forensic workflow. The AFF4 format extends to
-become a Global Distributed Evidence Management system, while
-maintaing compatibility with the old AFF file format.
+logical evidence, and forensic workflow. The AFF4 format extends the
+format to make it the basis of a global distributed evidence management system.
 
@@
 AFFLIB implementations with major version numbers 1, 2 and 3. We
-therefore call our system AFF4 to avoid confusion.}
+therefore call our system AFF4 to avoid confusion.} The AFF4
+implementation, when it is publicly released, will be able to read
+existing AFF files.
 
@@
   network packets.
-
-\item AFF1 (and the EWF format) are unable to recover from loss or
-  corruption. Due to the specialised nature of these formats,
-  specialised recovery software must be written to recover from
-  corruptions.
 
@@
 at a time. This allows us to extend evidence collection and management
-to a truely universal scale - allowing creation and management of a
+to a truely universal scale---allowing creation and management of a
 large corpus.
@@
 URN scheme, which we call the AFF4 URN. URN's of this scheme use the
-namespace\cite{RFC1737} \emph{\'aff4\'} and therefore begin with the 
+namespace\cite{RFC1737} ``\emph{aff4}'' and therefore begin with the 
 string(\emph{urn:aff4}). AFF4 URNs are then be made unique by use of 
@@
 Attribute, Value), where {\em Subject} is the URN of the object the
-statement is made about. Using this system we are able to store
+statement is made about. An \emph{Attribute{ can be any kind of value, such
+as the sector size of a device, a device capacity, or the name of the
+person who performed an imaging operation. A \emph{Value} is the value
+of the attribute. Using this system we are able to store
 arbitrary attributes about any object in the AFF4
@@
 information visibile to it. AFF4 provides a mechanism to use an
-external resolver stored on a MySQL database to provide for a
+external resolver as well---for example, we have implemented a
+resolved that stores Attributes in a MySQL database to provide for a
 Universal Resolver that shares information between different instances
@@
 \begin{itemize}
-\item Most Zip implementations do not implement the Zip64
-  extensions. These extensions are required to support Evidence
-  Volumes larger than 2GB.
+\item The commonly available Zip implementations written in C do not
+ implement the Zip64 extensions. These extensions are required to
+ support Evidence Volumes larger than 2GB.
 
@@
 that Alice does not need any hardware to obtain the image as it is
-done over the network - she therefore can respond rapidly.
+done over the network---she therefore can respond rapidly.
 

### 2009-05-13 paper/paper.tex
whoops Ignore-this: 8906b6af17650a9224713896ee9a11d3

@@
 Attribute, Value), where {\em Subject} is the URN of the object the
-statement is made about. An \emph{Attribute{ can be any kind of value, such
+statement is made about. An \emph{Attribute} can be any kind of value, such
 as the sector size of a device, a device capacity, or the name of the

### 2009-05-15 paper/paper.tex
More hacking on paper.

@@
 AFF4, was designed to be simple to implement, bult upon the well
-supported ZIP file format specification. AFF4 implements will have
+supported ZIP file format specification. AFF4 implementation have
 downward compatability with existing AFF files.
@@
 these formats in themselves have a limited metadata model which is
-unable to exist outside the context of the image itself.
-But AFF went beyond merely
-specifying metadata as an addition to the forensic evidence
-itself. Metadata forms an essential part of the file format, with the
-same metadata system storing both essential information required to
-read the image itself, as well as non-essential metadata. AFF allows
-the use of arbitrary name/value pairs to store metadata, but it also
-stored critical information such as sector size and device serial
-number using the same metadata representation system. \verb+Aimage+,
-the AFF hard  disk acquisition tool should, not only stored the image,
-but also stored a description of the tool
+unable to exist outside the context of the image itself.  But AFF went
+beyond merely specifying metadata as an addition to the forensic
+evidence itself. Metadata forms an essential part of the file format,
+with the same metadata system storing both essential information
+required to read the image itself, as well as non-essential
+metadata. AFF allows the use of arbitrary name/value pairs to store
+metadata, but it also stored critical information such as sector size
+and device serial number using the same metadata representation
+system. \verb+Aimage+, the AFF hard disk acquisition tool, not only
+stores the image, but additionally stores a description of the tool
 itself, the version of AFFLIB used to create the image, the computer
-on which the image was made, the operator of the tool, the user supplied parameters
-supplied to the tool, all in addition to the drive being imaged.
+on which the image was made, the operator of the tool, the user
+supplied parameters supplied to the tool.
 
@@
 We call the new system AFF4, and use the phrase AFF1 to refer to the
-legacy system developed by Garfinkel et al.\footnote{Although Garfinkel
-never changed the AFF bit-level specification, Garfinkel released
-AFFLIB implementations with major version numbers 1, 2 and 3. We
-therefore call our system AFF4 to avoid confusion.} The AFF4
-implementation, when it is publicly released, will be able to read
-existing AFF files.
+legacy system developed by Garfinkel et al.\footnote{Although
+Garfinkel never changed the AFF bit-level specification, Garfinkel
+released AFFLIB implementations with major version numbers 1, 2 and
+3. We therefore call our system AFF4 to avoid confusion.} The publicly
+released AFF4 implementation, is able to read existing AFF files.
 
@@
 
-\section{Global Distributed Evidence Management}
+\subsection{Global Distributed Evidence Management}
 While AFF1 was designed for use on a single machine that could both
@@
 
-\subsection{Migration from AFF1 to AFF4}
-We have addressed these issues in a number of ways. First, we have
-extended the AFF schema to support multiple forensic images (which are
-now called \emph{streams}). It is therefore possible now to store
-multiple related streams in the same volume, or to acquire mutiple
-streams at the same time (e.g. network and disk data).
-
-We replaced AFF's ``bit-level'' format with an improved format that is
-more flexible and has better existing support, namely the Zip64 file
-format is now adopted as a default storage format. This enhanced our
-compatibility with other tools and simplifies implementations for
-compatible systems.
-
-We have also defined an information model which addresses uniquely
-identifying evidence data, expressing arbitrary information and
-integrating information and image data from multiple evidence volumes
-at a time. This allows us to extend evidence collection and management
-to a truely universal scale---allowing creation and management of a
-large corpus.
-
-We have also released a library and toolset, \emph{LIBAFF4}
-implementing the new schema and file format. This new library provides
-legacy support for existing AFF1 volumes.
+%\subsection{Migration from AFF1 to AFF4}
+%We have addressed these issues in a number of ways. First, we have
+%extended the AFF schema to support multiple forensic images (which are
+%now called \emph{streams}). It is therefore possible now to store
+%multiple related streams in the same volume, or to acquire mutiple
+%streams at the same time (e.g. network and disk data).
+
+%We replaced AFF's ``bit-level'' format with an improved format that is
+%more flexible and has better existing support, namely the Zip64 file
+%format is now adopted as a default storage format. This enhanced our
+%compatibility with other tools and simplifies implementations for
+%compatible systems.
+
+%We have also defined an information model which addresses uniquely
+%identifying evidence data, expressing arbitrary information and
+%integrating information and image data from multiple evidence volumes
+%at a time. This allows us to extend evidence collection and management
+%to a truely universal scale - allowing creation and management of a
+%large corpus.
+
+%We have also released a library and toolset, \emph{LIBAFF4}
+%implementing the new schema and file format. This new library provides
+%legacy support for existing AFF1 volumes.
 
@@
 \item \emph{An AFF Object} is the basic building block of our
-file format. AFF Objects have a globally unique name (URN) as described in
-\cite{RFC1737}. The name is defined within the aff4 namespace, and is
-made unique by use of a unique identifier generated as per
-RFC4122\cite{RFC4122}.
-
-\item \emph{A Relation} is a statement about an AFF Object which
+  file format. AFF Objects have a globally unique name (URN) as
+  described in \cite{RFC1737}. The name is defined within the aff4
+  namespace, and is made unique by use of a unique identifier
+  generated as per RFC4122\cite{RFC4122}.
+
+\item \emph{A Relation or Fact} is a statement about an AFF Object which
   expresses some property of the object. The relation comprises of a
-triple of (Subject, Attribute, Value). All metadata is reduced to the
-triple notation.
+  tuple of (Subject, Attribute, Value). All metadata is reduced to
+  this tuple notation.
 
@@
 
-Management of evidence requires an effective identififaction, with
+Management of evidence requires an effective identification, with
 practitioners currently employing acquisition time metadata such as
@@
 independently from the data itself. We term {\em metadata} to be a set
-of statements about objects, written in triple notations (Subject,
+of statements about objects, written in tuple notations (Subject,
 Attribute, Value), where {\em Subject} is the URN of the object the
@@
 stream stored in a different volume, it is not generally possible to
-tell in which volume the stream exists, and where that volume is
-actually stored.
+tell where that volume is actually stored.
 
@@
 To this end we define a way for serializing metadata statements (or
-triples) into a standard format which implementations can load into
-their respective resolvers when parsing the volume. Triples can be
+tuples) into a standard format which implementations can load into
+their respective resolvers when parsing the volume. Relations can be
 stored in segments having a URN ending with ``{\em properties}''. The
@@
 
-Triples are stored within the properties segment one per line, with
+Relations are stored within the properties segment one per line, with
 the subject URN (encoded according to RFC1737), followed by whitespace
@@
 external resolver as well---for example, we have implemented a
-resolved that stores Attributes in a MySQL database to provide for a
+resolver that stores Attributes in a MySQL database to provide for a
 Universal Resolver that shares information between different instances
@@
   map aff4:type=link
+
+Identity Object:
+  urn:aff4:identity/41:13 aff4:common_name=/C=US/ST=CA/
+                          L=SanFrancisco/O=Fort-Funston/
+			  CN=client1/emailAddress=
+			  me@myhost.mydomain
+  urn:aff4:identity/41:13 aff4:type=identity
+  urn:aff4:identity/41:13 aff4:statement=00000000
+  urn:aff4:identity/41:13 aff4:x509=urn:aff4:identity/
+                          41:13/cert.pem
 \end{lstlisting}
@@
 
-\subsection{Unused Zip64 Features}
-There are a number of aspects in the Zip file format specifications
-which we do not utilize.
-\begin{itemize}
-\item We ignore Zip64's built-in support for splitting
-archives into multiple Zip files. Instead, our implementation treats
-each volume as a complete and stand-alone Zip file. The AFF4
-implementation then considers the segments contained within as
-belonging to the universal collection. We provide the ability to split
-a stream across volumes automatically, as each segment may be stored
-in a different volume, and the Universal Resolver is used to find it.
-
-\item Although Zip64 also defines encryption and authentication
-extensions, we do not use them due to the restrictions imposed on
-their use and because they lack the functionality that is important
-for a forensic user. Instead, we use AFF4's digital signature
-facilities for integrity and non-repudiation, and we introduce a new
-stream based encryption scheme for ensuring data privacy
-(Section~\ref{crypted_stream}).
-
-\end{itemize}
-
-\subsection{Implementing Zip64 in AFFLIB}
+We ignore Zip64's built-in support for splitting archives into
+multiple Zip files. Instead, our implementation treats each volume as
+a complete and stand-alone Zip file. The AFF4 implementation then
+considers the segments contained within as belonging to the universal
+collection. This provide the ability to split a stream across volumes
+automatically, as different segments within the same stream may be
+stored in different volumes.
+
+Zip64 also defines encryption and authentication extensions. We do not
+use them due to the restrictions imposed on their use and because they
+lack the functionality that is important for a forensic user. Instead,
+we use AFF4's digital signature facilities for integrity and
+non-repudiation, and we introduce a new stream based encryption scheme
+for ensuring data privacy (Section~\ref{crypted_stream}).
+
 Although there are numerous Zip implementations available today, we
@@
 volume may be specified as a URL
-(e.g. \emph{http://intranet/123453/}).  AFFLIB provides transparent
+(e.g. \emph{http://intranet/123453/}).  AFF4 provides transparent
 support for HTTP and FTP URLs by means of the curl HTTP
@@
 client is interested in. This allows efficient network transport as we
-do not need to download unnecessary data, we just request exactly the
-data the client application requires.
+do not need to download unnecessary data, we just request those chunks
+the client application requires.
 
@@
 supports the File Transfer Protocol (FTP) and HTTPS (Secure Sockets
-Layer---SSL) protocols transparently. This support is provided
-automatically by the curl library.
+Layer---SSL) protocols transparently, as provided by the curl library.
 
@@
 
-\subsection{Identity Object}
+\section{Identity Object}
 AFF4 defines a \emph{Statement} as a collection of facts, or
@@
 	 ]
-urn:aff4:34a62f06/00000000 aff4:sha256=+Xf4i....7rPCgo=
-urn:aff4:34a62f06/00000000.idx aff4:sha256=ptV7xOK6....C7R6Xs=
-urn:aff4:34a62f06/properties aff4:sha256=yoZ9dtv....eUignYMtk=
+urn:aff4:34a62f06/00000 aff4:sha256=+Xf4i....7rPCgo=
+urn:aff4:34a62f06/00000.idx aff4:sha256=ptV7xOK6....C7R6Xs=
+urn:aff4:34a62f06/properties aff4:sha256=yoZ....YMtk=
 urn:aff4:34a62f06 aff4:sha256=udajC5C...BVii7psU=
@@
 
-\subsection{Private Streams}
-Often forensic software needs to store internal state, such as the
-current state of the GUI, or internal data structures which may be
-required for caching.
-
-Currently, software create additional cache files to maintain this
-information. It is advantageous to be able to store these in the
-evidence volume itself. The evidence volume can then be re-opened by
-the software at a later stage and private data can be retrieved and
-used directly. Such private application data can be stored in a
-Private stream. There is no specification of what can be stored in the
-private stream, and applications with private streams should be able
-to store arbitrary segments in any format at all. Segments URNs should
-be prefixed by the private stream URNs, and the stream type should be
-specified as ``private''.
-
-It is recommended that applications create attributes for the private
-stream so they can reliably recognize their own private stream, and
-identify the version of the application which created it. The AFFLIB
-API can be used to render access to private segments and private
-attributes.
+%\section{Private Objects}
+%Often forensic software needs to store internal state, such as the
+%current state of the GUI, or internal data structures which may be
+%required for caching.
+
+%Currently, software create additional cache files to maintain this
+%information. It is advantageous to be able to store these in the
+%evidence volume itself. The evidence volume can then be re-opened by
+%the software at a later stage and private data can be retrieved and
+%used directly. Such private application data can be stored in a
+%Private object. There is no specification of what can be stored in the
+%private object, and applications with private objects should be able
+%to store arbitrary segments in any format at all. Segments URNs should
+%be prefixed by the private object's URN, and the stream type should be
+%specified as ``private''.
+
+%It is recommended that applications create attributes for the private
+%stream so they can reliably recognize their own private stream, and
+%identify the version of the application which created it. The AFF4
+%API can be used to render access to private segments and private
+%attributes.
 

### 2009-05-19 paper/paper.tex
bls-paper Ignore-this: da8191ba56f9ce08571c88872042044a

@@
 analysis results among different tools. The new specification, termed
-AFF4, was designed to be simple to implement, bult upon the well
-supported ZIP file format specification. AFF4 implementation have
+AFF4, is designed to be simple to implement, bult upon the well
+supported ZIP file format specification. Furthermore, the AFF4 implementation has
 downward compatability with existing AFF files.
@@
 Storing and managing digital evidence is becoming increasingly more
-difficult, as the volume and size of digital evidence increases. The
-type of what consitutes evidence has also evolved to include data
+difficult, as the volume and size of digital evidence increases. Evidence
+sources have also evolved to include data
 other than disk images, such as memory images, network images and
-regular files. Preserving digital evidence is an important part of
+regular files. Preserving such digital evidence is an important part of
 most digital investigations\cite{carrier:event-based}, and managing
@@
 software systems for making and authenticating ``images'' of digital
-evidence are commonly used
+evidence have become common
 (e.g. \cite{safeback,ilook,encase}). PyFlag\cite{pyflag} introduced a
@@
 for Encase\cite{encase}, but then adopted by other vendors
-\cite{libewf}. The EWF file format also compresses the image into 32kb
+\cite{libewf}. The EWF file format similarly compresses the image into 32kb
 chunks which are stored back to back in groupings inside the file. The
-format contains multiple tables containing relative indexes to these
-compressed chunks which are also stored within the volume. EWF volumes
-have a maximum size of 2Gb and therefore usually split an image across
-many volumes. EWF provides for a small number of predefined metadata
+format employs tables of relative indexes to the
+compressed chunks to improve random access efficienct. EWF volumes
+have a maximum size limit of 2Gb and therefore usually split an image across
+many files. EWF provides for a small number of predefined metadata
 fields to be stored within the file format.
@@
 these formats in themselves have a limited metadata model which is
-unable to exist outside the context of the image itself.  But AFF went
+unable to exist outside the context of the image itself.  But the AFF went
 beyond merely specifying metadata as an addition to the forensic
@@
 
-\item \emph{A Relation or Fact} is a statement about an AFF Object which
-  expresses some property of the object. The relation comprises of a
+\item \emph{A Relation} is a factual statement which is used either to describe
+  a relationship between two AFF Objects, or to describe some property of an object. The relation comprises of a
   tuple of (Subject, Attribute, Value). All metadata is reduced to
@@
 Attribute, Value), where {\em Subject} is the URN of the object the
-statement is made about. An \emph{Attribute} can be any kind of value, such
+statement is made about. An \emph{Attribute} can be any kind of value or relationship, such
 as the sector size of a device, a device capacity, or the name of the
 person who performed an imaging operation. A \emph{Value} is the value
-of the attribute. Using this system we are able to store
+of the attribute, which is either another URN, or some textual value. Using this system we are able to store
 arbitrary attributes about any object in the AFF4
@@
 
-It is important to emphesise that Volumes are merely containers which
+It is important to emphasise that Volumes are merely containers which
 provide storage for segments. There is no restriction of which
@@
 \section{Identity Object}
-AFF4 defines a \emph{Statement} as a collection of facts, or
+AFF4 defines a \emph{Statement} as a collection of relations, or
 (subject, attribute, value) tuples. Listing \ref{statement}
-illustrates a collection of facts encoded in the standard AFF4
+illustrates a collection of relations encoded in the standard AFF4
 notation (SHA254 hashes are base64 encoded).
@@
 statement is verified (i.e. deemed as correct according to the
-identity), the facts within it are checked. Note that it is possible
+identity), the relations within it are checked. Note that it is possible
 for multiple identities to sign the same data.

### 2009-05-19 include/aff4.h
Working encrypted implementation using passphrases.

@@ -6,7 +6,7 @@
 
 // This is the URI namespace for the AFF4 scheme
 #define NAMESPACE "aff4:"
-#define VOLATILE_NS "aff2volatile:"
+#define VOLATILE_NS "aff4volatile:"
 #define FQN "urn:" NAMESPACE
 #define FILE_NS "file://"
 
@@ -60,7 +60,7 @@
     information related to the current running instance **/
 #define AFF4_CONFIG_NAMESPACE FQN "urn:aff4config:"
 #define AFF4_CONFIG_PAD           "pad_on_read_error"
-
+#define AFF4_CONFIG_AUTOLOAD      "autoload_volumes"
 
 /** These are various properties */
 #define AFF4_AUTOLOAD         NAMESPACE "autoload"  /** Instructs the loader to
@@ -78,6 +78,19 @@
 
   // All identity URNs are prefixed with this:
 #define AFF4_IDENTITY_PREFIX  FQN AFF4_IDENTITY
+
+/** Volatile attributes are never written to the file but are used to
+    pass messages between different parts of the code.
+*/
+// Thats the passphrase that will be used to encrypt the session key
+#define AFF4_VOLATILE_PASSPHRASE VOLATILE_NS "passphrase"
+
+// This is the session key for encryption
+#define AFF4_VOLATILE_KEY   VOLATILE_NS "key"
+#define AFF4_CRYPTO_NAMESPACE NAMESPACE "crypto:"
+#define AFF4_CRYPTO_FORTIFICATION_COUNT AFF4_CRYPTO_NAMESPACE "fortification"
+#define AFF4_CRYPTO_IV  AFF4_CRYPTO_NAMESPACE "IV"
+#define AFF4_CRYPTO_PASSPHRASE AFF4_CRYPTO_NAMESPACE "passphrase"
 
 /** 
     Opens the URIs specified in images and uses them to populate the

### 2009-05-19 include/class.h

@@ -443,6 +443,10 @@ void *raise_errors(enum _error_type t, char *string,  ...);
 #define CheckError(error)			\
   (_global_error == error)
 
+  // These dont do anything but are useful to indicate when a function
+  // parameter is used purely to return a value
+#define OUT
+#define IN
 
 #endif
 #ifdef __cplusplus

### 2009-05-19 include/encode.h

@@ -1,0 +1,17 @@
+#ifndef _ENCODE_H
+#define _ENCODE_H
+#include <class.h>
+#include <misc.h>
+
+/* binary encoding functions.
+      
+   We assume output buffer to be large enough to accept the encoding.
+*/
+int encode64(unsigned char *inbuf, int len, unsigned char *outbuf, int outlen);
+int decode64(unsigned char *inbuf, int len, unsigned char *outbuf, int outlen);
+int encode32(unsigned char *inbuf, int len, unsigned char *outbuf, int outlen);
+int decode32(unsigned char *inbuf, int len, unsigned char *outbuf, int outlen);
+int encodehex(unsigned char *inbuf, int len, unsigned char *outbuf);
+void encode_init();
+
+#endif

### 2009-05-19 include/zip.h

@@ -418,9 +418,12 @@ END_CLASS
   the block number (as 32 bit int in little endian) appended to an 8
   byte salt.
 *************************************************************/
+#include <openssl/aes.h>
+
 CLASS(Encrypted, FileLikeObject)
      StringIO block_buffer;
-     char *key;
+     AES_KEY ekey;
+     AES_KEY dkey;
      char *salt;
      char *target_urn;
      // Our volume urn
@@ -428,6 +431,7 @@ CLASS(Encrypted, FileLikeObject)
 
      // The block size for CBC mode
      int block_size;
+     int block_number;
 END_CLASS
 
  char *resolver_get_with_default(Resolver self, char *urn, 

### 2009-05-19 lib/encode.c

@@ -1,0 +1,334 @@
+#include "encode.h"
+
+#define LUT_SIZE 256
+#define STUFF_CHAR 'a'
+
+static unsigned char *encode_lut;
+static unsigned char *decode_lut;
+
+static unsigned char *encode64_lut;
+static unsigned char *decode64_lut;
+
+static unsigned char *urlsafe_encode64_lut;
+static unsigned char *urlsafe_decode64_lut;
+
+/* Build the luts */
+void encode_init() {
+  unsigned char i;
+  int k,j=0;
+
+  //Prepare the LUT buffers and clear them:
+  encode_lut = (unsigned char *)talloc_size(NULL, LUT_SIZE);
+  decode_lut = (unsigned char *)talloc_size(NULL, LUT_SIZE);
+
+  encode64_lut = (unsigned char *)talloc_size(NULL, LUT_SIZE);
+  decode64_lut = (unsigned char *)talloc_size(NULL, LUT_SIZE);
+
+  urlsafe_encode64_lut = (unsigned char *)talloc_size(NULL, LUT_SIZE);
+  urlsafe_decode64_lut = (unsigned char *)talloc_size(NULL, LUT_SIZE);
+
+  memset(decode_lut,' ',LUT_SIZE);
+  memset(encode_lut,' ',LUT_SIZE);
+  memset(decode64_lut,' ',LUT_SIZE);
+  memset(encode64_lut,' ',LUT_SIZE);
+  memset(urlsafe_decode64_lut,' ',LUT_SIZE);
+  memset(urlsafe_encode64_lut,' ',LUT_SIZE);
+
+  for(i='z';i>='b';i--,j++) encode_lut[j]=i;
+  for(i='9';i>='0';i--,j++) encode_lut[j]=i;
+  for(i='A';i<='Z';i++,j++) encode_lut[j]=i;
+
+  j=0;
+  for(i='A';i<='Z';i++,j++) encode64_lut[j]=i;
+  for(i='a';i<='z';i++,j++) encode64_lut[j]=i;
+  for(i='0';i<='9';i++,j++) encode64_lut[j]=i;
+  j--;
+
+  j=0;
+  for(i='A';i<='Z';i++,j++) urlsafe_encode64_lut[j]=i;
+  for(i='a';i<='z';i++,j++) urlsafe_encode64_lut[j]=i;
+  for(i='0';i<='9';i++,j++) urlsafe_encode64_lut[j]=i;
+  j--;
+
[…]

### 2009-05-19 lib/encrypt.c

@@ -1,5 +1,191 @@
-#include "openssl/aes.h"
+#include <openssl/ssl.h>
+#include <openssl/aes.h>
+#include <openssl/rand.h>
 #include "zip.h"
+
+/** Saves the key in the resolver, using all the methods which are
+    available:
+
+    Passphrase is a secret password to encrypt the key with.
+    The following steps are taken:
+
+    1) A random number between 2^16 and 2^24 is used as round number
+       for passphrase fortification. This parameter is saved in the
+       resolver as AFF4_CRYPTO_FORTIFICATION_COUNT.
+
+    2) The password is hashed this many times using SHA256. This
+       produces a new 256 bit key.
+
+    3) The stream key is then encrypted using the above key
+       with AES, and the IV.
+
+    4) We save the IV in the resolver using AFF4_CRYPTO_IV
+*/
+static int prepare_passphrase(Encrypted self, OUT unsigned char *key, int key_length) {
+  char *passphrase = CALL(oracle, resolve, AFF4_CONFIG_NAMESPACE, AFF4_VOLATILE_PASSPHRASE);
+  
+  if(passphrase) {
+    int passphrase_len = strlen(passphrase);
+    uint32_t fortification_count = parse_int(CALL(oracle, resolve, URNOF(self),
+						  AFF4_CRYPTO_FORTIFICATION_COUNT));
+    EVP_MD_CTX digest;
+    unsigned char buffer[BUFF_SIZE];
+    unsigned int buffer_length=BUFF_SIZE;
+    int i;
+
+    EVP_DigestInit(&digest, EVP_sha256());
+
+    if(fortification_count == 0) {
+      // Some random data for the fortification_count
+      RAND_pseudo_bytes((unsigned char *)&fortification_count, 
+			sizeof(fortification_count));
+      fortification_count &= 0xFFFF;
+      fortification_count <<= 8;
+      
+      CALL(oracle, set, URNOF(self), AFF4_CRYPTO_FORTIFICATION_COUNT, 
+	   from_int(fortification_count));
+    };
+
+    /* Fortify the passphrase (This is basically done in order to
+       increase the cost of checking the passphrase by a huge factor
+       making dictionary attacks very ineffective). 
+    */
+    for(i=0; i<fortification_count; i++) {
+      EVP_DigestUpdate(&digest, passphrase, passphrase_len);
+    };
+
[…]

### 2009-05-19 lib/identity.c

@@ -427,7 +427,7 @@ static void verify_hashes(Identity self, int (*cb)(uint64_t progress, char *urn)
 
 	memset(buff, 0, BUFF_SIZE);
 	EVP_DigestFinal(&digest, buff,&len);
-	encode64(buff, len, hash_base64);
+	encode64(buff, len, hash_base64, sizeof(hash_base64));
 
 	CALL(oracle, set, urn, AFF4_SHA, (char *)hash_base64);
       };

### 2009-05-19 lib/image.c

@@ -241,7 +241,7 @@ static void Image_close(FileLikeObject self) {
   dump_stream_properties(self);
 
   EVP_DigestFinal(&this->digest, buff, &len);
-  encode64(buff, len, hash_base64);
+  encode64(buff, len, hash_base64, sizeof(hash_base64));
   CALL(oracle, set, URNOF(self), AFF4_SHA, (char *)hash_base64);
 
   this->__super__->close(self);

### 2009-05-19 lib/libaff4.c

@@ -5,8 +5,14 @@
 
 ZipFile open_volume(char *urn, char mode) {
   ZipFile result=NULL;
-  char *filename = normalise_url(urn);
+  char *filename;
   struct dispatch_t *i=volume_handlers;
+
+  if(startswith(urn, FQN)) {
+    filename = urn;
+  } else {
+    filename =  normalise_url(urn);
+  };
 
   ClearError();
 
@@ -25,6 +31,48 @@ ZipFile open_volume(char *urn, char mode) {
     i++;
   };
   PrintError();
+
+  // Check for autoload in this volume
+  if(result && mode=='r') {
+    char *autoload = resolver_get_with_default(oracle, AFF4_CONFIG_NAMESPACE, 
+					       AFF4_CONFIG_AUTOLOAD, "1");
+    char base_path[BUFF_SIZE]; 
+    char *dirpath;
+
+    strncpy(base_path, urn, BUFF_SIZE);
+    dirpath = dirname(base_path);
+
+    if(!strcmp(autoload, "1")) {
+      // Get a list of autoload volumes:
+      struct aff4_tripple **result_set = aff4_query(oracle, URNOF(result), AFF4_AUTOLOAD, NULL);
+      int i;
+      char *result_urn = talloc_strdup(result_set, URNOF(result));
+
+      CALL(oracle, cache_return, (AFFObject)result);
+
+      // Open them all and return to cache
+      for(i=0; result_set[i]; i++) {
+	// Note that autoloads are specified relative to the current
+	// urn path, or in a fully qualified way:
+	ZipFile fd;
+	char *path = fully_qualified_name(result_set[i]->value, dirpath);
+	
+	LogWarnings("Autoloading %s", path);
+	fd = open_volume(path, 'r');
+	if(fd) {
+	  CALL(oracle, cache_return, (AFFObject)fd);
+	};
+      };
+
+      // Get it back
+      result = (ZipFile)CALL(oracle, open, result_urn, mode);
+
+      // Remove autoloads from our resolver
+      CALL(oracle, del, URNOF(result), AFF4_AUTOLOAD);
+      // Done.
+      talloc_free(result_set);
+    };
+  };
 
   return result;
 };

### 2009-05-19 lib/resolver.c

@@ -527,7 +527,7 @@ static AFFObject Resolver_create(Resolver self, AFFObject *class_reference) {
   if(!class_reference || !*class_reference) return NULL;
 
   result = CONSTRUCT_FROM_REFERENCE((*class_reference), Con, self, NULL, 'w');
-
+  talloc_set_name(result, "%s: %s (" __location__ ")\n              ", NAMEOF(result), URNOF(result));
   return result;
 };
 
@@ -646,6 +646,8 @@ static AFFObject AFFObject_Con(AFFObject self, char *uri, char mode) {
 
   if(!self->urn)
     self->urn = uri;
+
+  self->mode = mode;
 
   return self;
 };

### 2009-05-19 lib/zip.c

@@ -867,12 +867,16 @@ static void ZipFile_close(ZipFile self) {
 	// This is a zip member
 	if(!strcmp(attribute, AFF4_CONTAINS)) {
 	  struct CDFileHeader cd;
+	  char *type = CALL(oracle, resolve, value, AFF4_TYPE);
 	  time_t epoch_time = parse_int(CALL(oracle, resolve, value, 
 					     AFF4_TIMESTAMP));
 	  struct tm *now = localtime(&epoch_time);
 	  // This gets the relative name of the fqn
 	  char *escaped_filename = escape_filename(relative_name(value, URNOF(self)));
 	  //	  LogWarnings("Writing archive member %s -> %s", escaped_filename, value);
+
+	  // Its not a segment after all
+	  if(type && strcmp(type, AFF4_SEGMENT)) continue;
 
 	  memset(&cd, 0, sizeof(cd));
 	  cd.magic = 0x2014b50;
@@ -1230,7 +1234,7 @@ static void ZipFileStream_close(FileLikeObject self) {
     memset(digest, 0, BUFF_SIZE);
     EVP_DigestFinal(&this->digest,digest,&digest_len);
     if(digest_len < BUFF_SIZE) {
-      encode64(digest, digest_len, digest_encoded);
+      encode64(digest, digest_len, digest_encoded, sizeof(digest_encoded));
       
       // Tell the resolver what the hash is:
       CALL(oracle, set, URNOF(self), AFF4_SHA, (char *)digest_encoded);

### 2009-05-19 tools/aff4imager.c

@@ -1,0 +1,658 @@
+#include "zip.h"
+#include "aff4.h"
+#include <openssl/bio.h>
+#include <openssl/pem.h>
+#include <openssl/x509.h>
+#include "common.h"
+
+#define IMAGE_BUFF_SIZE (1024*1024)
+
+void list_info() {
+  char *result =  CALL(oracle, export_all, "");
+  printf("%s", result);
+
+  talloc_free(result);
+};
+
+ZipFile create_volume(char *driver) {
+  if(!strcmp(driver, AFF4_ZIP_VOLUME)) {
+    return (ZipFile)CALL(oracle, create, (AFFObject *)&__ZipFile);
+  } else if(!strcmp(driver, AFF4_DIRECTORY_VOLUME)) {
+    return (ZipFile)CALL(oracle, create, (AFFObject *)&__DirVolume);
+  } else {
+    RaiseError(ERuntimeError, "Unknown driver %s", driver);
+    return NULL;
+  };
+};
+
+void aff2_open_volume(char *urn, char mode) {
+  // Try the different volume implementations in turn until one works:
+  ZipFile volume;
+
+  volume = open_volume(urn, mode);
+
+  CALL(oracle, cache_return, (AFFObject)volume);
+};
+
+void aff2_print_info(int verbose) {
+  Cache i,j;
+
+  list_for_each_entry(i, &oracle->urn->cache_list, cache_list) {
+    char *urn = (char *)i->key;
+    Cache attributes = (Cache)i->data;
+
+    printf("\n******** Object %s ***********\n",urn);
+    list_for_each_entry(j, &attributes->cache_list, cache_list) {
+      char *key = (char *)j->key;
+      char *value = (char *)j->data;
+      Resolver i;
+
+      // Ignore volatile data in default verbosity.
+      if(verbose==0 && !memcmp(key, ZSTRING_NO_NULL(VOLATILE_NS)))
+	continue;
+
+      printf("\t%s = %s\n", key, value);
+      /** Now print which identity vouches for this fact */
+      list_for_each_entry(i, &oracle->identities, identities) {
+	char *hash = CALL(i, resolve, urn, key);
+
+	if(hash) {
+	  if(!strcmp(hash, value)) {
+	    printf("\t\t---> Signed by %s\n", URNOF(i));
+	  } else {
+	    printf("\t\t---> Incorrect %s says it should be %s\n", URNOF(i), hash);
+	  };
+	};
+      };
+    };
+  };
+};
+
+/** When we want to use encryption we must create a container volume,
[…]

### 2009-05-19 tools/common.c

@@ -1,0 +1,50 @@
+#include "common.h"
+
+void print_help(struct option *opts) {
+  struct option *i;
+
+  printf("The following options are supported\n");
+
+  for(i=opts; i->name; i++) {
+    char *description = (char *)i->name + strlen(i->name) +1;
+    char short_opt[BUFF_SIZE];
+    char *prefix="\t";
+
+    if(description[0]=='*') {
+      prefix="\n";
+      description++;
+    };
+
+    if(i->val) 
+      snprintf(short_opt, BUFF_SIZE, ", -%c", i->val);
+    else
+      short_opt[0]=0;
+
+    if(i->has_arg) {
+      printf("%s--%s%s [ARG]\t%s\n", prefix, i->name, short_opt, description);
+    } else {
+      printf("%s--%s%s\t\t%s\n", prefix,i->name, short_opt, description);
+    };
+  };
+};
+
+
+char *generate_short_optargs(struct option *opts) {
+  static char short_opts[BUFF_SIZE];
+  int j,i;
+  
+  for(i=0, j=0; i<BUFF_SIZE && opts[j].name; j++) {
+    if(opts[j].val) {
+      short_opts[i]=opts[j].val;
+      i++;
+      if(opts[j].has_arg) {
+	short_opts[i]=':';
+	i++;
+      };
+    };
+  };
+  
+  short_opts[i]=0;
+
+  return short_opts;
+};

### 2009-05-20 paper/paper.tex
Polishing of paper - cleaned up code to remove warnings.

@@
 globally unique identifiers to ensure no name collisions can occur
-with disconnected systems. AFF1 assigns each piece of evidence a
+with disconnected locations. AFF1 assigns each piece of evidence a
 unique 128-bit identifier called a GID but did not make it clear when
@@
 
-In this case they can each create a new volume which extends to
+In this case they can each create a new volume which extends the
 original volume and save their analysis on this new volume. Now they
@@
 This is made possible because each volume is independent of one
-another, but can still be recombined into a bigger evidence set.
+another, but is still viewed as part of a bigger evidence set.
 
@@
 
-\item \emph{A Target Reference} is a way of referencing objects by use
+\item \emph{A Reference} is a way of referencing objects by use
   of a Uniform Resource Identifier (URI). The URI can be another AFF
@@
 namespace\cite{RFC1737} ``\emph{aff4}'' and therefore begin with the 
-string(\emph{urn:aff4}). AFF4 URNs are then be made unique by use of 
+string ``\emph{urn:aff4}''. AFF4 URNs are then be made unique by use of 
 a unique identifier generated
@@
 in some other way inside or outside the volume (e.g. SQLite
-databases).
+database files).
 
@@
 any properties files and populates its Universal Resolver with the
-information visibile to it. AFF4 provides a mechanism to use an
+information visible to it. AFF4 provides a mechanism to use an
 external resolver as well---for example, we have implemented a
 resolver that stores Attributes in a MySQL database to provide for a
-Universal Resolver that shares information between different instances
-on the same network.
+persistent Universal Resolver that shares information between
+different instances on the same network.
 
 Each URN within the AFF4 universe must have an ``\texttt{aff4:type}''
-attribute to denote the type of the Object.
+attribute to denote the type of the Object. Objects may also have a
+the ``\texttt{aff4:interface}'' attribute to denote what kind of
+interface they present (e.g. {\emph stream} or {\emph volume}).
 
@@
 convert from one implementation to another easily, without affecting
-any external references. This is because external references are to
-the Volume URN itself, so it is safe to convert from one
-representation to another. In this way its possible for example, to
-unzip a Zip64 volume into a directory and vice versa.
-
-It is important to emphasise that Volumes are merely containers which
+any external references.
+
+It is important to emphasize that Volumes are merely containers which
 provide storage for segments. There is no restriction of which
@@
 restriction on what type of URL this can be, so it may be a location
-on a filesystem (e.g. {\em file://some\_directory/}) or a location on a
-HTTP server (e.g. {\em http://intranet.server/some\_path}). In this
+on a filesystem (e.g. {\em file:///some/directory/}) or a location on a
+HTTP server (e.g. {\em http://intranet.server/some/path}). In this
 way its possible to move the entire volume from a filesystem to a web
@@
   by Java, Python and PERL).
-
-\item Zip64 already includes support for the ``table of contents''
-  envisioned but never implemented by Garfinkel; Zip64 calls this
-  table the ``Central Directory.'' It is written at the end of the
-  Zip64 volume.
 
@@
 be written to pipes for example, sending an image over the network
-using netcat or ssh.
+using netcat or ssh. AFF4 always uses the data description header to
+ensure volumes are written continuously without needing to seek in the
+output file.
 
@@
 mechanism, our system also supports legacy AFF1 volumes as well as
-segments that are stored in other storage systems---for example,
-Amazon.com's Simple Storage System (S3)\cite{s3-aws-home-page-money}.
+Expert Witness Evidence files \cite{libewf}.
 
@@
   (or even hundreds of thousands) of segments in an efficient data
-  structure. In fact our implementation uses the Universal Resolver to
-  store the parsed central directory information, which means that in
-  most cases we dont even need to scan the Central Directory at all.
+  structure. In fact our implementation uses the Universal Resolver
+  itself to store the parsed central directory information, which
+  means that in most cases we do not even need to scan the Central
+  Directory at all.
 
@@
   is possible to scan through the volume, and locate all the Zip64
-  file headers. Then its possible to repair and reconstruct the
+  file headers. Then it is possible to repair and reconstruct the
   central directory.
@@
 Image streams specify the \texttt{chunk\_size} attribute, as the
-number of image bytes each chunk contains (typically chunk sizes are
+number of image bytes each chunk contains (chunk size defaults to
 32kb). Also specified is the \texttt{chunks\_per\_segment} attribute
@@
 size results in a better match between requested size and the minimum
-size required for decompression. This means that less data is needed
-to be decompressed unnecessarily where reading small sectors randomly.
+size required for decompression. Less data is needed to be
+decompressed unnecessarily where reading small sectors randomly,
+leading to vast performance improvements.
 
@@
 mapping of the Physical Address Space through a page table
-transformation.  Zero Storage Carving\cite{Meijer2006} is a way of
-specifying carved files in terms of a sequence of blocks taken from
-the image; Cohen extended this concept to an arbitrary mapping
-function\cite{1363239,Cohen2007} which can be used to describe
-arbitrary mappings of carved files within a single image.
+transformation \cite{Tanenbaum2008}.  Zero Storage
+Carving\cite{Meijer2006} is a way of specifying carved files in terms
+of a sequence of blocks taken from the image; Cohen extended this
+concept to an arbitrary mapping function\cite{1363239,Cohen2007} which
+can be used to describe arbitrary mappings of carved files within a
+single image.
 
@@
 y = (x - X_i) + Y_i & &
-\forall x \in \left [X_i, X_i+1 \right )
+\forall x \in \left [X_i, X_{i+1} \right )
 \end{eqnarray}
@@
 RAID arrays, the Map stream may be provided with two optional
-parameters a {\em target\_period} ($T_p$), and {\em stream\_period}
+parameters: a {\em target\_period} ($T_p$), and {\em stream\_period}
 ($S_p$). If specified, the above relation becomes:
@@
 (e.g. \emph{http://intranet/123453/}).  AFF4 provides transparent
-support for HTTP and FTP URLs by means of the curl HTTP
+support for HTTP and FTP URLs by means of the Curl HTTP
 library\cite{libcurl}. The HTTP Stream, therefore satisfies read
@@
 supports the File Transfer Protocol (FTP) and HTTPS (Secure Sockets
-Layer---SSL) protocols transparently, as provided by the curl library.
+Layer---SSL) protocols transparently, as provided by the Curl library.
 
@@
 Encrypted stream. This effectively allows arbitrary access policies to
-be implemented as only volumes which can be accessed can be merged.
+be implemented as only volumes which can be accessed can be read.
 
@@
 certificate's fingerprint, and is therefore unique to the
-certificate. Identity objects contains \emph{aff4:statement}
+certificate. Identity objects contain \emph{aff4:statement}
 attributes which refer to AFF4 streams containing statements. The
@@
 An investigation is conducted by the New York team. The case relates
-to a hard disk image stream stored inside a volume, in turn stored on
-the NY evidence server at URL http://ny.wan/evidence1.aff4 . The team
-requires an analyst (Bob) in LA to assist with their analysis. The LA
-analyst types\footnote{fls is the file listing command which is part
-of the Sleuthkit}:
+to a hard disk {\emph Image stream} stored inside a volume, in turn
+stored on the NY evidence server at URL http://ny.wan/evidence1.aff4
+. The team requires an analyst (Bob) in LA to assist with their
+analysis. The LA analyst types\footnote{fls is the file listing
+command which is part of the Sleuthkit}:
 
@@
 AFF4 evidence universe by acquiring each evidence file using a unique
-URN. As explained in \ref{definitions}, segments are AFF4 stream
+URN. As explained in Section \ref{definitions}, segments are AFF4 stream
 objects which are implemented by storing the data in Zip archive
@@
 Alice uses windows explorer to obtain a Zip file of the files of
-interest. After taking the file back to the lab, she then signs the
+interest. After taking the archive back to the lab, she then signs the
 files, and adds a volume URN, making the Zip file a fully compliant

### 2009-05-20 include/aff4.h

@@ -58,7 +58,7 @@
 #define AFF4_IDENTITY         "identity"
 /** This is the configuration namespace - it stores configuration
     information related to the current running instance **/
-#define AFF4_CONFIG_NAMESPACE FQN "urn:aff4config:"
+#define AFF4_CONFIG_NAMESPACE     "urn:aff4config:"
 #define AFF4_CONFIG_PAD           "pad_on_read_error"
 #define AFF4_CONFIG_AUTOLOAD      "autoload_volumes"
 
@@ -91,6 +91,10 @@
 #define AFF4_CRYPTO_FORTIFICATION_COUNT AFF4_CRYPTO_NAMESPACE "fortification"
 #define AFF4_CRYPTO_IV  AFF4_CRYPTO_NAMESPACE "IV"
 #define AFF4_CRYPTO_PASSPHRASE AFF4_CRYPTO_NAMESPACE "passphrase"
+#define AFF4_CRYPTO_ALGORITHM AFF4_CRYPTO_NAMESPACE "algorithm"
+
+// Supported algorithms
+#define AFF4_CRYPTO_ALGORITHM_AES_SHA254 "AES256/SHA256"
 
 /** 
     Opens the URIs specified in images and uses them to populate the

### 2009-05-20 lib/encrypt.c

@@ -41,7 +41,8 @@ static int prepare_passphrase(Encrypted self, OUT unsigned char *key, int key_le
 			sizeof(fortification_count));
       fortification_count &= 0xFFFF;
       fortification_count <<= 8;
-      
+
+      // Store it
       CALL(oracle, set, URNOF(self), AFF4_CRYPTO_FORTIFICATION_COUNT, 
 	   from_int(fortification_count));
     };
@@ -90,6 +91,7 @@ static int save_AES_key(Encrypted self, unsigned char affkey[32]) {
     
     encode64((unsigned char *)encrypted_key, 32, buffer, sizeof(buffer));
     CALL(oracle, set, URNOF(self), AFF4_CRYPTO_PASSPHRASE, (char *)buffer);
+    CALL(oracle, set, URNOF(self), AFF4_CRYPTO_ALGORITHM, AFF4_CRYPTO_ALGORITHM_AES_SHA254);
     
     saved = 1;
   };
@@ -100,9 +102,7 @@ static int save_AES_key(Encrypted self, unsigned char affkey[32]) {
 /*
   Make a new random key. Note that we use 256 bits (32 bytes)
   of randomness, but for convenience we base64 encode it so
-  it can be stored in the resolver easily. Since openssl
-  doesnt care it can make AES keys from whatever string we
-  give it.
+  it can be stored in the resolver easily.
 
   Note that we store the key in the resolver for other instances to
   use.

### 2009-05-20 lib/image.c

@@ -123,7 +123,7 @@ static int dump_chunk(Image this, char *data, uint32_t length, int force) {
     memcpy(cbuffer, data, length);
     clength = length;
   } else {
-    if(compress((unsigned char *)cbuffer, (unsigned long int *)&clength, 
+    if(compress((unsigned char *)cbuffer, (unsigned long int *)(char *)&clength, 
 		(unsigned char *)data, (unsigned long int)length) != Z_OK) {
       RaiseError(ERuntimeError, "Compression error");
       return -1;
@@ -266,7 +266,7 @@ static int partial_read(FileLikeObject self, StringIO result, int length) {
 
   /* Temporary storage for the uncompressed chunk */
   char *uncompressed_chunk;
-  unsigned int uncompressed_length=this->chunk_size + 1024;
+  unsigned long int uncompressed_length=this->chunk_size + 1024;
 
   /** Now we need to figure out where the segment is */
   char buffer[BUFF_SIZE];
@@ -352,7 +352,7 @@ static int partial_read(FileLikeObject self, StringIO result, int length) {
   if(this->compression == 8) {
     // Try to decompress it:
     if(uncompress((unsigned char *)uncompressed_chunk, 
-		  (unsigned long int *)&uncompressed_length, 
+		  &uncompressed_length, 
 		  (unsigned char *)compressed_chunk,
 		  (unsigned long int)compressed_length) != Z_OK ) {
       RaiseError(ERuntimeError, "Unable to decompress chunk %d", chunk_id);
@@ -441,10 +441,11 @@ void dump_stream_properties(FileLikeObject self) {
 
 
 VIRTUAL(Image, FileLikeObject)
-     VMETHOD(super.super.Con) = Image_Con;
-     VMETHOD(super.super.finish) = Image_finish;
-     VMETHOD(super.read) = Image_read;
-     VMETHOD(super.write) = Image_write;
-     VMETHOD(super.close) = Image_close;
+     VMETHOD_BASE(AFFObject, Con) = Image_Con;
+     VMETHOD_BASE(AFFObject, finish) = Image_finish;
+
+     VMETHOD_BASE(FileLikeObject, read) = Image_read;
+     VMETHOD_BASE(FileLikeObject, write) = Image_write;
+     VMETHOD_BASE(FileLikeObject,close) = Image_close;
 END_VIRTUAL
 

### 2009-05-20 lib/libaff4.c

@@ -2,6 +2,7 @@
  */
 #include "zip.h"
 #include "aff4.h"
+#include <libgen.h>
 
 ZipFile open_volume(char *urn, char mode) {
   ZipFile result=NULL;

### 2009-05-20 lib/map.c

@@ -317,10 +317,10 @@ static int MapDriver_read(FileLikeObject self, char *buffer, unsigned long int l
 
 
 VIRTUAL(MapDriver, FileLikeObject)
-     VMETHOD(super.super.Con) = MapDriver_Con;
-     VMETHOD(super.super.finish) = MapDriver_finish;
+     VMETHOD_BASE(AFFObject, Con) = MapDriver_Con;
+     VMETHOD_BASE(AFFObject, finish) = MapDriver_finish;
      VMETHOD(add) = MapDriver_add;
      VMETHOD(save_map) = MapDriver_save_map;
-     VMETHOD(super.read) = MapDriver_read;  
-     VMETHOD(super.close) = MapDriver_close;
+     VMETHOD_BASE(FileLikeObject, read) = MapDriver_read;  
+     VMETHOD_BASE(FileLikeObject, close) = MapDriver_close;
 END_VIRTUAL

### 2009-05-20 lib/resolver.c

@@ -326,7 +326,7 @@ static char **Resolver_resolve_list(Resolver self, void *ctx, char *urn, char *a
   if(i->hash_table && i->hash_table[hash]) {
     list_for_each_entry(j, &i->hash_table[hash]->hash_list, hash_list) {
       char *key = (char *)j->key;
-      char **existing = (char **)&j->data;
+      char **existing = (char **)((char *)&j->data);
       if(!strcmp(key, attribute)) {
 	// Push the string onto the result - note that we rely on the
 	// fact that the existing string will never be freed.
@@ -356,7 +356,6 @@ static AFFObject Resolver_open(Resolver self, char *urn, char mode) {
   struct dispatch_t *dispatch_ptr=NULL;
   AFFObject result;
   Cache tmp;
-  char buff[BUFF_SIZE];
 
   if(!urn) return NULL;
 

### 2009-05-20 lib/zip.c

@@ -677,7 +677,7 @@ static char *ZipFile_read_member(ZipFile self, void *ctx,
 static FileLikeObject ZipFile_open_member(ZipFile self, char *filename, char mode,
 				   char *extra, uint16_t extra_field_len,
 				   int compression) {
-  FileLikeObject result;
+  FileLikeObject result=NULL;
 
   switch(mode) {
   case 'w': {
@@ -1118,7 +1118,7 @@ static int ZipFileStream_write(FileLikeObject self, char *buffer, unsigned long 
 static int ZipFileStream_read(FileLikeObject self, char *buffer,
 			      unsigned long int length) {
   ZipFileStream this = (ZipFileStream)self;
-  int len;
+  int len=0;
   FileLikeObject fd;
 
   // We only read as much data as there is

### 2009-05-20 paper/paper.tex
More paper polishing

@@
 analysis results among different tools. The new specification, termed
-AFF4, is designed to be simple to implement, bult upon the well
+AFF4, is designed to be simple to implement, built upon the well
 supported ZIP file format specification. Furthermore, the AFF4 implementation has
-downward compatability with existing AFF files.
+downward comparability with existing AFF files.
 \end{abstract}
@@
 efforts into a universal evidence management system. The detailed
-description of the AFF4 proposal is then followed by concerete real
+description of the AFF4 proposal is then followed by concrete real
 world use cases.
@@
 format employs tables of relative indexes to the
-compressed chunks to improve random access efficienct. EWF volumes
+compressed chunks to improve random access efficiency. EWF volumes
 have a maximum size limit of 2Gb and therefore usually split an image across
@@
 Both the AFF and EWF file formats are designed to store a single
-image, and any metadata implicitely refers to that image. Therefore,
+image, and any metadata implicitly refers to that image. Therefore,
 these formats in themselves have a limited metadata model which is
@@
 
-%Both Of These Approaches Are Limited In Their Expressiveness, By
-%Nature Of Their (Property, Value) Data Model. The Subject Of The
-%Properties Is Implicitly The Single Image Contained Within The
-%Evidence Container. Clearly An Implicit Subject Metadata Model Is
-%Inadequate For Storing And Managing Multiple Evidence Sources
-%Simultaneously.
-
 Schatz proposed a Sealed Digital Evidence Bags architecture,
@@
 
-%Recently the open source AFF implementation library (AFFLIB) was
-%expanded to support both encryption and digital signatures of digital
-%evidence\cite{garfinkel:affcrypto}. Clearly, the implementation of
-%encryption and authentication is critical for forensic file formats.
-
 \subsection{This paper}
@@
 transformation functions to the forensic storage container. Prior
-works simply focused on forensically sound storage of bitstreams,
+works simply focused on forensically sound storage of bit-streams,
 leaving the necessary activities of translating low level storage into
@@
 multiple transformations of the same data with little additional storage
-cost. This mechanism is an important enabler for interoperable
+cost. This mechanism is an important enabler for inter-operable
 forensic tools. For example, carved files may be described in terms of
@@
 
-%\begin{figure}
-%\fbox{
-%\begin{minipage}{.95\linewidth}
-%\begin{itemize}
-%\item The ability to store evidence from one or more physical
-%  devices.
-%\item Schema allowing evidence files to contain memory
-%  images, network packet intercepts, Netflow data, and extracted
-%  files.
-
-%\item Direct encoding of arbitrary mapping transformations of one or
-%more sources to allow multiple logical views. This allows 
-%RAID volumes to be automatically reassembled from evidence files made
-%of the physical RAID devices. Maps can be created using Sleuthkit and
-%stored in the evidence file, thereafter allowing arbitrary programs to directly
-%read individual files stored in a disk image.
-
-%\item A revised compression system that allows individual chunks
-%  within a 16MB segment to be decompressed without decompressing the
-%  entire segment.
-
-%\item Support for Global Distributed Evidence Management.
-%\end{itemize}
-%\end{minipage}}
-%\caption{Specific design goals for the next generation forensic file
-%  format.}
-%\end{figure}
-
 \section{The Need for an Improved Forensic Format}
@@
 
-%(Once the file is opened, the
-%  sectors corresponding to the segment headers are stored in the
-%  computer's disk cache.)
-
 \item AFF1's bit-level specification is essentially a simple container
@@
 multiple locations and is performed by multiple individuals.
-
-%This is a
-%state of affairs which mirrors common practice in the field.
 
@@
 
-%\subsection{Migration from AFF1 to AFF4}
-%We have addressed these issues in a number of ways. First, we have
-%extended the AFF schema to support multiple forensic images (which are
-%now called \emph{streams}). It is therefore possible now to store
-%multiple related streams in the same volume, or to acquire mutiple
-%streams at the same time (e.g. network and disk data).
-
-%We replaced AFF's ``bit-level'' format with an improved format that is
-%more flexible and has better existing support, namely the Zip64 file
-%format is now adopted as a default storage format. This enhanced our
-%compatibility with other tools and simplifies implementations for
-%compatible systems.
-
-%We have also defined an information model which addresses uniquely
-%identifying evidence data, expressing arbitrary information and
-%integrating information and image data from multiple evidence volumes
-%at a time. This allows us to extend evidence collection and management
-%to a truely universal scale - allowing creation and management of a
-%large corpus.
-
-%We have also released a library and toolset, \emph{LIBAFF4}
-%implementing the new schema and file format. This new library provides
-%legacy support for existing AFF1 volumes.
-
 \section{Introducing AFF4}
@@
 design is object oriented, in that a few generic objects are presented
-with externaly accessible behaviour. We discuss a number of
+with externally accessible behavior. We discuss a number of
 implementations of these high level concepts and show how these can be
@@
 \item \emph{A Relation} is a factual statement which is used either to describe
-  a relationship between two AFF Objects, or to describe some property of an object. The relation comprises of a
-  tuple of (Subject, Attribute, Value). All metadata is reduced to
-  this tuple notation.
+  a relationship between two AFF Objects, or to describe some property
+  of an object. The relation comprises of a tuple of (Subject,
+  Attribute, Value). All metadata is reduced to this tuple notation.
 
@@
 \label{resolver}
-%One of the fundamental tasks of forensics is in ensuring the
-%provenance of any particular piece of digital evidence, for example by
-%employing a chain of custody process. To assist in this, acquisition
-%methodologies emphasise the process of individuating evidence, and
-%storing it in a manner that is easily identifiable.
 
@@
 managing evidence. For example, at acquisition time a suitably unique
-individuator may not be selected. If that occured, at analysis time
+individuator may not be selected. If that occurred, at analysis time
 evidence container files may need to be renamed to avoid name
@@
 
-%In this way we can identify uniquely any evidence without the chance
-%for name collisions. Once a universal identification regime is
-%created, it is now possible to collect metadata about each object.
-
 The AFF4 model treats metadata as an abstract concept which may exist
@@
 Attribute, Value), where {\em Subject} is the URN of the object the
-statement is made about. An \emph{Attribute} can be any kind of value or relationship, such
-as the sector size of a device, a device capacity, or the name of the
-person who performed an imaging operation. A \emph{Value} is the value
-of the attribute, which is either another URN, or some textual value. Using this system we are able to store
-arbitrary attributes about any object in the AFF4
+statement is made about. An \emph{Attribute} can be any kind of value
+or relationship, such as the sector size of a device, a device
+capacity, or the name of the person who performed an imaging
+operation. A \emph{Value} is the value of the attribute, which is
+either another URN, or some textual value. Using this system we are
+able to store arbitrary attributes about any object in the AFF4
 universe. Additionally, as these statements are universally scoped,
@@
 different instances on the same network.
+
+Although the Universal Resolver should be thought of as a truly
+universal entity, the library provides a local resolver which is
+available to the running instance. As the library explores different
+volumes, relations are added to the local resolver. This means that
+the AFF4 library does not necessarily need to have an ideal Universal
+Resolver, but can approximate this by use of a local resolver. The
+local resolver can be {\emph primed} in advance by the user, by
+loading various volumes which may be needed to resolve internal
+references.
 
@@
 
-%% FIXME - hashing may apply to any segment not just bevies
-%Bevies can optionally be individually hashed, allowing for detection
-%of modifications. The attribute \texttt{hash\_type} specifies the type
-%of hash used (implementations will typically support md5,sha1 and
-%sha256, but can define others). The hashes are stored in any number of
-%segments within the stream, of arbitrary name. The hash segment names
-%is provided by using the \texttt{hash\_name} stream attribute.
-
-%Hashes are calculated in an identical way to that defined by AFF1's
-%crypto layer\cite{garfinkel:affcrypto}---that is, the full name of the
-%segment (including stream component) is followed by a single NULL
-%byte, followed by the (uncompressed) content of the chunk. The
-%inclusion of the segment name ensure that even identical chunks in
-%content will have a different hash and prevents reordering attacks.
-
-%Each hash segment is a list of a chunk id encoded in 32 bit
-%little endian integer format followed by the hash (the length of which
-%depends on the hash itself). This approach is illustrated in Figure
-%\ref{hash_index}.
-
-%Hash segments can contain any number of chunk hashes in them. It is
-%recommended that a hash segment be created when a volume is closed to
-%cover all the chunks within that volume. This allows for hashes to be
-%verified even if some volumes are missing. The AFFLIB implementation
-%may update a hash of a chunk previously set by an earlier hash
-%segment---this would be required when a chunk is updated. The hash
-%segment can then be signed to assure the integrity of the updated segment.
-
-%\begin{figure}[tb]
-%  \begin{center}
-%  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
-%  \caption{The structure of a hashing segment. The attributes {\em
-%  hash} and {\em hash\_type} declare the hash segment. The segment
-%  contains a list of integer encoded chunk id followed by an unencoded
-%  hash for each chunk.}
-%  \label{hash_index}
-%  \end{center}
-%\end{figure}
-
-%This design differs from AFF1, which creates a separate segment for
-%each hash. AFF4 uses headers that are more verbose than the AFF
-%segment headers. Therefore, creating a new segment for each hash value
-%would result in too much overhead: Packing the hashes into a few
-%segments it is possible to reduce this overhead to a minimum.
-
 \subsection{The Map Stream}
@@
 
-%\Subsection{The Overlay Stream}
-%The Overlay Stream Is A Variant Of The {\Em Image} Stream That
-%Provides A Layer Of Indirection Between Requests Made For Specific
-%Bytes (Or Chunks Or Bevys) And The Actual Location Where The Requested
-%Information Is Stored. The Main Utility Of The Overlay Stream Is In
-%Providing Transparent Access To Legacy Forensic Formats, Such As Ewf.
-
-%To Access These Legacy File Formats We Only Need To Have An Index Of
-%Each Chunk Offset, And Then Directly Use The Overlayed Files.  The {\Em
-%Ewf2aff} Tool Provides Such An Overlay Feature For Ewf Volumes By
-%Employing Libewf To Scan The Internal Data Structures.
-
-%Overlay Stream Require That A \Texttt{Chunk\_Size} Attribute Be Specified. In
-%Addition The {\Em Compression\_Type} Attribute May Be Specified. The
-%Attribute Should Contain An Integer Corresponding To A Suitable Zip
-%Compression Scheme (E.G. Deflate). If Not Specified It Is Assumed
-%That No Compression Is Used.
-
-%The ``Target'' Stream Attribute May Be Specified Multiple Times And
-%Refer To External Files (See Section \Ref{Target_Reference}). There
-%Can Be A Number Of Overlay Segments Named ``Stream Name/Overlay.00''
-%Which Are All Merged To A Single Overlay. If More Than One Overlay Is
-%Present, The Attribute {\Em Overlays} Must Be Specify How Many.
-
-%The Overlay Segment Contains A Series Of Lines With Comma Delimited,
-%Decimal Encoded, 64 Bit Integers Representing Chunk Number, Target
-%Offset, Target Length And Target Number. Target Number Is An Index
-%Into The Targets Defined In The Attributes (Count Starts With Zero).
-%Chunks Are Then Fetched From The Specified Target And Decompressed If
-%Necessary.
-
-%It Is Possible To Refer To The Current File By Providing A Target Of
-%``File://.''. This Can Be Useful For Modifying An Existing Ewf File To
-%Be A Aff4 File. Since Zip Files Are Normally Read From The End Of The
-%File, But Ewf Are Both Read From The Front Of The File, It May Be
-%Possible To Append A Aff4 Central Directory To The End Of An Ewf File
-%Without Interfering With The Overlayed File. In This Case The
-%Converted File Can Still Be Used As An Ewf File Without Change, But
-%Could Also Be Used As An Aff4 File.
-
-%In Image Streams Some Chunks May Compress Very Well. In That Case, The
-%Overheads Introduced By The Zip File Header For Each Chunk In The
-%Image Stream May Become Unacceptably High. In That Case Its Possible
-%To Use An {\Em Overlay Stream} To Coalesce Chunks Into Larger
-%Segments. This Approach Is Illustrated In Figure \Ref{Overlay}.
-
-%\Begin{Figure}[Tb]
-%  \Begin{Center}
-%  \Mbox{\Epsfxsize=0.6\Columnwidth \Epsffile{Overlay.Eps}}
-%  \Caption{The Use Of An Overlay Stream To Overlay An Image With
-%Coalesced Chunks.}
-%  \Label{Overlay}
-%  \End{Center}
-%\End{Figure}
-
-%In The Above Figure A Segment Is Written To The Aff4 Volume Which
-%Contains A Number Of Chunks Back To Back. An Overlay Stream Then Uses
-%Direct References To Each Compressed Chunk Within The Coalesced
-%Segment. The Target Is Then Specified As The Segment Name. Although
-%This Technique Is Valid, It Is Generally Discouraged Since It Breaks
-%The Appealing Simplicity Of The Standard Image Stream.  It Is No
-%Longer Possible To Simply Unzip All The Chunks And Concatenate Them
-%Together To Recover The Original Image.
-
 \subsection{The HTTP Stream}
@@
 
-A Link is an object which has a single attribute
-\emph{aff4:target}. When the Link object is opened, it returns the
-object named by this attribute. This allows images with complex names
-to be referred to via short, meaningful names. Note that in practice
-both Image Streams and Link Objects are created by the imaging tools,
-so as far as the operators are concerned they can always refer to the
-Image stream via the Link name.
+A Link object has a \emph{aff4:target} attribute. When the Link object
+is opened, the object named by this attribute is returned. This allows
+images with complex names to be referred to via short, meaningful
+names. In practice both Image Streams and Link Objects are
+automatically created by imaging tools, so users can always refer to
+the Image stream via the simplified Link name.
 
@@
 for multiple identities to sign the same data.
-
-%\subsection{Encryption Schemes}
-%Our implementation provides a number of encryption schemes.
-
-%\subsubsection{Null Encryption}
-%This scheme ({\em scheme=null}) specifies no encryption at all. This
-%may be useful for testing an implementation, but obviously provides
-%no real security.
-
-%\subsubsection{AES-SHA-PSK}
-%This scheme ({\em scheme=aes-sha-psk}) uses AES for block level
-%encryption, with the following SHA based ESSIV scheme, and Pre-Shared
-%Key (passphrase).
-
-%When creating a new encryption stream, a master key is generated using
-%the first 128 bits of a SHA1 hash of the Pre-Shared-Key appended to a
-%64 bit random salt:
-%\begin{eqnarray}
-%Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
-%\end{eqnarray}
-
-%For each chunk, the chunk IV is obtained by taking the first 128 bits
-%of the SHA1 hash of the chunk id encoded as a 32 bit little endian
-%integer, followed by the master key:
-%\begin{eqnarray}
-%IV = \left | SHA1(chunk\_id + Key_{Master}) \right | _{128}
-%\end{eqnarray}
-
-%The block is then encrypted or decrypted using AES in CBC mode with
-%this IV.
-
-%The salt is specified as a stream attribute named {\em aff4:salt} in
-%base64 encoding. Implementations are free to specify any API for
-%passing the Pre-Shared Key, for example the PSK can be specified in a
-%configuration file, typed in at the terminal, passed in using process
-%environment variables, or encoded within the AFF4 URL itself.
-
-%\section{Private Objects}
-%Often forensic software needs to store internal state, such as the
-%current state of the GUI, or internal data structures which may be
-%required for caching.
-
-%Currently, software create additional cache files to maintain this
-%information. It is advantageous to be able to store these in the
-%evidence volume itself. The evidence volume can then be re-opened by
-%the software at a later stage and private data can be retrieved and
-%used directly. Such private application data can be stored in a
-%Private object. There is no specification of what can be stored in the
-%private object, and applications with private objects should be able
-%to store arbitrary segments in any format at all. Segments URNs should
-%be prefixed by the private object's URN, and the stream type should be
-%specified as ``private''.
-
-%It is recommended that applications create attributes for the private
-%stream so they can reliably recognize their own private stream, and
-%identify the version of the application which created it. The AFF4
-%API can be used to render access to private segments and private
-%attributes.
 
@@
 This command requests an image be created directly on the evidence
-server (it will be uploaded using webdav). The image is signed using
+server (it will be uploaded using WebDav). The image is signed using
 Alice's certificate and key (which might need to be unlocked). Note
@@
 or RAID-0. Previously, if disks were acquired independently, they
-would need to be analysed using a tool which was able to reassemble
+would need to be analyzed using a tool which was able to reassemble
 RAID devices.
@@
 
-%\subsection{In-place encryption of AFF4 volumes}
-%With the large size of a typical forensic image it is often difficult
-%to have to create a second copy of a volume which needs to be
-%encrypted. It is usually more convenient to encrypt the file in place,
-%overwriting the plain text data with the equivalent encrypted data.
-
-%This can be achieved using the technique shown in
-%Figure~\ref{inplace_crypto}. The first step is to encrypt the first
-%block from the front of the file and add it to the end of the archive
-%as a new encrypted segment. Then a ZipFile header is written to the
-%front which encapsulates the rest of the plain text volume. The plain
-%text volume is then encrypted in place. Finally a map stream is used
-%to re-map the first and second fragments into place.
-
-%\begin{figure}[tb]
-%  \begin{center}
-%  \mbox{\epsfxsize=0.8\columnwidth \epsffile{inplace.eps}}
-
-%  \caption{The process of encrypting a volume in place. The first
-%  block from the volume is encrypted and appended to the end of the
-%  volume as a new segment (A). Now a zip file header can be written in
-%  its place to form the remainder of the volume as segment B. The rest
-%  of the volume is encrypted in-place, block by block (Step 4). Now we
-%  create a Map stream which combines segment A and B into a single
-%  logical stream (M). An Encrypted stream which is backed by this map
-%  stream (E) is now added. Finally a Central Directory is appended to
-%  the end of the new volume. We end up with a new Zip File which
-%  contains our old volume stored within an encrypted stream.}
-
-%  \label{inplace_crypto}
-%  \end{center}
-%\end{figure}
-
 \subsection{Cryptographic management of evidence}
@@
 different access to the different streams. However, the analysts may
-still store the results of their analysis in an unencrypted form, or
+still store the results of their analysis in an un-encrypted form, or
 assign others permissions to decrypt their analysis results, without
@@
 underlying data.
-
-%\subsection{Sharing evidence}
-%When managing a large corpus of evidence it is common to require these
-%to be shared across the network. Many laboratories create a
-%centralized Network Attached Storage (NAS) facility to retain their
-%evidence corpus. This is typically shared via some file sharing
-%protocol, commonly Windows sharing (CIFS) or NFS.
-
-%It is advantageous to be able to move evidence around between several
-%servers and locations. For example, consider a disk image stored in a
-%Volume with a given URN. Normally, when a user wishes to open the
-%image, they will specify the path and filename of the volume which
-%contains it. The library will then open the volume and retrieve the
-%stream.
-
-%The AFF4 model is different however. In AFF4 metadata is managed in a location
-%independant manner 
-%through a Universal Resolver (perhaps implemented in a database). Relevant information
-%is found not only in Volumes, but may additionally come from other sources.
-%This means that in the
-%above example, the user will query the resolver about the location of
-%the volume. The resolver returns a URL to the current location of the
-%volume. The user's library will then open a HTTP Stream instance to
-%the location returned and the volume may be opened directly over the
-%intra-web connection.
-
-%If the volume needs to be moved, the volume's \emph{aff4:stored}
-%attribute can be adjusted to the new location, and all clients will
-%automatically know where to find it. Furthermore, if several copies of
-%the same volume exist, the resolver may have multiple values for the
-%\emph{aff4:stored} property, some locations may be faster to access than
-%others. The Resolver can then return the most suitable attribute based
-%on locality.
-
-%Another interesting side effect of the sharing model is that it is far
-%more efficient to allow remote analysts to peruse subsets of the
-%evidence. For example, Alice is in the laboratory and wants Bob who is
-%on a remote site to view several files she found on her
-%image. Traditionally she would either need to copy the entire volume
-%and send it to him. Or maybe extract the files she thinks are relevant
-%and send those to him.
-
-%With the new AFF4 architecture, Bob can open the volume directly,
-%using the HTTP protocol, and view the files he needs. Very little
-%extra data is actually transferred in excess of the data he wishes to
-%see. This is a much more efficient way to allow remote analysts to
-%collaborate.
 
@@
 \end{document}
-
-% BLS - The evidence set as identified by a particular UUID is no longer really in play 
-% as below. Evidence can be grouped into a case by defining a case object and stating
-% what objects (or volumes) are a part of it. Or it could be grouped into someting more 
-% arbitraty depending on what you want - its all just objects and relationships.
-
-%\item \emph{An Evidence Set UUID} is a unique identifier generated
-%  according to RFC4122\cite{RFC4122} for each new Evidence
-%  Set. Multiple Evidence Volumes that make up the same Evidence Set
-%  should all have the same UUID. When an Evidence Set is opened all its volumes are
-%  loaded, their UUIDs are checked, and all of their contents are
-%  logically merged so as to be visible as if they were all found in
-% the same volume.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2009-06-24 python/aff4.py
Started implementing python aff4 implementation. Ignore-this: cb66e19703fe23e8346dca571023289d

@@ -2,10 +2,14 @@
 
 import unittest
 import uuid
-import urllib, os, re, struct, zlib
+import urllib, os, re, struct, zlib, time
+import StringIO
+from zipfile import ZIP_STORED, ZIP_DEFLATED, ZIP64_LIMIT, structCentralDir, stringCentralDir, ZIP_FILECOUNT_LIMIT, structEndArchive64, stringEndArchive64, structEndArchive, stringEndArchive
+import threading, mutex
+
 
 NAMESPACE = "aff4:"
-VOLATILE_NS = "aff2volatile:"
+VOLATILE_NS = "aff4volatile:"
 FQN = "urn:" + NAMESPACE
 
 #/** These are standard aff4 attributes */
@@ -22,13 +26,13 @@ AFF4_VOLATILE_COMPRESSED_SIZE = VOLATILE_NS + "compress_size"
 AFF4_VOLATILE_CRC             = VOLATILE_NS + "crc32"
 AFF4_VOLATILE_COMPRESSION     = VOLATILE_NS + "compression"
 AFF4_VOLATILE_FILE_OFFSET     = VOLATILE_NS + "file_offset"
+AFF4_VOLATILE_DIRTY           = VOLATILE_NS + "dirty"
 
 #/** Image attributes */
 AFF4_CHUNK_SIZE =NAMESPACE + "chunk_size"
 AFF4_COMPRESSION =NAMESPACE + "compression"
 AFF4_CHUNKS_IN_SEGMENT =NAMESPACE + "chunks_in_segment"
 AFF4_DIRECTORY_OFFSET =VOLATILE_NS + "directory_offset"
-AFF4_DIRTY       =VOLATILE_NS + "dirty"
 
 #/** Link, encryption attributes */
 AFF4_TARGET= NAMESPACE + "target"
@@ -57,8 +61,15 @@ AFF4_IDENTITY         ="identity"
 #/** These are various properties */
 AFF4_AUTOLOAD         =NAMESPACE +"autoload"
 
-def DEBUG(fmt, *args):
-    print fmt % args
+_DEBUG = 1
+_INFO = 5
+_WARNING = 10
+
+VERBOSITY=_INFO
+
+def DEBUG(verb, fmt, *args):
+    if verb >= VERBOSITY:
+        print fmt % args
 
 import urllib
 def escape_filename(filename):
@@ -73,9 +84,20 @@ def fully_qualified_name(filename, context_name):
 
     return filename
 
+def relative_name(filename, context_name):
[…]

### 2009-06-24 python/aff4imager.py

@@ -1,0 +1,56 @@
+#!/usr/bin/python
+
+from aff4 import *
+import optparse
+
+parser = optparse.OptionParser()
+parser.add_option("-i", "--image", default=None,
+                  action='store_true',
+                  help = "Imaging mode")
+
+parser.add_option("-o", "--output", default=None,
+                  help="Create the output volume on this file or URL (using webdav)")
+
+(options, args) = parser.parse_args()
+
+if not options.image:
+    print "You must specify a mode (try -h)"
+    sys.exit(1)
+
+if options.image:
+    ## Store the volume here:
+    output = options.output
+    if "://" not in output:
+        output = "file://" + output
+
+    ## Make a new volume
+    volume_fd = ZipVolume(None, "w")
+    volume_urn = volume_fd.urn
+    oracle.add(volume_urn, AFF4_STORED, output)
+    volume_fd.finish()
+    oracle.cache_return(volume_fd)
+
+    for image in args:
+        ## Make an image
+        image_fd = Image(None, "w")
+        oracle.add(image_fd.urn, AFF4_STORED, volume_urn)
+        image_urn = image_fd.urn
+        image_fd.finish()
+
+        ## Make sure its a fully qualified url
+        if "://" not in image:
+            image = "file://%s" % image
+            
+        in_fd = oracle.open(image)
+
+        while 1:
+            data = in_fd.read(1024 * 1024)
+            if not data: break
+
+            image_fd.write(data)
+
+        image_fd.close()
+        oracle.cache_return(in_fd)
+    
+    volume_fd = oracle.open(volume_urn, 'w')
+    volume_fd.close()
